### 2016-10-27 paper.tex
First outline

\documentclass{article}

\usepackage{pdfcomment}
\usepackage[margin=1in]{geometry}

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
\newcommand{\todo}[2]{\pdfmargincomment[color=red,author=#1,open=true]{#2}}

\title{Reproducibility of HCP pipelines across operating systems}

\author{Lalet Scaria, Lindsay B. Lewis, Tristan Glatard}

\begin{document}

\maketitle

\abstract{This paper is a reproducibility study of the work in~\cite{glasser2015multi}.}

\section{Introduction}

Groenschild \emph{et al.} first identified the effect of operating
systems on Freesurfer
results~\cite{Gronenschild2012}. In~\cite{10.3389/fninf.2015.00012} we
quantified this effect on some of the main neuroimaging pipelines
including several FSL pipelines, CIVET and Freesurfer. In this work we
aim at evaluating this effect on the work in~\cite{glasser2015multi}.

\section{Materials and Methods}

\subsection{General framework}

We installed the pipelines in three versions of Linux using Docker containers.

Data: report on versions and subjects used.

We integrated the pipelines in CBRAIN through the Boutiques system.

We executed the pipelines on Compute Canada.

\subsection{Minimal pre-processing pipelines}

Described in~\cite{glasser2013minimal}.

Report on versions used and any other specificity such as dependencies (e.g. explain why they depend on gcc).

\subsection{Results comparisons}

Describe the scripts that will be written to compare results.

\section{Results}

Show results of the comparison.

\section{Discussion}

Discuss the results.

\section{Conclusion}

Conclude and highlight future work.

\section{Acknowledgments}

CBRAIN team. Compute Canada.

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}

### 2016-10-28 paper.tex
Added a comment on FSL installation
More comments in sections 2 and 3.

@@
 
-\author{Lalet Scaria, Lindsay B. Lewis, Tristan Glatard}
+\author{Lalet Scaria, Lindsay B. Lewis, Alan C. Evans, Tristan Glatard}
 
@@
 
-\subsection{General framework}
-
-We installed the pipelines in three versions of Linux using Docker containers.
+\subsection{Dataset}
 
 Data: report on versions and subjects used.
-
-We integrated the pipelines in CBRAIN through the Boutiques system.
-
-We executed the pipelines on Compute Canada.
 
@@
 
-Report on versions used and any other specificity such as dependencies (e.g. explain why they depend on gcc).
+We installed the pipelines in three versions of Linux using Docker
+containers. Refer to (immutable) version of containers used. Make sure
+the exact same containers are used for all subjects (no rebuild as
+rebuild would trigger package update). For each execution, we
+monitored the list of packages used with versions.
+
+We integrated the pipelines in CBRAIN through the Boutiques system.
+
+\paragraph{Hardware.} We executed the pipelines on Compute Canada
+using the Guillimin cluster of Calcul Qu\'ebec. We monitored the
+hardware used (CPU and memory).
+
+\paragraph{FSL installation.} We developed our own FSL installer to be
+able to customize easily the FSL version and build (the script
+provided by FSL installs the latest version by default). We used FSL
+5.0.6, built for CentOS5 and CentOS6.
+
+\paragraph{Execution.} The pipeline was launched twice in the same
+environment to detect any use of pseudo-random numbers. We
+made sure that the results produced by these two executions were
+identical.
 
@@
 \section{Results}
+
+Show lists of packages used with version. Focus on important
+differences (e.g.: python interpreter, gcc if relevant) and explain
+why these packages are important.
 

### 2016-11-01 paper.tex
Added references

@@
 aim at evaluating this effect on the work in~\cite{glasser2015multi}.
+
+The operating system is defined here as a consistent set of software
+packages organized in a trusted repository.
+
+The operating system is not the only part of the computational
+environment that may hamper reproducibility. The work
+in~\cite{diethelm2012limits} mentions reproducibility issues coming
+from parallelization.
+
+Reproducibility has several other aspects, see, e.g.,
+\url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}/
 

### 2017-01-06 paper.tex
Added references

@@
 Reproducibility has several other aspects, see, e.g.,
-\url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}/
+\url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}
+
+For an overview of variability accross analysis methods in fMRI, see \url{http://journal.frontiersin.org/article/10.3389/fnins.2012.00149/full}.
+
+For a general overview in fMRI, see \url{http://www.nature.com/nrn/journal/vaop/ncurrent/box/nrn.2016.167_BX3.html}
 

### 2017-01-07 paper.tex
Some comments on safeguards

@@
 
+Safeguards:
+\begin{itemize}
+\item Checksums to detect file corruption.
+\item Each task works on its own copy of the input files to avoid race conditions in case pipeline modifies its inputs.
+\item Hardware information is captured to make sure that differences are not coming from, e.g., different CPUs (when on cluster).
+\item Checksum of Docker container is recorded to make sure that the same container was used all over.
+\item Multiple runs are executed per condition to detect potential intra-condition differences (e.g., use of pseudo-random numbers).
+\end{itemize}
+
+Describe how metrics (distances?) are associated to file types (based on regexp) and used to compute distance matrices.
+
 \section{Results}

### 2017-01-07 paper.tex
Some comments on pipeline analysis

@@
 
-\subsection{Results comparisons}
+\subsection{File comparisons}
 
@@
 \item Hardware information is captured to make sure that differences are not coming from, e.g., different CPUs (when on cluster).
-\item Checksum of Docker container is recorded to make sure that the same container was used all over.
+\item Checksum of Docker container is recorded to make sure that the same container was used all over. Also, list of packages with versions is printed for every task. 
 \item Multiple runs are executed per condition to detect potential intra-condition differences (e.g., use of pseudo-random numbers).
@@
 Describe how metrics (distances?) are associated to file types (based on regexp) and used to compute distance matrices.
+Including metrics to implement the safeguards.
+Report comparison algorithm (see Google Drive).
+
+\subsection{Pipeline analysis}
+
+Once differences between conditions are quantified, we automatically identify the steps in the pipeline responsible for such differences as follows:
+\begin{itemize}
+\item Capture the execution trace using reprozip (based on system-call interception).
+\item Query the reprozip trace to retrieve the executable(s) that wrote in a given file.
+\end{itemize}
+
+% Could also identify the steps in the pipeline that amplified/reduced differences.
 

### 2017-05-25 paper.tex
Added a few notes
Added not in intro
Added some indicators on subsections

--- 2017-01-06 paper.tex
@@
 
-\author{Lalet Scaria, Lindsay B. Lewis, Alan C. Evans, Tristan Glatard}
+\author{Lalet Scaria, Lindsay B. Lewis, Najmeh Khalili-Mahani, Alan C. Evans, Tristan Glatard}
 
@@
 Reproducibility has several other aspects, see, e.g.,
-\url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}
-
-For an overview of variability accross analysis methods in fMRI, see \url{http://journal.frontiersin.org/article/10.3389/fnins.2012.00149/full}.
-
-For a general overview in fMRI, see \url{http://www.nature.com/nrn/journal/vaop/ncurrent/box/nrn.2016.167_BX3.html}
+\url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}/
 
@@
 5.0.6, built for CentOS5 and CentOS6.
+
+\paragraph{Algorithms}
+A brief description of the algorithms used in the script for analysis
+
+\paragraph{Softwares and Libraries}
+A brief description of the softwares and libraries used
 

### 2017-06-14 paper.tex
Added Freesurfer installation details

@@
 
+\subsection{FreeSurfer Docker container}
+
+For the creation of Docker container contianer the FreeSurfer Pipelines, there are a few prerequisites. They are
+
+\begin{itemize}
+  \item 64 Bit Operating System (We chose Centos 5, 6 and 7)
+  \item FMRIB Software Library(a.k.a FSL) version 5.0.6 (An installation script was developed for this)
+  \item Freesurfer version 5.3.0-HCP
+  \item Connectome Workbench
+\end{itemize}
+
+Out of the images having PreFreeSurfer and environment variables available in Dockerhub(https://hub.docker.com/u/bigdatalabteam/dashboard/), we installed FreeSurfer.
+FreeSurfer(ftp://surfer.nmr.mgh.harvard.edu/pub/dist/freesurfer/5.3.0-HCP/freesurfer-Linux-centos4\_x86\_64-stable-pub-v5.3.0-HCP.tar.gz) was used for installation.
+SetUpFreeSurfer script should be sourced before running the FreeSurferPipelineBatch script.
+There were a few missing libraries that are added using yum to the base image o f the centos since these libraries are used by FreeSurfer Pipeline.
+The images which are input to the FreeSurfer should be processed using PreFreeSurfer.
+FreeSurfer needs a license number which is obtained from the FreeSurfer site and it is essential for smooth functioning of FreeSurfer.
+
 \subsection{Results comparisons}

### 2017-06-14 paper.tex
PreFreesurfer installation

@@
 
-\subsection{FreeSurfer Docker container}
+\paragraph{HCP Pipelines}
 
-For the creation of Docker container contianer the FreeSurfer Pipelines, there are a few prerequisites. They are
+For the creation of Docker container contianer the HCP Pipelines, there are a few prerequisites. They are
 
@@
 
+\paragraph{PreFreeSurfer Installation}
+From the base image of the centos versions 5, 6 and 7, we created docker containers containing all the necessary libraries and softwares needed for the PreFreeSurfer Pipelines.A list of software packages installed in the docker container is available at (https://github.com/lalet/Dockerfiles/blob/master/Dockerfile.exec-centos5.fslbuild-centos5).
+CentOS-5 , yum repos are no longer supported, so we had to bypass the repo url to enable the software installation.
+Numpy library had to be installed separately after RPM rebuilt command since normal installation of numpy was throwing error.
+To install FMRIB Library a separate installation script was written (https://github.com/lalet/Dockerfiles/blob/master/myFslInstallerScript.sh) which would install the required FMRIB Library based on the input parameters.
+HCP pipeline version 3.19.0 is used for the study.
+Humanconnectome workbenchversion 1.2.3 is also installed in the docker directory.
+Environment variables needed for the functioning of FMRBI and HCP Pipelines are added later on.
+A few scripts are also added to the image. They are,
+\begin{itemize}
+  \item monitor.sh - For monitoring the hardware details and software library versions
+  \item create-execution.sh for keeping the input directories unmodified and thus preventing modification of input files.
+  \item checksums.sh - To make sure that the files are not corrupted while transferring or processing the subjects.
+  \item command-line-script.sh - For submitting the PreFreeSurfer Pipeline script with the right paramets.
+\end{itemize}
+
+\paragraph{FreeSurfer Installation}
 Out of the images having PreFreeSurfer and environment variables available in Dockerhub(https://hub.docker.com/u/bigdatalabteam/dashboard/), we installed FreeSurfer.
@@
 SetUpFreeSurfer script should be sourced before running the FreeSurferPipelineBatch script.
-There were a few missing libraries that are added using yum to the base image o f the centos since these libraries are used by FreeSurfer Pipeline.
+There were a few missing libraries that are added using yum to the base image of the centos since these libraries are used by FreeSurfer Pipeline.
 The images which are input to the FreeSurfer should be processed using PreFreeSurfer.

### 2017-06-14 paper.tex
Updated text
Added bullet points

@@
 \section{Materials and Methods}
-
+\begin{itemize}
+  \item{verifyFiles Script}
+  \item{Input parameters}
+  \item{Working}
+  \item{Reference File}
+  \item{SQLite}
+  \item{Reprozip}
+  \item{Query to find out the process that created a particular file}
+  \item{Ignored files}
+  \item{checksums-before-after}
+  \item{monitor.txt}
+  \item{metrices}
+  \item{Describe how the reference file was processed,"docker run -it --rm --security-opt seccomp=unconfined -w /data -v /home/lscaria/BigDataLab:/data 8eeec0517441 /bin/bash"}
+  \item{hardware-monitor metric}
+  \item{Distance using fslmats,nibabel}
+  \item{corruption check flag}
+\end{itemize}
 \subsection{Dataset}
@@
 Environment variables needed for the functioning of FMRBI and HCP Pipelines are added later on.
-A few scripts are also added to the image. They are,
+Below listed scripts are also added to the image. They are,
 \begin{itemize}

### 2017-06-22 paper.tex
Ack
Added os and hcp details

@@
 Groenschild \emph{et al.} first identified the effect of operating
-systems on Freesurfer
-results~\cite{Gronenschild2012}. In~\cite{10.3389/fninf.2015.00012} we
-quantified this effect on some of the main neuroimaging pipelines
+systems on Freesurfer results~\cite{Gronenschild2012}. In~\cite{10.3389/fninf.2015.00012} 
+we quantified this effect on some of the main neuroimaging pipelines
 including several FSL pipelines, CIVET and Freesurfer. In this work we
@@
 \begin{itemize}
+\subsection{Operating System and Applications} 
+Since we are conducting a study on the influence of the changes in operating system and the underlying libraries on the biological images,we chose different versions of CentOS.
+CentOS is a community-driven  free software effort focused on delivering a roubust open source ecosystem.
+CentOS focuses on development of framework for cloud providers, the hosting community, and scientific data processing.\cite{https://www.centos.org/about/} The different versions chosen we selected for our study are centos5.11, centos6.8 and centos7.2.1511.
+These images are available in the official centos \url{https://hub.docker.com/\_/centos/} docker repository.
+
+These operating system versions are base images.
+A base image is the image that is used to create all of your container images\url{https://www.ibm.com/support/knowledgecenter/en/SS6PEW\_9.5.0/com.ibm.help.install.omsoftware.doc/installation/c\_DockerBaseImage.html}.
+On top of the base images, the required software libraries were installed to create containers having all the required prerequisites and softwares.
+
+Human Connectome Project\cite{The WU-Minn Human Connectome Project: An Overview} is a consortium led by Washington University, University of Minnesota, and Oxford University is undertaking a systematic effort to map macroscopic human brain circuits and their relationship to behavior in a large population of healthy adults.
+
   \item{verifyFiles Script}
@@
   \item{monitor.txt}
-  \item{metrices}
+  \item{Metrices}
   \item{Describe how the reference file was processed,"docker run -it --rm --security-opt seccomp=unconfined -w /data -v /home/lscaria/BigDataLab:/data 8eeec0517441 /bin/bash"}
-  \item{hardware-monitor metric}
+  \item{Hardware-monitor metric}
   \item{Distance using fslmats,nibabel}
-  \item{corruption check flag}
+  \item{Corruption check flag}
 \end{itemize}
@@
 
-CBRAIN team. Compute Canada.
+CBRAIN team. Compute Canada(Calcul Quebec).
 

### 2017-06-23 paper.tex
Added some more lines about HCP

@@
 \subsection{Operating System and Applications} 
-Since we are conducting a study on the influence of the changes in operating system and the underlying libraries on the biological images,we chose different versions of CentOS.
+The study on the influence of the changes in operating system and the underlying libraries on the biological images is the topic of this paper.
+For conducting this study, we chose different versions of CentOS.
 CentOS is a community-driven  free software effort focused on delivering a roubust open source ecosystem.
@@
 
-These operating system versions are base images.
+The operating system versions used for the study are base images available in the official repository of centos docker repository.
 A base image is the image that is used to create all of your container images\url{https://www.ibm.com/support/knowledgecenter/en/SS6PEW\_9.5.0/com.ibm.help.install.omsoftware.doc/installation/c\_DockerBaseImage.html}.
@@
 
-Human Connectome Project\cite{The WU-Minn Human Connectome Project: An Overview} is a consortium led by Washington University, University of Minnesota, and Oxford University is undertaking a systematic effort to map macroscopic human brain circuits and their relationship to behavior in a large population of healthy adults.
+The data used for the study is referred in this document as subjects.
+These subjects are publicly available from the human connectome project dataset repository.
+Human Connectome Project\cite{The WU-Minn Human Connectome Project: An Overview} is a consortium led by Washington University, University of Minnesota, and Oxford University is undertaking a systematic effort to map macroscopic human brain circuits andtheir relationship to behavior in a large population of healthy adults.
 

### 2017-08-12 paper.tex

@@
 Reproducibility has several other aspects, see, e.g.,
-\url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}/
+\url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}
+
+James P Turner, speaker in INCF 2016 Track B, is looking at numerical stability on GPUs. And reproducibility.
+
+For an overview of variability accross analysis methods in fMRI, see \url{http://journal.frontiersin.org/article/10.3389/fnins.2012.00149/full}.
+
+For a general overview in fMRI, see \url{http://www.nature.com/nrn/journal/vaop/ncurrent/box/nrn.2016.167_BX3.html}
 
@@
 \subsection{Results comparisons}
+\subsection{File comparisons}
 
 Describe the scripts that will be written to compare results.
+
+Safeguards:
+\begin{itemize}
+\item Checksums to detect file corruption.
+\item Each task works on its own copy of the input files to avoid race conditions in case pipeline modifies its inputs.
+\item Hardware information is captured to make sure that differences are not coming from, e.g., different CPUs (when on cluster).
+\item Checksum of Docker container is recorded to make sure that the same container was used all over. Also, list of packages with versions is printed for every task. 
+\item Multiple runs are executed per condition to detect potential intra-condition differences (e.g., use of pseudo-random numbers).
+\end{itemize}
+
+Describe how metrics (distances?) are associated to file types (based on regexp) and used to compute distance matrices.
+Including metrics to implement the safeguards.
+Report comparison algorithm (see Google Drive).
+
+\subsection{Pipeline analysis}
+
+Once differences between conditions are quantified, we automatically identify the steps in the pipeline responsible for such differences as follows:
+\begin{itemize}
+\item Capture the execution trace using reprozip (based on system-call interception).
+\item Query the reprozip trace to retrieve the executable(s) that wrote in a given file.
+\end{itemize}
+
+% Could also identify the steps in the pipeline that amplified/reduced differences.
 
@@
 
+Mention that this was only possible because the unprocessed data was shared in the first place.
+
+DICOM to Nifti conversion was out of scope and may introduce other issues.
+
 \section{Conclusion}

### 2017-08-12 paper.tex
Added a note

@@
 
-\author{Lalet Scaria, Lindsay B. Lewis, Najmeh Khalili-Mahani, Alan C. Evans, Tristan Glatard}
+\author{Lalet Scaria, Lindsay B. Lewis, Alan C. Evans, Tristan Glatard}
 
@@
 Groenschild \emph{et al.} first identified the effect of operating
-systems on Freesurfer results~\cite{Gronenschild2012}. In~\cite{10.3389/fninf.2015.00012} 
-we quantified this effect on some of the main neuroimaging pipelines
+systems on Freesurfer
+results~\cite{Gronenschild2012}. In~\cite{10.3389/fninf.2015.00012} we
+quantified this effect on some of the main neuroimaging pipelines
 including several FSL pipelines, CIVET and Freesurfer. In this work we
 aim at evaluating this effect on the work in~\cite{glasser2015multi}.
+
+\todo{Link this to Lindsay's HBM 2017 poster and to Redolfi et al.}
 
@@
 \section{Materials and Methods}
-\begin{itemize}
-\subsection{Operating System and Applications} 
-The study on the influence of the changes in operating system and the underlying libraries on the biological images is the topic of this paper.
-For conducting this study, we chose different versions of CentOS.
-CentOS is a community-driven  free software effort focused on delivering a roubust open source ecosystem.
-CentOS focuses on development of framework for cloud providers, the hosting community, and scientific data processing.\cite{https://www.centos.org/about/} The different versions chosen we selected for our study are centos5.11, centos6.8 and centos7.2.1511.
-These images are available in the official centos \url{https://hub.docker.com/\_/centos/} docker repository.
 
-The operating system versions used for the study are base images available in the official repository of centos docker repository.
-A base image is the image that is used to create all of your container images\url{https://www.ibm.com/support/knowledgecenter/en/SS6PEW\_9.5.0/com.ibm.help.install.omsoftware.doc/installation/c\_DockerBaseImage.html}.
-On top of the base images, the required software libraries were installed to create containers having all the required prerequisites and softwares.
-
-The data used for the study is referred in this document as subjects.
-These subjects are publicly available from the human connectome project dataset repository.
-Human Connectome Project\cite{The WU-Minn Human Connectome Project: An Overview} is a consortium led by Washington University, University of Minnesota, and Oxford University is undertaking a systematic effort to map macroscopic human brain circuits andtheir relationship to behavior in a large population of healthy adults.
-
-  \item{verifyFiles Script}
-  \item{Input parameters}
-  \item{Working}
-  \item{Reference File}
-  \item{SQLite}
-  \item{Reprozip}
-  \item{Query to find out the process that created a particular file}
-  \item{Ignored files}
-  \item{checksums-before-after}
-  \item{monitor.txt}
-  \item{Metrices}
-  \item{Describe how the reference file was processed,"docker run -it --rm --security-opt seccomp=unconfined -w /data -v /home/lscaria/BigDataLab:/data 8eeec0517441 /bin/bash"}
-  \item{Hardware-monitor metric}
-  \item{Distance using fslmats,nibabel}
-  \item{Corruption check flag}
-\end{itemize}
 \subsection{Dataset}
@@
 
-\paragraph{Algorithms}
-A brief description of the algorithms used in the script for analysis
-
-\paragraph{Softwares and Libraries}
-A brief description of the softwares and libraries used
-
 \paragraph{Execution.} The pipeline was launched twice in the same
@@
 
-\paragraph{HCP Pipelines}
-
-For the creation of Docker container contianer the HCP Pipelines, there are a few prerequisites. They are
-
-\begin{itemize}
-  \item 64 Bit Operating System (We chose Centos 5, 6 and 7)
-  \item FMRIB Software Library(a.k.a FSL) version 5.0.6 (An installation script was developed for this)
-  \item Freesurfer version 5.3.0-HCP
-  \item Connectome Workbench
-\end{itemize}
-
-\paragraph{PreFreeSurfer Installation}
-From the base image of the centos versions 5, 6 and 7, we created docker containers containing all the necessary libraries and softwares needed for the PreFreeSurfer Pipelines.A list of software packages installed in the docker container is available at (https://github.com/lalet/Dockerfiles/blob/master/Dockerfile.exec-centos5.fslbuild-centos5).
-CentOS-5 , yum repos are no longer supported, so we had to bypass the repo url to enable the software installation.
-Numpy library had to be installed separately after RPM rebuilt command since normal installation of numpy was throwing error.
-To install FMRIB Library a separate installation script was written (https://github.com/lalet/Dockerfiles/blob/master/myFslInstallerScript.sh) which would install the required FMRIB Library based on the input parameters.
-HCP pipeline version 3.19.0 is used for the study.
-Humanconnectome workbenchversion 1.2.3 is also installed in the docker directory.
-Environment variables needed for the functioning of FMRBI and HCP Pipelines are added later on.
-Below listed scripts are also added to the image. They are,
-\begin{itemize}
-  \item monitor.sh - For monitoring the hardware details and software library versions
-  \item create-execution.sh for keeping the input directories unmodified and thus preventing modification of input files.
-  \item checksums.sh - To make sure that the files are not corrupted while transferring or processing the subjects.
-  \item command-line-script.sh - For submitting the PreFreeSurfer Pipeline script with the right paramets.
-\end{itemize}
-
-\paragraph{FreeSurfer Installation}
-Out of the images having PreFreeSurfer and environment variables available in Dockerhub(https://hub.docker.com/u/bigdatalabteam/dashboard/), we installed FreeSurfer.
-FreeSurfer(ftp://surfer.nmr.mgh.harvard.edu/pub/dist/freesurfer/5.3.0-HCP/freesurfer-Linux-centos4\_x86\_64-stable-pub-v5.3.0-HCP.tar.gz) was used for installation.
-SetUpFreeSurfer script should be sourced before running the FreeSurferPipelineBatch script.
-There were a few missing libraries that are added using yum to the base image of the centos since these libraries are used by FreeSurfer Pipeline.
-The images which are input to the FreeSurfer should be processed using PreFreeSurfer.
-FreeSurfer needs a license number which is obtained from the FreeSurfer site and it is essential for smooth functioning of FreeSurfer.
-
-\subsection{Results comparisons}
 \subsection{File comparisons}
@@
 
-CBRAIN team. Compute Canada(Calcul Quebec).
+CBRAIN team. Compute Canada.
 

### 2018-01-26 paper_OHBM.tex
adding the OHBM paper (LaTex files)

\documentclass{article}
Gronenschild2012,Glatard2015,Scaria2017,Glasser2013,Chirigati2016
\usepackage{pdfcomment}
\usepackage[margin=1in]{geometry}
\usepackage{xcolor}
\usepackage{graphicx}

\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
\newcommand{\todo}[2]{\pdfmargincomment[color=red,author=#1,open=true]{#2}}

\title{Numerical error propagation in the HCP structural pre-processing pipelines}

\author{M. Ali Salari, Lalet Scaria, Gregory Kiar, Tristan Glatard}

\begin{document}

\maketitle

\abstract{}

\section{Introduction}

Operating systems are known to have an effect on the results produced by neuroimaging pipelines~\cite{Gronenschild2012, Glatard2015}, 
presumably due to the creation, propagation and amplification of small numerical errors across the pipelines. 
Such errors highlight numerical instability which is also likely to appear as a result of other types of small 
perturbations such as acquisition and parametric noise. In a previous study~\cite{Scaria2017}, 
we showed that pre-processing pipelines of the Human Connectome Project~\cite{Glasser2013} were sensitive to operating system variation (see Figure \ref{fig:1}). 
However, the precise causes of such instabilities and the path along which they propagate in the pipelines are unclear. 
We present a technique to identify the processes in a pipeline that create numerical errors along the execution, 
and we apply this technique to the HCP structural pre-processing pipelines.

\begin{figure}
  \includegraphics[width=\linewidth]{brain_classificatio}
  \caption{Tissue classification produced by the HCP pre-processing pipelines on subject 105216 (CentOS 6 vs CentOS 7).}
  \label{fig:1}
\end{figure}

 \emph{et al.}

\todo{Link this to Lindsay's HBM 2017 poster and to Redolfi et al.}

\section{Methods}

We randomly selected 6 subjects from the HCP data release S500 and processed them with 
the HCP structural pre-processing pipelines v3.19.0 (PreFreeSurfer and FreeSurfer)~\cite{Glasser2013} 
in CentOS 6.8 and CentOS 7.2 using Docker containers. We collected the provenance trace 
(tree of executed processes and files accessed in read or write mode) for one subject 
using system-call interception as provided by the reprozip tool~\cite{5}. We computed a file difference matrix 
between files produced in each operating system based on file checksums as in~\cite{Scaria2017}.
From the reprozip trace and file error matrix, we classified processes depending whether they created errors, 
removed errors, or did not have any impact. To control the propagation of errors in the pipeline and allow 
for the classification of all processes, process classification was done in an iterative manner in which we 
artificially corrected errors by replacing the original process by a file copy from the results obtained 
in the other operating system. In this way, we were able to pinpoint the origin of the errors measured in the pipeline.

\section{Results}

Among the 117 data files produced by PreFreeSurfer, 23 did not have any error for any subject, 92 had errors 
for all subjects and 2 had errors for 1 subject only. 
Figure \ref{fig:2} shows the annotated provenance graph of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. 
Processes that created errors are shown in red, processes that removed errors are in blue, and other processes are in green. 
Squares denote processes for which the classification is uncertain. Black edges link sub-processes to their parents while 
dashed edges denote file dependencies between processes (green edges: files with no errors; red edges: files with errors; yellow edges: temporary files).  
The processes that introduce errors in PreFreeSurfer are: linear registration with “\emph{FLIRT}” 
(8 occurrences, in ACPC Alignment, BrainExtraction, DistortionCorrection, AtlasRegistration), 
non-linear registration with “\emph{FNIRT}” (3 occurrences, in BrainExtraction and AltasRegistration), 
image warping with “\emph{new_invwarp}” (3 occurrences, in BrainExtraction and AtlasRegistration). 
In addition, errors were observed in image mean and standard-deviation computations with “\emph{fslstats}” (3 occurrences in BiasFieldCorrection), 
and in masked image extrapolation with “\emph{fslmaths}” (1 occurrence, in BiasFieldCorrection). 
Besides, transformation format conversion with “\emph{convertwarp}” (2 occurrences, in DistortionCorrection) was able to remove errors.

\begin{figure}
  \includegraphics[width=\linewidth]{graph}
  \caption{An annotated provenance graph from the PreFreesurfer pipeline. Processes that created errors are in red. 
Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
  \label{fig:2}
\end{figure}

\section{Conclusion}

The numerical instability in the PreFreesurfer HCP pipeline arises mainly from linear and non-linear registration processes 
implemented in FSL FLIRT and FNIRT. These processes need to be reviewed to understand and correct the cause of instabilities. 
In this correction process, accuracy has to be considered in addition to stability. Our technique is able to 
characterize the stability of a pipeline’s components automatically, but it also suffers from limitations as it cannot deal with: 
(1) results that are not written to disk (values processed in memory); (2) temporary files (3) files that are written by more than one process.

\section{Acknowledgments}

\bibliographystyle{plain}
\bibliography{biblio_OHBM}

\end{document}

### 2018-01-30 paper.tex
Fixed minor issues
Merge HBM abstract in paper

@@
 \usepackage[margin=1in]{geometry}
+\usepackage{xcolor}
+\usepackage{graphicx}
 
 \newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
-\newcommand{\todo}[2]{\pdfmargincomment[color=red,author=#1,open=true]{#2}}
+\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
 
-\title{Reproducibility of HCP pipelines across operating systems}
+\title{Numerical error propagation in the HCP structural pre-processing pipelines}
 
-\author{Lalet Scaria, Lindsay B. Lewis, Alan C. Evans, Tristan Glatard}
+\author{M. Ali Salari, Lalet Scaria, Gregory Kiar, Lindsay B. Lewis,
+  Alan C. Evans, Tristan Glatard}
 
@@
 
+Operating systems are known to have an effect on the results produced
+by neuroimaging pipelines~\cite{Gronenschild2012, Glatard2015},
+presumably due to the creation, propagation and amplification of small
+numerical errors across the pipelines.  Such errors highlight
+numerical instability which is also likely to appear as a result of
+other types of small perturbations such as acquisition and parametric
+noise. However, the precise causes of
+such instabilities and the path along which they propagate in the
+pipelines are unclear.  We present a technique to identify the
+processes in a pipeline that create numerical errors along the
+execution, and we apply this technique to the HCP structural
+pre-processing pipelines.
+
 Groenschild \emph{et al.} first identified the effect of operating
@@
 aim at evaluating this effect on the work in~\cite{glasser2015multi}.
-
 \todo{Link this to Lindsay's HBM 2017 poster and to Redolfi et al.}
@@
 
-Reproducibility has several other aspects, see, e.g.,
+Notes:
+\begin{itemize}
+  \item Reproducibility has several other aspects, see, e.g.,
 \url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}
-
-James P Turner, speaker in INCF 2016 Track B, is looking at numerical stability on GPUs. And reproducibility.
-
-For an overview of variability accross analysis methods in fMRI, see \url{http://journal.frontiersin.org/article/10.3389/fnins.2012.00149/full}.
-
-For a general overview in fMRI, see \url{http://www.nature.com/nrn/journal/vaop/ncurrent/box/nrn.2016.167_BX3.html}
+ \item James P Turner, speaker in INCF 2016 Track B, is looking at numerical stability on GPUs. And reproducibility.
+ \item For an overview of variability accross analysis methods in fMRI, see \url{http://journal.frontiersin.org/article/10.3389/fnins.2012.00149/full}.
+ \item For a general overview in fMRI, see \url{http://www.nature.com/nrn/journal/vaop/ncurrent/box/nrn.2016.167_BX3.html}
+\end{itemize}
 
@@
 
-\subsection{Dataset}
+\subsection{Data and Processing}
 
-Data: report on versions and subjects used.
+We randomly selected 6 subjects from the HCP data release S500 and
+processed them with the HCP structural pre-processing pipelines
+v3.19.0 (PreFreeSurfer and FreeSurfer)~\cite{Glasser2013} in CentOS
+6.8 and CentOS 7.2 using Docker containers. We collected the
+provenance trace (tree of executed processes and files accessed in
+read or write mode) for one subject using system-call interception as
+provided by the reprozip tool~\cite{5}.
 
-\subsection{Minimal pre-processing pipelines}
+Notes:
+\begin{itemize}
+\item Point to the Docker images
+\end{itemize}
+
+\paragraph{Pipelines}
 
@@
 
-We installed the pipelines in three versions of Linux using Docker
-containers. Refer to (immutable) version of containers used. Make sure
-the exact same containers are used for all subjects (no rebuild as
-rebuild would trigger package update). For each execution, we
-monitored the list of packages used with versions.
+\paragraph{Processing environment}
 
-We integrated the pipelines in CBRAIN through the Boutiques system.
+See Lalet's thesis. Boutiques, CBRAIN, wrappers scripts, etc.
 
-\paragraph{Hardware.} We executed the pipelines on Compute Canada
-using the Guillimin cluster of Calcul Qu\'ebec. We monitored the
-hardware used (CPU and memory).
-
-\paragraph{FSL installation.} We developed our own FSL installer to be
-able to customize easily the FSL version and build (the script
-provided by FSL installs the latest version by default). We used FSL
-5.0.6, built for CentOS5 and CentOS6.
-
-\paragraph{Execution.} The pipeline was launched twice in the same
+The pipeline was launched twice in the same
 environment to detect any use of pseudo-random numbers. We
@@
 
-\subsection{File comparisons}
-
-Describe the scripts that will be written to compare results.
-
-Safeguards:
 \begin{itemize}
+  \item monitor.sh - For monitoring the hardware details and software library versions
+  \item create-execution.sh for keeping the input directories unmodified and thus preventing modification of input files.
+  \item checksums.sh - To make sure that the files are not corrupted while transferring or processing the subjects.
+  \item command-line-script.sh - For submitting the PreFreeSurfer Pipeline script with the right paramets.
+  \item command-line-script.sh - For submitting the PreFreeSurfer Pipeline script with the right paramets.
   \item Checksums to detect file corruption.
-\item Each task works on its own copy of the input files to avoid race conditions in case pipeline modifies its inputs.
   \item Hardware information is captured to make sure that differences are not coming from, e.g., different CPUs (when on cluster).
@@
 
-Describe how metrics (distances?) are associated to file types (based on regexp) and used to compute distance matrices.
-Including metrics to implement the safeguards.
-Report comparison algorithm (see Google Drive).
+\subsection{Measuring differences}
 
-\subsection{Pipeline analysis}
+ We computed a file difference matrix between files produced in each
+ operating system based on file checksums as in~\cite{Scaria2017}.
+
+Binary and non-binary.
+
+See Lalet's thesis
+
+Describe the scripts that will be written to compare results.
+
+Describe how metrics (distances?) are associated to file types (based
+on regexp) and used to compute distance matrices. Including metrics to
+implement the safeguards. Report comparison algorithm (see Google
+Drive).
+
+\subsection{Pipeline analysis: Characterizing differences}
+
+From the reprozip trace and file error matrix, we classified
+ processes depending whether they created errors, removed errors, or
+ did not have any impact. To control the propagation of errors in the
+ pipeline and allow for the classification of all processes, process
+ classification was done in an iterative manner in which we
+ artificially corrected errors by replacing the original process by a
+ file copy from the results obtained in the other operating system. In
+ this way, we were able to pinpoint the origin of the errors measured
+ in the pipeline.
 
@@
 \end{itemize}
-
-% Could also identify the steps in the pipeline that amplified/reduced differences.
 
@@
 Show results of the comparison.
+
+\subsection{Inter-OS differences}
+
+From Lalet's thesis
+
+Among the 117 data files produced by PreFreeSurfer, 23 did not have any error for any subject, 92 had errors 
+for all subjects and 2 had errors for 1 subject only.
+
+In a previous study~\cite{Scaria2017}, we showed that
+pre-processing pipelines of the Human Connectome
+Project~\cite{Glasser2013} were sensitive to operating system
+variation (see Figure \ref{fig:1}).
+\begin{figure}
+%  \includegraphics{brain\_classification}
+  \caption{Tissue classification produced by the HCP pre-processing
+    pipelines on subject 105216 (CentOS 6 vs CentOS 7). \todo{Represent this in a different way.}}
+  \label{fig:1}
+\end{figure}
+
+\subsection{Pipeline analysis}
+
+Ali's results
+
+Figure \ref{fig:2} shows the annotated provenance graph of the
+PreFreeSurfer pipeline executed on CentOS6 and CentOS7.  Processes
+that created errors are shown in red, processes that removed errors
+are in blue, and other processes are in green.  Squares denote
+processes for which the classification is uncertain. Black edges link
+sub-processes to their parents while dashed edges denote file
+dependencies between processes (green edges: files with no errors; red
+edges: files with errors; yellow edges: temporary files).  The
+processes that introduce errors in PreFreeSurfer are: linear
+registration with “\emph{FLIRT}” (8 occurrences, in ACPC Alignment,
+BrainExtraction, DistortionCorrection, AtlasRegistration), non-linear
+registration with “\emph{FNIRT}” (3 occurrences, in BrainExtraction
+and AltasRegistration), image warping with “\emph{new\_invwarp}” (3
+occurrences, in BrainExtraction and AtlasRegistration).  In addition,
+errors were observed in image mean and standard-deviation computations
+with “\emph{fslstats}” (3 occurrences in BiasFieldCorrection), and in
+masked image extrapolation with “\emph{fslmaths}” (1 occurrence, in
+BiasFieldCorrection).  Besides, transformation format conversion with
+“\emph{convertwarp}” (2 occurrences, in DistortionCorrection) was able
+to remove errors.
+
+\begin{figure}
+  \includegraphics[width=\linewidth]{graph}
+  \caption{An annotated provenance graph from the PreFreesurfer pipeline. Processes that created errors are in red. 
+Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
+  \label{fig:2}
+\end{figure}
 
@@
 
+The numerical instability in the PreFreesurfer HCP pipeline arises mainly from linear and non-linear registration processes 
+implemented in FSL FLIRT and FNIRT. These processes need to be reviewed to understand and correct the cause of instabilities. 
+In this correction process, accuracy has to be considered in addition to stability. Our technique is able to 
+characterize the stability of a pipeline’s components automatically, but it also suffers from limitations as it cannot deal with: 
+(1) results that are not written to disk (values processed in memory); (2) temporary files (3) files that are written by more than one process.
+
 \section{Acknowledgments}
 
-CBRAIN team. Compute Canada.
+CBRAIN team. Compute Canada(Calcul Quebec).
 

### 2018-03-09 paper.tex
Added references and worked on results

@@
 
-We randomly selected 6 subjects from the HCP data release S500 and
-processed them with the HCP structural pre-processing pipelines
-v3.19.0 (PreFreeSurfer and FreeSurfer)~\cite{Glasser2013} in CentOS
-6.8 and CentOS 7.2 using Docker containers. We collected the
-provenance trace (tree of executed processes and files accessed in
-read or write mode) for one subject using system-call interception as
-provided by the reprozip tool~\cite{5}.
+We randomly selected 10 subjects from the HCP data release S500 and processed them on a 
+containerized version of HCP pipelines. The Docker images are built for the HCP structural 
+pre-processing pipelines v3.19.0 (PreFreeSurfer and FreeSurfer)~\cite{Glasser2013} 
+in CentOS 6.8 and CentOS 7.2 for the ease of conducting the experiment and reproducibility issues. 
+We collected the provenance trace (tree of executed processes and files accessed in read or write mode) 
+for one subject using system-call interception as provided by the reprozip tool~\cite{5}.
 
-Notes:
-\begin{itemize}
-\item Point to the Docker images
-\end{itemize}
+\subsection{HCP Minimal preprocessing Pipelines}
 
-\paragraph{Pipelines}
+The Human Connectom Project (HCP)~\cite{Gla13} is a set of pipelines To help extraction of structural, functional or diffusion MRI data
+across a large set of high resolution MR images.these preprocessing pipelines creates results that are available in standard
+volume and combined surface and volume spaces which makes it easier for researchers to compare
+the images across the neuroimaging spectrum.
 
-Described in~\cite{glasser2013minimal}.
+\paragraph{Structural Pipelines} The structural pre-processing consists of PreFreeSurfer, FreeSurfer and PostFreeSurfer. 
+\paragraph{Functional Pipelines} Functional pre-processing consists of fMRIVolume and fMRISurface~\cite{FSL}. 
+\paragraph{Diffusion Preprocessing pipeline} preparing...
 
@@
 
-See Lalet's thesis. Boutiques, CBRAIN, wrappers scripts, etc.
+To ensure that the origin of differences are generated exactly across the operating system updates (inter-os differences) 
+on HCP preprocessing pipelines, some pre-executions have been done in a wrapper script; 
+(a) the checksum of the files are computed in each subject before and after a multiple execution to detect 
+potential intra-condition differences (e.g., use of pseudo-random numbers). 
+(b) Checksum of Docker container is recorded to make sure that the same container was used all over. Also, list of packages 
+with versions is printed for every task. (c) Hardware information is captured to make sure that differences are not coming from, for instance, different CPUs (when on cluster).
 
-The pipeline was launched twice in the same
-environment to detect any use of pseudo-random numbers. We
-made sure that the results produced by these two executions were
-identical.
+HCP preprocessing pipelines are both time-consuming and computationally intensive which might even run for days, depending upon 
+the type and size of dataset. We made use of a web platform (CBRAIN) for managing the containers and data with the help of Boutiques technology. 
+The Canadian Brain Imaging Research Platform (CBRAIN), is a web platform developed at the Montreal Neurological Institute (MNI) to tackle 
+the Big-Data research and heavy computational challenges faced by neuroimaging researchers. ~\cite{DBLP:journals/fini/DasGRSPMSRSKMKR17}.
+CBRAIN supports container technologies like Docker, Singularity and it also extends its support to Boutiques. Thus, 
+with the help of container technologies and Boutiques, applications can be ported to CBRAIN. Any user who has access to this application 
+can try to reproduce the experiments since the data and the application are accessible within the framework. However, one caveat is that, 
+the neuroimaginng pipelines might have different results due to differences in computing platforms on which the images are getting processed ~\cite{10.3389/conf.fninf.2014.18.00076}
+therefore, the execution results of the pipelines across different operating systems are considered to find out 
+the cause of the reproducibility issues using an interposition techniques for the provenance capture.
 
@@
 
- We computed a file difference matrix between files produced in each
- operating system based on file checksums as in~\cite{Scaria2017}.
+We computed a file difference matrix between files produces in each operating system based on file checksum as in~\cite{Scaria2017}.
+ The checksums are computed before and after execution so that corruption check can be done with the use of recorded checksums.
+ We have used MD5 {\url{https://tools.ietf.org/html/rfc1321}} algorithm for calculating the checksum of files. 
+The output of a MD5 algorithm is a 128-bit ``fingerprint" or ``message digest"~\cite{md5}. 
+Though MD5 is not completely secure against collision attack, Repro-tools focus more on the data integrity than security 
+and how fast the algorithm is able to create the checksum. These qualities make MD5 a good choice for checksum generation in Repro-tools.
+\paragraph{} For the files that are identified to have differences, different kind of metrics are used base on the file 
+type to quantify the differences. Normalized root mean square error, Dice Coefficient, Text filter etc. are the 
+various metrics used for quantifying the differences. As defined in \cite{khosrow2017handbook}, ``The Root Mean Square Deviation (RMSD) 
+or root-mean-square error (RMSE) is a frequently used measure of the difference between values predicted by a model or 
+an estimator and the values actually observed". Normalizing the RMSD value makes it easier to compare between models or datasets.
+\paragraph{} Dice Similarity Coefficient can be used as a statistical validation metric for measuring the reproducibility
+ of magnetic resonance images~\cite{Zou2004}. To measure the similarity of image files in reproducibility study, we used dice similarity coefficient. 
+The coefficient ranges from 0.0 to 1. 1 meaning the images are exactly similiar and 0 meaning there is zero similarity.
+\paragraph{} These metric values help us understand how big or small the differences are. Apart from quantifying 
+the differences using type specific metrics, Repro-tools can also be used to trace the provenance of these differences.
+ It is able to identify all the processes and associated parameters that wrote the files having differences. 
+These details about various processes is helpful in debugging the pipelines. This information helps in 
+recreating the processing step by step and also to identify the processes that creates the differences.
 
 Binary and non-binary.
-
-See Lalet's thesis
 
@@
 
-From the reprozip trace and file error matrix, we classified
- processes depending whether they created errors, removed errors, or
- did not have any impact. To control the propagation of errors in the
- pipeline and allow for the classification of all processes, process
- classification was done in an iterative manner in which we
- artificially corrected errors by replacing the original process by a
- file copy from the results obtained in the other operating system. In
- this way, we were able to pinpoint the origin of the errors measured
- in the pipeline.
-
- Once differences between conditions are quantified, we automatically identify the steps in the pipeline responsible for such differences as follows:
-\begin{itemize}
-\item Capture the execution trace using reprozip (based on system-call interception).
-\item Query the reprozip trace to retrieve the executable(s) that wrote in a given file.
-\end{itemize}
+Once differences between conditions are quantified, we automatically identify the steps in the pipeline responsible for such differences.
+ we use the reprozip tool [Chirigati, 2016] to record: (1) the tree of processes executed by the pipeline and
+ (2) the list of files read and written by each process. This information is collected by system call interceptions, 
+through the ptrace system call in Unix-like operating system, and stored in a sqlite database. The database has three 
+tables including table of processes to store information of all the processes which are created by clone() 
+or fork() Linux system calls and the table of opened files which keeps the detail of all files including the pipeline 
+and temporary files. This table has an attribute to specify the file mode, namely files with mode 0X01 and 0X02 are read 
+from and written on respectively; the final table indicated the executed files in the pipeline through execv() system call.
+ In fact, the tree of processes is retrieved by query the root process at the first level and then all its child processes with 
+dependency files in the next levels from database. The process tree of two subjects has tested in the same pipeline conditions
+ which are similar, so we collect such data only for one subject (The term condition refers to the variation of operating systems
+ on which the data processing takes place). Furthermore, the process tree is acyclic “with the respect of this assumption that 
+there is no race conditions between processes to write on the same file.”
+\paragraph{} In the next step, we classify the processes into three categories considering to the obtained process tree 
+and the error matrix file. First, processes that read file(s) without error but write different file(s) create error in 
+the pipeline. The second group of the processes are those which remove errors in pipeline, read file(s) with error and
+ write “nothing or” file(s) without error. Most of the processes are in the third category beyond any fault which
+ read and write file(s) without error. The processes in each category may read or write temporary file(s) that implies 
+the uncertainty of the process.
+\paragraph{} Pipeline is conducted of a bunch of processes sequentially, setting a modification step after classification
+ of processes iteratively enables us to control the propagation of errors in the pipeline correctly. In each iteration
+ we detect all the processes which are create error in the pipeline and then by fixing those processes artificially, 
+we’ll find the other processes in the pipeline that introduce an error. This procedure will be terminated by fixing all the processes. 
+To fix the processes that create error, we replace their dependency files with error by the same file copy from the results 
+obtained in the other operating system. In this way, we were able to pinpoint the origin of the errors measured in the pipeline.
 
@@
 
-From Lalet's thesis
+\paragraph{Binary differences}
 
-Among the 117 data files produced by PreFreeSurfer, 23 did not have any error for any subject, 92 had errors 
-for all subjects and 2 had errors for 1 subject only.
+Among the 117 data files produced by PreFreeSurfer, 21 did not have any error for any subject, 92 had errors 
+for all subjects and 4 had errors for 3 subjects only.
 
@@
 
+\paragraph{}
+Two types of differences can occur in the subjects due to the differences in the operating systems. 
+One is inter-OS difference caused by the operating system library updates and the other type,
+ intra-OS differences occurs as a result of the pseudo-random processes used in the pipelines.
+ An example of a pseudo-random process function is, a random number generator that would get initialized 
+using a seed state. the proposed method can be used to identify both kind of differences. 
+The files that are common to all the subjects only are taken into consideration for comparison. The first step
+ is identification of files with differences in their checksums. This is identified using the checksums that are
+ recorded after the processing. Intra-OS differences are identified using the run-number added as the suffix for
+ the conditions. For example, the two batches of subjects processed under the same condition (CentOS6) are stored 
+as run-1 and run-2. The files belonging to the subjects stored under the above mentioned conditions are treated as intra-OS runs.
+
 \subsection{Pipeline analysis}
-
-Ali's results
 
@@
 are in blue, and other processes are in green.  Squares denote
-processes for which the classification is uncertain. Black edges link
+processes for which the classification is uncertain, due to temporary
+files that were removed during the execution. Black edges link
 sub-processes to their parents while dashed edges denote file
@@
 
+\todo{Check if a given file is written by only one process (add a safeguard in your code).}
+
 \begin{figure}
-  \includegraphics[width=\linewidth]{graph}
+  \includegraphics[width=\linewidth]{images/graph}
   \caption{An annotated provenance graph from the PreFreesurfer pipeline. Processes that created errors are in red. 

### 2018-03-09 paper.tex
Edited the paper during meeting with @ali4006

@@
 \newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
+\newcommand{\reprozip}[0]{ReproZip}
 
@@
 
+% There is a computational challenge
+HCP preprocessing pipelines are both time-consuming and
+computationally intensive which might even run for days, depending
+upon the type and size of dataset.
+
+Any
+user who has access to this application can try to reproduce the
+experiments since the data and the application are accessible within
+the framework. However, one caveat is that, the neuroimaginng
+pipelines might have different results due to differences in computing
+platforms on which the images are getting processed
+~\cite{10.3389/conf.fninf.2014.18.00076} therefore, the execution
+results of the pipelines across different operating systems are
+considered to find out the cause of the reproducibility issues using
+an interposition techniques for the provenance capture.
+
 \section{Materials and Methods}
 
-\subsection{Data and Processing}
+\subsection{Data}
 
-We randomly selected 10 subjects from the HCP data release S500 and processed them on a 
-containerized version of HCP pipelines. The Docker images are built for the HCP structural 
-pre-processing pipelines v3.19.0 (PreFreeSurfer and FreeSurfer)~\cite{Glasser2013} 
-in CentOS 6.8 and CentOS 7.2 for the ease of conducting the experiment and reproducibility issues. 
-We collected the provenance trace (tree of executed processes and files accessed in read or write mode) 
-for one subject using system-call interception as provided by the reprozip tool~\cite{5}.
+We randomly selected 10 subjects from the HCP data release
+S500. \todo{list the subjects.}
 
@@
 
-The Human Connectom Project (HCP)~\cite{Gla13} is a set of pipelines To help extraction of structural, functional or diffusion MRI data
-across a large set of high resolution MR images.these preprocessing pipelines creates results that are available in standard
-volume and combined surface and volume spaces which makes it easier for researchers to compare
-the images across the neuroimaging spectrum.
+The Human Connectome Project (HCP)~\cite{Gla13} developed a set of
+pipelines to help extraction of structural, functional or diffusion
+MRI data across a large set of high resolution MR images. These
+preprocessing pipelines create results that are available in standard
+volume and combined surface and volume spaces which makes it easier
+for researchers to compare the images across the neuroimaging
+spectrum.
 
@@
 
-\paragraph{Processing environment}
+\todo{Write a short summary of what the pipelines do.}
 
-To ensure that the origin of differences are generated exactly across the operating system updates (inter-os differences) 
-on HCP preprocessing pipelines, some pre-executions have been done in a wrapper script; 
-(a) the checksum of the files are computed in each subject before and after a multiple execution to detect 
-potential intra-condition differences (e.g., use of pseudo-random numbers). 
-(b) Checksum of Docker container is recorded to make sure that the same container was used all over. Also, list of packages 
-with versions is printed for every task. (c) Hardware information is captured to make sure that differences are not coming from, for instance, different CPUs (when on cluster).
+\subsection{Processing environment}
 
-HCP preprocessing pipelines are both time-consuming and computationally intensive which might even run for days, depending upon 
-the type and size of dataset. We made use of a web platform (CBRAIN) for managing the containers and data with the help of Boutiques technology. 
-The Canadian Brain Imaging Research Platform (CBRAIN), is a web platform developed at the Montreal Neurological Institute (MNI) to tackle 
-the Big-Data research and heavy computational challenges faced by neuroimaging researchers. ~\cite{DBLP:journals/fini/DasGRSPMSRSKMKR17}.
-CBRAIN supports container technologies like Docker, Singularity and it also extends its support to Boutiques. Thus, 
-with the help of container technologies and Boutiques, applications can be ported to CBRAIN. Any user who has access to this application 
-can try to reproduce the experiments since the data and the application are accessible within the framework. However, one caveat is that, 
-the neuroimaginng pipelines might have different results due to differences in computing platforms on which the images are getting processed ~\cite{10.3389/conf.fninf.2014.18.00076}
-therefore, the execution results of the pipelines across different operating systems are considered to find out 
-the cause of the reproducibility issues using an interposition techniques for the provenance capture.
+We executed the pipelines using Docker
+containers to simplify the deployment of different operating
+system versions on execution platforms. The Docker images were built for the HCP
+ pre-processing pipelines v3.19.0 (PreFreeSurfer and
+FreeSurfer)~\cite{Glasser2013} in CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for reuse \todo{Add URL}. We
+collected the provenance trace (tree of executed processes and files
+accessed in read or write mode) for subject \todo{XXX} using system-call
+interception as provided by the \reprozip tool~\cite{Chirigati2016}.
 
+To support the computational challenge, we used the Canadian Brain
+Imaging Research Platform (CBRAIN)~\cite{cbrain} to manage the
+containers and data. CBRAIN is a Web platform developed at the
+Montreal Neurological Institute (MNI) to tackle the computational
+challenges faced by neuroimaging researchers.  Applications can be
+ported to CBRAIN using the Docker and Singularity container
+technologies through the Boutiques framework~\cite{boutiques}. Our
+CBRAIN HCP plugins are publicly available at \url{github url} for
+installation and re-use in CBRAIN instances.
+
+We processed subjects twice in every condition to identify possible \emph{intra-condition} errors coming from the use of pseudo-random numbers or other external
+factors.
+
+To further ensure that the observed errors are caused by operating system
+updates (inter-os differences) and not by other factors, we
+implemented the following sanity checks:
 \begin{itemize}
-  \item monitor.sh - For monitoring the hardware details and software library versions
-  \item create-execution.sh for keeping the input directories unmodified and thus preventing modification of input files.
-  \item checksums.sh - To make sure that the files are not corrupted while transferring or processing the subjects.
-  \item command-line-script.sh - For submitting the PreFreeSurfer Pipeline script with the right paramets.
-  \item command-line-script.sh - For submitting the PreFreeSurfer Pipeline script with the right paramets.
-  \item Checksums to detect file corruption.
-  \item Hardware information is captured to make sure that differences are not coming from, e.g., different CPUs (when on cluster).
-  \item Checksum of Docker container is recorded to make sure that the same container was used all over. Also, list of packages with versions is printed for every task. 
-  \item Multiple runs are executed per condition to detect potential intra-condition differences (e.g., use of pseudo-random numbers).
+\item The file checksums of every
+subject are computed before and after processing to detect potential corruption during file transfers and other manipulations.
+\item The checksum of the Docker container images is recorded with
+  every execution to identify any unintentional package update. In
+  addition, the list of packages with versions is printed by every
+  task and checked for consistency.
+\item Hardware information is captured to make sure that errors are
+  not coming from different CPU architectures.
+\item Every task is executed on its own copy of the data to ensure
+  that input data is not modified by the pipeline. 
 \end{itemize}
+
+\todo{Check ~\cite{DBLP:journals/fini/DasGRSPMSRSKMKR17}}
 
@@
 
-We computed a file difference matrix between files produces in each operating system based on file checksum as in~\cite{Scaria2017}.
- The checksums are computed before and after execution so that corruption check can be done with the use of recorded checksums.
- We have used MD5 {\url{https://tools.ietf.org/html/rfc1321}} algorithm for calculating the checksum of files. 
-The output of a MD5 algorithm is a 128-bit ``fingerprint" or ``message digest"~\cite{md5}. 
-Though MD5 is not completely secure against collision attack, Repro-tools focus more on the data integrity than security 
-and how fast the algorithm is able to create the checksum. These qualities make MD5 a good choice for checksum generation in Repro-tools.
-\paragraph{} For the files that are identified to have differences, different kind of metrics are used base on the file 
-type to quantify the differences. Normalized root mean square error, Dice Coefficient, Text filter etc. are the 
-various metrics used for quantifying the differences. As defined in \cite{khosrow2017handbook}, ``The Root Mean Square Deviation (RMSD) 
-or root-mean-square error (RMSE) is a frequently used measure of the difference between values predicted by a model or 
-an estimator and the values actually observed". Normalizing the RMSD value makes it easier to compare between models or datasets.
-\paragraph{} Dice Similarity Coefficient can be used as a statistical validation metric for measuring the reproducibility
- of magnetic resonance images~\cite{Zou2004}. To measure the similarity of image files in reproducibility study, we used dice similarity coefficient. 
-The coefficient ranges from 0.0 to 1. 1 meaning the images are exactly similiar and 0 meaning there is zero similarity.
-\paragraph{} These metric values help us understand how big or small the differences are. Apart from quantifying 
-the differences using type specific metrics, Repro-tools can also be used to trace the provenance of these differences.
- It is able to identify all the processes and associated parameters that wrote the files having differences. 
-These details about various processes is helpful in debugging the pipelines. This information helps in 
-recreating the processing step by step and also to identify the processes that creates the differences.
+We computed a file difference matrix between files produced for every
+subject in every operating system based on MD5 checksums~\cite{md5}
+(\url{https://tools.ietf.org/html/rfc1321}) as
+in~\cite{Scaria2017}. MD5 is suitable to identify incidental file
+differences, although it is not cryptographically secure. Files
+produced only in some subjects or in some operating systems were left out.
 
-Binary and non-binary.
-
-Describe the scripts that will be written to compare results.
-
-Describe how metrics (distances?) are associated to file types (based
-on regexp) and used to compute distance matrices. Including metrics to
-implement the safeguards. Report comparison algorithm (see Google
-Drive).
+We quantified the error in the image files using the Dice coefficient and normalized root mean square error (NRMSE) computed as follows:
+\todo{put the equation, see Lalet's thesis}
 
@@
 
-Once differences between conditions are quantified, we automatically identify the steps in the pipeline responsible for such differences.
- we use the reprozip tool [Chirigati, 2016] to record: (1) the tree of processes executed by the pipeline and
- (2) the list of files read and written by each process. This information is collected by system call interceptions, 
-through the ptrace system call in Unix-like operating system, and stored in a sqlite database. The database has three 
-tables including table of processes to store information of all the processes which are created by clone() 
-or fork() Linux system calls and the table of opened files which keeps the detail of all files including the pipeline 
-and temporary files. This table has an attribute to specify the file mode, namely files with mode 0X01 and 0X02 are read 
-from and written on respectively; the final table indicated the executed files in the pipeline through execv() system call.
- In fact, the tree of processes is retrieved by query the root process at the first level and then all its child processes with 
-dependency files in the next levels from database. The process tree of two subjects has tested in the same pipeline conditions
- which are similar, so we collect such data only for one subject (The term condition refers to the variation of operating systems
- on which the data processing takes place). Furthermore, the process tree is acyclic “with the respect of this assumption that 
-there is no race conditions between processes to write on the same file.”
+Once differences between conditions are quantified, we automatically
+identify the steps in the pipeline responsible for such differences.
+We use the \reprozip tool~\cite{reprozip} to record: (1) the tree of
+processes executed by the pipeline and (2) the list of files read and
+written by each process. This information is collected by system call
+interceptions, through the \texttt{ptrace} system call \todo{is it
+  POSIX or UNIX or Linux?}, and stored in a \texttt{SQLite}
+database. The database stores information on all the processes which
+are created by the \texttt{clone()} or \texttt{fork()} Linux
+\todo{check if POSIX} system calls. It also records the files opened by each process (in read, write or execution mode), including the final and
+temporary files.
+
+We reconstruct
+the tree of processes starting from the first process created by the pipeline and retrieving its child processes with dependency files in
+the next levels from database. \todo{Start here.} The process tree of two subjects has
+tested in the same pipeline conditions which are similar, so we
+collect such data only for one subject (The term condition refers to
+the variation of operating systems on which the data processing takes
+place). Furthermore, the process tree is acyclic “with the respect of
+this assumption that there is no race conditions between processes to
+write on the same file.”
 \paragraph{} In the next step, we classify the processes into three categories considering to the obtained process tree 
@@
 obtained in the other operating system. In this way, we were able to pinpoint the origin of the errors measured in the pipeline.
+
+\todo{Check that} We also recorded the provenance trace of the execution using \reprozip. \reprozip identifies all the
+processes and associated parameters that wrote the files having
+differences.  These details about various processes is helpful in
+debugging the pipelines. This information helps in recreating the
+processing step by step and also to identify the processes that
+creates the differences.
 

### 2018-03-13 paper.tex
Updates following meeting with @ali4006

@@
 
-We reconstruct
-the tree of processes starting from the first process created by the pipeline and retrieving its child processes with dependency files in
-the next levels from database. \todo{Start here.} The process tree of two subjects has
-tested in the same pipeline conditions which are similar, so we
-collect such data only for one subject (The term condition refers to
-the variation of operating systems on which the data processing takes
-place). Furthermore, the process tree is acyclic “with the respect of
-this assumption that there is no race conditions between processes to
-write on the same file.”
+We reconstruct the tree of processes starting from the first process
+created by the pipeline and defining its child processes as the ones
+that were created through \texttt{clone()} or \texttt{fork()}. We
+create a process graph from this tree by adding edges corresponding to
+file dependencies between processes. A file dependency is defined
+between processes A and B if a file written by A is read by
+B. Figure~\ref{fig:graph-example} shows an example of a process graph
+constructed at this stage \todo{Build the graph obtained on
+  bin/example.sh and add it to the Figure}.
+\begin{figure}
+%  \includegraphics{brain\_classification}
+  \caption{Process graph constructed from an example brain extraction script.}
+  \label{fig:graph-example}
+\end{figure}
+
+A different process graph may be produced for every subject in every
+execution condition. Among our dataset, we identified 4 types of
+subjects with different numbers of T1w and T2w images. We verified
+that the process graphs were identical for all subjects of the same
+type, for all execution conditions. The remainder of the analysis is
+done separately for each subject type. \todo{Check that}
+
+Cycles may be present in the process graph in case a file was written
+by more than one process.  We remove the cycles in the process graph
+by removing file edges between processes A and B when A's process
+creation timestamp is posterior to B's or when A=B.
+
+\todo{Start from here}
 \paragraph{} In the next step, we classify the processes into three categories considering to the obtained process tree 

### 2018-03-13 bin/example.sh

@@ -1,0 +1,19 @@
+#!/usr/bin/env bash
+
+if [ $# != 1 ]
+then
+    echo "usage: $0 <input_image.nii.gz>"
+    exit 1
+fi
+
+input_image=$1
+bet_output="$(basename ${input_image} .nii.gz)_brain.nii.gz"
+bet_output_binarized="$(basename ${input_image} .nii.gz)_brain_bin.nii.gz"
+
+bet ${input_image} ${bet_output}
+echo "Voxels / volume in brain mask:"
+fslstats ${bet_output} -V
+fslmaths ${bet_output} ${bet_output_binarized}
+echo "Voxels / volume in binarized brain mask:"
+fslstats ${bet_output_binarized} -V
+\rm ${bet_output}

### 2018-03-15 paper.tex

@@
 We randomly selected 10 subjects from the HCP data release
-S500. \todo{list the subjects.}
+S500. \todo{ list the subjects.}
 
@@
  pre-processing pipelines v3.19.0 (PreFreeSurfer and
-FreeSurfer)~\cite{Glasser2013} in CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for reuse \todo{Add URL}. We
+FreeSurfer)~\cite{Glasser2013} in CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}. We
 collected the provenance trace (tree of executed processes and files
-accessed in read or write mode) for subject \todo{XXX} using system-call
+accessed in read or write mode) for each type of subjects using system-call
 interception as provided by the \reprozip tool~\cite{Chirigati2016}.
@@
 technologies through the Boutiques framework~\cite{boutiques}. Our
-CBRAIN HCP plugins are publicly available at \url{github url} for
+CBRAIN HCP plugins are publicly available at \url{https://github.com/glatard/cbrain-plugins-hcp} for
 installation and re-use in CBRAIN instances.
@@
 To further ensure that the observed errors are caused by operating system
-updates (inter-os differences) and not by other factors, we
+updates (\emph{inter-os differences}) and not by other factors, we
 implemented the following sanity checks:
@@
 We quantified the error in the image files using the Dice coefficient and normalized root mean square error (NRMSE) computed as follows:
-\todo{put the equation, see Lalet's thesis}
+\begin{center}
+\begin{equation}
+RMSE = {\sqrt {\frac{1} {n}{\sum\limits_{t = 1}^n {(\hat{y}_{t} - {y}_{t} } })^{2} } }
+\end{equation}
+\end{center}
+
+\begin{center}
+\begin{equation}
+NRMSE = {\frac{RMSE} {y_{max} - y_{min}}}
+\end{equation}
+\end{center}
 
@@
 written by each process. This information is collected by system call
-interceptions, through the \texttt{ptrace} system call \todo{is it
-  POSIX or UNIX or Linux?}, and stored in a \texttt{SQLite}
+interceptions, through the \texttt{ptrace} as a UNIX base system call, and stored in a \texttt{SQLite}
 database. The database stores information on all the processes which
-are created by the \texttt{clone()} or \texttt{fork()} Linux
-\todo{check if POSIX} system calls. It also records the files opened by each process (in read, write or execution mode), including the final and
+are created by the \texttt{clone()} in Linux or \texttt{fork()} in UNIX-like
+operating system calls. It also records the files opened by each process (in read, write or execution mode), including the final and
 temporary files.
@@
 B. Figure~\ref{fig:graph-example} shows an example of a process graph
-constructed at this stage \todo{Build the graph obtained on
-  bin/example.sh and add it to the Figure}.
+constructed at this stage.
+
 \begin{figure}
 %  \includegraphics{brain\_classification}
+  \includegraphics[scale=0.3]{images/simple_graph}
   \caption{Process graph constructed from an example brain extraction script.}
@@
 type, for all execution conditions. The remainder of the analysis is
-done separately for each subject type. \todo{Check that}
+done separately for each subject type.
 
@@
 %  \includegraphics{brain\_classification}
+  \includegraphics[width=0.5\textwidth, height=6cm]{images/brain_classification.png}
   \caption{Tissue classification produced by the HCP pre-processing
-    pipelines on subject 105216 (CentOS 6 vs CentOS 7). \todo{Represent this in a different way.}}
+    pipelines on subject 105216 (CentOS 6 vs CentOS 7).}
   \label{fig:1}

### 2018-03-15 paper.tex
Wrote in section 2.5

@@
 an interposition techniques for the provenance capture.
+
+\todo{Check everywhere that differences are called error. And define that in the intro.}
 
@@
 
-\subsection{Pipeline analysis: Characterizing differences}
+\subsection{Pipeline analysis: Characterizing errors}
 
-Once differences between conditions are quantified, we automatically
+Once errors between conditions are quantified, we automatically
 identify the steps in the pipeline responsible for such differences.
@@
 written by each process. This information is collected by system call
-interceptions, through the \texttt{ptrace} as a UNIX base system call, and stored in a \texttt{SQLite}
-database. The database stores information on all the processes which
-are created by the \texttt{clone()} in Linux or \texttt{fork()} in UNIX-like
-operating system calls. It also records the files opened by each process (in read, write or execution mode), including the final and
-temporary files.
+interceptions, through the \texttt{ptrace} Unix system call, and
+stored in a \texttt{SQLite} database. The database stores information
+on all the processes which are created by the \texttt{clone()} or
+\texttt{fork()} system calls. It also records the files opened by each
+process (in read, write or execution mode), including the result files
+and the temporary files.
 
 We reconstruct the tree of processes starting from the first process
-created by the pipeline and defining its child processes as the ones
+created by the pipeline and identifying its child processes as the ones
 that were created through \texttt{clone()} or \texttt{fork()}. We
@@
   \includegraphics[scale=0.3]{images/simple_graph}
-  \caption{Process graph constructed from an example brain extraction script.}
+  \caption{Process graph constructed from an example brain extraction
+    script. Every node in the graph is labeled using (1) a process id
+    created by our reconstruction, (2) the name of the executable run
+    by the process. Processes that read or write temporary files are
+    represented with squares; other ones with circles. Plain edges
+    represent the process tree (\texttt{fork()} or \texttt{clone()}
+    system calls). Dashed edges represent file dependencies: temporary
+    files are in yellow and result files are in green.\todo{Add the
+      script next to the graph. Remove node colors.}}
   \label{fig:graph-example}
@@
 that the process graphs were identical for all subjects of the same
-type, for all execution conditions. The remainder of the analysis is
-done separately for each subject type.
+type, for all execution conditions \todo{Check that}. The remainder of
+the analysis is done separately for each subject type.
 
@@
 by removing file edges between processes A and B when A's process
-creation timestamp is posterior to B's or when A=B.
+creation timestamp is posterior to B's or when A=B \todo{to be done}.
 
-\todo{Start from here}
-\paragraph{} In the next step, we classify the processes into three categories considering to the obtained process tree 
-and the error matrix file. First, processes that read file(s) without error but write different file(s) create error in 
-the pipeline. The second group of the processes are those which remove errors in pipeline, read file(s) with error and
- write “nothing or” file(s) without error. Most of the processes are in the third category beyond any fault which
- read and write file(s) without error. The processes in each category may read or write temporary file(s) that implies 
-the uncertainty of the process.
-\paragraph{} Pipeline is conducted of a bunch of processes sequentially, setting a modification step after classification
- of processes iteratively enables us to control the propagation of errors in the pipeline correctly. In each iteration
- we detect all the processes which are create error in the pipeline and then by fixing those processes artificially, 
-we’ll find the other processes in the pipeline that introduce an error. This procedure will be terminated by fixing all the processes. 
-To fix the processes that create error, we replace their dependency files with error by the same file copy from the results 
-obtained in the other operating system. In this way, we were able to pinpoint the origin of the errors measured in the pipeline.
+\paragraph{} In the next step, we classify the processes into three categories based on the process graph
+and the error matrix file mentioned in the previous Section:
+\begin{enumerate}
+\item Processes that read files that do not have errors and write files that do not have errors are \emph{transparent}.
+\item Processes that read files that do not have any error but write files that have errors \emph{create} errors in the pipeline. 
+\item Processes that read files that have errors and write files that do not have errors \emph{remove} errors from the pipeline.
+\end{enumerate}
 
-\todo{Check that} We also recorded the provenance trace of the execution using \reprozip. \reprozip identifies all the
-processes and associated parameters that wrote the files having
-differences.  These details about various processes is helpful in
-debugging the pipelines. This information helps in recreating the
-processing step by step and also to identify the processes that
-creates the differences.
+The classification of processes that read or write temporary files is
+uncertain when these files are deleted during the execution. In such
+situations, our classification marks processes as uncertain as we were
+unable to find a technical solution to prevent temporary files from
+being deleted by the pipeline \todo{investigate that a bit more}.
+
+Another challenge is to classify processes that read files that have
+errors and write files that also have errors. In such situations, it
+is not possible to determine from the pipeline results whether
+the process created errors, removed some errors, or was
+transparent. We label such processes \emph{unknown}. To address this issue, we developed an iterative approach that consists of the following steps:
+\begin{enumerate}
+  \item Run the pipeline in conditions 1 and 2; classify the
+    processes as \emph{transparent}, \emph{create errors},
+    \emph{remove errors} or \emph{unknown}.
+  \item If there are \emph{unknown} processes then: replace each process P that create errors by a process Q that
+    copies the results produced by process P in condition 1 to the pipeline output in condition 2.
+  \item Repeat steps 1 and 2 until there is no \emph{unknown} process left.
+\end{enumerate}
+The replacement of process P at step 2 is done by overwriting P's
+executable with a custom script produced by step 1. This custom script
+copies the results obtained in condition 1 if it is invoked with the
+arguments of a process that created errors, and calls P's original
+executable otherwise. This algorithm converges to a process graph
+without any \emph{unknown} process after a finite number of iterations.
+
+Figure~\ref{fig:iterations} illustrates this iterative process on the previous
+exemplar pipeline.
 

### 2018-04-19 paper.tex
Worked on section 2.5 with @ali4006

@@
 Cycles may be present in the process graph in case a file was written 
-by more than one process.  We remove the cycles in the process graph
-by removing file edges between processes A and B when A's process
-creation timestamp is posterior to B's or when A=B \todo{to be done}.
+by more than one process. We remove such cycles by removing file edges 
+between processes A and B when A's process creation timestamp is 
+posterior to B's or when A=B. Indeed, such edges cannot happen in 
+practice unless A and B were running concurrently, which we assume is 
+not the case (we also checked that on the workflow we tested).
 
-\paragraph{} In the next step, we classify the processes into three categories based on the process graph
-and the error matrix file mentioned in the previous Section:
+\paragraph{} In the next step, we classify the processes into three 
+categories based on the process graph and the error matrix file 
+mentioned in the previous Section:
 \begin{enumerate}
-\item Processes that read files that do not have errors and write files that do not have errors are \emph{transparent}.
-\item Processes that read files that do not have any error but write files that have errors \emph{create} errors in the pipeline. 
-\item Processes that read files that have errors and write files that do not have errors \emph{remove} errors from the pipeline.
+\item Processes 
+that read files that do not have errors and write files that do not 
+have errors are \emph{transparent}.
+\item Processes that read files 
+that do not have any error but write files that have errors 
+\emph{create} errors in the pipeline.
+\item Processes that read files 
+that have errors and write files that do not have errors\emph{remove} 
+errors from the pipeline.
 \end{enumerate}
@@
 The classification of processes that read or write temporary files is 
-uncertain when these files are deleted during the execution. In such
-situations, our classification marks processes as uncertain as we were
-unable to find a technical solution to prevent temporary files from
-being deleted by the pipeline \todo{investigate that a bit more}.
+uncertain when these files are deleted during the execution. To address 
+this issue, we replace every process P that writes temporary files with 
+a modification of P that first calls P and then backs up all its output 
+files to a read-only directory. This replacement is done by modifying 
+the Unix PATH variable to point to a directory containing the modified
+versions of the processes. This solution does not cover the temporary files
+that are removed by P itself; this is not a problem since these files do not
+play any role in the subsequent steps of the pipeline, by definition.
+
+Files written by multiple processes also lead to unknown classification 
+labels. To address this issue for a file F written by processes in 
+\textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we (1) check that processes 
+in \textbf{P} do not write concurrently to F, (2) we establish an order 
+on \textbf{P} based on the creation timestamp of the processes, (3) we 
+replace ever process $P_{i}$ in \textbf{P} by a wrapper that first 
+calls $P_{i}$ and then backs up F to a read-only directory. Thus, 
+multiple versions of F are saved and used in the analysis.
 
@@
 \end{enumerate}
-The replacement of process P at step 2 is done by overwriting P's
-executable with a custom script produced by step 1. This custom script
+The replacement of process P at step 2 is done by replacing P
+with a custom script produced by step 1. This custom script
 copies the results obtained in condition 1 if it is invoked with the
 arguments of a process that created errors, and calls P's original
-executable otherwise. This algorithm converges to a process graph
+executable otherwise. The replacement is done through the PATH variable, as before. This algorithm converges to a process graph
 without any \emph{unknown} process after a finite number of iterations.
@@
 exemplar pipeline.
+
+\todo{Explain how graphs are represented.}
 

### 2018-04-25 paper.tex

@@
 
-Figure~\ref{fig:iterations} illustrates this iterative process on the previous
-exemplar pipeline.
+\begin{figure}
+  \includegraphics[scale=0.3]{images/iterative_modif}
+  \caption{iteration of proposed modification.}
+  \label{fig:iterations}
+\end{figure}
 
-\todo{Explain how graphs are represented.}
+Figure~\ref{fig:iterations} illustrates the modification process iteratively in three steps for the previous example.
+ Processes that created errors are shown in red, processes that removed errors are in blue,
+ yellow nodes are the processes which propagate errors, and other processes are in green.
+ For instance, in step one after correcting bet process, fslmaths is recognized as a red process in the next step.
+In fact, the created graph in a real pipeline execution is much more complex. 
+To simplicity of intricate graph and tracing processes comprehensible, 
+the original graph compacted to several subgraphs using a compression technique. 
+The compression technique will apply, (1) build subgraphs for every interesting node (e.g. red and blue nodes with their children)
+ (2) remove the interesting nodes from graph (3) finding largest connected components in the remaining nodes (iteratively).
 

### 2018-04-25 paper.tex
Improved description of Figure 2

--- 2018-04-19 paper.tex
@@
 
-Figure~\ref{fig:iterations} illustrates this iterative process on the previous
-exemplar pipeline.
+\begin{figure}
+  \includegraphics[scale=0.3]{images/iterative_modif}
+  \caption{iteration of proposed modification.}
+  \label{fig:iterations}
+\end{figure}
 
-\todo{Explain how graphs are represented.}
+Figure~\ref{fig:iterations} illustrates our iterative classification 
+process for the example in Figure~\ref{fig:graph-example}. At every 
+step, processes that created errors are shown in red, processes that 
+removed errors are in blue, processes that propagated errors are in 
+yellow, and other processes (transparent processes) are in green. As in 
+Figure~\ref{fig:graph-example}, plain black edges represent the process 
+tree and dashed edges represent file dependencies: green edges 
+represent files with no errors, while red edges represent files with 
+errors. Temporary files are not represented because they have been 
+backed up as previously described.
+
+The three steps in Figure~\ref{fig:iterations} correspond to the 
+iterations of the classification algorithm At step one \texttt{2\#bet2} 
+is classified as error creator (red) as it produced files with errors 
+from files without error. \texttt{3\#fslstats} is classified as error 
+remover (blue) as it produced files without errors from files with 
+errors. \texttt{4\#fslmaths} and \texttt{5\#fslstats} are classified as 
+unknown (yellow) as they produce files with errors from files with errors.
+
+At step 2, the files produced by \texttt{2\#bet2} in the tested 
+condition are replaced with the files produced by \texttt{2\#bet2} in 
+the other condition. \texttt{3\#fslstats} is now classified as transparent, \texttt{4\#fslmaths} is now classified
+as error creator and \texttt{5\#fslstats} is still unknown.
+
+ At step 3, the files produced by \texttt{4\#fslmaths} in the tested 
+ condition are replaced with the files produced by \texttt{4\#fslmaths} 
+ in the other condition. \texttt{5\#fslstats} is now transparent.
+
+ As a result of those 3 steps, the final process classification is: 
+ \texttt{2\#bet2} and \texttt{4\#fslmath} are error creators (red), 
+ \texttt{3\#fslstats} is error remover (blue) and the other processes 
+ are transparent.
+
+The process graphs of real pipelines are much larger than the one in 
+Figure~\ref{fig:iterations}. To represent such large graphs, we 
+summarize them by (1) building subgraphs for every interesting node 
+(e.g. red and blue nodes with their children) (2) remove the 
+interesting nodes from graph (3) finding largest connected components 
+in the remaining nodes (iteratively).
 

### 2018-04-30 paper.tex
Update figures

--- 2018-04-25 paper.tex
@@
 \usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{caption}
+\usepackage{listings}
+\usepackage{verbatim}
 
@@
 
-\begin{figure}
-%  \includegraphics{brain\_classification}
-  \includegraphics[scale=0.3]{images/simple_graph}
-  \caption{Process graph constructed from an example brain extraction
+\medskip
+\noindent
+\begin{minipage}[]{.5\linewidth}
+  \centering
+  \includegraphics[scale=0.2,width=.4\linewidth]{images/simple_graph}
+%  \captionof{figure}{This is a figure caption.} \label{myfig1}
+\end{minipage}
+\begin{minipage}[]{.3\linewidth}
+\small\captionof{algorithm}{sample script of a process graph}\label{sample_script}
+%\begin{algorithmic}[1]
+%\begin{lstlisting}[frame=single]  % Start your code-block
+%
+%#!/bin/bash
+%
+%if [ # !=1 ]
+%then
+%    echo "usage: 0 <inputimage.nii.gz>"
+%    exit 1
+%fi
+%\end{lstlisting}
+
+{\scriptsize
+\begin{verbatim}[frame=single]
+#!/bin/bash
+if [ # !=1 ]
+then
+    echo "usage: 0 <inputimage.nii.gz>"
+    exit 1
+fi
+input_image=$1
+bet_output="$(basename ${input_image} .nii.gz)_brain.nii.gz"
+bet_output_binarized="$(basename ${input_image} .nii.gz)_brain_bin.nii.gz"
+
+bet ${input_image} ${bet_output}
+echo "Voxels / Volume in brain mask:"
+fslstats ${bet_output} -V
+fslmaths ${bet_output} ${bet_output_binarized}
+echo "Voxels / Volum in binarized brain mask:"
+\rm ${bet_output}
+\end{verbatim}
+}
+
+%input_image=$1
+%bet_output="$(basename ${input_image} .nii.gz)_brain.nii.gz"
+%bet_output_binarized="$(basename ${input_image} .nii.gz)_brain_bin.nii.gz"
+%
+%bet ${input_image} ${bet_output}
+%echo "Voxels / Volume in brain mask:"
+%fslstats ${bet_output} -V
+%fslmaths ${bet_output} ${bet_output_binarized}
+%echo "Voxels / Volum in binarized brain mask:"
+%\rm ${bet_output}
+
+%  \Procedure{Euclid}{$a,b$}\Comment{The g.c.d. of a and b}
+%    \State $r\gets a\bmod b$
+%    \While{$r\not=0$}\Comment{We have the answer if r is 0}
+%      \State $a\gets b$
+%      \State $b\gets r$
+%      \State $r\gets a\bmod b$
+%    \EndWhile\label{euclidendwhile}
+%    \State \textbf{return} $b$\Comment{The gcd is b}
+%  \EndProcedure
+
+%\end{algorithmic}
+\end{minipage}
+
+\captionof{figure}{Process graph constructed from an example brain extraction
     script. Every node in the graph is labeled using (1) a process id
@@
     system calls). Dashed edges represent file dependencies: temporary
-    files are in yellow and result files are in green.\todo{Add the
-      script next to the graph. Remove node colors.}}
-  \label{fig:graph-example}
-\end{figure}
+    files are in yellow and result files are in green.} \label{myfig1}
+
+\medskip
 
@@
 \begin{figure}
-  \includegraphics[scale=0.3]{images/iterative_modif}
+  \includegraphics[scale=0.6]{images/iterative_modif}
   \caption{iteration of proposed modification.}

### 2018-05-02 paper.tex

--- 2018-04-25 paper.tex
@@
 \usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage{caption}
+\usepackage{listings}
+\usepackage{verbatim}
 
@@
 
-\begin{figure}
-%  \includegraphics{brain\_classification}
-  \includegraphics[scale=0.3]{images/simple_graph}
-  \caption{Process graph constructed from an example brain extraction
+\medskip
+\noindent
+\begin{minipage}[]{.5\linewidth}
+  \centering
+  \includegraphics[scale=0.2,width=.4\linewidth]{images/simple_graph}
+%  \captionof{figure}{This is a figure caption.} \label{myfig1}
+\end{minipage}
+\begin{minipage}[]{.3\linewidth}
+\small\captionof{algorithm}{sample script of a process graph}\label{sample_script}
+{\scriptsize
+\begin{verbatim}
+#!/bin/bash
+if [ # !=1 ]
+then
+    echo "usage: 0 <inputimage.nii.gz>"
+    exit 1
+fi
+input_image=$1
+bet_output="$(basename ${input_image} .nii.gz)_brain.nii.gz"
+bet_output_binarized="$(basename ${input_image} .nii.gz)_brain_bin.nii.gz"
+
+bet ${input_image} ${bet_output} > bet_temp.out
+echo "Voxels / Volume in brain mask:"
+fslstats ${bet_output} -V
+fslmaths ${bet_output} ${bet_output_binarized}
+echo "Voxels / Volum in binarized brain mask:"
+\rm bet_temp.out
+\end{verbatim}
+}
+\end{minipage}
+\captionof{figure}{Process graph constructed from an example brain extraction
     script. Every node in the graph is labeled using (1) a process id
@@
     system calls). Dashed edges represent file dependencies: temporary
-    files are in yellow and result files are in green.\todo{Add the
-      script next to the graph. Remove node colors.}}
-  \label{fig:graph-example}
-\end{figure}
+    files are in yellow and result files are in green.} \label{myfig1}
+\medskip
 
@@
 summarize them by (1) building subgraphs for every interesting node 
-(e.g. red and blue nodes with their children) (2) remove the 
-interesting nodes from graph (3) finding largest connected components 
-in the remaining nodes (iteratively).
+(e.g. red and blue nodes with their children) (2) building connected components 
+in the absence of interesting nodes for the remaining nodes in the process graph. 
+Each node in the summarized graph represent either a shrinked interesting subgraph or 
+connected component and edges represent any connection between two shrinked nodes.
 
@@
 
-Figure \ref{fig:2} shows the annotated provenance graph of the
+Figure \ref{fig:summarized-graph} shows the annotated provenance graph of the
 PreFreeSurfer pipeline executed on CentOS6 and CentOS7.  Processes
@@
 \begin{figure}
+  \includegraphics[width=\linewidth]{images/summarized_graph}
+  \caption{An annotated summarized process graph from the PreFreesurfer pipeline.
+Each node in the graph is represented in red, blue or green colors and is labeled 
+using the name of the executable run by the interesting process. The close up nodes 
+reveal each shrinked subgraph in details.}
+  \label{fig:summarized-graph}
+\end{figure}
+
+\begin{figure}
   \includegraphics[width=\linewidth]{images/graph}
-  \caption{An annotated provenance graph from the PreFreesurfer pipeline. Processes that created errors are in red. 
+  \caption{A complete process graph from the PreFreesurfer pipeline.
 Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
-  \label{fig:2}
+  \label{fig:complete-graph}
 \end{figure}

### 2018-10-01 paper.tex

@@
 \usepackage{caption}
+\usepackage{subcaption}
 \usepackage{listings}
 \usepackage{verbatim}
+\usepackage{booktabs}
+\usepackage[flushleft]{threeparttable}
 
@@
 We randomly selected 10 subjects from the HCP data release
-S500. \todo{ list the subjects.}
+S500 which are described in table~\ref{table:data}. \todo{ list the subjects.}
+
+\begin{table}[H]\hspace*{-1cm}
+\centering
+\begin{threeparttable}
+\caption{An overview of the HCP data (Humman Connectome DB).}
+
+\begin{tabular}{@{}llllll@{}}
+\toprule
+Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
+103515  & Q1      & Q02         & F      & 26-30    & type1         \\
+105216  & Q3      & Q03         & M      & 26-30    & type4         \\
+103414  & Q2      & Q02         & F      & 22-25    & type1         \\
+125525  & Q1      & Q01         & F      & 31-35    & type1         \\
+142828  & Q1      & Q01         & M      & 31-35    & type1         \\
+129533  & Q3      & Q04         & F      & 31-35    & type1         \\
+103818  & Q1      & Q01         & F      & 31-35    & type3         \\
+133928  & Q2      & Q03         & M      & 26-30    & type3         \\
+148032  & Q3      & Q03         & F      & 31-35    & type2         \\
+139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
+\end{tabular}
+\begin{tablenotes}
+      \small
+      \item *Subj\_Type refers to the graph topology type of the subjects,
+      we will discuss further on the next sections.
+\end{tablenotes}
+\end{threeparttable}
+\label{table:data}
+\end{table}
 
@@
 \todo{Write a short summary of what the pipelines do.}
+
+we focused on HCP structural pre-processing pipelines to:
+\begin{itemize}
+\item Quantify the effect of OS on HCP pre-processing pipelines.
+\item Identify the process(es) in the pipelines that are responsible for such effects.
+\end{itemize}
 
@@
 
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.9]{images/overview.png}
+  \caption{An overview of proposed technique.}
+  \label{fig:overview}
+\end{figure}
+
+Figure~\ref{fig:overview} illustrate an overview of the proposed technique. First, we execute pipelines in a
+consistence environments and then the origin of errors in the pipeline is detected through a
+classification approach. We will explain each step in detail in the following sections.
+
 \subsection{Measuring differences}
@@
 between processes A and B if a file written by A is read by
-B. Figure~\ref{fig:graph-example} shows an example of a process graph
+B. Figure~\ref{fig:simple_script} shows an example of a process graph
 constructed at this stage.
@@
     system calls). Dashed edges represent file dependencies: temporary
-    files are in yellow and result files are in green.} \label{myfig1}
+    files are in yellow and result files are in green.} \label{fig:simple_script}
 \medskip
@@
 \begin{enumerate}
-\item Processes 
-that read files that do not have errors and write files that do not 
-have errors are \emph{transparent}.
+\item Processes that read files that do not have errors and write files that do not 
+have errors are \emph{transparent} (Figure~\ref{fig:processes}.a).
 \item Processes that read files 
 that do not have any error but write files that have errors 
-\emph{create} errors in the pipeline.
+\emph{create} errors in the pipeline (Figure~\ref{fig:processes}.b).
 \item Processes that read files 
-that have errors and write files that do not have errors\emph{remove} 
-errors from the pipeline.
+that have errors and write files that do not have errors \emph{remove} 
+errors from the pipeline (Figure~\ref{fig:processes}.c).
+\item Processes that read files that have errors and write files that also have errors are 
+\emph{unknown} (Figure~\ref{fig:processes}.d).
 \end{enumerate}
 
-The classification of processes that read or write temporary files is 
+\begin{figure}[H]%\centering
+    \begin{subfigure}{0.15\textwidth}
+        \includegraphics[scale=0.35]{images/green.png}
+        \caption{Transparent}
+        \label{fig:green}
+    \end{subfigure}
+\hfill%\hfil
+    \begin{subfigure}{0.15\textwidth}
+    \includegraphics[scale=0.35]{images/red.png}
+    \caption{Creates error}
+    \label{fig:red}
+\end{subfigure}
+\hfill%\hfil
+    \begin{subfigure}{0.15\textwidth}
+    \includegraphics[scale=0.35]{images/blue.png}
+    \caption{Removes error}
+    \label{fig:blue}
+\end{subfigure}
+\hfill%\hfil
+    \begin{subfigure}{0.15\textwidth}
+    \includegraphics[scale=0.35]{images/yellow.png}
+    \caption{Unknown}
+    \label{fig:yellow}
+\end{subfigure}
+    \caption{Different type of classified process based on the input/output files.
+  Dashed edges refer to the file dependencies between processes A and B if a file written bu A is read bu B.
+  Solid black edges refer ti the relationship between parent and child processes.}
+    \label{fig:processes}
+\end{figure}
+
+\paragraph{Challenge1:} The classification of processes that read or write temporary files is 
 uncertain when these files are deleted during the execution. To address 
@@
 
-Files written by multiple processes also lead to unknown classification 
+\paragraph{Challenge2:} Files written by multiple processes also lead to unknown classification 
 labels. To address this issue for a file F written by processes in 
@@
 
-Another challenge is to classify processes that read files that have
+\paragraph{Challenge3:} Another challenge is to classify processes that read files that have
 errors and write files that also have errors. In such situations, it
@@
 \end{enumerate}
+
 The replacement of process P at step 2 is done by replacing P
@@
 arguments of a process that created errors, and calls P's original
-executable otherwise. The replacement is done through the PATH variable, as before. This algorithm converges to a process graph
-without any \emph{unknown} process after a finite number of iterations.
+executable otherwise. The replacement is done through the PATH variable, as before.
+This algorithm converges to a process graph without any \emph{unknown} process after
+a finite number of iterations.
 
-\begin{figure}
-  \includegraphics[scale=0.6]{images/iterative_modif}
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.5]{images/iterative_modif}
   \caption{iteration of proposed modification.}
@@
 variation (see Figure \ref{fig:1}).
-\begin{figure}
+\begin{figure}[H]
 %  \includegraphics{brain\_classification}
@@
   \label{fig:1}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.8]{images/fnirt_result.png}
+  \caption{Differences between FNIRT results (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (centos6 vs. centos7).}
+  \label{fig:fnirt_result}
 \end{figure}
@@
 Figure \ref{fig:summarized-graph} shows the annotated provenance graph of the
-PreFreeSurfer pipeline executed on CentOS6 and CentOS7.  Processes
-that created errors are shown in red, processes that removed errors
+PreFreeSurfer pipeline executed on CentOS6 and CentOS7.
+Each node in the graph represent an executed process in the pipeline.
+Processes that created errors are shown in red, processes that removed errors
 are in blue, and other processes are in green.  Squares denote
@@
 dependencies between processes (green edges: files with no errors; red
-edges: files with errors; yellow edges: temporary files).  The
-processes that introduce errors in PreFreeSurfer are: linear
+edges: files with errors; yellow edges: temporary files).
+
+The processes that introduce errors in PreFreeSurfer are: linear
 registration with “\emph{FLIRT}” (8 occurrences, in ACPC Alignment,
@@
 
-\begin{figure}
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.3]{images/pfs_chart.png}
+  \caption{Number of occurrences of errors in PreFreeSurfer, red and blue bars 
+indicate the processes which create and remove errors respectively.}
+  \label{fig:pfs_chart}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.5]{images/pfs_table.png}
+  \caption{Process name along with the pipeline elements that inbtroduce errors.}
+  \label{fig:pfs_table}
+\end{figure}
+
+The processes that introduce errors in FreeSurfer are shown on Figure~\ref{fig:fs_error_table}.
+Also, processes that remove errors in FreeSurfer represented on Figure~\ref{fig:fs_remove_table}.
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.5]{images/fs_error_table.png}
+  \caption{Number of occurrences of errors in FreeSurfer pipeline, subject 103818, 
+indicate the processes that create errors.}
+  \label{fig:fs_error_table}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.5]{images/fs_remove_table.png}
+  \caption{Number of occurrences of errors in FreeSurfer pipeline, subject 103818, 
+indicate the processes that remove errors.}
+  \label{fig:fs_remove_table}
+\end{figure}
+
+\section{Discussion}
+
+pipeline amplify small numerical differences, they are numerically unstable.
+Furthermore, math libraries evolve over time, leading to different numerical errors.
+we listed some of the irreproducibility causes of the pipelines as we mentioned
+in the previous section overally along with exact command arguments.
+
+\begin{itemize}
+\item FSLSTATS: Calculate statistics on an image (mean, standard deviation and etc.)
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+STD=`${FSLDIR}/bin/fslstats $WD/T1wmulT2w_brain_norm_modulate.nii.gz -S` 
+MEAN=`${FSLDIR}/bin/fslstats $WD/T1wmulT2w_brain_norm_modulate.nii.gz -M` 
+    \end{lstlisting}
+\item FSLMATH: Create and apply a binary mask to filling of holes to remove  any non-grey/white tissue
+using a threshold created by FSLSTATS at Mean.
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+{FSLDIR}/bin/fslmaths $WD/T1wmulT2w_brain_norm_modulate -thr $Lower -bin -ero -mul 255 
+                      $WD/T1wmulT2w_brain_norm_modulate_mask ${CARET7DIR}/wb_command -volume-remove-islands 
+                      $WD/T1wmulT2w_brain_norm_modulate_mask.nii.gz $WD/T1wmulT2w_brain_norm_modulate_mask.nii.gz
+    \end{lstlisting}
+
+\item FLIRT: Register cropped image to MNI152 (12 DOF)
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+{FSLDIR}/bin/flirt -interp spline -in "$WD"/robustroi.nii.gz -ref "$Reference"
+                   -omat "$WD"/roi2std.mat -out "$WD"/acpc_final.nii.gz
+                   -searchrx -30 30 -searchry -30 30 -searchrz -30 30
+
+{FSLDIR}/bin/flirt -interp spline -dof 6 -in ${WD}/Magnitude_warpped${TXw}
+                    -ref ${TXwImage} -out ${WD}/Magnitude_warpped${TXw}2${TXwImageBasename}
+                    -omat ${WD}/Fieldmap2${TXwImageBasename}.mat
+                    -searchrx -30 30 -searchry -30 30 -searchrz -30 30
+    \end{lstlisting}
+
+\item FNIRT: non-linear registration to MNI
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+{FSLDIR}/bin/fnirt --in=${T1wRestore} --ref=${Reference2mm} --aff=${WD}/xfms/acpc2MNILinear.mat
+                   --refmask=${Reference2mmMask} --fout=${OutputTransform}
+                   --jout=${WD}/xfms/NonlinearRegJacobians.nii.gz --refout=${WD}/xfms/IntensityModulatedT1.nii.gz
+                   --iout=${WD}/xfms/2mmReg.nii.gz --logout=${WD}/xfms/NonlinearReg.txt
+                   --intout=${WD}/xfms/NonlinearIntensities.nii.gz --cout=${WD}/xfms/NonlinearReg.nii.gz --config=${FNIRTConfig}
+
+%~ Register to 2mm reference image (linear then non-linear)
+%~ ${FSLDIR}/bin/flirt -interp spline -dof 12 -in "$Input" -ref "$Reference2mm" 
+                    %~ -omat "$WD"/roughlin.mat -out "$WD"/"$BaseName"_to_MNI_roughlin.nii.gz -nosearch
+
+%~ ${FSLDIR}/bin/fnirt --in="$Input" --ref="$Reference2mm" --aff="$WD"/roughlin.mat
+                    %~ --refmask="$Reference2mmMask" --fout="$WD"/str2standard.nii.gz
+                    %~ --jout="$WD"/NonlinearRegJacobians.nii.gz --refout="$WD"/IntensityModulatedT1.nii.gz
+                    %~ --iout="$WD"/"$BaseName"_to_MNI_nonlin.nii.gz
+                    %~ --logout="$WD"/NonlinearReg.txt --intout="$WD"/NonlinearIntensities.nii.gz
+                    %~ --cout="$WD"/NonlinearReg.nii.gz --config="$FNIRTConfig"
+    \end{lstlisting}
+
+\item INVWARP \& APPLYWARP: Invert warp and transform dilated brain mask back into native space, and use it to mask input image
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+${FSLDIR}/bin/invwarp --ref="$Reference2mm" -w "$WD"/str2standard.nii.gz -o "$WD"/standard2str.nii.gz
+${FSLDIR}/bin/applywarp --rel --interp=nn --in="$ReferenceMask" --ref="$Input" -w "$WD"/standard2str.nii.gz -o "$OutputBrainMask"
+    \end{lstlisting}
+
+\item MRIS\_ANATOMICAL\_STATS: Measures a number of anatomical properties
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+mris_anatomical_stats -mgz -cortex ../label/lh.cortex.label -f ../stats/lh.aparc.a2009s.stats -b -a ../label/lh.aparc.a2009s.annot -c ../label/aparc.annot.a2009s.ctab 0050299 lh white
+    \end{lstlisting}
+
+\item MRIS\_MAKE\_SURFACES: Generate pial surfaces, first pass, no T2 adjustment. Files created are Xh.pial, Xh.curv.pial, Xh.area.pial, Xh.thickness
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+mris_make_surfaces -variablesigma ${VARIABLESIGMA} -white NOWRITE -aseg aseg.hires -orig white.deformed -filled filled.hires -wm wm.hires -sdir $SubjectDIR -mgz -T1 T1w_hires.norm "$SubjectID" lh
+    \end{lstlisting}
+
+\item MRI\_CONVERT:Normalize T1w and T2w images for the benefit of mris\_make\_surfaces
+
+\#Resample the native space binary ROI file into "brainmask" space
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+mri_convert -rl /subA/mri/T1w_hires.nii.gz -rt nearest /subA/mri/filled.mgz /subA/mri/filled.hires.mgz
+    \end{lstlisting}
+
+\item MRI\_NORMALIZE: Normalize the white-matter, optionally based on control points.
+The input volume is converted into a new volume where white matter image values all range around 110.
+
+\#Uses the norm volume, and creates the brain volume, making use of the aseg and masking with brainmask
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+mri_normalize -aseg aseg.mgz -mask brainmask.mgz norm.mgz brain.mgz
+    \end{lstlisting}
+
+\item MRI\_APARC2ASEG:The volumes of the cortical labels will be different than
+         reported by mris\_anatomical\_stats because partial volume information
+         is lost when mapping the surface to the volume. The values reported by
+         mris\_anatomical\_stats will be more accurate than the volumes from the
+         aparc+aseg volume.
+
+Label the cortex of a subject's segmented volume according to the edited surface labels
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+mri_aparc2aseg --s SUBJ_NUM --volmask
+\#slice labeling
+mri_aparc2aseg --s ${subject} --labelwm --hypo-as-wm --rip-unknown --volmask --o ${subj_dir}/mri/wmdivided --annot aparc.split
+    \end{lstlisting}
+
+\item MRI\_EM\_REGISTER: creates a tranform in lta format
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+mri_em_register -skull -t transforms/talairach.lta nu_noneck.mgz $FREESURFER_HOME/average/RB_all_withskull_2008-03-26.gca transforms/talairach_with_skull.lta
+    \end{lstlisting}
+
+\item MRIS\_SMOOTH: Smooths the tessellation of a cortical surface.
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+mris_smooth -nw -seed 1234 ../surf/lh.orig.nofix ../surf/lh.smoothwm.nofix
+    \end{lstlisting}
+
+\item MRIS\_INFLATE: inflate cortical surface.
+\#The command generates inflated and sulc files.
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+mris_inflate -no-save-sulc surf/lh.smoothwm.nofix surf/lh.inflated.nofix
+    \end{lstlisting}
+
+\item MRIS\_CALC: mris\_calc' is a simple calculator that operates on
+FreeSurfer curvatures and volumes. In most cases, the calculator functions
+with three arguments: two inputs and an <ACTION> linking them.
+Some actions, however, operate with only one input <file1>.
+In all cases, the first input <file1> is the name of a FreeSurfer curvature overlay (e.g. rh.curv)
+or volume file (e.g. orig.mgz). For two inputs, the calculator first assumes
+that the second input is a file. If, however, this second input file doesn't exist,
+the calculator assumes it refers to a float number, which is then processed according to <ACTION>.Note:
+<file1> and <file2> should typically be generated on the same subject.
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+
+mris_calc -o lh.area.mid lh.area.mid div2
+    \end{lstlisting}
+
+\item CONVERTWARP: Forward warp the fieldmap magnitude and register to TXw image (transform phase image as well).
+    \begin{lstlisting}[language=bash, basicstyle=\tiny]
+${FSLDIR}/bin/convertwarp --relout --rel --ref=${WD}/Magnitude --shiftmap=${WD}/FieldMap_ShiftMap${TXw}.nii.gz 
+                          --shiftdir=${UnwarpDir} --out=${WD}/FieldMap_Warp${TXw}.nii.gz
+    \end{lstlisting}
+\item MRI\_CONVERT:
+
+\item MRI\_SURF2SURF:
+
+\item MRIS\_MAKE\_SURFACE:
+
+\item MRI\_CA\_REGISTER:
+
+\item MRIS\_CONVERT:
+
+\end{itemize}
+
+Mention that this was only possible because the unprocessed data was shared in the first place.
+DICOM to Nifti conversion was out of scope and may introduce other issues.
+
+\section{Conclusion}
+
+Our thechnique is able to characterize the stability of a pipeline's components automatically.
+The numerical instability in the PreFreesurfer HCP pipeline arises mainly from
+linear and non-linear registration processes implemented in FSL FLIRT and FNIRT.
+
+There are a few ways to impede such instabilities:
+\begin{itemize}
+\item Use a single operating system
+\item Containerize pipelines
+\item Increase numerical precision
+\item Be stricter on truncation and rounding standards (IEEE 754)
+\item Build static executable
+\end{itemize}
+
+The results still suffer from small pertubations literally because of the fact that pipeline are not numerically stable.
+The preferred solution is to detect and fix numerical instability of the pipeline instead of masking the problem.
+These processes need to be reviewed to understand and correct the cause of instabilities. 
+In this correction process, accuracy has to be considered in addition to stability.
+
+\item In the future works we aim to measure the impact of such errors
+on the specific domain of the neuroscience and how show the significant it is.
+
+\item The purpose of the project is generalizing pipelines to get a reproducible execution 
+on every operating system instead of choosing one main OS to.
+
+\section{Acknowledgments}
+
+CBRAIN team. Compute Canada(Calcul Quebec).
+
+\begin{figure}[H]
   \includegraphics[width=\linewidth]{images/summarized_graph}
@@
 
-\begin{figure}
+\begin{figure}[H]
   \includegraphics[width=\linewidth]{images/graph}
@@
 \end{figure}
-
-\section{Discussion}
-
-Discuss the results.
-
-Mention that this was only possible because the unprocessed data was shared in the first place.
-
-DICOM to Nifti conversion was out of scope and may introduce other issues.
-
-\section{Conclusion}
-
-Conclude and highlight future work.
-
-The numerical instability in the PreFreesurfer HCP pipeline arises mainly from linear and non-linear registration processes 
-implemented in FSL FLIRT and FNIRT. These processes need to be reviewed to understand and correct the cause of instabilities. 
-In this correction process, accuracy has to be considered in addition to stability. Our technique is able to 
-characterize the stability of a pipeline’s components automatically, but it also suffers from limitations as it cannot deal with: 
-(1) results that are not written to disk (values processed in memory); (2) temporary files (3) files that are written by more than one process.
-
-\section{Acknowledgments}
-
-CBRAIN team. Compute Canada(Calcul Quebec).
-
 \bibliographystyle{plain}

### 2019-01-02 paper.tex
revise the paper structure and focus on the tool

--- 2018-05-02 paper.tex
@@
 \usepackage{caption}
+\usepackage{subcaption}
 \usepackage{listings}
 \usepackage{verbatim}
+\usepackage{booktabs}
+\usepackage{makecell}
+
+\usepackage[flushleft]{threeparttable}
 
@@
 
-\title{Numerical error propagation in the HCP structural pre-processing pipelines}
+\title{Numerical error propagation in the HCP structural pre-processing 
+pipelines}
+\title{Reprotool: a debugging tool to identify numerical errors 
+and assess reproducibility of the neuroimaging pipelines}
 
@@
 
-\abstract{This paper is a reproducibility study of the work in~\cite{glasser2015multi}.}
+\abstract{
+
+Many experiments show that computational analysis results are still not 
+completely reproducible. Although new technologies such as 
+virtualization , version control systems , and provenance management 
+tools provide more reliable environment and significantly improve the 
+reproducibility crisis. 
+In particular, computational environments are known to have an effect 
+on the results produced by neuroimaging pipelines, presumably due to 
+the creation, propagation and amplification of small numerical errors 
+across the pipelines. Similarly, small perturbations that originated 
+from computational environments including diverse operating systems, 
+hardware categories, software version, etc. highlight the numerical 
+errors.
+However, the precise causes of such instabilities and the path along 
+which they propagate in the pipelines are unclear.  We present a 
+technique to identify the processes in the pipeline that create numerical 
+errors along the execution, and we apply this technique to the HCP 
+structural pre-processing pipelines.
 
@@
 
-Operating systems are known to have an effect on the results produced
-by neuroimaging pipelines~\cite{Gronenschild2012, Glatard2015},
-presumably due to the creation, propagation and amplification of small
-numerical errors across the pipelines.  Such errors highlight
-numerical instability which is also likely to appear as a result of
-other types of small perturbations such as acquisition and parametric
-noise. However, the precise causes of
-such instabilities and the path along which they propagate in the
-pipelines are unclear.  We present a technique to identify the
-processes in a pipeline that create numerical errors along the
-execution, and we apply this technique to the HCP structural
-pre-processing pipelines.
+Research findings are expected to be reproducible to evaluate the 
+authenticity and reliability of the claims.There are different 
+terminologies for the terms reproducibility, repeatability, and 
+replicability over the time. For example, the Association for Computing 
+Machinery (ACM)~\cite{acm2016terminology} proposes three different 
+definitions. From the ACM's points of view, repeatability is defined as 
+repeating the computation by the same experimental setup including 
+operator team, operating conditions, location, and measuring system. 
+Replicability has almost the similar definition with repeatability 
+which use the identical experimental conditions except performer team 
+which means that an independent group can achieve the same results. 
+Beside, reproducibility is measured by performing different 
+experimental setup via different laboratory team independently. In this 
+paper, we follow Peng's~\cite{peng2011reproducible} definition of 
+reproducibility which is assessed by performing the experiment using 
+the same data and the same analytic methods and expecting the same 
+results. A reproducible study provides a context in which, one is able 
+to get the consistence results to the original work. In addition, it 
+enables others to use the existing work as a part of their 
+experiments~\cite{plesser2018reproducibility}. It not only saves a 
+great deal of time, but also contributes to focusing on the other 
+parts.
 
-Groenschild \emph{et al.} first identified the effect of operating
-systems on Freesurfer
-results~\cite{Gronenschild2012}. In~\cite{10.3389/fninf.2015.00012} we
-quantified this effect on some of the main neuroimaging pipelines
-including several FSL pipelines, CIVET and Freesurfer. In this work we
-aim at evaluating this effect on the work in~\cite{glasser2015multi}.
-\todo{Link this to Lindsay's HBM 2017 poster and to Redolfi et al.}
+Reproducibility of computational pipelines has been repeatedly 
+investigated so far. The comparisons show that the analysis results are 
+not completely reproducible across the small perturbations which have 
+emerged by changing the execution environments including operating 
+system, software and analysis method, and hardware configuration. There 
+are different studies that discovered factors that introduce such tiny 
+perturbations and consequently hamper the reproducibility of the 
+experiments. In particular, Groenschild \emph{et al.} first identified 
+the effect of operating systems on Freesurfer 
+results~\cite{Gronenschild2012}. Similarly, the effect of OS is 
+quantified in~\cite{Glatard2015} more specifically on some 
+of the main neuroimaging pipelines including several FSL pipelines, 
+CIVET and Freesurfer. The results of comparisons indicate a significant 
+errors between analysis results of the same experiment on different 
+operating system.
 
-The operating system is defined here as a consistent set of software
+It should be noted that \emph{error} is defined as the statistical 
+differences between results of two experiments in this work. Moreover, 
+the operating system is defined here as a consistent set of software 
 packages organized in a trusted repository.
@@
 The operating system is not the only part of the computational 
-environment that may hamper reproducibility. The work
-in~\cite{diethelm2012limits} mentions reproducibility issues coming
-from parallelization.
+environment that may hamper reproducibility. Many other computational 
+factors including hardware configuration, software version, and 
+parallelization techniques are reported~\cite{diethelm2012limits} as 
+the main influential computing infrastructure on reproducibility.
 
-Notes:
+It has revealed that the majority of analysis results are not 
+numerically reproducible because of arithmetic operations which is 
+associated with the type of execution such as sequential or parallel 
+executions, floating-point operations, different standard calculations 
+like IEEE-754 or other standards, etc.
+
+In order to overcome these issues and making reproducible analysis, a 
+comprehensive solution requires which not only encapsulates and 
+re-executes all the software/hardware dependencies but also fixes the 
+numerical instabilities. Therefore, we introduce a framework to 
+identify the processes in the pipeline which are responsible for such 
+numerical errors along the execution. In this paper, the execution 
+results of the pipelines across different operating systems are 
+considered to find out the cause of the reproducibility issues using an 
+interposition techniques.
+
+Figure~\ref{fig:overview} illustrate an overview of the proposed 
+technique. First, we execute pipelines in a consistence environment 
+using docker container and Boutiques framework. In the next step, the 
+origin of errors, processes that make differences, are detected through 
+a classification approach. Processes are classified by several 
+modifications iteratively and finally, it creates corresponding process 
+graph of the pipeline execution. We will explain each step in detail in 
+the following sections.
+
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.5]{images/overview.png}
+  \caption{An overview of proposed technique.}
+  \label{fig:overview}
+\end{figure}
+
+\paragraph{} The reminder of this paper explains the proposed debugging tool in two 
+parts. First, the execution environment of the pipelines and then the 
+classification procedure of the processes are clarified. We describe 
+the HCP preprocessing pipelines as a popular neuroimaging pipelines to 
+test and apply reprotool. Finally, we report the analytic process which 
+are responsible for the errors in the pipelines.
+
+\section{Processing Environment}
+
+We build the containerized version of pipelines for each operating 
+system using docker technology which encapsulated whole the pipeline 
+dependencies. The docker images consists of an specific computing 
+environment such as operating system version, hardware configuration, 
+software version, etc. Furthermore, derived images can be stored on 
+the dockerhub repositories and be version controlled in the same way as 
+the analysis code so that the exact same environment can be used to run 
+an experiment.
+
+To simplify computational reproducibility of the analysis, we execute 
+docker images using Boutiques framework based on the command-line 
+descriptions~\cite{glatard2017boutiques}. The Boutiques descriptor 
+specifies the parameters and implementation of the experiment, and the 
+invocation schema describes the parameter values such as input/output 
+data path.
+
+We processed subjects twice in every condition to identify possible 
+\emph{intra-condition} errors coming from the use of pseudo-random 
+numbers or other external factors. To further ensure that the observed 
+errors are caused by operating system updates (\emph{inter-os 
+differences}) and not by other factors, we implemented the following 
+sanity checks:
+
 \begin{itemize}
-  \item Reproducibility has several other aspects, see, e.g.,
-\url{https://medium.com/@lorenaabarba/barba-group-reproducibility-syllabus-e3757ee635cf#.ty3zmgd4k}
- \item James P Turner, speaker in INCF 2016 Track B, is looking at numerical stability on GPUs. And reproducibility.
- \item For an overview of variability accross analysis methods in fMRI, see \url{http://journal.frontiersin.org/article/10.3389/fnins.2012.00149/full}.
- \item For a general overview in fMRI, see \url{http://www.nature.com/nrn/journal/vaop/ncurrent/box/nrn.2016.167_BX3.html}
-\end{itemize}
 
-% There is a computational challenge
-HCP preprocessing pipelines are both time-consuming and
-computationally intensive which might even run for days, depending
-upon the type and size of dataset.
-
-Any
-user who has access to this application can try to reproduce the
-experiments since the data and the application are accessible within
-the framework. However, one caveat is that, the neuroimaginng
-pipelines might have different results due to differences in computing
-platforms on which the images are getting processed
-~\cite{10.3389/conf.fninf.2014.18.00076} therefore, the execution
-results of the pipelines across different operating systems are
-considered to find out the cause of the reproducibility issues using
-an interposition techniques for the provenance capture.
-
-\todo{Check everywhere that differences are called error. And define that in the intro.}
-
-\section{Materials and Methods}
-
-\subsection{Data}
-
-We randomly selected 10 subjects from the HCP data release
-S500. \todo{ list the subjects.}
-
-\subsection{HCP Minimal preprocessing Pipelines}
-
-The Human Connectome Project (HCP)~\cite{Gla13} developed a set of
-pipelines to help extraction of structural, functional or diffusion
-MRI data across a large set of high resolution MR images. These
-preprocessing pipelines create results that are available in standard
-volume and combined surface and volume spaces which makes it easier
-for researchers to compare the images across the neuroimaging
-spectrum.
-
-\paragraph{Structural Pipelines} The structural pre-processing consists of PreFreeSurfer, FreeSurfer and PostFreeSurfer. 
-\paragraph{Functional Pipelines} Functional pre-processing consists of fMRIVolume and fMRISurface~\cite{FSL}. 
-\paragraph{Diffusion Preprocessing pipeline} preparing...
-
-\todo{Write a short summary of what the pipelines do.}
-
-\subsection{Processing environment}
-
-We executed the pipelines using Docker
-containers to simplify the deployment of different operating
-system versions on execution platforms. The Docker images were built for the HCP
- pre-processing pipelines v3.19.0 (PreFreeSurfer and
-FreeSurfer)~\cite{Glasser2013} in CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}. We
-collected the provenance trace (tree of executed processes and files
-accessed in read or write mode) for each type of subjects using system-call
-interception as provided by the \reprozip tool~\cite{Chirigati2016}.
-
-To support the computational challenge, we used the Canadian Brain
-Imaging Research Platform (CBRAIN)~\cite{cbrain} to manage the
-containers and data. CBRAIN is a Web platform developed at the
-Montreal Neurological Institute (MNI) to tackle the computational
-challenges faced by neuroimaging researchers.  Applications can be
-ported to CBRAIN using the Docker and Singularity container
-technologies through the Boutiques framework~\cite{boutiques}. Our
-CBRAIN HCP plugins are publicly available at \url{https://github.com/glatard/cbrain-plugins-hcp} for
-installation and re-use in CBRAIN instances.
-
-We processed subjects twice in every condition to identify possible \emph{intra-condition} errors coming from the use of pseudo-random numbers or other external
-factors.
-
-To further ensure that the observed errors are caused by operating system
-updates (\emph{inter-os differences}) and not by other factors, we
-implemented the following sanity checks:
-\begin{itemize}
-\item The file checksums of every
-subject are computed before and after processing to detect potential corruption during file transfers and other manipulations.
+\item The file checksums of every subject are computed before and after 
+processing to detect potential corruption during file transfers and 
+other manipulations. 
 \item The checksum of the Docker container images is recorded with
   every execution to identify any unintentional package update. In 
-  addition, the list of packages with versions is printed by every
-  task and checked for consistency.
+  addition, the list of packages with versions is printed by every task 
+  and checked for consistency.
 \item Hardware information is captured to make sure that errors are
@@
 
-\todo{Check ~\cite{DBLP:journals/fini/DasGRSPMSRSKMKR17}}
+\section{Measuring Differences}
 
-\subsection{Measuring differences}
+We compute a file difference matrix between files produced for every 
+subject in every operating system based on MD5 
+checksums~\cite{md51992url} as in~\cite{Scaria2017}. MD5 is suitable to 
+identify incidental file differences, although it is not 
+cryptographically secure. Files produced only in some subjects or in 
+some operating systems were left out.
 
-We computed a file difference matrix between files produced for every
-subject in every operating system based on MD5 checksums~\cite{md5}
-(\url{https://tools.ietf.org/html/rfc1321}) as
-in~\cite{Scaria2017}. MD5 is suitable to identify incidental file
-differences, although it is not cryptographically secure. Files
-produced only in some subjects or in some operating systems were left out.
+In addition, we are able to quantify errors in the image files using 
+the Dice coefficient and normalized root mean square error (NRMSE) 
+computed as follows:
 
-We quantified the error in the image files using the Dice coefficient and normalized root mean square error (NRMSE) computed as follows:
 \begin{center}
@@
 
-\subsection{Pipeline analysis: Characterizing errors}
+\section{Pipeline Analysis: Characterizing Errors}
 
@@
 identify the steps in the pipeline responsible for such differences.
-We use the \reprozip tool~\cite{reprozip} to record: (1) the tree of
+We use the \reprozip tool~\cite{Chirigati2016} to record: (1) the tree of
 processes executed by the pipeline and (2) the list of files read and
@@
 file dependencies between processes. A file dependency is defined 
-between processes A and B if a file written by A is read by
-B. Figure~\ref{fig:graph-example} shows an example of a process graph
+between processes A and B if a file written by A is read by B. 
+Figure~\ref{fig:simple_script} shows an example of a process graph 
 constructed at this stage.
@@
 	system calls). Dashed edges represent file dependencies: temporary 
-    files are in yellow and result files are in green.} \label{myfig1}
+	files are in yellow and result files are in green.} 
+	\label{fig:simple_script}
 \medskip
@@
 execution condition. Among our dataset, we identified 4 types of 
-subjects with different numbers of T1w and T2w images. We verified
-that the process graphs were identical for all subjects of the same
-type, for all execution conditions \todo{Check that}. The remainder of
-the analysis is done separately for each subject type.
+subjects with different numbers of T1w and T2w images. We verified that 
+the process graphs were identical for all subjects of the same type, 
+for all execution conditions \todo{Check that}. The remainder of the 
+analysis is done separately for each subject type.
 
@@
 \begin{enumerate}
-\item Processes 
-that read files that do not have errors and write files that do not 
-have errors are \emph{transparent}.
+\item Processes that read files that do not have errors and write files that do not 
+have errors are \emph{transparent} (Figure~\ref{fig:processes}.a).
 \item Processes that read files 
 that do not have any error but write files that have errors 
-\emph{create} errors in the pipeline.
+\emph{create} errors in the pipeline (Figure~\ref{fig:processes}.b).
 \item Processes that read files 
-that have errors and write files that do not have errors\emph{remove} 
-errors from the pipeline.
+that have errors and write files that do not have errors \emph{remove} 
+errors from the pipeline (Figure~\ref{fig:processes}.c).
+\item Processes that read files that have errors and write files that also have errors are 
+\emph{unknown} (Figure~\ref{fig:processes}.d).
 \end{enumerate}
 
-The classification of processes that read or write temporary files is 
-uncertain when these files are deleted during the execution. To address 
-this issue, we replace every process P that writes temporary files with 
-a modification of P that first calls P and then backs up all its output 
-files to a read-only directory. This replacement is done by modifying 
-the Unix PATH variable to point to a directory containing the modified
-versions of the processes. This solution does not cover the temporary files
-that are removed by P itself; this is not a problem since these files do not
-play any role in the subsequent steps of the pipeline, by definition.
+\begin{figure}[H]%\centering
+    \begin{subfigure}{0.15\textwidth}
+        \includegraphics[scale=0.35]{images/green.png}
+        \caption{Transparent}
+        \label{fig:green}
+    \end{subfigure}
+\hfill%\hfil
+    \begin{subfigure}{0.15\textwidth}
+    \includegraphics[scale=0.35]{images/red.png}
+    \caption{Creates error}
+    \label{fig:red}
+\end{subfigure}
+\hfill%\hfil
+    \begin{subfigure}{0.15\textwidth}
+    \includegraphics[scale=0.35]{images/blue.png}
+    \caption{Removes error}
+    \label{fig:blue}
+\end{subfigure}
+\hfill%\hfil
+    \begin{subfigure}{0.15\textwidth}
+    \includegraphics[scale=0.35]{images/yellow.png}
+    \caption{Unknown}
+    \label{fig:yellow}
+\end{subfigure}
+    \caption{Different type of classified process based on the input/output files.
+  Dashed edges refer to the file dependencies between processes A and B 
+  if a file written by A is read by B. Solid black edges refer to the 
+  relationship between parent and child processes.}
+    \label{fig:processes}
+\end{figure}
 
-Files written by multiple processes also lead to unknown classification 
-labels. To address this issue for a file F written by processes in 
-\textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we (1) check that processes 
-in \textbf{P} do not write concurrently to F, (2) we establish an order 
-on \textbf{P} based on the creation timestamp of the processes, (3) we 
-replace ever process $P_{i}$ in \textbf{P} by a wrapper that first 
-calls $P_{i}$ and then backs up F to a read-only directory. Thus, 
-multiple versions of F are saved and used in the analysis.
+\paragraph{Challenge1:} The classification of processes that read or 
+write temporary files is uncertain when these files are deleted during 
+the execution. To address this issue, we replace every process P that 
+writes temporary files with a modification of P that first calls P and 
+then backs up all its output files to a read-only directory. This 
+replacement is done by modifying the Unix PATH variable to point to a 
+directory containing the modified versions of the processes. This 
+solution does not cover the temporary files that are removed by P 
+itself; this is not a problem since these files do not play any role in 
+the subsequent steps of the pipeline, by definition.
 
-Another challenge is to classify processes that read files that have
-errors and write files that also have errors. In such situations, it
-is not possible to determine from the pipeline results whether
-the process created errors, removed some errors, or was
-transparent. We label such processes \emph{unknown}. To address this issue, we developed an iterative approach that consists of the following steps:
+\paragraph{Challenge2:} Files written by multiple processes also lead 
+to unknown classification labels. To address this issue for a file F 
+written by processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
+(1) check that processes in \textbf{P} do not write concurrently to F, 
+(2) we establish an order on \textbf{P} based on the creation timestamp 
+of the processes, (3) we replace ever process $P_{i}$ in \textbf{P} by 
+a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
+directory. Thus, multiple versions of F are saved and used in the 
+analysis.
+
+\paragraph{Challenge3:} Another challenge is to classify processes that 
+read files that have errors and write files that also have errors. In 
+such situations, it is not possible to determine from the pipeline 
+results whether the process created errors, removed some errors, or was 
+transparent. We label such processes \emph{unknown}. To address this 
+issue, we developed an iterative approach that consists of the 
+following steps:
+
 \begin{enumerate}
@@
 \end{enumerate}
-The replacement of process P at step 2 is done by replacing P
-with a custom script produced by step 1. This custom script
-copies the results obtained in condition 1 if it is invoked with the
-arguments of a process that created errors, and calls P's original
-executable otherwise. The replacement is done through the PATH variable, as before. This algorithm converges to a process graph
-without any \emph{unknown} process after a finite number of iterations.
 
-\begin{figure}
-  \includegraphics[scale=0.6]{images/iterative_modif}
+The replacement of process P at step 2 is done by replacing P with a 
+custom script produced by step 1. This custom script copies the results 
+obtained in condition 1 if it is invoked with the arguments of a 
+process that created errors, and calls P's original executable 
+otherwise. The replacement is done through the PATH variable, as 
+before. This algorithm converges to a process graph without any 
+\emph{unknown} process after a finite number of iterations.
+
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.5]{images/iterative_modif}
   \caption{iteration of proposed modification.}
@@
 errors. \texttt{4\#fslmaths} and \texttt{5\#fslstats} are classified as 
-unknown (yellow) as they produce files with errors from files with errors.
+unknown (yellow) as they produce files with errors from files with 
+errors.
 
@@
 condition are replaced with the files produced by \texttt{2\#bet2} in 
-the other condition. \texttt{3\#fslstats} is now classified as transparent, \texttt{4\#fslmaths} is now classified
-as error creator and \texttt{5\#fslstats} is still unknown.
+the other condition. \texttt{3\#fslstats} is now classified as 
+transparent, \texttt{4\#fslmaths} is now classified as error creator 
+and \texttt{5\#fslstats} is still unknown.
 
@@
 summarize them by (1) building subgraphs for every interesting node 
-(e.g. red and blue nodes with their children) (2) building connected components 
-in the absence of interesting nodes for the remaining nodes in the process graph. 
-Each node in the summarized graph represent either a shrinked interesting subgraph or 
-connected component and edges represent any connection between two shrinked nodes.
+(e.g. red and blue nodes with their children) (2) building connected 
+components in the absence of interesting nodes for the remaining nodes 
+in the process graph. Each node in the summarized graph represent 
+either a shrinked interesting subgraph or connected component and edges 
+represent any connection between two shrinked nodes.
+
+\section{HCP Minimal preprocessing Pipelines}
+
+The Human Connectome Project (HCP)~\cite{glasser2013minimal} developed 
+a set of pipelines to help extraction of structural, functional or 
+diffusion MRI data across a large set of high resolution MR images. 
+These pre-processing pipelines create results that are available in 
+standard volume and combined surface and volume spaces which makes it 
+easier for researchers to compare the images across the neuroimaging 
+spectrum.
+
+We focused on HCP structural pre-processing pipelines to (1)quantify 
+the effect of OS on HCP pre-processing pipelines (2) identify the 
+process(es) in the pipelines that are responsible for such effects.
+
+\paragraph{Structural Pre-processing Pipelines} The structural 
+pre-processing consists of PreFreeSurfer, FreeSurfer and 
+PostFreeSurfer~\cite{glasser2013minimal}. In this paper, we analyzed two 
+pipelines, PreFreeSurfer and FreeSurfer.
+
+PreFreeSurfer consist of various steps including correction of MR 
+gradient nonlinearity distortions, align the T1w and T2w images using 
+FSL FLIRT, Align native space to MNI template using ACPC and FSL FLIRT, 
+brain extraction usin FSL FLIRT and FNIRT to MNI template, register T2w 
+to T1w using FLIRT's BBR, perform a bias field correction, and register 
+the subject's native structural space to MNI space.
+
+FreeSurfer steps are downsample the images if it exceeds the resolution limit, 
+generation of white matter surface using a segmentation of downsampled image, 
+adjust T1w colum intensity and white matter surface position based on 0.7mm 
+T1w images, fine tune the T2w to T1w registration using FreeSurfer's BBRegister 
+algorithm, accurately place the white and pial surface with high-resolution 0.7mm 
+T1w and T2w data, these inputs are fed back into recon-all at each stage.
+
+We randomly selected 10 subjects from the HCP data release S500 which 
+are described in table~\ref{table:data}.
+
+\begin{table}[H]\hspace*{-1cm}
+\centering
+\begin{threeparttable}
+\caption{An overview of the HCP data (Humman Connectome DB).}
+
+\begin{tabular}{@{}llllll@{}}
+\toprule
+Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
+103515  & Q1      & Q02         & F      & 26-30    & type1         \\
+105216  & Q3      & Q03         & M      & 26-30    & type4         \\
+103414  & Q2      & Q02         & F      & 22-25    & type1         \\
+125525  & Q1      & Q01         & F      & 31-35    & type1         \\
+142828  & Q1      & Q01         & M      & 31-35    & type1         \\
+129533  & Q3      & Q04         & F      & 31-35    & type1         \\
+103818  & Q1      & Q01         & F      & 31-35    & type3         \\
+133928  & Q2      & Q03         & M      & 26-30    & type3         \\
+148032  & Q3      & Q03         & F      & 31-35    & type2         \\
+139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
+\end{tabular}
+\begin{tablenotes}
+      \small
+      \item *Subj\_Type refers to the graph topology type of the subjects,
+      we will discuss further on the next sections.
+\end{tablenotes}
+\end{threeparttable}
+\label{table:data}
+\end{table}
 
@@
 
-Show lists of packages used with version. Focus on important
-differences (e.g.: python interpreter, gcc if relevant) and explain
-why these packages are important.
+We executed the pipelines using Docker containers to simplify the 
+deployment of different operating system versions on execution 
+platforms. The Docker images were built for the HCP pre-processing 
+pipelines v3.19.0 (PreFreeSurfer and FreeSurfer) in 
+CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for 
+reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}. 
+We collected the provenance trace (tree of executed processes and files 
+accessed in read or write mode) for each type of 
+subjects(Table~\ref{table:data}) using system-call interception as 
+provided by the \reprozip tool~\cite{Chirigati2016}.
 
-Show results of the comparison.
+\paragraph{} Two types of errors can occur in the subjects due to the 
+errors in the operating systems. One is inter-OS error caused by the 
+operating system library updates and the other type, intra-OS errors 
+occurs as a result of the pseudo-random processes used in the 
+pipelines. An example of a pseudo-random process function is, a random 
+number generator that would get initialized using a seed state. the 
+proposed method can be used to identify both kind of errors. The files 
+that are common to all the subjects only are taken into consideration 
+for comparison. The first step is identification of files with 
+differences in their checksums. This is identified using the checksums 
+that are recorded after the processing. Intra-OS errors are identified 
+using the run-number added as the suffix for the conditions. For 
+example, the two batches of subjects processed under the same condition 
+(CentOS6) are stored as run-1 and run-2. The files belonging to the 
+subjects stored under the above mentioned conditions are treated as 
+intra-OS runs.
 
@@
 
-Among the 117 data files produced by PreFreeSurfer, 21 did not have any error for any subject, 92 had errors 
-for all subjects and 4 had errors for 3 subjects only.
+In a previous study~\cite{Scaria2017}, we showed that pre-processing 
+pipelines of the Human Connectome Project~\cite{Glasser2013} were 
+sensitive to operating system variation so that Figure 
+\ref{fig:tissue_class} illustrates the binarized errors after brain tissue 
+segmentation process in PreFreeSurfer. In addition, 
+Figure~\ref{fig:fnirt_result} shows the same errors occurred in 
+non-linear registration process between CentOS6 and CentOS7. 
+Figure~\ref{fig:pipeline_errors}.a.
 
-In a previous study~\cite{Scaria2017}, we showed that
-pre-processing pipelines of the Human Connectome
-Project~\cite{Glasser2013} were sensitive to operating system
-variation (see Figure \ref{fig:1}).
-\begin{figure}
-%  \includegraphics{brain\_classification}
-  \includegraphics[width=0.5\textwidth, height=6cm]{images/brain_classification.png}
-  \caption{Tissue classification produced by the HCP pre-processing
-    pipelines on subject 105216 (CentOS 6 vs CentOS 7).}
-  \label{fig:1}
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.6]{images/fnirt_result.png} 
+  \caption{Absolute errors between FNIRT results from PreFreeSurfer 
+  (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (centos6 vs. 
+  centos7)~\cite{Scaria2017}. } 
+  \label{fig:fnirt_result}
 \end{figure}
 
-\paragraph{}
-Two types of differences can occur in the subjects due to the differences in the operating systems. 
-One is inter-OS difference caused by the operating system library updates and the other type,
- intra-OS differences occurs as a result of the pseudo-random processes used in the pipelines.
- An example of a pseudo-random process function is, a random number generator that would get initialized 
-using a seed state. the proposed method can be used to identify both kind of differences. 
-The files that are common to all the subjects only are taken into consideration for comparison. The first step
- is identification of files with differences in their checksums. This is identified using the checksums that are
- recorded after the processing. Intra-OS differences are identified using the run-number added as the suffix for
- the conditions. For example, the two batches of subjects processed under the same condition (CentOS6) are stored 
-as run-1 and run-2. The files belonging to the subjects stored under the above mentioned conditions are treated as intra-OS runs.
+\begin{figure}[H]
+%  \includegraphics{brain\_classification}
+\centering
+  \includegraphics[scale=0.18]{images/brain_classification.png} 
+  \caption{Binarized error between brain segmentation results from 
+  FreeSurfer, subject 105216 (CentOS6 vs. CentOS7)~\cite{Scaria2017}.
+    } 
+  \label{fig:tissue_class}
+\end{figure}
 
@@
 
-Figure \ref{fig:summarized-graph} shows the annotated provenance graph of the
-PreFreeSurfer pipeline executed on CentOS6 and CentOS7.  Processes
-that created errors are shown in red, processes that removed errors
-are in blue, and other processes are in green.  Squares denote
+We identified processes that introduce errors in both PreFreeSurfer and 
+FreeSurfer pipelines. The result of our experiment is described in the 
+following paragraphs.
+
+\paragraph{PreFreeSurfer.} Among the 117 data files produced by 
+PreFreeSurfer, 21 did not have any error for any subject, 92 had errors 
+for all subjects and 4 had errors for 3 subjects only.
+
+Figure \ref{fig:summarized-graph} shows the annotated provenance graph 
+of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
+node in the graph represent an executed process in the pipeline. 
+Processes that created errors are shown in red, processes that removed 
+errors are in blue, and other processes are in green.  Squares denote 
 processes for which the classification is uncertain, due to temporary 
@@
 dependencies between processes (green edges: files with no errors; red 
-edges: files with errors; yellow edges: temporary files).  The
-processes that introduce errors in PreFreeSurfer are: linear
-registration with “\emph{FLIRT}” (8 occurrences, in ACPC Alignment,
-BrainExtraction, DistortionCorrection, AtlasRegistration), non-linear
-registration with “\emph{FNIRT}” (3 occurrences, in BrainExtraction
-and AltasRegistration), image warping with “\emph{new\_invwarp}” (3
-occurrences, in BrainExtraction and AtlasRegistration).  In addition,
-errors were observed in image mean and standard-deviation computations
-with “\emph{fslstats}” (3 occurrences in BiasFieldCorrection), and in
-masked image extrapolation with “\emph{fslmaths}” (1 occurrence, in
+edges: files with errors; yellow edges: temporary files).
+
+The processes that introduce errors in PreFreeSurfer along with the 
+number of occurrences are depicted in Figures~\ref{fig:pfs_table, 
+fig:pfs_chart} including linear registration with “\emph{FLIRT}” (in 
+ACPC Alignment, BrainExtraction, DistortionCorrection, 
+AtlasRegistration), non-linear registration with “\emph{FNIRT}” (in 
+BrainExtraction and AltasRegistration), image warping with 
+“\emph{new\_invwarp}” (in BrainExtraction and AtlasRegistration).  In 
+addition, errors were observed in image mean and standard-deviation 
+computations with “\emph{fslstats}” (in BiasFieldCorrection), and in 
+masked image extrapolation with “\emph{fslmaths}” (in 
 BiasFieldCorrection).  Besides, transformation format conversion with 
-“\emph{convertwarp}” (2 occurrences, in DistortionCorrection) was able
-to remove errors.
+“\emph{convertwarp}” (in DistortionCorrection) was able to remove 
+errors.
 
-\todo{Check if a given file is written by only one process (add a safeguard in your code).}
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.5]{images/pfs_table.png}
+  \caption{Process name along with the pipeline elements that introduce errors.}
+  \label{fig:pfs_table}
+\end{figure}
 
-\begin{figure}
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.38]{images/pfs_chart.png} \caption{Number 
+  of occurrences of errors in PreFreeSurfer on subject 103515, red and 
+  blue bars indicate the processes which create and remove errors 
+  respectively.} 
+  \label{fig:pfs_chart}
+\end{figure}
+
+\paragraph{FreeSurfer.} In the second analysis, the result of 
+PreFreeSurfer is considered as the input files of FreeSurfer pipeline. 
+From the FreeSurfer results, total 61 files were found to be different 
+between CentOS6 and CentOS7. The processes that introduce errors in 
+FreeSurfer are shown on Figure~\ref{fig:fs_error_table}. Most frequent 
+processes that create errors are \emph{mris\_anatomical\_stats} which 
+measures a number of anatomical properties, \emph{mris\_make\_surface} 
+to generate pial surfaces, first pass, no T2 adjustment, 
+\emph{mri\_convert} which normalize T1w and T2w images for the benefit 
+of \emph{mris\_make\_surfaces}, and \emph{fslmaths} create and apply a 
+binary mask to filling of holes to remove  any non-grey/white tissue 
+using a threshold created by FSLSTATS at Mean. In addition, all the 
+other processes that created errors in the pipeline at least once are 
+categorized as others such as \emph{mri\_ca\_register}, 
+\emph{mri\_ca\_normalize}, and \emph{mri\_watershed}. Furthermore, 
+processes that remove errors in FreeSurfer represented on 
+Figure~\ref{fig:fs_remove_table} including \emph{mri\_convert} and 
+\emph{mri\_surf2surf}.
+
+\begin{figure}[H] 
+\centering
+  \includegraphics[scale=0.6]{images/fs_error_table.png} 
+  \caption{Number of occurrences of errors in FreeSurfer pipeline, 
+subject 103818, indicate the processes that create errors.}
+  \label{fig:fs_error_table}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.5]{images/fs_remove_table.png} 
+  \caption{Number of occurrences of errors in FreeSurfer pipeline, 
+subject 103818, indicate the processes that remove errors.}
+  \label{fig:fs_remove_table}
+\end{figure}
+
+\section{Discussion}
+
+pipeline amplify small numerical differences, they are numerically 
+unstable. Furthermore, math libraries evolve over time, leading to 
+different numerical errors. we listed some of the irreproducibility 
+causes of the pipelines as we mentioned in the previous section 
+overally along with exact command arguments.
+
+Mention that this was only possible because the unprocessed data was shared in the first place.
+DICOM to Nifti conversion was out of scope and may introduce other issues.
+
+\section{Conclusion}
+
+Our thechnique is able to characterize the stability of a pipeline's components automatically.
+The numerical instability in the PreFreesurfer HCP pipeline arises mainly from
+linear and non-linear registration processes implemented in FSL FLIRT and FNIRT.
+
+There are a few ways to impede such instabilities:
+\begin{itemize}
+\item Use a single operating system
+\item Containerize pipelines
+\item Increase numerical precision
+\item Be stricter on truncation and rounding standards (IEEE 754)
+\item Build static executable
+\end{itemize}
+
+The results still suffer from small pertubations literally because of the fact that pipeline are not numerically stable.
+The preferred solution is to detect and fix numerical instability of the pipeline instead of masking the problem.
+These processes need to be reviewed to understand and correct the cause of instabilities. 
+In this correction process, accuracy has to be considered in addition to stability.
+
+\item In the future works we aim to measure the impact of such errors
+on the specific domain of the neuroscience and how show the significant it is.
+
+\item The purpose of the project is generalizing pipelines to get a reproducible execution 
+on every operating system instead of choosing one main OS to.
+
+\section{Acknowledgments}
+
+CBRAIN team. Compute Canada(Calcul Quebec).
+
+\begin{figure}[H]
   \includegraphics[width=\linewidth]{images/summarized_graph}

### 2019-03-19 paper.tex

@@
 
-\newcommand{\note}[2]{\pdfmargincomment[color=yellow,author=#1,open=true]{#2}}
 \newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
+\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
+
 \newcommand{\reprozip}[0]{ReproZip}
 
-\title{Numerical error propagation in the HCP structural pre-processing 
-pipelines}
 \title{Reprotool: a debugging tool to identify numerical errors 
-and assess reproducibility of the neuroimaging pipelines}
+and assess reproducibility of pipelines}
 
@@
 Many experiments show that computational analysis results are still not 
-completely reproducible. Although new technologies such as 
-virtualization , version control systems , and provenance management 
-tools provide more reliable environment and significantly improve the 
-reproducibility crisis. 
-In particular, computational environments are known to have an effect 
-on the results produced by neuroimaging pipelines, presumably due to 
-the creation, propagation and amplification of small numerical errors 
-across the pipelines. Similarly, small perturbations that originated 
-from computational environments including diverse operating systems, 
-hardware categories, software version, etc. highlight the numerical 
-errors.
+completely reproducible. In particular, computational environments 
+including different operating systems, hardware categories, and 
+software versions are known to have effects on the results produced by 
+analysis pipelines. These effects are presumably due to the creation, 
+propagation and amplification of small numerical errors across the 
+pipelines. 
+Although new techniques such as virtualization , version 
+control systems , and provenance management tools provide more reliable 
+environment and significantly improve the reproducibility of scientific 
+findings. 
 However, the precise causes of such instabilities and the path along 
 which they propagate in the pipelines are unclear.  We present a 
-technique to identify the processes in the pipeline that create numerical 
-errors along the execution, and we apply this technique to the HCP 
-structural pre-processing pipelines.
+technique to identify the processes in the pipeline that create 
+numerical errors along the execution, and we apply this technique to 
+the HCP structural pre-processing pipelines.
 
@@
 
-Research findings are expected to be reproducible to evaluate the 
-authenticity and reliability of the claims.There are different 
-terminologies for the terms reproducibility, repeatability, and 
-replicability over the time. For example, the Association for Computing 
-Machinery (ACM)~\cite{acm2016terminology} proposes three different 
-definitions. From the ACM's points of view, repeatability is defined as 
-repeating the computation by the same experimental setup including 
-operator team, operating conditions, location, and measuring system. 
-Replicability has almost the similar definition with repeatability 
-which use the identical experimental conditions except performer team 
-which means that an independent group can achieve the same results. 
-Beside, reproducibility is measured by performing different 
-experimental setup via different laboratory team independently. In this 
-paper, we follow Peng's~\cite{peng2011reproducible} definition of 
-reproducibility which is assessed by performing the experiment using 
-the same data and the same analytic methods and expecting the same 
-results. A reproducible study provides a context in which, one is able 
-to get the consistence results to the original work. In addition, it 
-enables others to use the existing work as a part of their 
-experiments~\cite{plesser2018reproducibility}. It not only saves a 
-great deal of time, but also contributes to focusing on the other 
-parts.
+Research findings are expected to be reproducible so that 
+their authenticity and reliability can be evaluated. 
+There are different definitions for the terms reproducibility, 
+and replicability, which leads to confusion since the 
+same words are used for different concepts.
 
-Reproducibility of computational pipelines has been repeatedly 
-investigated so far. The comparisons show that the analysis results are 
-not completely reproducible across the small perturbations which have 
-emerged by changing the execution environments including operating 
-system, software and analysis method, and hardware configuration. There 
-are different studies that discovered factors that introduce such tiny 
-perturbations and consequently hamper the reproducibility of the 
-experiments. In particular, Groenschild \emph{et al.} first identified 
-the effect of operating systems on Freesurfer 
-results~\cite{Gronenschild2012}. Similarly, the effect of OS is 
-quantified in~\cite{Glatard2015} more specifically on some 
-of the main neuroimaging pipelines including several FSL pipelines, 
-CIVET and Freesurfer. The results of comparisons indicate a significant 
-errors between analysis results of the same experiment on different 
-operating system.
+According to Peng's definition~\cite{peng2011reproducible}, 
+reproducibility is defined as the ability to regenerate the same 
+results as the original findings when the experiment is reanalyzed 
+given exactly the same analytic methods and data. Reproducibility 
+ensures that independent scientists can reproduce the same results 
+using the same data and procedure as published in the original 
+publication. On the contrary, replicability is defined as the ability 
+to obtain similar results as published in the original study when 
+the experiment is reimplemented using independent data and analytic 
+methods. Replicability confirms scientific claims and ensures that 
+independent investigators can produce consistent results, using new 
+data and methods.
 
-It should be noted that \emph{error} is defined as the statistical 
-differences between results of two experiments in this work. Moreover, 
-the operating system is defined here as a consistent set of software 
-packages organized in a trusted repository.
+In addition, numerical reproducibility is defined as the ability to 
+regenerate bit for bit identical results from multiple 
+runs~\cite{hill2017numerical}. Two files will be considered numerically 
+reproducible if they have identical binary contents. Binary 
+comparison is calculated by comparing checksums. It must be pointed out 
+that a computation might be reproducible based on Peng's definition, 
+but not be numerically reproducible. For instance, small numerical 
+errors created during the pipeline execution may hamper numerical 
+reproducibility, but be negligible in the final results.
 
-The operating system is not the only part of the computational 
+Reproducibility, as the cornerstone of scientific research, guarantees 
+the reliability of results. A reproducible study provides a context in 
+which one is able to get results that are consistent with the original 
+work. In addition, it not only saves a great deal of time, but also 
+enables others to use existing works as a part of their 
+experiments~\cite{plesser2018reproducibility}. 
+In this paper, we follow Peng's definition of reproducibility 
+unless we directly refer to numerical reproducibility. 
+We seek to identify the reasons why such reproducibility may not be 
+ensured, focusing in particular on computational aspects.
+
+Recently, scientists began to realize that the results of many 
+scientific experiments were neither replicable nor reproducible. 
+Results show that the variety of computational infrastructures 
+including workstation types, parallelization methods, operating 
+systems, and analysis packages are known to influence reproducibility 
+because of the creation of small numerical 
+errors~\cite{Gronenschild2012, diethelm2012limits, Glatard2015}. The 
+propagation and amplification of these tiny errors by analysis 
+pipelines may cause reproducibility issues. 
+\note{In this case, the analysis pipelines are said to be numerically 
+unstable. Numerical instability is a characteristic of the pipelines 
+which amplify small numerical errors and then hamper the 
+reproducibility of the analyses depending on the length of the pipeline 
+and magnitude of the errors. In many cases, numerical instability is an 
+important issue for reproducibility. }
+In particular, the effect of operating system is quantified 
+in~\cite{Glatard2015, Gronenschild2012} more specifically on some of 
+the main neuroimaging pipelines including FSL, CIVET and Freesurfer 
+pipelines. The results of comparisons indicate a significant errors 
+between analysis results of the same experiment on different operating 
+system. It should be noted that \emph{error} is defined as the binary 
+differences between results of two analyses in this paper, which is 
+calculated by the checksum method.
+
+\note{The operating system is not the only part of the computational 
 environment that may hamper reproducibility. Many other computational 
 factors including hardware configuration, software version, and 
-parallelization techniques are reported~\cite{diethelm2012limits} as 
-the main influential computing infrastructure on reproducibility.
+parallelization techniques are reported as 
+the main influential computing infrastructure on reproducibility. }
 
-It has revealed that the majority of analysis results are not 
-numerically reproducible because of arithmetic operations which is 
-associated with the type of execution such as sequential or parallel 
-executions, floating-point operations, different standard calculations 
-like IEEE-754 or other standards, etc.
+There are different solutions to make reproducible analysis including 
+containerization techniques to encapsulate software/hardware 
+dependencies, provenance capturing tools, and version control systems. 
+However, a comprehensive solution requires to fix the numerical 
+instabilities. For this purpose, we introduce a framework to identify 
+the processes in pipeline that create numerical errors across different 
+runs. In this paper, the execution results of pipelines across 
+different operating systems are considered to find out the cause of the 
+irreproducibility using an interposition techniques.
 
-In order to overcome these issues and making reproducible analysis, a 
-comprehensive solution requires which not only encapsulates and 
-re-executes all the software/hardware dependencies but also fixes the 
-numerical instabilities. Therefore, we introduce a framework to 
-identify the processes in the pipeline which are responsible for such 
-numerical errors along the execution. In this paper, the execution 
-results of the pipelines across different operating systems are 
-considered to find out the cause of the reproducibility issues using an 
-interposition techniques.
-
-Figure~\ref{fig:overview} illustrate an overview of the proposed 
+Figure~\ref{fig:overview} illustrates an overview of the proposed 
 technique. First, we execute pipelines in a consistence environment 
 using docker container and Boutiques framework. In the next step, the 
-origin of errors, processes that make differences, are detected through 
-a classification approach. Processes are classified by several 
-modifications iteratively and finally, it creates corresponding process 
-graph of the pipeline execution. We will explain each step in detail in 
-the following sections.
+processes that create differences are classified using an iterative 
+approach. Finally, we make corresponding process graph of the 
+pipeline execution that shows the processes related to the creation of 
+errors. We will explain each step in detail in the following sections.
 
@@
   \includegraphics[scale=0.5]{images/overview.png}
-  \caption{An overview of proposed technique.}
+  \caption{An overview of the proposed technique.}
   \label{fig:overview}
@@
 
-\paragraph{} The reminder of this paper explains the proposed debugging tool in two 
-parts. First, the execution environment of the pipelines and then the 
-classification procedure of the processes are clarified. We describe 
-the HCP preprocessing pipelines as a popular neuroimaging pipelines to 
-test and apply reprotool. Finally, we report the analytic process which 
-are responsible for the errors in the pipelines.
+The reminder of this paper explains the Reprotool in different parts as 
+the execution environment of the pipelines, methods to measure 
+differences, and how to characterize errors. We describe the HCP 
+preprocessing pipelines as a popular neuroimaging pipelines to test and 
+apply the Reprotool on. Finally, we report the processes that are 
+responsible for the errors in the pipelines.
 
@@
 the process graphs were identical for all subjects of the same type, 
-for all execution conditions \todo{Check that}. The remainder of the 
+for all execution conditions. The remainder of the 
 analysis is done separately for each subject type.
@@
 
-\paragraph{} In the next step, we classify the processes into three 
+In the next step, we classify the processes into three 
 categories based on the process graph and the error matrix file 
@@
 results whether the process created errors, removed some errors, or was 
-transparent. We label such processes \emph{unknown}. To address this 
-issue, we developed an iterative approach that consists of the 
-following steps:
+transparent. We label such processes \emph{unknown}.
+
+To address Challenge 3, we developed an iterative approach that 
+consists of the following steps:
 
@@
 The three steps in Figure~\ref{fig:iterations} correspond to the 
-iterations of the classification algorithm At step one \texttt{2\#bet2} 
+iterations of the classification algorithm. At step one, \texttt{2\#bet2} 
 is classified as error creator (red) as it produced files with errors 
@@
 
-\paragraph{Structural Pre-processing Pipelines} The structural 
-pre-processing consists of PreFreeSurfer, FreeSurfer and 
-PostFreeSurfer~\cite{glasser2013minimal}. In this paper, we analyzed two 
-pipelines, PreFreeSurfer and FreeSurfer.
-
+The structural pre-processing pipelines consists of PreFreeSurfer, 
+FreeSurfer and PostFreeSurfer. In this paper, 
+we analyzed two pipelines, PreFreeSurfer and FreeSurfer. 
 PreFreeSurfer consist of various steps including correction of MR 
@@
 
-FreeSurfer steps are downsample the images if it exceeds the resolution limit, 
-generation of white matter surface using a segmentation of downsampled image, 
-adjust T1w colum intensity and white matter surface position based on 0.7mm 
-T1w images, fine tune the T2w to T1w registration using FreeSurfer's BBRegister 
-algorithm, accurately place the white and pial surface with high-resolution 0.7mm 
-T1w and T2w data, these inputs are fed back into recon-all at each stage.
+FreeSurfer steps are downsample the images if it exceeds the resolution 
+limit, generation of white matter surface using a segmentation of 
+downsampled image, adjust T1w colum intensity and white matter surface 
+position based on 0.7mm T1w images, fine tune the T2w to T1w 
+registration using FreeSurfer's BBRegister algorithm, accurately place 
+the white and pial surface with high-resolution 0.7mm T1w and T2w data, 
+these inputs are fed back into recon-all at each stage.
 
@@
 
-\paragraph{} Two types of errors can occur in the subjects due to the 
+Two types of errors can occur in the subjects due to the 
 errors in the operating systems. One is inter-OS error caused by the 
@@
 
-\subsection{Inter-OS differences}
+\subsection{Inter-OS binary differences}
 
-\paragraph{Binary differences}
-
-In a previous study~\cite{Scaria2017}, we showed that pre-processing 
+In the previous study~\cite{Scaria2017}, we showed that pre-processing 
 pipelines of the Human Connectome Project~\cite{Glasser2013} were 
 sensitive to operating system variation so that Figure 
-\ref{fig:tissue_class} illustrates the binarized errors after brain tissue 
-segmentation process in PreFreeSurfer. In addition, 
+\ref{fig:tissue_class} illustrates the binarized errors after brain 
+tissue segmentation process in FreeSurfer. In addition, 
 Figure~\ref{fig:fnirt_result} shows the same errors occurred in 
-non-linear registration process between CentOS6 and CentOS7. 
-Figure~\ref{fig:pipeline_errors}.a.
-
-\begin{figure}[H]
-\centering
-  \includegraphics[scale=0.6]{images/fnirt_result.png} 
-  \caption{Absolute errors between FNIRT results from PreFreeSurfer 
-  (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (centos6 vs. 
-  centos7)~\cite{Scaria2017}. } 
-  \label{fig:fnirt_result}
-\end{figure}
+non-linear registration process between CentOS6 and CentOS7 in PreFreeSurfer.
 
@@
 
+\begin{figure}[H]
+\centering
+  \includegraphics[scale=0.6]{images/fnirt_result.png} 
+  \caption{Absolute errors between FNIRT results from PreFreeSurfer 
+  (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (CentOS6 vs. 
+  CentOS7)~\cite{Scaria2017}. } 
+  \label{fig:fnirt_result}
+\end{figure}
+
 \subsection{Pipeline analysis}
@@
 
-\paragraph{PreFreeSurfer.} Among the 117 data files produced by 
+\subsubsection{PreFreeSurfer pipeline}
+
+Among the 117 data files produced by 
 PreFreeSurfer, 21 did not have any error for any subject, 92 had errors 
@@
   \includegraphics[scale=0.5]{images/pfs_table.png}
-  \caption{Process name along with the pipeline elements that introduce errors.}
+  \caption{Process name along with the pipeline elements that introduce 
+  errors.}
   \label{fig:pfs_table}
@@
 
-\paragraph{FreeSurfer.} In the second analysis, the result of 
+\subsubsection{FreeSurfer pipeline}
+
+In the second analysis, the result of 
 PreFreeSurfer is considered as the input files of FreeSurfer pipeline. 
@@
 
-pipeline amplify small numerical differences, they are numerically 
+\note{Pipeline amplify small numerical differences because they are numerically 
 unstable. Furthermore, math libraries evolve over time, leading to 
@@
 
-Mention that this was only possible because the unprocessed data was shared in the first place.
-DICOM to Nifti conversion was out of scope and may introduce other issues.
+Mention that this was only possible because the unprocessed data was 
+shared in the first place. DICOM to Nifti conversion was out of scope 
+and may introduce other issues.}
 
@@
 
-Our thechnique is able to characterize the stability of a pipeline's components automatically.
-The numerical instability in the PreFreesurfer HCP pipeline arises mainly from
-linear and non-linear registration processes implemented in FSL FLIRT and FNIRT.
+\note{Our technique is able to characterize the stability of a pipeline's 
+components automatically. The numerical instability in the 
+PreFreesurfer HCP pipeline arises mainly from linear and non-linear 
+registration processes implemented in FSL FLIRT and FNIRT.
 
@@
 
-The results still suffer from small pertubations literally because of the fact that pipeline are not numerically stable.
-The preferred solution is to detect and fix numerical instability of the pipeline instead of masking the problem.
-These processes need to be reviewed to understand and correct the cause of instabilities. 
-In this correction process, accuracy has to be considered in addition to stability.
-
-\item In the future works we aim to measure the impact of such errors
-on the specific domain of the neuroscience and how show the significant it is.
-
-\item The purpose of the project is generalizing pipelines to get a reproducible execution 
-on every operating system instead of choosing one main OS to.
+The results still suffer from small perturbations literally because of 
+the fact that pipeline are not numerically stable. The preferred 
+solution is to detect and fix numerical instability of the pipeline 
+instead of masking the problem. These processes need to be reviewed to 
+understand and correct the cause of instabilities. }
 

### 2019-04-11 paper.tex
re-write in GigaScience journal format

@@
-\documentclass{article}
+% GigaScience template
+\documentclass[a4paper,num-refs]{oup-contemporary}
 
-\usepackage{pdfcomment}
-\usepackage[margin=1in]{geometry}
+\journal{gigascience}
+
+%%%% Packages %%%%
+\usepackage{siunitx}
+\usepackage{minted} % Used for JSON highlighting
+\usepackage{algpseudocode} % Algorithmic environment
+\usepackage{xspace}
+\usepackage{booktabs}
+
+%%%%%%
+
 \usepackage{xcolor}
@@
 \usepackage{verbatim}
-\usepackage{booktabs}
 \usepackage{makecell}
-
 \usepackage[flushleft]{threeparttable}
 
+%%%% Commands %%%%
 \newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
 \newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
-
 \newcommand{\reprozip}[0]{ReproZip}
 
-\title{Reprotool: a debugging tool to identify numerical errors 
+\title{NRM-tool (Numerical Reproducibility Measurement): 
+a debugging tool to identify numerical errors 
 and assess reproducibility of pipelines}
-
-\author{M. Ali Salari, Lalet Scaria, Gregory Kiar, Lindsay B. Lewis,
-  Alan C. Evans, Tristan Glatard}
 
@@
 
+\author[1]{Ali Salari}
+\author[1]{Lalet Scaria}
+\author[2,3]{Gregory Kiar}
+\author[2]{Lindsay Lewis}
+\author[1]{Tristan Glatard}
+\author[2,3]{Alan C. Evans}
+
+\affil[1]{Department of Computer Science and Software Engineering, Concordia University, Montreal, Canada}
+\affil[2]{McGill University, Montreal, Canada}
+\affil[3]{Montreal Neurological Institute, Montreal, Canada}
+
 \maketitle
 
-\abstract{
-
+\begin{abstract}
 Many experiments show that computational analysis results are still not 
@@
 
+\end{abstract}
+
+\begin{keywords}
+Reproducibility; Numerical Instability; Neuroimaging.
+\end{keywords}
+
 \section{Introduction}
 
-Research findings are expected to be reproducible so that 
-their authenticity and reliability can be evaluated. 
-There are different definitions for the terms reproducibility, 
-and replicability, which leads to confusion since the 
-same words are used for different concepts.
-
-According to Peng's definition~\cite{peng2011reproducible}, 
-reproducibility is defined as the ability to regenerate the same 
-results as the original findings when the experiment is reanalyzed 
-given exactly the same analytic methods and data. Reproducibility 
-ensures that independent scientists can reproduce the same results 
-using the same data and procedure as published in the original 
-publication. On the contrary, replicability is defined as the ability 
-to obtain similar results as published in the original study when 
-the experiment is reimplemented using independent data and analytic 
-methods. Replicability confirms scientific claims and ensures that 
-independent investigators can produce consistent results, using new 
-data and methods.
+A reproducible study provides a context in which one is able to get 
+results that are consistent with the original 
+work~\cite{plesser2018reproducibility}. Research findings are expected 
+to be reproducible so that their authenticity and reliability can be 
+evaluated. 
+There are different definitions for the term reproducibility, we follow 
+Peng's definition~\cite{peng2011reproducible} in which reproducibility 
+is defined as the ability to regenerate the same results as the 
+original findings when the experiment is reanalyzed given exactly the 
+same analytic methods, software package, parameters, and data.
 
@@
 runs~\cite{hill2017numerical}. Two files will be considered numerically 
-reproducible if they have identical binary contents. Binary 
-comparison is calculated by comparing checksums. It must be pointed out 
-that a computation might be reproducible based on Peng's definition, 
-but not be numerically reproducible. For instance, small numerical 
-errors created during the pipeline execution may hamper numerical 
+reproducible if they have identical binary contents. Binary comparison 
+is calculated by comparing checksums. It must be pointed that a 
+computation might be reproducible based on Peng's definition, but not 
+be numerically reproducible. For instance, small numerical errors 
+created during the pipeline execution may hamper numerical 
 reproducibility, but be negligible in the final results.
 
-Reproducibility, as the cornerstone of scientific research, guarantees 
-the reliability of results. A reproducible study provides a context in 
-which one is able to get results that are consistent with the original 
-work. In addition, it not only saves a great deal of time, but also 
-enables others to use existing works as a part of their 
-experiments~\cite{plesser2018reproducibility}. 
-In this paper, we follow Peng's definition of reproducibility 
-unless we directly refer to numerical reproducibility. 
-We seek to identify the reasons why such reproducibility may not be 
-ensured, focusing in particular on computational aspects.
-
 Recently, scientists began to realize that the results of many 
-scientific experiments were neither replicable nor reproducible. 
-Results show that the variety of computational infrastructures 
-including workstation types, parallelization methods, operating 
-systems, and analysis packages are known to influence reproducibility 
-because of the creation of small numerical 
-errors~\cite{Gronenschild2012, diethelm2012limits, Glatard2015}. The 
-propagation and amplification of these tiny errors by analysis 
-pipelines may cause reproducibility issues. 
-\note{In this case, the analysis pipelines are said to be numerically 
+scientific experiments were not reproducible. Results show that the 
+variety of computational infrastructures including workstation types, 
+parallelization methods, operating systems, and analysis packages are 
+known to influence reproducibility because of the creation of small 
+numerical errors~\cite{Gronenschild2012, diethelm2012limits, 
+Glatard2015, bowring2018exploring}. The propagation and amplification 
+of these tiny errors by analysis pipelines may cause reproducibility 
+issues. 
+In this case, the analysis pipelines are said to be numerically 
 unstable. Numerical instability is a characteristic of the pipelines 
@@
 and magnitude of the errors. In many cases, numerical instability is an 
-important issue for reproducibility. }
+important issue for reproducibility.
+
 In particular, the effect of operating system is quantified 
@@
 calculated by the checksum method. 
-
-\note{The operating system is not the only part of the computational 
+The operating system is not the only part of the computational 
 environment that may hamper reproducibility. Many other computational 
 factors including hardware configuration, software version, and 
-parallelization techniques are reported as 
-the main influential computing infrastructure on reproducibility. }
+parallelization techniques are reported as the main influential 
+computing infrastructure on reproducibility.
 
@@
 However, a comprehensive solution requires to fix the numerical 
-instabilities. For this purpose, we introduce a framework to identify 
-the processes in pipeline that create numerical errors across different 
-runs. In this paper, the execution results of pipelines across 
-different operating systems are considered to find out the cause of the 
-irreproducibility using an interposition techniques.
+instabilities. For this purpose, we introduce a Numerical 
+Reproducibility Measurement (NRM) tool to identify the processes in 
+pipeline that create numerical errors across different runs. In this 
+paper, the execution results of pipelines across different operating 
+systems are considered to find out the cause of the irreproducibility 
+using an interposition techniques.
 
-Figure~\ref{fig:overview} illustrates an overview of the proposed 
-technique. First, we execute pipelines in a consistence environment 
-using docker container and Boutiques framework. In the next step, the 
-processes that create differences are classified using an iterative 
-approach. Finally, we make corresponding process graph of the 
-pipeline execution that shows the processes related to the creation of 
-errors. We will explain each step in detail in the following sections.
+The reminder of this paper describe the NRM-tool in different parts as 
+the way of capturing provenance information and their representation, 
+the influence of different type of subjects, and how to classify 
+processes and characterize errors. We describe the HCP preprocessing 
+pipelines as a popular neuroimaging pipelines to test and apply the 
+NRM-tool on. Finally, we report the experimental results, the processes 
+that are responsible for the errors in the pipelines.
 
-\begin{figure}[H]
+\section{Tool description}
+
+We present a Numerical Reproducibility Measurement tool (NRM-tool) to 
+measure numerical reproducibility of the pipelines across operating 
+systems automatically (see Figure~\ref{fig:overview}). We build the 
+containerized version of pipelines for each operating system using 
+docker technology which encapsulated whole the pipeline dependencies. 
+The docker images consists of an specific computing environment such as 
+operating system version, hardware configuration, software version, 
+etc. Furthermore, to simplify computational reproducibility of the 
+analysis, we execute docker images using Boutiques framework based on 
+the command-line descriptions~\cite{glatard2017boutiques}. The 
+Boutiques descriptor specifies the parameters and implementation of the 
+experiment, and the invocation schema describes the parameter values 
+such as input/output data path.
+
+NRM-tool produces the error JSON schema in which recognizes output files 
+with errors using the checksum comparison of the results. 
+We consider that a file contains errors if 
+its binary comparison is different, otherwise file has no error. Also, 
+we use \reprozip tool to capture the provenance information. 
+Subsequently, using a iterative approach, we classify processes into 
+different categories based on the state of read/write files 
+with/without error. NRM-tool identifies relevant processes in the 
+pipeline creating, propagating, or removing errors. We will explain 
+each steps in detail in the following sections.
+
+\begin{figure}
 \centering
-  \includegraphics[scale=0.5]{images/overview.png}
+  \includegraphics[width=\columnwidth]{images/overview.png}
   \caption{An overview of the proposed technique.}
@@
 
-The reminder of this paper explains the Reprotool in different parts as 
-the execution environment of the pipelines, methods to measure 
-differences, and how to characterize errors. We describe the HCP 
-preprocessing pipelines as a popular neuroimaging pipelines to test and 
-apply the Reprotool on. Finally, we report the processes that are 
-responsible for the errors in the pipelines.
+\subsection{Provenance capture and representation}
 
-\section{Processing Environment}
+We use the \reprozip tool~\cite{Chirigati2016} to record: (1) 
+the tree of processes executed by the pipeline and (2) the list of 
+files read and written by each process. This information is collected 
+by system call interceptions, through the \texttt{ptrace} Unix system 
+call, and stored in a \texttt{SQLite} database. The database stores 
+information on all the processes which are created by the 
+\texttt{clone()} or \texttt{fork()} system calls. It also records the 
+files opened by each process (in read, write or execution mode), 
+including the result files and the temporary files.
 
-We build the containerized version of pipelines for each operating 
-system using docker technology which encapsulated whole the pipeline 
-dependencies. The docker images consists of an specific computing 
-environment such as operating system version, hardware configuration, 
-software version, etc. Furthermore, derived images can be stored on 
-the dockerhub repositories and be version controlled in the same way as 
-the analysis code so that the exact same environment can be used to run 
-an experiment.
-
-To simplify computational reproducibility of the analysis, we execute 
-docker images using Boutiques framework based on the command-line 
-descriptions~\cite{glatard2017boutiques}. The Boutiques descriptor 
-specifies the parameters and implementation of the experiment, and the 
-invocation schema describes the parameter values such as input/output 
-data path.
-
-We processed subjects twice in every condition to identify possible 
-\emph{intra-condition} errors coming from the use of pseudo-random 
-numbers or other external factors. To further ensure that the observed 
-errors are caused by operating system updates (\emph{inter-os 
-differences}) and not by other factors, we implemented the following 
-sanity checks:
-
-\begin{itemize}
-
-\item The file checksums of every subject are computed before and after 
-processing to detect potential corruption during file transfers and 
-other manipulations. 
-\item The checksum of the Docker container images is recorded with
-  every execution to identify any unintentional package update. In 
-  addition, the list of packages with versions is printed by every task 
-  and checked for consistency.
-\item Hardware information is captured to make sure that errors are
-  not coming from different CPU architectures.
-\item Every task is executed on its own copy of the data to ensure
-  that input data is not modified by the pipeline. 
-\end{itemize}
-
-\section{Measuring Differences}
-
-We compute a file difference matrix between files produced for every 
-subject in every operating system based on MD5 
-checksums~\cite{md51992url} as in~\cite{Scaria2017}. MD5 is suitable to 
-identify incidental file differences, although it is not 
-cryptographically secure. Files produced only in some subjects or in 
-some operating systems were left out.
-
-In addition, we are able to quantify errors in the image files using 
-the Dice coefficient and normalized root mean square error (NRMSE) 
-computed as follows:
-
-\begin{center}
-\begin{equation}
-RMSE = {\sqrt {\frac{1} {n}{\sum\limits_{t = 1}^n {(\hat{y}_{t} - {y}_{t} } })^{2} } }
-\end{equation}
-\end{center}
-
-\begin{center}
-\begin{equation}
-NRMSE = {\frac{RMSE} {y_{max} - y_{min}}}
-\end{equation}
-\end{center}
-
-\section{Pipeline Analysis: Characterizing Errors}
-
-Once errors between conditions are quantified, we automatically
-identify the steps in the pipeline responsible for such differences.
-We use the \reprozip tool~\cite{Chirigati2016} to record: (1) the tree of
-processes executed by the pipeline and (2) the list of files read and
-written by each process. This information is collected by system call
-interceptions, through the \texttt{ptrace} Unix system call, and
-stored in a \texttt{SQLite} database. The database stores information
-on all the processes which are created by the \texttt{clone()} or
-\texttt{fork()} system calls. It also records the files opened by each
-process (in read, write or execution mode), including the result files
-and the temporary files.
-
-We reconstruct the tree of processes starting from the first process 
+NRM-tool reconstructs the tree of processes starting from the first process 
 created by the pipeline and identifying its child processes as the ones 
-that were created through \texttt{clone()} or \texttt{fork()}. We 
-create a process graph from this tree by adding edges corresponding to 
+that were created through \texttt{clone()} or \texttt{fork()}. It 
+creates a process graph from this tree by adding edges corresponding to 
 file dependencies between processes. A file dependency is defined 
@@
 Figure~\ref{fig:simple_script} shows an example of a process graph 
-constructed at this stage.
+constructed for the simple bash script illustrated in 
+Algorithm~\ref{algo:sample-script} from the brain extraction procedure.
 
-\medskip
-\noindent
-\begin{minipage}[]{.5\linewidth}
-  \centering
-  \includegraphics[scale=0.2,width=.4\linewidth]{images/simple_graph}
-%  \captionof{figure}{This is a figure caption.} \label{myfig1}
-\end{minipage}
-\begin{minipage}[]{.3\linewidth}
-\small\captionof{algorithm}{sample script of a process graph}\label{sample_script}
-{\scriptsize
+\begin{algorithm}[h!]
+\caption{Sample script from brain extraction process}
+\label{algo:sample-script}
 \begin{verbatim}
@@
 input_image=$1
-bet_output="$(basename ${input_image} .nii.gz)_brain.nii.gz"
-bet_output_binarized="$(basename ${input_image} .nii.gz)_brain_bin.nii.gz"
+bet_output="$(basename ${input_image} 
+                       .nii.gz)_brain.nii.gz"
+bet_output_binarized="$(basename ${input_image} 
+                       .nii.gz)_brain_bin.nii.gz"
 
@@
 \end{verbatim}
-}
-\end{minipage}
-\captionof{figure}{Process graph constructed from an example brain extraction
+\end{algorithm}
+
+\begin{figure}
+\centering
+  \includegraphics[scale=0.3]{images/simple_graph}
+  \caption{Process graph constructed from an example brain extraction
 	script. Every node in the graph is labeled using (1) a process id 
@@
   \label{fig:simple_script}
-\medskip
+\end{figure}
 
-A different process graph may be produced for every subject in every 
-execution condition. Among our dataset, we identified 4 types of 
-subjects with different numbers of T1w and T2w images. We verified that 
-the process graphs were identical for all subjects of the same type, 
-for all execution conditions. The remainder of the 
-analysis is done separately for each subject type.
+\subsection{Subject type}
 
-Cycles may be present in the process graph in case a file was written 
-by more than one process. We remove such cycles by removing file edges 
-between processes A and B when A's process creation timestamp is 
-posterior to B's or when A=B. Indeed, such edges cannot happen in 
-practice unless A and B were running concurrently, which we assume is 
-not the case (we also checked that on the workflow we tested).
+Depending on the variety of computing environments, the produced 
+process graph can be different. In particular, different analysis 
+software package produce different process graphs for the similar 
+experiments because of following different methods. We used a fixed 
+analysis software package in different conditions, however, we found 
+that different subjects may produce a different process graph. Among 
+our dataset, we identified 4 types of subjects with different numbers 
+of T1w and T2w images.
 
-In the next step, we classify the processes into three 
-categories based on the process graph and the error matrix file 
-mentioned in the previous Section:
+In addition, we verified that the process graphs were identical for all 
+subjects of the same type in different versions of CentOS operating 
+system. The remainder of the analysis is done separately for each 
+subject type.
+
+\subsection{Graph analysis}
+
+NRM-tool classifies the processes into four 
+categories based on the process graph and the error JSON file 
+mentioned in the previous sections:
 \begin{enumerate}
-\item Processes that read files that do not have errors and write files that do not 
-have errors are \emph{transparent} (Figure~\ref{fig:processes}.a).
+\item Processes that read files that do not have errors and write files 
+that do not have errors are \emph{transparent} (Figure~\ref{fig:processes}.a).
 \item Processes that read files 
@@
 
-\begin{figure}[H]%\centering
-    \begin{subfigure}{0.15\textwidth}
-        \includegraphics[scale=0.35]{images/green.png}
+\begin{figure}%\centering
+\centering
+    \begin{subfigure}{0.4\linewidth}
+        \includegraphics[scale=0.34]{images/green.png}
         \caption{Transparent}
@@
     \end{subfigure}
-\hfill%\hfil
-    \begin{subfigure}{0.15\textwidth}
-    \includegraphics[scale=0.35]{images/red.png}
+    \hfill
+    \begin{subfigure}{0.4\linewidth}
+    \includegraphics[scale=0.34]{images/red.png}
     \caption{Creates error}
@@
     \end{subfigure}
-\hfill%\hfil
-    \begin{subfigure}{0.15\textwidth}
-    \includegraphics[scale=0.35]{images/blue.png}
+    \hfill
+    \begin{subfigure}{0.4\linewidth}
+    \includegraphics[scale=0.34]{images/blue.png}
     \caption{Removes error}
@@
     \end{subfigure}
-\hfill%\hfil
-    \begin{subfigure}{0.15\textwidth}
-    \includegraphics[scale=0.35]{images/yellow.png}
+    \hfill
+    \begin{subfigure}{0.4\linewidth}
+    \includegraphics[scale=0.34]{images/yellow.png}
     \caption{Unknown}
@@
 
-\paragraph{Challenge1:} The classification of processes that read or 
+There are some issues that should be addressed to classify process 
+properly, including capturing temporary files that have been removed 
+during the execution, capturing files written by multiple processes, 
+and unknown process.
+
+\subsubsection{Capture temporary files} 
+The classification of processes that read or 
 write temporary files is uncertain when these files are deleted during 
@@
 
-\paragraph{Challenge2:} Files written by multiple processes also lead 
+\subsubsection{Capture multiple write files} 
+Files written by multiple processes also lead 
 to unknown classification labels. To address this issue for a file F 
@@
 
-\paragraph{Challenge3:} Another challenge is to classify processes that 
+Furthermore, cycles may be present in the process graph in case a file 
+was written by more than one process. We remove such cycles by removing 
+file edges between processes A and B when A's process creation 
+timestamp is posterior to B's or when A=B. Indeed, such edges cannot 
+happen in practice unless A and B were running concurrently, which we 
+assume is not the case (we also checked that on the workflow we 
+tested).
+
+\subsubsection{Unknown process} 
+Another challenge is to classify processes that 
 read files that have errors and write files that also have errors. In 
@@
 transparent. We label such processes \emph{unknown}.
-
-To address Challenge 3, we developed an iterative approach that 
+To address this issue, we developed an iterative approach that 
 consists of the following steps:
@@
     \emph{remove errors} or \emph{unknown}.
-  \item If there are \emph{unknown} processes then: replace each process P that create errors by a process Q that
-    copies the results produced by process P in condition 1 to the pipeline output in condition 2.
+  \item If there are \emph{unknown} processes then: replace each 
+  process P that create errors by a process Q that copies the results 
+  produced by process P in condition 1 to the pipeline output in 
+  condition 2. 
   \item Repeat steps 1 and 2 until there is no \emph{unknown} process left.
@@
 
-\begin{figure}[H]
+\begin{figure}
   \centering
-  \includegraphics[scale=0.5]{images/iterative_modif}
+  \includegraphics[width=\columnwidth]{images/iterative_modif}
   \caption{iteration of proposed modification.}
@@
 Figure~\ref{fig:iterations} illustrates our iterative classification 
-process for the example in Figure~\ref{fig:graph-example}. At every 
+process for the example in Figure~\ref{fig:simple_script}. At every 
 step, processes that created errors are shown in red, processes that 
@@
 yellow, and other processes (transparent processes) are in green. As in 
-Figure~\ref{fig:graph-example}, plain black edges represent the process 
+Figure~\ref{fig:simple_script}, plain black edges represent the process 
 tree and dashed edges represent file dependencies: green edges 
@@
 
-The process graphs of real pipelines are much larger than the one in 
-Figure~\ref{fig:iterations}. To represent such large graphs, we 
-summarize them by (1) building subgraphs for every interesting node 
-(e.g. red and blue nodes with their children) (2) building connected 
-components in the absence of interesting nodes for the remaining nodes 
-in the process graph. Each node in the summarized graph represent 
-either a shrinked interesting subgraph or connected component and edges 
-represent any connection between two shrinked nodes.
-
-\section{HCP Minimal preprocessing Pipelines}
-
-The Human Connectome Project (HCP)~\cite{glasser2013minimal} developed 
-a set of pipelines to help extraction of structural, functional or 
-diffusion MRI data across a large set of high resolution MR images. 
-These pre-processing pipelines create results that are available in 
-standard volume and combined surface and volume spaces which makes it 
-easier for researchers to compare the images across the neuroimaging 
-spectrum.
-
-We focused on HCP structural pre-processing pipelines to (1)quantify 
-the effect of OS on HCP pre-processing pipelines (2) identify the 
-process(es) in the pipelines that are responsible for such effects.
-
-The structural pre-processing pipelines consists of PreFreeSurfer, 
-FreeSurfer and PostFreeSurfer. In this paper, 
-we analyzed two pipelines, PreFreeSurfer and FreeSurfer. 
-PreFreeSurfer consist of various steps including correction of MR 
-gradient nonlinearity distortions, align the T1w and T2w images using 
-FSL FLIRT, Align native space to MNI template using ACPC and FSL FLIRT, 
-brain extraction usin FSL FLIRT and FNIRT to MNI template, register T2w 
-to T1w using FLIRT's BBR, perform a bias field correction, and register 
-the subject's native structural space to MNI space.
-
-FreeSurfer steps are downsample the images if it exceeds the resolution 
-limit, generation of white matter surface using a segmentation of 
-downsampled image, adjust T1w colum intensity and white matter surface 
-position based on 0.7mm T1w images, fine tune the T2w to T1w 
-registration using FreeSurfer's BBRegister algorithm, accurately place 
-the white and pial surface with high-resolution 0.7mm T1w and T2w data, 
-these inputs are fed back into recon-all at each stage.
-
-We randomly selected 10 subjects from the HCP data release S500 which 
-are described in table~\ref{table:data}.
-
-\begin{table}[H]\hspace*{-1cm}
-\centering
-\begin{threeparttable}
-\caption{An overview of the HCP data (Humman Connectome DB).}
-
-\begin{tabular}{@{}llllll@{}}
-\toprule
-Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
-103515  & Q1      & Q02         & F      & 26-30    & type1         \\
-105216  & Q3      & Q03         & M      & 26-30    & type4         \\
-103414  & Q2      & Q02         & F      & 22-25    & type1         \\
-125525  & Q1      & Q01         & F      & 31-35    & type1         \\
-142828  & Q1      & Q01         & M      & 31-35    & type1         \\
-129533  & Q3      & Q04         & F      & 31-35    & type1         \\
-103818  & Q1      & Q01         & F      & 31-35    & type3         \\
-133928  & Q2      & Q03         & M      & 26-30    & type3         \\
-148032  & Q3      & Q03         & F      & 31-35    & type2         \\
-139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
-\end{tabular}
-\begin{tablenotes}
-      \small
-      \item *Subj\_Type refers to the graph topology type of the subjects,
-      we will discuss further on the next sections.
-\end{tablenotes}
-\end{threeparttable}
-\label{table:data}
-\end{table}
-
-\section{Results}
-
-We executed the pipelines using Docker containers to simplify the 
-deployment of different operating system versions on execution 
-platforms. The Docker images were built for the HCP pre-processing 
-pipelines v3.19.0 (PreFreeSurfer and FreeSurfer) in 
-CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for 
-reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}. 
-We collected the provenance trace (tree of executed processes and files 
-accessed in read or write mode) for each type of 
-subjects(Table~\ref{table:data}) using system-call interception as 
-provided by the \reprozip tool~\cite{Chirigati2016}.
-
-Two types of errors can occur in the subjects due to the 
-errors in the operating systems. One is inter-OS error caused by the 
-operating system library updates and the other type, intra-OS errors 
-occurs as a result of the pseudo-random processes used in the 
-pipelines. An example of a pseudo-random process function is, a random 
-number generator that would get initialized using a seed state. the 
-proposed method can be used to identify both kind of errors. The files 
-that are common to all the subjects only are taken into consideration 
-for comparison. The first step is identification of files with 
-differences in their checksums. This is identified using the checksums 
-that are recorded after the processing. Intra-OS errors are identified 
-using the run-number added as the suffix for the conditions. For 
-example, the two batches of subjects processed under the same condition 
-(CentOS6) are stored as run-1 and run-2. The files belonging to the 
-subjects stored under the above mentioned conditions are treated as 
-intra-OS runs.
-
-\subsection{Inter-OS binary differences}
-
-In the previous study~\cite{Scaria2017}, we showed that pre-processing 
-pipelines of the Human Connectome Project~\cite{Glasser2013} were 
-sensitive to operating system variation so that Figure 
-\ref{fig:tissue_class} illustrates the binarized errors after brain 
-tissue segmentation process in FreeSurfer. In addition, 
-Figure~\ref{fig:fnirt_result} shows the same errors occurred in 
-non-linear registration process between CentOS6 and CentOS7 in PreFreeSurfer.
-
-\begin{figure}[H]
-%  \includegraphics{brain\_classification}
-\centering
-  \includegraphics[scale=0.18]{images/brain_classification.png} 
-  \caption{Binarized error between brain segmentation results from 
-  FreeSurfer, subject 105216 (CentOS6 vs. CentOS7)~\cite{Scaria2017}.
-    } 
-  \label{fig:tissue_class}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-  \includegraphics[scale=0.6]{images/fnirt_result.png} 
-  \caption{Absolute errors between FNIRT results from PreFreeSurfer 
-  (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (CentOS6 vs. 
-  CentOS7)~\cite{Scaria2017}. } 
-  \label{fig:fnirt_result}
-\end{figure}
-
-\subsection{Pipeline analysis}
-
-We identified processes that introduce errors in both PreFreeSurfer and 
-FreeSurfer pipelines. The result of our experiment is described in the 
-following paragraphs.
-
-\subsubsection{PreFreeSurfer pipeline}
-
-Among the 117 data files produced by 
-PreFreeSurfer, 21 did not have any error for any subject, 92 had errors 
-for all subjects and 4 had errors for 3 subjects only.
-
-Figure \ref{fig:summarized-graph} shows the annotated provenance graph 
-of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
-node in the graph represent an executed process in the pipeline. 
-Processes that created errors are shown in red, processes that removed 
-errors are in blue, and other processes are in green.  Squares denote 
-processes for which the classification is uncertain, due to temporary 
-files that were removed during the execution. Black edges link 
-sub-processes to their parents while dashed edges denote file 
-dependencies between processes (green edges: files with no errors; red 
-edges: files with errors; yellow edges: temporary files).
-
-The processes that introduce errors in PreFreeSurfer along with the 
-number of occurrences are depicted in Figures~\ref{fig:pfs_table, 
-fig:pfs_chart} including linear registration with “\emph{FLIRT}” (in 
-ACPC Alignment, BrainExtraction, DistortionCorrection, 
-AtlasRegistration), non-linear registration with “\emph{FNIRT}” (in 
-BrainExtraction and AltasRegistration), image warping with 
-“\emph{new\_invwarp}” (in BrainExtraction and AtlasRegistration).  In 
-addition, errors were observed in image mean and standard-deviation 
-computations with “\emph{fslstats}” (in BiasFieldCorrection), and in 
-masked image extrapolation with “\emph{fslmaths}” (in 
-BiasFieldCorrection).  Besides, transformation format conversion with 
-“\emph{convertwarp}” (in DistortionCorrection) was able to remove 
-errors.
-
-\begin{figure}[H]
-\centering
-  \includegraphics[scale=0.5]{images/pfs_table.png}
-  \caption{Process name along with the pipeline elements that introduce 
-  errors.}
-  \label{fig:pfs_table}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-  \includegraphics[scale=0.38]{images/pfs_chart.png} \caption{Number 
-  of occurrences of errors in PreFreeSurfer on subject 103515, red and 
-  blue bars indicate the processes which create and remove errors 
-  respectively.} 
-  \label{fig:pfs_chart}
-\end{figure}
-
-\subsubsection{FreeSurfer pipeline}
-
-In the second analysis, the result of 
-PreFreeSurfer is considered as the input files of FreeSurfer pipeline. 
-From the FreeSurfer results, total 61 files were found to be different 
-between CentOS6 and CentOS7. The processes that introduce errors in 
-FreeSurfer are shown on Figure~\ref{fig:fs_error_table}. Most frequent 
-processes that create errors are \emph{mris\_anatomical\_stats} which 
-measures a number of anatomical properties, \emph{mris\_make\_surface} 
-to generate pial surfaces, first pass, no T2 adjustment, 
-\emph{mri\_convert} which normalize T1w and T2w images for the benefit 
-of \emph{mris\_make\_surfaces}, and \emph{fslmaths} create and apply a 
-binary mask to filling of holes to remove  any non-grey/white tissue 
-using a threshold created by FSLSTATS at Mean. In addition, all the 
-other processes that created errors in the pipeline at least once are 
-categorized as others such as \emph{mri\_ca\_register}, 
-\emph{mri\_ca\_normalize}, and \emph{mri\_watershed}. Furthermore, 
-processes that remove errors in FreeSurfer represented on 
-Figure~\ref{fig:fs_remove_table} including \emph{mri\_convert} and 
-\emph{mri\_surf2surf}.
-
-\begin{figure}[H] 
-\centering
-  \includegraphics[scale=0.6]{images/fs_error_table.png} 
-  \caption{Number of occurrences of errors in FreeSurfer pipeline, 
-subject 103818, indicate the processes that create errors.}
-  \label{fig:fs_error_table}
-\end{figure}
-
-\begin{figure}[H]
-\centering
-  \includegraphics[scale=0.5]{images/fs_remove_table.png} 
-  \caption{Number of occurrences of errors in FreeSurfer pipeline, 
-subject 103818, indicate the processes that remove errors.}
-  \label{fig:fs_remove_table}
-\end{figure}
+\section{Experiments}
+\subsection{HCP pipelines and datasets}
+\subsection{Results}
 
@@
 
-\begin{figure}[H]
-  \includegraphics[width=\linewidth]{images/summarized_graph}
-  \caption{An annotated summarized process graph from the PreFreesurfer pipeline.
-Each node in the graph is represented in red, blue or green colors and is labeled 
-using the name of the executable run by the interesting process. The close up nodes 
-reveal each shrinked subgraph in details.}
-  \label{fig:summarized-graph}
-\end{figure}
-
-\begin{figure}[H]
-  \includegraphics[width=\linewidth]{images/graph}
-  \caption{A complete process graph from the PreFreesurfer pipeline.
-Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
-  \label{fig:complete-graph}
-\end{figure}
 \bibliographystyle{plain}

### 2019-04-25 paper.tex
add resuls

@@
 
-\title{NRM-tool (Numerical Reproducibility Measurement): 
-a debugging tool to identify numerical errors 
-and assess reproducibility of pipelines}
+\title{NURM-tool: a quantitative reproducibility measurement tool to 
+identify numerical differences in pipelines}
 
@@
 \author[2]{Lindsay Lewis}
+\author[2,3]{Alan C. Evans}
 \author[1]{Tristan Glatard}
-\author[2,3]{Alan C. Evans}
 
@@
 analysis pipelines. These effects are presumably due to the creation, 
-propagation and amplification of small numerical errors across the 
+propagation and amplification of small numerical differences across the 
 pipelines. 
@@
 technique to identify the processes in the pipeline that create 
-numerical errors along the execution, and we apply this technique to 
+numerical differences along the execution, and we apply this technique to 
 the HCP structural pre-processing pipelines.
@@
 computation might be reproducible based on Peng's definition, but not 
-be numerically reproducible. For instance, small numerical errors 
+be numerically reproducible. For instance, small numerical differences 
 created during the pipeline execution may hamper numerical 
@@
 known to influence reproducibility because of the creation of small 
-numerical errors~\cite{Gronenschild2012, diethelm2012limits, 
+numerical differences~\cite{Gronenschild2012, diethelm2012limits, 
 Glatard2015, bowring2018exploring}. The propagation and amplification 
-of these tiny errors by analysis pipelines may cause reproducibility 
+of these slight differences by analysis pipelines may cause reproducibility 
 issues. 
@@
 unstable. Numerical instability is a characteristic of the pipelines 
-which amplify small numerical errors and then hamper the 
+which amplify small numerical differences and then hamper the 
 reproducibility of the analyses depending on the length of the pipeline 
-and magnitude of the errors. In many cases, numerical instability is an 
+and magnitude of the differences. In many cases, numerical instability is an 
 important issue for reproducibility.
@@
 the main neuroimaging pipelines including FSL, CIVET and Freesurfer 
-pipelines. The results of comparisons indicate a significant errors 
+pipelines. The results of comparisons indicate a significant differences 
 between analysis results of the same experiment on different operating 
-system. It should be noted that \emph{error} is defined as the binary 
-differences between results of two analyses in this paper, which is 
-calculated by the checksum method. 
+system. \note{It should be noted that the binary 
+differences are measured between results of two analyses in this paper, which is 
+calculated by the checksum method. }
 The operating system is not the only part of the computational 
@@
 instabilities. For this purpose, we introduce a Numerical 
-Reproducibility Measurement (NRM) tool to identify the processes in 
-pipeline that create numerical errors across different runs. In this 
+Reproducibility Measurement (NURM) tool to identify the processes in 
+pipeline that create numerical differences across several runs. In this 
 paper, the execution results of pipelines across different operating 
@@
 
-The reminder of this paper describe the NRM-tool in different parts as 
+The reminder of this paper describe the NURM-tool in different parts as 
 the way of capturing provenance information and their representation, 
 the influence of different type of subjects, and how to classify 
-processes and characterize errors. We describe the HCP preprocessing 
+processes and characterize differences. We describe the HCP preprocessing 
 pipelines as a popular neuroimaging pipelines to test and apply the 
-NRM-tool on. Finally, we report the experimental results, the processes 
-that are responsible for the errors in the pipelines.
+NURM-tool on. Finally, we report the experimental results, the processes 
+that are responsible for the differences in the pipelines.
 
@@
 
-We present a Numerical Reproducibility Measurement tool (NRM-tool) to 
+We present a Numerical Reproducibility Measurement tool (NURM-tool) to 
 measure numerical reproducibility of the pipelines across operating 
@@
 
-NRM-tool produces the error JSON schema in which recognizes output files 
+NURM-tool produces the error JSON schema in which recognizes output files 
 with errors using the checksum comparison of the results. 
@@
 different categories based on the state of read/write files 
-with/without error. NRM-tool identifies relevant processes in the 
+with/without error. NURM-tool identifies relevant processes in the 
 pipeline creating, propagating, or removing errors. We will explain 
@@
 
-NRM-tool reconstructs the tree of processes starting from the first process 
+NURM-tool reconstructs the tree of processes starting from the first process 
 created by the pipeline and identifying its child processes as the ones 
@@
 between processes A and B if a file written by A is read by B. 
-Figure~\ref{fig:simple_script} shows an example of a process graph 
+Figure~\ref{fig:simple_script} shows an example of a process tree 
 constructed for the simple bash script illustrated in 
@@
   \includegraphics[scale=0.3]{images/simple_graph}
-  \caption{Process graph constructed from an example brain extraction
+  \caption{Process tree constructed from an example brain extraction
 	script. Every node in the graph is labeled using (1) a process id 
@@
 
-Depending on the variety of computing environments, the produced 
-process graph can be different. In particular, different analysis 
-software package produce different process graphs for the similar 
-experiments because of following different methods. We used a fixed 
-analysis software package in different conditions, however, we found 
-that different subjects may produce a different process graph. Among 
-our dataset, we identified 4 types of subjects with different numbers 
-of T1w and T2w images.
+Depending on the variety of computing environments, different  
+process tree can be produced. In particular, different analysis 
+software package can construct different process trees for the 
+similar experiments because of following different methods. We used a 
+fixed analysis software package in different conditions, however, we 
+found that different subjects may produce a different process trees.
 
-In addition, we verified that the process graphs were identical for all 
-subjects of the same type in different versions of CentOS operating 
-system. The remainder of the analysis is done separately for each 
-subject type.
+The challenging problem is that of comparing various process trees with 
+different topological structures from different subjects. To address 
+this issue, we categorize subjects into different types based on the 
+distances, topologically differences, between corresponding process 
+trees. 
+We cluster subjects into various types using the bottom-up 
+hierarchical clustering approach. This approach groups together the 
+subjects with similar distance, less than a threshold.
+
+We use the tree edit distance~\cite{zhang1989simple} method to measure 
+the distance between labeled trees. The distance between two labeled 
+trees is defined as the minimum weighted number of edit operations that 
+it takes to transform one tree to another. There are three edit 
+operations that can be used to transform a tree including modify, 
+remove, and insert. Modifying node \textit{n} means changing the label 
+on \textit{n}. Removing a node \textit{n} means making the children of 
+\textit{n} become the children of the parent of \textit{n} and then 
+deleting \textit{n}. Inserting is the the complement of remove, this 
+means that inserting \textit{n} as the child of \textit{n}$'$ will make 
+\textit{n} the parent of a consecutive subsequence of the current 
+children of \textit{n}$'$.
+Each operation has an associated cost---insertions and removals each 
+cost 1, and modifications cost either 1 or 2, depending on whether or 
+not you consider a modification to essentially be just a removal 
+followed by an insertion.
+
+The clustering algorithm starts by finding the two subjects that are 
+closest to each other on the basis of edit distance method and form the 
+first cluster between these two subjects. The next step join this 
+cluster to the next nearest cluster or subject which in turn results in 
+another cluster. This process continues until all the subjects are 
+joined together to form one big cluster or terminates when the distances 
+are less than the threshold. The threshold is defined as the minimum 
+distance required to separate clusters.
 
@@
 
-NRM-tool classifies the processes into four 
+NURM-tool classifies the processes into four 
 categories based on the process graph and the error JSON file 
@@
 \section{Experiments}
+
 \subsection{HCP pipelines and datasets}
+
+The Human Connectome Project (HCP)~\cite{glasser2013minimal} developed 
+a set of pipelines to help extraction of structural, functional or 
+diffusion MRI data across a large set of high resolution MR images. 
+These pre-processing pipelines create results that are available in 
+standard volume and combined surface and volume spaces which makes it 
+easier for researchers to compare the images across the neuroimaging 
+spectrum.
+
+We focused on HCP structural pre-processing pipelines to (1)quantify 
+the effect of OS on HCP pre-processing pipelines (2) identify the 
+process(es) in the pipelines that are responsible for such effects.
+
+The structural pre-processing pipelines consists of PreFreeSurfer, 
+FreeSurfer and PostFreeSurfer. In this paper, 
+we analyzed PreFreeSurfer pipeline which consist of various steps 
+including correction of MR 
+gradient nonlinearity distortions, align the T1w and T2w images using 
+FSL FLIRT, Align native space to MNI template using ACPC and FSL FLIRT, 
+brain extraction usin FSL FLIRT and FNIRT to MNI template, register T2w 
+to T1w using FLIRT's BBR, perform a bias field correction, and register 
+the subject's native structural space to MNI space.
+
+We randomly selected 10 subjects from the HCP data release S500 which 
+are described in table~\ref{table:data}.
+
+%\begin{figure}
+%\centering
+%  \includegraphics[width=\columnwidth]{images/overview.png}
+%  \caption{An overview of the proposed technique.}
+%  \label{fig:overview}
+%\end{figure}
+
+\begin{table}
+\centering
+\begin{threeparttable}
+\caption{An overview of the HCP data (Humman Connectome DB).}
+
+\begin{tabular}{@{}llllll@{}}
+\toprule
+Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
+103515  & Q1      & Q02         & F      & 26-30    & type1         \\
+105216  & Q3      & Q03         & M      & 26-30    & type4         \\
+103414  & Q2      & Q02         & F      & 22-25    & type1         \\
+125525  & Q1      & Q01         & F      & 31-35    & type1         \\
+142828  & Q1      & Q01         & M      & 31-35    & type1         \\
+129533  & Q3      & Q04         & F      & 31-35    & type1         \\
+103818  & Q1      & Q01         & F      & 31-35    & type3         \\
+133928  & Q2      & Q03         & M      & 26-30    & type3         \\
+148032  & Q3      & Q03         & F      & 31-35    & type2         \\
+139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
+\end{tabular}
+\begin{tablenotes}
+      \small
+      \item *Subj\_Type refers to the graph topology type of the subjects.
+\end{tablenotes}
+\end{threeparttable}
+\label{table:data}
+\end{table}
+
+\note{Among our dataset, we identified 4 types of subjects with 
+different numbers of T1w and T2w images. In addition, we verified that 
+the process graphs were identical for all subjects of the same type in 
+different versions of CentOS operating system. The remainder of the 
+analysis is done separately for each subject type.}
+
 \subsection{Results}
+
+We executed the pipelines using Docker containers to simplify the 
+deployment of different operating system versions on execution 
+platforms. The Docker images were built for the HCP PreFreeSurfer 
+pipeline v3.19.0 in 
+CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for 
+reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}. 
+We collected the provenance trace (tree of executed processes and files 
+accessed in read or write mode) for each type of 
+subjects(Table~\ref{table:data}) using system-call interception as 
+provided by the \reprozip tool~\cite{Chirigati2016}.
+
+Two types of errors can occur in the subjects due to the 
+errors in the operating systems. One is inter-OS error caused by the 
+operating system library updates and the other type, intra-OS errors 
+occurs as a result of the pseudo-random processes used in the 
+pipelines. An example of a pseudo-random process function is, a random 
+number generator that would get initialized using a seed state. the 
+proposed method can be used to identify both kind of errors. The files 
+that are common to all the subjects only are taken into consideration 
+for comparison. The first step is identification of files with 
+differences in their checksums. This is identified using the checksums 
+that are recorded after the processing. Intra-OS errors are identified 
+using the run-number added as the suffix for the conditions. For 
+example, the two batches of subjects processed under the same condition 
+(CentOS6) are stored as run-1 and run-2. The files belonging to the 
+subjects stored under the above mentioned conditions are treated as 
+intra-OS runs.
+
+\subsubsection{Inter-OS binary differences}
+
+In the previous study~\cite{Scaria2017}, we showed that pre-processing 
+pipelines of the Human Connectome Project~\cite{Glasser2013} were 
+sensitive to operating system variation so that Figure 
+\ref{fig:tissue_class} illustrates the binarized errors after brain 
+tissue segmentation process in FreeSurfer. In addition, 
+Figure~\ref{fig:fnirt_result} shows the same errors occurred in 
+non-linear registration process between CentOS6 and CentOS7 in PreFreeSurfer.
+
+\begin{figure}
+%  \includegraphics{brain\_classification}
+\centering
+  \includegraphics[width=\columnwidth]{images/brain_classification.png} 
+  \caption{Binarized error between brain segmentation results from 
+  FreeSurfer, subject 105216 (CentOS6 vs. CentOS7)~\cite{Scaria2017}.
+    } 
+  \label{fig:tissue_class}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=\columnwidth]{images/fnirt_result.png} 
+  \caption{Absolute errors between FNIRT results from PreFreeSurfer 
+  (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (CentOS6 vs. 
+  CentOS7)~\cite{Scaria2017}. } 
+  \label{fig:fnirt_result}
+\end{figure}
+
+\subsubsection{PreFreeSurfer pipeline analysis}
+
+We identified processes that introduce errors in PreFreeSurfer 
+pipeline. Among the 117 data files produced by PreFreeSurfer, 21 did 
+not have any error for any subject, 92 had errors for all subjects and 
+4 had errors for 3 subjects only.
+
+Figure \ref{fig:complete-graph} shows the annotated provenance graph 
+of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
+node in the graph represent an executed process in the pipeline. 
+Processes that created errors are shown in red, processes that removed 
+errors are in blue, and other processes are in green.  Squares denote 
+processes for which the classification is uncertain, due to temporary 
+files that were removed during the execution. Black edges link 
+sub-processes to their parents while dashed edges denote file 
+dependencies between processes (green edges: files with no errors; red 
+edges: files with errors; yellow edges: temporary files).
+
+The processes that introduce errors in PreFreeSurfer along with the 
+number of occurrences are depicted in Figures~\ref{fig:pfs_table, 
+fig:pfs_chart} including linear registration with “\emph{FLIRT}” (in 
+ACPC Alignment, BrainExtraction, DistortionCorrection, 
+AtlasRegistration), non-linear registration with “\emph{FNIRT}” (in 
+BrainExtraction and AltasRegistration), image warping with 
+“\emph{new\_invwarp}” (in BrainExtraction and AtlasRegistration).  In 
+addition, errors were observed in image mean and standard-deviation 
+computations with “\emph{fslstats}” (in BiasFieldCorrection), and in 
+masked image extrapolation with “\emph{fslmaths}” (in 
+BiasFieldCorrection).  Besides, transformation format conversion with 
+“\emph{convertwarp}” (in DistortionCorrection) was able to remove 
+errors.
+
+\begin{figure}
+\centering
+  \includegraphics[width=\columnwidth]{images/pfs_table.png}
+  \caption{Process name along with the pipeline elements that introduce 
+  errors.}
+  \label{fig:pfs_table}
+\end{figure}
+
+\begin{figure}
+\centering
+  \includegraphics[width=\columnwidth]{images/pfs_chart.png} \caption{Number 
+  of occurrences of errors in PreFreeSurfer on subject 103515, red and 
+  blue bars indicate the processes which create and remove errors 
+  respectively.} 
+  \label{fig:pfs_chart}
+\end{figure}
 
@@
 
+\begin{figure}
+  \includegraphics[width=1\textwidth]{images/graph}
+  \caption{A complete process graph from the PreFreesurfer pipeline.
+Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
+  \label{fig:complete-graph}
+\end{figure}
+
 \bibliographystyle{plain}

### 2019-05-02 paper.tex
add clustering section

@@
 the main neuroimaging pipelines including FSL, CIVET and Freesurfer 
-pipelines. The results of comparisons indicate a significant differences 
-between analysis results of the same experiment on different operating 
-system. \note{It should be noted that the binary 
-differences are measured between results of two analyses in this paper, which is 
-calculated by the checksum method. }
+pipelines. 
+The binary differences are measured by the checksum method between 
+results of the same analyses on different operating system, which indicate 
+a significant disparity. 
 The operating system is not the only part of the computational 
@@
 
-We present a Numerical Reproducibility Measurement tool (NURM-tool) to 
+We present NURM-tool to automatically
 measure numerical reproducibility of the pipelines across operating 
-systems automatically (see Figure~\ref{fig:overview}). We build the 
-containerized version of pipelines for each operating system using 
+systems (see Figure~\ref{fig:overview}). We build the 
+containerized version of the pipelines for each operating system using 
 docker technology which encapsulated whole the pipeline dependencies. 
@@
 
-NURM-tool produces the error JSON schema in which recognizes output files 
-with errors using the checksum comparison of the results. 
-We consider that a file contains errors if 
-its binary comparison is different, otherwise file has no error. Also, 
+As shown in the left side of Figure~\ref{fig:overview}, NURM-tool clusters 
+the input subjects into various types based on the similarity of their 
+process trees which are achieved by \reprozip tool. The right side of this figure 
+shows the iterative classification of the process based on the binary 
+comparison of the outputs from various conditions.
+NURM-tool produces the \textit{difference} JSON schema in which 
+recognizes output files with differences using the checksum comparison 
+of the results. We consider that a file contains difference if only 
+its binary comparison shows difference. Also, 
 we use \reprozip tool to capture the provenance information. 
 Subsequently, using a iterative approach, we classify processes into 
-different categories based on the state of read/write files 
-with/without error. NURM-tool identifies relevant processes in the 
-pipeline creating, propagating, or removing errors. We will explain 
-each steps in detail in the following sections.
+different categories regarding the state of read/write files 
+with/without differences. NURM-tool identifies relevant processes in the 
+pipeline creating, propagating, or removing differences. 
+All the steps are explained in detail in the following sections.
 
@@
 process tree can be produced. In particular, different analysis 
-software package can construct different process trees for the 
+software package can construct different process tree for the 
 similar experiments because of following different methods. We used a 
 fixed analysis software package in different conditions, however, we 
-found that different subjects may produce a different process trees.
+found that different subjects may produce a different process tree.
 
@@
 trees.
-We cluster subjects into various types using the bottom-up 
-hierarchical clustering approach. This approach groups together the 
-subjects with similar distance, less than a threshold.
 
@@
 
+We cluster subjects into various types using the bottom-up hierarchical 
+clustering approach implemented in \textit{scipy} python package (see 
+Algorithm~\ref{also:hclustering}). This approach groups together the 
+subjects with similar distance, less than a threshold.
+
+\begin{algorithm}[h!]
+\caption{Hierarchical clustering algorithm}
+\label{algo:hclustering}
+\begin{algorithmic}
+
+  \State /* Input: a list of n trees; \texttt{T = [t${_1}$, t${_2}$,...,t${_n}$]}
+  \State /* Output: a list of clusters; \texttt{C = [c${_1}$,c${_2}$,...,c${_k}$]}
+  \For{\texttt{i=1} in \texttt{n}}
+  \State \texttt{C${_i}$} = \texttt{T${_i}$} 
+  \State \texttt{\# Create a cluster for each tree in input list; \texttt{agglomerative approach}}
+  \EndFor
+  \While{\texttt{len(C)} > \texttt{1}}  
+  \For{\texttt{i=1} to \texttt{n}}
+  \For{\texttt{j=i+1} to \texttt{n}}
+  \State \texttt{D${_{ij}}$} = argmin\{\texttt{edit\_distance(C${_i}$, C${_j}$)}\}
+  \State \texttt{\# Select two nearest clusters based on \texttt{nearest\_neighbors} method}
+  \EndFor
+  \EndFor
+  \If{\texttt{D${_{ij}}$} > \texttt{thre}}
+  \State Return \texttt{C}
+  \Else {\State Merge \texttt{C${_i}$ and C${_j}$}}
+  \EndIf
+  \EndWhile
+%  \State Function edit\_distance (C${_1}$, C${_2}$)
+%  \State EndFunction
+\end{algorithmic}
+\end{algorithm}
+
 The clustering algorithm starts by finding the two subjects that are 
 closest to each other on the basis of edit distance method and form the 
-first cluster between these two subjects. The next step join this 
+new cluster between these two subjects. The next step join this 
 cluster to the next nearest cluster or subject which in turn results in 
@@
 
-\subsection{Graph analysis}
+\subsection{Tree analysis}
 
 NURM-tool classifies the processes into four 
-categories based on the process graph and the error JSON file 
+categories based on the process tree and the \textit{difference} JSON file 
 mentioned in the previous sections:
 \begin{enumerate}
-\item Processes that read files that do not have errors and write files 
-that do not have errors are \emph{transparent} (Figure~\ref{fig:processes}.a).
+\item Processes that read and write files that do not have differences 
+are \emph{transparent} (Figure~\ref{fig:processes}.a).
 \item Processes that read files 
-that do not have any error but write files that have errors 
-\emph{create} errors in the pipeline (Figure~\ref{fig:processes}.b).
+that do not have any differences but write files that have differences 
+\emph{create} differences in the pipeline (Figure~\ref{fig:processes}.b).
 \item Processes that read files 
-that have errors and write files that do not have errors \emph{remove} 
-errors from the pipeline (Figure~\ref{fig:processes}.c).
-\item Processes that read files that have errors and write files that also have errors are 
+that have differences and write files that do not have differences \emph{remove} 
+differences from the pipeline (Figure~\ref{fig:processes}.c).
+\item Processes that read files that have differences and write files that 
+also have differences are 
 \emph{unknown} (Figure~\ref{fig:processes}.d).
@@
     \includegraphics[scale=0.34]{images/red.png}
-    \caption{Creates error}
+    \caption{Creates differences}
     \label{fig:red}
@@
     \includegraphics[scale=0.34]{images/blue.png}
-    \caption{Removes error}
+    \caption{Removes differences}
     \label{fig:blue}
@@
 
-Furthermore, cycles may be present in the process graph in case a file 
+Furthermore, cycles may be present in the process tree in case a file 
 was written by more than one process. We remove such cycles by removing 
@@
 Another challenge is to classify processes that 
-read files that have errors and write files that also have errors. In 
+read files that have differences and write files that also have differences. In 
 such situations, it is not possible to determine from the pipeline 
-results whether the process created errors, removed some errors, or was 
+results whether the process created differences, removed some differences, or was 
 transparent. We label such processes \emph{unknown}.
@@
   \item Run the pipeline in conditions 1 and 2; classify the
-    processes as \emph{transparent}, \emph{create errors},
-    \emph{remove errors} or \emph{unknown}.
+    processes as \emph{transparent}, \emph{create differences},
+    \emph{remove differences} or \emph{unknown}.
   \item If there are \emph{unknown} processes then: replace each 
-  process P that create errors by a process Q that copies the results 
+  process P that create differences by a process Q that copies the results 
   produced by process P in condition 1 to the pipeline output in 
@@
 obtained in condition 1 if it is invoked with the arguments of a 
-process that created errors, and calls P's original executable 
+process that created differences, and calls P's original executable 
 otherwise. The replacement is done through the PATH variable, as 
-before. This algorithm converges to a process graph without any 
+before. This algorithm converges to a process tree without any 
 \emph{unknown} process after a finite number of iterations.
@@
 process for the example in Figure~\ref{fig:simple_script}. At every 
-step, processes that created errors are shown in red, processes that 
-removed errors are in blue, processes that propagated errors are in 
+step, processes that created differences are shown in red, processes that 
+removed differences are in blue, processes that propagated differences are in 
 yellow, and other processes (transparent processes) are in green. As in 
@@
 tree and dashed edges represent file dependencies: green edges 
-represent files with no errors, while red edges represent files with 
-errors. Temporary files are not represented because they have been 
+represent files with no differences, while red edges represent files with 
+differences. Temporary files are not represented because they have been 
 backed up as previously described.
@@
 iterations of the classification algorithm. At step one, \texttt{2\#bet2} 
-is classified as error creator (red) as it produced files with errors 
-from files without error. \texttt{3\#fslstats} is classified as error 
-remover (blue) as it produced files without errors from files with 
-errors. \texttt{4\#fslmaths} and \texttt{5\#fslstats} are classified as 
-unknown (yellow) as they produce files with errors from files with 
-errors.
+is classified as difference creator (red) as it produced files with differences 
+from files without differences. \texttt{3\#fslstats} is classified as difference  
+remover (blue) as it produced files without differences from files with 
+differences. \texttt{4\#fslmaths} and \texttt{5\#fslstats} are classified as 
+unknown (yellow) as they produce files with differences from files with 
+differences.
 
@@
 the other condition. \texttt{3\#fslstats} is now classified as 
-transparent, \texttt{4\#fslmaths} is now classified as error creator 
+transparent, \texttt{4\#fslmaths} is now classified as difference creator 
 and \texttt{5\#fslstats} is still unknown.
@@
 As a result of those 3 steps, the final process classification is: 
-\texttt{2\#bet2} and \texttt{4\#fslmath} are error creators (red), 
-\texttt{3\#fslstats} is error remover (blue) and the other processes 
+\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red), 
+\texttt{3\#fslstats} is difference remover (blue) and the other processes 
 are transparent.
@@
 
-We randomly selected 10 subjects from the HCP data release S500 which 
-are described in table~\ref{table:data}.
+We randomly selected 100 subjects from the HCP data release S500 and then 
+using the NURM-tool we cluster them into different groups. 
+Figure~\ref{fig:subj-clusters} shows the result of data clustering. 
+Among our dataset, we identified 4 types of subjects with 
+different topology of process trees. In addition, we verified that 
+the process trees were identical for all subjects of the same type in 
+different versions of CentOS operating system. The remainder of the 
+analysis is done separately for each subject type.
 
@@
 
-\begin{table}
-\centering
-\begin{threeparttable}
-\caption{An overview of the HCP data (Humman Connectome DB).}
-
-\begin{tabular}{@{}llllll@{}}
-\toprule
-Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
-103515  & Q1      & Q02         & F      & 26-30    & type1         \\
-105216  & Q3      & Q03         & M      & 26-30    & type4         \\
-103414  & Q2      & Q02         & F      & 22-25    & type1         \\
-125525  & Q1      & Q01         & F      & 31-35    & type1         \\
-142828  & Q1      & Q01         & M      & 31-35    & type1         \\
-129533  & Q3      & Q04         & F      & 31-35    & type1         \\
-103818  & Q1      & Q01         & F      & 31-35    & type3         \\
-133928  & Q2      & Q03         & M      & 26-30    & type3         \\
-148032  & Q3      & Q03         & F      & 31-35    & type2         \\
-139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
-\end{tabular}
-\begin{tablenotes}
-      \small
-      \item *Subj\_Type refers to the graph topology type of the subjects.
-\end{tablenotes}
-\end{threeparttable}
-\label{table:data}
-\end{table}
-
-\note{Among our dataset, we identified 4 types of subjects with 
-different numbers of T1w and T2w images. In addition, we verified that 
-the process graphs were identical for all subjects of the same type in 
-different versions of CentOS operating system. The remainder of the 
-analysis is done separately for each subject type.}
+%\begin{table}
+%\centering
+%\begin{threeparttable}
+%\caption{An overview of the HCP data (Humman Connectome DB).}
+%
+%\begin{tabular}{@{}llllll@{}}
+%\toprule
+%Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
+%103515  & Q1      & Q02         & F      & 26-30    & type1         \\
+%105216  & Q3      & Q03         & M      & 26-30    & type4         \\
+%103414  & Q2      & Q02         & F      & 22-25    & type1         \\
+%125525  & Q1      & Q01         & F      & 31-35    & type1         \\
+%142828  & Q1      & Q01         & M      & 31-35    & type1         \\
+%129533  & Q3      & Q04         & F      & 31-35    & type1         \\
+%103818  & Q1      & Q01         & F      & 31-35    & type3         \\
+%133928  & Q2      & Q03         & M      & 26-30    & type3         \\
+%148032  & Q3      & Q03         & F      & 31-35    & type2         \\
+%139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
+%\end{tabular}
+%\begin{tablenotes}
+%      \small
+%      \item *Subj\_Type refers to the tree topology type of the subjects.
+%\end{tablenotes}
+%\end{threeparttable}
+%\label{table:data}
+%\end{table}
 
@@
 
-CBRAIN team. Compute Canada(Calcul Quebec).
+Data were provided by the Human Connectome Project, WU-Minn 
+Consortium (Principal Investigators: David Van Essen and Kamil Ugurbil; 
+1U54MH091657) funded by the 16 NIH Institutes and Centers that support 
+the NIH Blueprint for Neuroscience Research; and by the McDonnell 
+Center for Systems Neuroscience at Washington University.
+
+\note{CBRAIN team. Compute Canada(Calcul Quebec).}
 
@@
 
+\begin{figure}
+  \includegraphics[width=\textwidth]{images/hclusters}
+  \caption{Different data types clustered among 100 subjects.}
+  \label{fig:subj-clusters}
+\end{figure}
+
 \bibliographystyle{plain}

### 2019-05-05 paper.tex

@@
 \begin{abstract}
+
 Many experiments show that computational analysis results are still not 
-completely reproducible. In particular, computational environments 
-including different operating systems, hardware categories, and 
-software versions are known to have effects on the results produced by 
-analysis pipelines. These effects are presumably due to the creation, 
-propagation and amplification of small numerical differences across the 
-pipelines. 
-Although new techniques such as virtualization , version 
-control systems , and provenance management tools provide more reliable 
+completely reproducible. Computational environments including different 
+operating systems, hardware categories, and software versions are known 
+to have effects on the results produced by analysis pipelines. These 
+effects are presumably due to the creation, propagation and 
+amplification of small numerical differences across the pipelines. 
+Although new techniques such as virtualization, version control 
+systems, and provenance management tools provide more reliable 
 environment and significantly improve the reproducibility of scientific 
-findings. 
-However, the precise causes of such instabilities and the path along 
-which they propagate in the pipelines are unclear.  We present a 
-technique to identify the processes in the pipeline that create 
-numerical differences along the execution, and we apply this technique to 
-the HCP structural pre-processing pipelines.
+findings. However, the precise causes of such instabilities and the 
+path along which they propagate in the pipelines are unclear.  We 
+present a technique to identify the processes in the pipeline that 
+create numerical differences along the execution, and we apply this 
+technique to the HCP structural pre-processing pipelines.
 
@@
 evaluated.
-There are different definitions for the term reproducibility, we follow 
-Peng's definition~\cite{peng2011reproducible} in which reproducibility 
-is defined as the ability to regenerate the same results as the 
-original findings when the experiment is reanalyzed given exactly the 
-same analytic methods, software package, parameters, and data.
+
+Reproducibility is defined as the ability to regenerate the same 
+results as the original findings when the experiment is reanalyzed by 
+exactly the same analytic methods, software package, parameters, and 
+data~\cite{peng2011reproducible}.
 
@@
 regenerate bit for bit identical results from multiple 
-runs~\cite{hill2017numerical}. Two files will be considered numerically 
-reproducible if they have identical binary contents. Binary comparison 
-is calculated by comparing checksums. It must be pointed that a 
-computation might be reproducible based on Peng's definition, but not 
+runs~\cite{hill2017numerical}. This Numerical reproducibility can be 
+checked by comparing the binary content of the results which is 
+computed by the checksum method.
+Regarding these definitions, it must be pointed that a
+computation might be reproducible, but not 
 be numerically reproducible. For instance, small numerical differences 
@@
 Recently, scientists began to realize that the results of many 
-scientific experiments were not reproducible. Results show that the 
-variety of computational infrastructures including workstation types, 
-parallelization methods, operating systems, and analysis packages are 
-known to influence reproducibility because of the creation of small 
-numerical differences~\cite{Gronenschild2012, diethelm2012limits, 
-Glatard2015, bowring2018exploring}. The propagation and amplification 
-of these slight differences by analysis pipelines may cause reproducibility 
-issues. 
-In this case, the analysis pipelines are said to be numerically 
-unstable. Numerical instability is a characteristic of the pipelines 
-which amplify small numerical differences and then hamper the 
-reproducibility of the analyses depending on the length of the pipeline 
-and magnitude of the differences. In many cases, numerical instability is an 
-important issue for reproducibility.
+scientific experiments were not reproducible. According to the previous 
+studies, variety of computational infrastructures including workstation 
+types, parallelization methods, operating systems, and analysis 
+packages are known to influence reproducibility~\cite{Gronenschild2012, 
+diethelm2012limits, Glatard2015, bowring2019exploring}.
 
 In particular, the effect of operating system is quantified 
-in~\cite{Glatard2015, Gronenschild2012} more specifically on some of 
-the main neuroimaging pipelines including FSL, CIVET and Freesurfer 
-pipelines. 
+in~\cite{Glatard2015, Gronenschild2012} specially on some of the main 
+neuroimaging pipelines including FSL, CIVET and Freesurfer pipelines. 
 The binary differences are measured by the checksum method between 
-results of the same analyses on different operating system, which indicate 
-a significant disparity. 
-The operating system is not the only part of the computational 
-environment that may hamper reproducibility. Many other computational 
-factors including hardware configuration, software version, and 
-parallelization techniques are reported as the main influential 
-computing infrastructure on reproducibility.
+results of the same analyses on different operating systems, which 
+indicate a significant disparity.
+
+This irreproducibility issues are reported as the result of the 
+creation, propagation, and amplification of small numerical 
+differences~\cite{Gronenschild2012, diethelm2012limits, Glatard2015, 
+bowring2019exploring}. In this case, the analysis pipelines are said to 
+be numerically unstable. Numerical instability is a characteristic of 
+the pipelines which amplify small numerical differences and then hamper 
+the reproducibility of the analyses depending on the length of the 
+pipeline and magnitude of the differences. In many cases, numerical 
+instability is an important issue for reproducibility.
 
 There are different solutions to make reproducible analysis including 
-containerization techniques to encapsulate software/hardware 
+containerization techniques that encapsulate software/hardware 
 dependencies, provenance capturing tools, and version control systems. 
@@
 systems are considered to find out the cause of the irreproducibility 
-using an interposition techniques.
+using the interposition techniques.
 
-The reminder of this paper describe the NURM-tool in different parts as 
-the way of capturing provenance information and their representation, 
-the influence of different type of subjects, and how to classify 
-processes and characterize differences. We describe the HCP preprocessing 
-pipelines as a popular neuroimaging pipelines to test and apply the 
-NURM-tool on. Finally, we report the experimental results, the processes 
-that are responsible for the differences in the pipelines.
+The remainder of this paper describes the NURM-tool in different parts 
+including the way of capturing and representation of the provenance 
+information, the method of clustering of different subject types, and 
+process classification to characterize differences. We describe the HCP 
+preprocessing pipelines as a popular neuroimaging project to test the 
+NURM-tool and evaluate its functionality. Finally, we reported the 
+experimental results which describe the processes that are responsible 
+for the differences in the pipelines.
 
@@
 
-We present NURM-tool to automatically
-measure numerical reproducibility of the pipelines across operating 
-systems (see Figure~\ref{fig:overview}). We build the 
+We present NURM-tool to automatically measure numerical reproducibility 
+of the pipelines across operating systems (see 
+Figures~\ref{fig:overview-cluster}, \ref{fig:overview-classify}). We build the 
 containerized version of the pipelines for each operating system using 
@@
 
-As shown in the left side of Figure~\ref{fig:overview}, NURM-tool clusters 
-the input subjects into various types based on the similarity of their 
-process trees which are achieved by \reprozip tool. The right side of this figure 
-shows the iterative classification of the process based on the binary 
-comparison of the outputs from various conditions.
-NURM-tool produces the \textit{difference} JSON schema in which 
-recognizes output files with differences using the checksum comparison 
-of the results. We consider that a file contains difference if only 
-its binary comparison shows difference. Also, 
-we use \reprozip tool to capture the provenance information. 
-Subsequently, using a iterative approach, we classify processes into 
-different categories regarding the state of read/write files 
-with/without differences. NURM-tool identifies relevant processes in the 
-pipeline creating, propagating, or removing differences. 
-All the steps are explained in detail in the following sections.
+As shown in Figure~\ref{fig:overview-cluster}, NURM-tool clusters the 
+input subjects into various types based on the similarity of their 
+process trees which are achieved by \reprozip tool. 
+Figure~\ref{fig:overview-classify} shows the iterative classification 
+of the process based on the binary comparison of the outputs from 
+various conditions. NURM-tool produces the \textit{difference} JSON 
+schema in which recognizes output files with differences using the 
+checksum comparison of the results. We consider that a file contains 
+difference if only its binary comparison shows difference. Also, we use 
+\reprozip tool to capture the provenance information. Subsequently, 
+using a iterative approach, we classify processes into different 
+categories regarding the state of read/write files with/without 
+differences. NURM-tool identifies relevant processes in the pipeline 
+creating, propagating, or removing differences. All the steps are 
+explained in detail in the following sections.
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{images/overview.png}
-  \caption{An overview of the proposed technique.}
-  \label{fig:overview}
+  \includegraphics[width=.7\columnwidth]{images/Slide1}
+  \caption{An overview of the subject type clustering method.}
+  \label{fig:overview-cluster}
 \end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=.75\columnwidth]{images/Slide2}
+  \caption{An overview of the process classification method.}
+  \label{fig:overview-classify}
+\end{figure}
+
+%\begin{figure*}
+%\centering
+%    \begin{subfigure}{\columnwidth}
+%        \includegraphics[width=\textwidth]{images/Slide1.png}
+%        \caption{Subject clustering}
+%        \label{fig:green}
+%    \end{subfigure}
+%\hfill%\hfil
+%    \begin{subfigure}{\columnwidth}
+%    \includegraphics[width=\textwidth]{images/Slide2.png}
+%    \caption{Process classification}
+%    \label{fig:red}
+%\end{subfigure}
+%    \caption{An overview of the proposed technique.}
+%    \label{fig:processes}
+%\end{figure*}
 
@@
 call, and stored in a \texttt{SQLite} database. The database stores 
-information on all the processes which are created by the 
+information of all the processes which are created by the 
 \texttt{clone()} or \texttt{fork()} system calls. It also records the 
@@
 
-NURM-tool reconstructs the tree of processes starting from the first process 
-created by the pipeline and identifying its child processes as the ones 
-that were created through \texttt{clone()} or \texttt{fork()}. It 
-creates a process graph from this tree by adding edges corresponding to 
-file dependencies between processes. A file dependency is defined 
-between processes A and B if a file written by A is read by B. 
-Figure~\ref{fig:simple_script} shows an example of a process tree 
-constructed for the simple bash script illustrated in 
+NURM-tool reconstructs the tree of processes starting from the first 
+process created by the pipeline and through identifying its child 
+processes as the ones that were created through \texttt{clone()} or 
+\texttt{fork()}. It creates a process graph from this tree by adding 
+edges corresponding to file dependencies between processes. A file 
+dependency is defined between processes A and B if a file written by A 
+is read by B. Figure~\ref{fig:simple_script} shows an example of a 
+process tree constructed for the simple bash script illustrated in 
 Algorithm~\ref{algo:sample-script} from the brain extraction procedure.
@@
 \centering
-  \includegraphics[scale=0.3]{images/simple_graph}
+  \includegraphics[scale=0.25]{images/simple_graph}
   \caption{Process tree constructed from an example brain extraction
@@
 
-Depending on the variety of computing environments, different  
-process tree can be produced. In particular, different analysis 
-software package can construct different process tree for the 
-similar experiments because of following different methods. We used a 
-fixed analysis software package in different conditions, however, we 
-found that different subjects may produce a different process tree.
+Depending on the variety of computing environments, in particular, 
+different analysis software packages, different process trees can be 
+constructed for the same experiment. It was expected to achieve the 
+same process trees since we used the same computing environments 
+specially the same analysis software package in different conditions. 
+However, we found that different subjects may result in different 
+process trees.
 
-The challenging problem is that of comparing various process trees with 
-different topological structures from different subjects. To address 
-this issue, we categorize subjects into different types based on the 
-distances, topologically differences, between corresponding process 
-trees.
+Therefore, all subjects should be grouped into different types based on 
+the topological structure of the process trees.
+To this order, the distance between trees, topological differences, 
+need to be calculated.
 
-We use the tree edit distance~\cite{zhang1989simple} method to measure 
+We used the tree edit distance~\cite{zhang1989simple} method to measure 
 the distance between labeled trees. The distance between two labeled 
@@
 remove, and insert. Modifying node \textit{n} means changing the label 
-on \textit{n}. Removing a node \textit{n} means making the children of 
-\textit{n} become the children of the parent of \textit{n} and then 
-deleting \textit{n}. Inserting is the the complement of remove, this 
+on \textit{n}. Removing node \textit{n} means making the children of 
+\textit{n} become the children of its parent and then 
+deleting it. Inserting is the complement of remove; this 
 means that inserting \textit{n} as the child of \textit{n}$'$ will make 
@@
 Each operation has an associated cost---insertions and removals each 
-cost 1, and modifications cost either 1 or 2, depending on whether or 
-not you consider a modification to essentially be just a removal 
-followed by an insertion.
+cost 1, and modifications cost either 1 or 2, depending on whether 
+you consider a modification to essentially be just a removal 
+or followed by an insertion.
 
-We cluster subjects into various types using the bottom-up hierarchical 
+We clustered subjects into various types using the bottom-up hierarchical 
 clustering approach implemented in \textit{scipy} python package (see 
-Algorithm~\ref{also:hclustering}). This approach groups together the 
+Algorithm~\ref{algo:hclustering}). This approach groups together the 
 subjects with similar distance, less than a threshold.
 
+% \texttt{agglomerative approach}
+%  based on \texttt{nearest\_neighbors} method
+
 \begin{algorithm}[h!]
-\caption{Hierarchical clustering algorithm}
+\caption{Hierarchical clustering algorithm from \textit{scipy}}
 \label{algo:hclustering}
@@
   \State \texttt{C${_i}$} = \texttt{T${_i}$} 
-  \State \texttt{\# Create a cluster for each tree in input list; \texttt{agglomerative approach}}
+  \State \texttt{\# Create a cluster for each tree in input list;}
   \EndFor
@@
   \For{\texttt{j=i+1} to \texttt{n}}
-  \State \texttt{D${_{ij}}$} = argmin\{\texttt{edit\_distance(C${_i}$, C${_j}$)}\}
-  \State \texttt{\# Select two nearest clusters based on \texttt{nearest\_neighbors} method}
+  \State \texttt{D${_{ij}}$} = \texttt{edit\_distance(C${_i}$, C${_j}$)}
+  \State \texttt{\# Select two nearest clusters;}
   \EndFor
   \EndFor
-  \If{\texttt{D${_{ij}}$} > \texttt{thre}}
+  \State \texttt{d\_min${_{ij}}$} = \texttt{MIN(D)}
+  \If{\texttt{d\_min${_{ij}}$} > \texttt{thre}}
   \State Return \texttt{C}
-  \Else {\State Merge \texttt{C${_i}$ and C${_j}$}}
+  \Else {
+  \State \texttt{C${_i}$} = Merge(\texttt{C${_i}$ and C${_j}$})
+  \State Remove \texttt{C${_j}$}}
   \EndIf
@@
 
-The clustering algorithm starts by finding the two subjects that are 
-closest to each other on the basis of edit distance method and form the 
-new cluster between these two subjects. The next step join this 
-cluster to the next nearest cluster or subject which in turn results in 
-another cluster. This process continues until all the subjects are 
-joined together to form one big cluster or terminates when the distances 
-are less than the threshold. The threshold is defined as the minimum 
-distance required to separate clusters.
+The clustering algorithm starts by finding two closest subjects to each 
+other on the basis of edit distance method and form a new cluster 
+between them. In the next iteration, this cluster aggregates with the 
+next nearest cluster or subject and turns to another cluster. This 
+process continues until all the subjects are aggregated to form one big 
+cluster or the distances are less than an specified threshold. The 
+threshold is defined as the minimum distance required to separate 
+clusters.
 
@@
 
-\begin{figure}
-  \includegraphics[width=1\textwidth]{images/graph}
+\begin{figure*}
+\centering
+  \includegraphics[width=.9\textwidth]{images/graph}
   \caption{A complete process graph from the PreFreesurfer pipeline.
@@
   \label{fig:complete-graph}
-\end{figure}
+\end{figure*}
 
-\begin{figure}
-  \includegraphics[width=\textwidth]{images/hclusters}
+\begin{figure*}
+\centering
+  \includegraphics[width=.8\textwidth]{images/hclusters}
   \caption{Different data types clustered among 100 subjects.}
   \label{fig:subj-clusters}
-\end{figure}
+\end{figure*}
 

### 2019-05-17 paper.tex
Edited tool description

@@
 
+\usepackage{subcaption}
+\usepackage{xspace}
+
 %%%% Commands %%%%
@@
 \newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
-\newcommand{\reprozip}[0]{ReproZip}
+\newcommand{\reprozip}[0]{ReproZip\xspace}
+\newcommand{\tristan}[1]{\color{blue}\textbf{From Tristan:}#1\color{black}}
 
-\title{NURM-tool: a quantitative reproducibility measurement tool to 
-identify numerical differences in pipelines}
+\title{NURM: a tool to 
+locate numerical differences in pipelines}
 
@@
 
-We present NURM-tool to automatically measure numerical reproducibility 
-of the pipelines across operating systems (see 
-Figures~\ref{fig:overview-cluster}, \ref{fig:overview-classify}). We build the 
-containerized version of the pipelines for each operating system using 
-docker technology which encapsulated whole the pipeline dependencies. 
-The docker images consists of an specific computing environment such as 
-operating system version, hardware configuration, software version, 
-etc. Furthermore, to simplify computational reproducibility of the 
-analysis, we execute docker images using Boutiques framework based on 
-the command-line descriptions~\cite{glatard2017boutiques}. The 
-Boutiques descriptor specifies the parameters and implementation of the 
-experiment, and the invocation schema describes the parameter values 
-such as input/output data path.
-
-As shown in Figure~\ref{fig:overview-cluster}, NURM-tool clusters the 
-input subjects into various types based on the similarity of their 
-process trees which are achieved by \reprozip tool. 
-Figure~\ref{fig:overview-classify} shows the iterative classification 
-of the process based on the binary comparison of the outputs from 
-various conditions. NURM-tool produces the \textit{difference} JSON 
-schema in which recognizes output files with differences using the 
-checksum comparison of the results. We consider that a file contains 
-difference if only its binary comparison shows difference. Also, we use 
-\reprozip tool to capture the provenance information. Subsequently, 
-using a iterative approach, we classify processes into different 
-categories regarding the state of read/write files with/without 
-differences. NURM-tool identifies relevant processes in the pipeline 
-creating, propagating, or removing differences. All the steps are 
-explained in detail in the following sections.
-
+Figure~\ref{fig:overview} shows an overview of our method.
+For each tested condition, the
+pipeline is containerized using Docker, and its interface is
+described using Boutiques~\cite{glatard2017boutiques}. \reprozip
+is added to the container images for provenance tracking. Our tool first
+clusters the subjects based on the similarity of their process trees
+produced by \reprozip (Fig.~\ref{fig:overview-cluster}). This results in a
+reduced set of representative process trees to evaluate. The processes in
+each representative process tree are labeled \tristan{Replace
+``classification'' by ``labelling'' everyrwhere} from potential differences
+found in their outputs produced in the tested conditions (Fig.~\ref{fig:overview-classify}).
+To do that, a difference file is first produced that contains all the
+files with differences across conditions
+(Fig.~\ref{fig:overview-classify} - (A)\tristan{Add labels to the figure}).
+Differences are identified simply by comparing file checksums. Using the
+difference file, we then classify processes in 4 categories (Fig.~\ref{fig:overview-classify} - (B)) depending
+whether they create differences (represented in red), remove differences
+(blue), neither create nor remove differences (green), or have uncertain
+behavior (yellow). When processes with an uncertain behavior are found, the
+pipeline in condition 2 is modified (as explained later) and re-executed. This
+procedure iterates until all processes are classified. 
 \begin{figure}
   \centering
+  \begin{subfigure}{\columnwidth}
+    \centering
     \includegraphics[width=.7\columnwidth]{images/Slide1}
-  \caption{An overview of the subject type clustering method.}
+    \caption{Subject clustering.}
     \label{fig:overview-cluster}
-\end{figure}
-
-\begin{figure}
+  \end{subfigure}
+   \begin{subfigure}{\columnwidth}
     \centering
      \includegraphics[width=.75\columnwidth]{images/Slide2}
-  \caption{An overview of the process classification method.}
+     \caption{Process classification.}
      \label{fig:overview-classify}
+   \end{subfigure}
+   \caption{Method overview}
+   \label{fig:overview}
   \end{figure}
-
-%\begin{figure*}
-%\centering
-%    \begin{subfigure}{\columnwidth}
-%        \includegraphics[width=\textwidth]{images/Slide1.png}
-%        \caption{Subject clustering}
-%        \label{fig:green}
-%    \end{subfigure}
-%\hfill%\hfil
-%    \begin{subfigure}{\columnwidth}
-%    \includegraphics[width=\textwidth]{images/Slide2.png}
-%    \caption{Process classification}
-%    \label{fig:red}
-%\end{subfigure}
-%    \caption{An overview of the proposed technique.}
-%    \label{fig:processes}
-%\end{figure*}
 

### 2019-05-17 paper.tex
Edited tool description

@@
 Differences are identified simply by comparing file checksums. Using the
-difference file, we then classify processes in 4 categories (Fig.~\ref{fig:overview-classify} - (B)) depending
+difference file, we then label processes in 4 categories (Fig.~\ref{fig:overview-classify} - (B)) depending
 whether they create differences (represented in red), remove differences
@@
 pipeline in condition 2 is modified (as explained later) and re-executed. This
-procedure iterates until all processes are classified. 
+procedure iterates until all processes are labeled. 
 \begin{figure}

### 2019-05-27 paper.tex

@@
 information, the method of clustering of different subject types, and 
-process classification to characterize differences. We describe the HCP 
+process labelling to characterize differences. We describe the HCP 
 preprocessing pipelines as a popular neuroimaging project to test the 
@@
 
-Figure~\ref{fig:overview} shows an overview of our method.
-For each tested condition, the
-pipeline is containerized using Docker, and its interface is
-described using Boutiques~\cite{glatard2017boutiques}. \reprozip
-is added to the container images for provenance tracking. Our tool first
-clusters the subjects based on the similarity of their process trees
-produced by \reprozip (Fig.~\ref{fig:overview-cluster}). This results in a
-reduced set of representative process trees to evaluate. The processes in
-each representative process tree are labeled \tristan{Replace
-``classification'' by ``labelling'' everyrwhere} from potential differences
-found in their outputs produced in the tested conditions (Fig.~\ref{fig:overview-classify}).
-To do that, a difference file is first produced that contains all the
-files with differences across conditions
-(Fig.~\ref{fig:overview-classify} - (A)\tristan{Add labels to the figure}).
-Differences are identified simply by comparing file checksums. Using the
-difference file, we then label processes in 4 categories (Fig.~\ref{fig:overview-classify} - (B)) depending
-whether they create differences (represented in red), remove differences
-(blue), neither create nor remove differences (green), or have uncertain
-behavior (yellow). When processes with an uncertain behavior are found, the
-pipeline in condition 2 is modified (as explained later) and re-executed. This
-procedure iterates until all processes are labeled. 
+Figure~\ref{fig:overview} shows an overview of our method. For each 
+tested condition, the pipeline is containerized using Docker, and its 
+interface is described using Boutiques~\cite{glatard2017boutiques}. 
+\reprozip is added to the container images for provenance tracking. Our 
+tool first clusters the subjects based on the similarity of their 
+process trees produced by \reprozip (Fig.~\ref{fig:overview-cluster}). 
+This results in a reduced set of representative process trees to 
+evaluate. The processes in each representative process tree are labeled 
+from potential differences found in their outputs produced in the 
+tested conditions (Fig.~\ref{fig:overview-classify}). To do that, a 
+difference file is first produced that contains all the files with 
+differences across conditions (Fig.~\ref{fig:overview-classify} - (A)). 
+Differences are identified simply by comparing file checksums. 
+Using difference file and process tree, we first capture temporary and 
+overwritten files in condition 1. In addition, we set various 
+breakpoints in the pipeline in condition 2 to check the output files 
+produced by each process immediately after running. We then label processes in 4 
+categories (Fig.~\ref{fig:overview-classify} - (B)) depending whether 
+they create differences (represented in red), remove differences 
+(blue), neither create nor remove differences (green), or have 
+uncertain behavior (yellow). When processes that create differences are 
+found, their written files in condition 2 are replaced (as explained 
+later) and then re-labelling processes. This procedure 
+continues until the end of the pipeline execution when all processes 
+are labeled correctly.
+
 \begin{figure}
@@
     \centering
-     \includegraphics[width=.75\columnwidth]{images/Slide2}
-     \caption{Process classification.}
+     \includegraphics[width=1\columnwidth]{images/Slide2}
+     \caption{Process labelling.}
      \label{fig:overview-classify}
@@
 
-NURM-tool classifies the processes into four 
+NURM-tool labels the processes into four 
 categories based on the process tree and the \textit{difference} JSON file 
@@
 
-There are some issues that should be addressed to classify process 
+There are some issues that should be addressed to label process 
 properly, including capturing temporary files that have been removed 
@@
 \subsubsection{Capture temporary files} 
-The classification of processes that read or 
+The labelling of processes that read or 
 write temporary files is uncertain when these files are deleted during 
@@
 Files written by multiple processes also lead 
-to unknown classification labels. To address this issue for a file F 
+to unknown labels. To address this issue for a file F 
 written by processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
@@
 
+we first capture temporary and overwritten files in condition 1
+we set various 
+breakpoints in the pipeline in condition 2 to check the output files 
+produced by each process after running. We then label processes in 4 
+categories. When processes that create differences are 
+found, their written files in condition 2 are replaced (as explained 
+later) and then re-label processes. This procedure (stop/resume) 
+continues until the end of the pipeline execution when all processes 
+are labeled correctly.
+
 \subsubsection{Unknown process}
-Another challenge is to classify processes that 
-read files that have differences and write files that also have differences. In 
-such situations, it is not possible to determine from the pipeline 
-results whether the process created differences, removed some differences, or was 
-transparent. We label such processes \emph{unknown}.
-To address this issue, we developed an iterative approach that 
-consists of the following steps:
+
+Another challenge is to label processes that read files that have 
+differences and write files that also have differences. In such 
+situations, it is not possible to determine from the pipeline results 
+whether the process created differences, removed some differences, or 
+was transparent. We label such processes \emph{unknown}.
+To address this issue, we first set various breakpoints in the pipeline 
+in condition 2. Breakpoints are set by replacing all the pipeline 
+processes with a custom script. This custom script labels each process 
+as \emph{transparent}, \emph{create differences}, \emph{remove 
+differences} or \emph{unknown} immediately after running original executable. 
+The replacement is done through the PATH variable, as before. Moreover, 
+this custom script copies the results obtained from the same process in 
+condition 1 to the pipeline output in condition 2 if it is invoked with 
+the arguments of a process that created differences. 
+We developed an incremental approach that consists of the following steps:
 
 \begin{enumerate}
-  \item Run the pipeline in conditions 1 and 2; classify the
-    processes as \emph{transparent}, \emph{create differences},
-    \emph{remove differences} or \emph{unknown}.
-  \item If there are \emph{unknown} processes then: replace each 
-  process P that create differences by a process Q that copies the results 
-  produced by process P in condition 1 to the pipeline output in 
-  condition 2. 
-  \item Repeat steps 1 and 2 until there is no \emph{unknown} process left.
+  \item Run the pipeline in conditions 2; after running each process 
+  label it as \emph{transparent}, \emph{create differences} or
+  \emph{remove differences}.
+
+  \item If there is \emph{create differences} processes then: 
+  copies the results produced by the same process in condition 1 to the 
+  pipeline output in condition 2.
+
+  \item Continue steps 1 and 2 until the end of the pipeline execution.
 \end{enumerate}
 
-The replacement of process P at step 2 is done by replacing P with a 
-custom script produced by step 1. This custom script copies the results 
-obtained in condition 1 if it is invoked with the arguments of a 
-process that created differences, and calls P's original executable 
-otherwise. The replacement is done through the PATH variable, as 
-before. This algorithm converges to a process tree without any 
-\emph{unknown} process after a finite number of iterations.
+This algorithm converges to a process tree without any 
+\emph{unknown} process at the end of the pipeline execution.
 
@@
   \includegraphics[width=\columnwidth]{images/iterative_modif}
-  \caption{iteration of proposed modification.}
+  \caption{A simple example of steps of the proposed approach.}
   \label{fig:iterations}
@@
 
-Figure~\ref{fig:iterations} illustrates our iterative classification 
+Figure~\ref{fig:iterations} illustrates our incremental labelling 
 process for the example in Figure~\ref{fig:simple_script}. At every 
@@
 The three steps in Figure~\ref{fig:iterations} correspond to the 
-iterations of the classification algorithm. At step one, \texttt{2\#bet2} 
+steps of the labelling algorithm. At step one, \texttt{2\#bet2} 
 is classified as difference creator (red) as it produced files with differences 
@@
 
-As a result of those 3 steps, the final process classification is: 
+As a result of those 3 steps, the final process labelling is: 
 \texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red), 
@@
 reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}.
-We collected the provenance trace (tree of executed processes and files 
-accessed in read or write mode) for each type of 
-subjects(Table~\ref{table:data}) using system-call interception as 
-provided by the \reprozip tool~\cite{Chirigati2016}.
 
@@
 errors are in blue, and other processes are in green.  Squares denote 
-processes for which the classification is uncertain, due to temporary 
+processes for which the labelling is uncertain, due to temporary 
 files that were removed during the execution. Black edges link 

### 2019-05-27 paper.tex
Some updates

--- 2019-05-17 paper.tex
@@
 
+\tristan{define subjects, explain neuroimaging, bring context.}
 \section{Tool description}
@@
 \subsection{Provenance capture and representation}
+\tristan{All these sections should refer to labels on Figure 1.}
 
-We use the \reprozip tool~\cite{Chirigati2016} to record: (1) 
+We use \reprozip~\cite{Chirigati2016} to record: (1) 
 the tree of processes executed by the pipeline and (2) the list of 
 files read and written by each process. This information is collected 
-by system call interceptions, through the \texttt{ptrace} Unix system 
+by \reprozip through the \texttt{ptrace} Unix system 
 call, and stored in a \texttt{SQLite} database. The database stores 
@@
 files opened by each process (in read, write or execution mode), 
-including the result files and the temporary files.
+including result files and temporary files.
 
-NURM-tool reconstructs the tree of processes starting from the first 
-process created by the pipeline and through identifying its child 
-processes as the ones that were created through \texttt{clone()} or 
-\texttt{fork()}. It creates a process graph from this tree by adding 
-edges corresponding to file dependencies between processes. A file 
-dependency is defined between processes A and B if a file written by A 
+Our tool reconstructs the tree of processes starting from the first process
+created by the pipeline and traversing the tree through occurrences of
+\texttt{clone()} and \texttt{fork()}. It also creates a process
+\emph{graph} from this tree by adding edges corresponding to file
+dependencies between processes. A file dependency is defined between
+processes A and B if a file written by A 
 is read by B. Figure~\ref{fig:simple_script} shows an example of a 
-process tree constructed for the simple bash script illustrated in 
-Algorithm~\ref{algo:sample-script} from the brain extraction procedure.
-
-\begin{algorithm}[h!]
-\caption{Sample script from brain extraction process}
-\label{algo:sample-script}
-\begin{verbatim}
+process tree and graph constructed from the example pipeline in
+Listing~\ref{listing:sample-script}.
+\begin{listing}
+\begin{minted}[frame=single,
+  framesep=3mm,
+  linenos=false,
+  xleftmargin=0pt,
+  tabsize=4]{bash}
 #!/bin/bash
-if [ # !=1 ]
+if [ #$ !=1 ]
 then
@@
 fi
+
+# Parse argument, set file names
 input_image=$1
@@
 
+# Run FSL bet, put result ${bet_output}
 bet ${input_image} ${bet_output} > bet_temp.out
 echo "Voxels / Volume in brain mask:"
+# Run FSL stats on ${bet_output}
 fslstats ${bet_output} -V
+# Run FSL maths on ${bet_output}
 fslmaths ${bet_output} ${bet_output_binarized}
-echo "Voxels / Volum in binarized brain mask:"
+echo "Voxels / Volume in binarized brain mask:"
+
 \rm bet_temp.out
-\end{verbatim}
-\end{algorithm}
+\end{minted}
+  \caption{Example pipeline}
+  \label{listing:sample-script}
+\end{listing}
 
@@
   \includegraphics[scale=0.25]{images/simple_graph}
-  \caption{Process tree constructed from an example brain extraction
-	script. Every node in the graph is labeled using (1) a process id 
-	created by our reconstruction, (2) the name of the executable run 
-	by the process. Processes that read or write temporary files are 
-	represented with squares; other ones with circles. Plain edges 
+  \caption{Process tree and graph
+  constructed from the example pipeline in
+  Listing~\ref{listing:sample-script}.
+  Processes that read or write
+  temporary files are 
+  represented with squares. Plain edges 
   represent the process tree (\texttt{fork()} or \texttt{clone()} 
   system calls). Dashed edges represent file dependencies: temporary 
-	files are in yellow and result files are in green.}
+  files are in yellow and result files are in green.
+  Every node in the tree is labeled using (1) a process id created by our
+  reconstruction, (2) the name of the executable run by the process.
+  Process 0 is the initial call to the \texttt{bash} interpreter, and
+  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
+  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
+  was captured by \reprozip while it did not appear in
+  Listing~\ref{listing:sample-script}. \tristan{The script has only 1 call to fslstats, this should be fixed.}
+}
   \label{fig:simple_script}
@@
 
-\subsection{Subject type}
+\subsection{Subject types}
+
+Different subjects may lead to different process trees and graphs, due to
+variations in the processing pipeline. Such variations may be coming from differences
+in data types or cardinality, for instance, some subj
 

### 2019-05-27 paper.tex
Updates in tool description

@@
 
-Figure~\ref{fig:overview} shows an overview of our method. For each 
-tested condition, the pipeline is containerized using Docker, and its 
-interface is described using Boutiques~\cite{glatard2017boutiques}. 
-\reprozip is added to the container images for provenance tracking. Our 
-tool first clusters the subjects based on the similarity of their 
-process trees produced by \reprozip (Fig.~\ref{fig:overview-cluster}). 
-This results in a reduced set of representative process trees to 
-evaluate. The processes in each representative process tree are labeled 
-from potential differences found in their outputs produced in the 
-tested conditions (Fig.~\ref{fig:overview-classify}). To do that, a 
-difference file is first produced that contains all the files with 
-differences across conditions (Fig.~\ref{fig:overview-classify} - (A)). 
-Differences are identified simply by comparing file checksums. 
-Using difference file and process tree, we first capture temporary and 
-overwritten files in condition 1. In addition, we set various 
-breakpoints in the pipeline in condition 2 to check the output files 
-produced by each process immediately after running. We then label processes in 4 
-categories (Fig.~\ref{fig:overview-classify} - (B)) depending whether 
-they create differences (represented in red), remove differences 
-(blue), neither create nor remove differences (green), or have 
-uncertain behavior (yellow). When processes that create differences are 
-found, their written files in condition 2 are replaced (as explained 
-later) and then re-labelling processes. This procedure 
-continues until the end of the pipeline execution when all processes 
-are labeled correctly.
+Figure~\ref{fig:overview} shows an overview of our method \tristan{add reprozip
+tags to 1.b too. Add something to match ``(fixed)'' in 1.b -- cond 2}. For
+each tested condition, the pipeline is containerized with Docker, and its
+interface is described with Boutiques~\cite{glatard2017boutiques}.
+\reprozip~\cite{Chirigati2016} is added to the container images for
+provenance tracking. Our tool first clusters the subjects based on the
+similarity of their process trees produced by \reprozip
+(Fig.~\ref{fig:overview-cluster}) \tristan{``Hclustering'' isn't clear, just
+mention ``clustering''}. This results in a reduced set of representative
+process trees to evaluate. The processes in each representative process
+tree are then labeled from the differences found in their outputs in the
+two tested conditions (Fig.~\ref{fig:overview-classify}). To do that, a
+difference file is first produced that contains all the differences between
+conditions obtained by comparing file checksums
+(Fig.~\ref{fig:overview-classify} - (A)). Using the difference file and the
+process tree in Condition 1, we first identify temporary files and files
+written by more than a single process -- these files result in undetermined
+process labels \tristan{Add label in the Figure. Here and for any other step}.
+We then label processes in 4 categories (Fig.~\ref{fig:overview-classify} -
+(B)) depending whether they create differences (represented in red), remove
+differences (blue), neither create nor remove differences (green), or are
+undetermined (yellow). Process labelling is done in an iterative way,
+through breakpoints and incremental updates in Condition 2, until no
+undetermined process remains.
 
@@
 
-We use \reprozip~\cite{Chirigati2016} to record: (1) 
+We use \reprozip to record: (1) 
 the tree of processes executed by the pipeline and (2) the list of 

### 2019-05-27 paper.tex
Edits and comments on subject types

@@
 \usepackage{xspace}
+\usepackage{stmaryrd} % for llbracket and rrbracket
+\usepackage{amsmath}
+\DeclareMathOperator*{\argmin}{argmin}
 
@@
 
-We use \reprozip to record: (1) 
-the tree of processes executed by the pipeline and (2) the list of 
-files read and written by each process. This information is collected 
-by \reprozip through the \texttt{ptrace} Unix system 
-call, and stored in a \texttt{SQLite} database. The database stores 
-information of all the processes which are created by the 
-\texttt{clone()} or \texttt{fork()} system calls. It also records the 
-files opened by each process (in read, write or execution mode), 
-including result files and temporary files.
+We use \reprozip to record: (1) the set of processes created by the
+pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
+(2) the set of files read, written and executed by each process, including
+temporary files. \reprozip collects this information through the
+\texttt{ptrace()} system call and stores it in a SQLite database.
 
-Our tool reconstructs the tree of processes starting from the first process
-created by the pipeline and traversing the tree through occurrences of
-\texttt{clone()} and \texttt{fork()}. It also creates a process
-\emph{graph} from this tree by adding edges corresponding to file
-dependencies between processes. A file dependency is defined between
-processes A and B if a file written by A 
-is read by B. Figure~\ref{fig:simple_script} shows an example of a 
-process tree and graph constructed from the example pipeline in
+Our tool reconstructs a process tree starting from the first process
+created by the pipeline and traversing processes through \texttt{clone()}
+and \texttt{fork()}. It also creates a process \emph{graph} from this tree
+by adding edges corresponding to file dependencies between processes. A
+file dependency is defined between processes A and B if a file written by A
+is read by B. Figure~\ref{fig:simple_script} shows an example of a process
+tree and graph constructed from the example pipeline in
 Listing~\ref{listing:sample-script}.
@@
 Different subjects may lead to different process trees and graphs, due to
-variations in the processing pipeline. Such variations may be coming from differences
-in data types or cardinality, for instance, some subj
+variations that may be coming from differences in data types or
+cardinality. For instance, some images may have been acquired multiple
+times in some subjects. Some of these differences can be neglected, for
+instance when a data decompression step is present at the beginning of the
+execution for some subjects only, and other ones cannot, when different
+processing paths are used in different subjects.
 
-Depending on the variety of computing environments, in particular, 
-different analysis software packages, different process trees can be 
-constructed for the same experiment. It was expected to achieve the 
-same process trees since we used the same computing environments 
-specially the same analysis software package in different conditions. 
-However, we found that different subjects may result in different 
-process trees.
+To address this issue, we use the tree edit distance~\cite{zhang1989simple}
+to measure the distance between labeled trees \tristan{how about graphs?
+also, why between \emph{labeled} trees and not just trees?}, defined as the
+minimum weighted \tristan{weighted by what?} number of edit operations
+required to transform one tree into the other one. Three edit operations
+are considered: node label modification, node removal, and node insertion.
+Each operation has an associated cost: 1 for insertion and removal, and
+either 1 or 2 for label modification depending on the type of modification. \tristan{Please check my edits.}
 
-Therefore, all subjects should be grouped into different types based on 
-the topological structure of the process trees.
-To this order, the distance between trees, topological differences, 
-need to be calculated.
-
-We used the tree edit distance~\cite{zhang1989simple} method to measure 
-the distance between labeled trees. The distance between two labeled 
-trees is defined as the minimum weighted number of edit operations that 
-it takes to transform one tree to another. There are three edit 
-operations that can be used to transform a tree including modify, 
-remove, and insert. Modifying node \textit{n} means changing the label 
-on \textit{n}. Removing node \textit{n} means making the children of 
-\textit{n} become the children of its parent and then 
-deleting it. Inserting is the complement of remove; this 
-means that inserting \textit{n} as the child of \textit{n}$'$ will make 
-\textit{n} the parent of a consecutive subsequence of the current 
-children of \textit{n}$'$.
-Each operation has an associated cost---insertions and removals each 
-cost 1, and modifications cost either 1 or 2, depending on whether 
-you consider a modification to essentially be just a removal 
-or followed by an insertion.
-
-We clustered subjects into various types using the bottom-up hierarchical 
-clustering approach implemented in \textit{scipy} python package (see 
-Algorithm~\ref{algo:hclustering}). This approach groups together the 
-subjects with similar distance, less than a threshold.
-
-% \texttt{agglomerative approach}
-%  based on \texttt{nearest\_neighbors} method
+With this distance, we cluster subjects using agglomerative hierarchical
+clustering as implemented in SciPy~\cite{oliphant2007scipy}
+(Algorithm~\ref{algo:hclustering}). As a result, the distance between
+any two subjects in a cluster is lower than a threshold parameter.
 
 \begin{algorithm}[h!]
-\caption{Hierarchical clustering algorithm from \textit{scipy}}
+\caption{Hierarchical clustering algorithm from SciPy}
 \label{algo:hclustering}
@@
 
-  \State /* Input: a list of n trees; \texttt{T = [t${_1}$, t${_2}$,...,t${_n}$]}
+  \State /* Input: - a list of n trees; \texttt{T = [t${_1}$, t${_2}$,...,t${_n}$]}
+  \State /*\quad \quad \quad \quad - a distance threshold value; \texttt{threshold}
   \State /* Output: a list of clusters; \texttt{C = [c${_1}$,c${_2}$,...,c${_k}$]}
-  \For{\texttt{i=1} in \texttt{n}}
-  \State \texttt{C${_i}$} = \texttt{T${_i}$} 
   \State \texttt{\# Create a cluster for each tree in input list;}
-  \EndFor
+  \State \texttt{C = [t${_1}$, t${_2}$,...,t${_n}$]}
+  \State \texttt{\# Define a 2 $\times$ 2 array to store distances}
+  \State \texttt{D} = \texttt{array(n, n)}
   \While{\texttt{len(C)} > \texttt{1}}  
+  \State \texttt{\# Select the two nearest clusters}
   \For{\texttt{i=1} to \texttt{n}}
-  \For{\texttt{j=i+1} to \texttt{n}}
-  \State \texttt{D${_{ij}}$} = \texttt{edit\_distance(C${_i}$, C${_j}$)}
-  \State \texttt{\# Select two nearest clusters;}
+  \For{\texttt{j=i+1} to \texttt{len(C)}}
+  \State \texttt{D[i][j]} = \texttt{edit\_distance(C[i], C[j])}
     \EndFor
   \EndFor
-  \State \texttt{d\_min${_{ij}}$} = \texttt{MIN(D)}
-  \If{\texttt{d\_min${_{ij}}$} > \texttt{thre}}
-  \State Return \texttt{C}
-  \Else {
-  \State \texttt{C${_i}$} = Merge(\texttt{C${_i}$ and C${_j}$})
-  \State Remove \texttt{C${_j}$}}
+  \State \texttt{i, j} = $\argmin \left\{ \texttt{D[i][j]}, \ \texttt{i, j} \in \llbracket 1, \texttt{len(C)}\rrbracket^2 \right\}$
+  \If{\texttt{D[i][j]} > \texttt{threshold}}
+  \State \Return \texttt{C}
+  \Else
+  \State \texttt{\# Merge C[j] into C[i]}
+  \State \texttt{C[i]} = \texttt{merge}(\texttt{C[i] and C[j]})
+  \State \texttt{\# Remove C[j] from C}
+  \State \texttt{remove(C[j], C)}
   \EndIf
@@
 \end{algorithm}
-
-The clustering algorithm starts by finding two closest subjects to each 
-other on the basis of edit distance method and form a new cluster 
-between them. In the next iteration, this cluster aggregates with the 
-next nearest cluster or subject and turns to another cluster. This 
-process continues until all the subjects are aggregated to form one big 
-cluster or the distances are less than an specified threshold. The 
-threshold is defined as the minimum distance required to separate 
-clusters.
 

### 2019-05-27 paper.tex
More edits

@@
 
-\subsection{Subject types}
+\subsection{Subject clustering}
 
@@
   \For{\texttt{j=i+1} to \texttt{len(C)}}
-  \State \texttt{D[i][j]} = \texttt{edit\_distance(C[i], C[j])}
+  \State \texttt{D[i][j]} = \texttt{edit\_distance(C[i], C[j])} \tristan{how do you define the distance between two clusters?}
     \EndFor
@@
 
+\tristan{how do you choose the representative process graph in each cluster?}
+
 \subsection{Tree analysis}
 
-NURM-tool labels the processes into four 
-categories based on the process tree and the \textit{difference} JSON file 
-mentioned in the previous sections:
+We label in four categories the processes of the representative graph of each subject:
 \begin{enumerate}
-\item Processes that read and write files that do not have differences 
-are \emph{transparent} (Figure~\ref{fig:processes}.a).
+\item Processes that read files that do not have differences and write files that do not have differences 
+are labeled \emph{transparent} (Figure~\ref{fig:processes}.a).
 \item Processes that read files 
-that do not have any differences but write files that have differences 
-\emph{create} differences in the pipeline (Figure~\ref{fig:processes}.b).
+that do not have differences but write files that have differences 
+\emph{create} differences (Figure~\ref{fig:processes}.b).
 \item Processes that read files 
 that have differences and write files that do not have differences \emph{remove} 
-differences from the pipeline (Figure~\ref{fig:processes}.c).
+differences (Figure~\ref{fig:processes}.c).
 \item Processes that read files that have differences and write files that 
-also have differences are 
-\emph{unknown} (Figure~\ref{fig:processes}.d).
+also have differences are called
+\emph{undetermined} (Figure~\ref{fig:processes}.d).
 \end{enumerate}
@@
     \includegraphics[scale=0.34]{images/yellow.png}
-    \caption{Unknown}
+    \caption{Undetermined}
     \label{fig:yellow}
@@
 
-There are some issues that should be addressed to label process 
-properly, including capturing temporary files that have been removed 
-during the execution, capturing files written by multiple processes, 
-and unknown process.
+Undetermined processes are due to temporary files, and files that were
+written by more than a single process. In both situations, breakpoints must
+be added to the pipeline to capture intermediary file states and label
+undetermined processes.
 
-\subsubsection{Capture temporary files} 
+\subsubsection{Capturing temporary files}
+
 The labelling of processes that read or 
-write temporary files is uncertain when these files are deleted during 
-the execution. To address this issue, we replace every process P that 
-writes temporary files with a modification of P that first calls P and 
-then backs up all its output files to a read-only directory. This 
-replacement is done by modifying the Unix PATH variable to point to a 
-directory containing the modified versions of the processes. This 
+write temporary files is undetermined because these files are deleted during 
+the execution. To address this issue, we replace in Condition 1 every process P that 
+writes temporary files with a modification of P that first calls P and then
+backs up all its output files to a read-only directory. This replacement is
+done by modifying the PATH environment variable to point to a directory
+containing a modified version of the executable called by the process. The
+modified executable will run the original executable and then backup its
+output files. \tristan{Add an example modified script (listing), and replace input checks on file names by checksums.}
+This 
 solution does not cover the temporary files that are removed by P 
-itself; this is not a problem since these files do not play any role in 
-the subsequent steps of the pipeline, by definition.
+itself, but this is not a problem since these files do not play any role in 
+the subsequent steps of the pipeline.
 
-\subsubsection{Capture multiple write files} 
+\subsubsection{Capturing multiple write files}
+
 Files written by multiple processes also lead 

### 2019-05-28 paper.tex
More comments and edits

@@
 
-\subsection{Tree analysis}
+\subsection{Graph analysis}
 
@@
 Files written by multiple processes also lead 
-to unknown labels. To address this issue for a file F 
-written by processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
+to undetermined labels. For a file F 
+written by all processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
 (1) check that processes in \textbf{P} do not write concurrently to F, 
 (2) we establish an order on \textbf{P} based on the creation timestamp 
-of the processes, (3) we replace ever process $P_{i}$ in \textbf{P} by 
+of the processes, (3) we replace every process $P_{i}$ in \textbf{P} by 
 a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
-directory. Thus, multiple versions of F are saved and used in the 
+directory \tristan{add example of wrapper}. Thus, multiple versions of F are saved and used in the 
 analysis.
 
-Furthermore, cycles may be present in the process tree in case a file 
+Furthermore, cycles may be present in the process graph in case a file 
 was written by more than one process. We remove such cycles by removing 
@@
 happen in practice unless A and B were running concurrently, which we 
-assume is not the case (we also checked that on the workflow we 
-tested).
+assume is not the case (the tool checks for that).
 
@@
 continues until the end of the pipeline execution when all processes 
-are labeled correctly.
+are labeled correctly. \tristan{this looks completely out of place}
 
-\subsubsection{Unknown process}
+\tristan{Explain that condition 1 is re-executed after all these modifications are done?}
 
-Another challenge is to label processes that read files that have 
-differences and write files that also have differences. In such 
+\subsubsection{Labelling undetermined processes}
+
+Another challenge is to label undetermined processes that read files with 
+differences and write files with differences too. In such 
 situations, it is not possible to determine from the pipeline results 
 whether the process created differences, removed some differences, or 
-was transparent. We label such processes \emph{unknown}.
-To address this issue, we first set various breakpoints in the pipeline 
-in condition 2. Breakpoints are set by replacing all the pipeline 
+was transparent.
+To address this issue, we set breakpoints in the pipeline 
+in Condition 2. Breakpoints are set by replacing all the pipeline 
 processes with a custom script. This custom script labels each process 
 as \emph{transparent}, \emph{create differences}, \emph{remove 
-differences} or \emph{unknown} immediately after running original executable. 
+differences} or \emph{undetermined} immediately after running the original executable. 
 The replacement is done through the PATH variable, as before. Moreover, 
 this custom script copies the results obtained from the same process in 
-condition 1 to the pipeline output in condition 2 if it is invoked with 
+Condition 1 to the pipeline output in Condition 2 if it is invoked with 
 the arguments of a process that created differences. 
@@
 
+\tristan{Not sure that we should call that breakpoints now.}
+
+\tristan{Why do you need to modify condition 2 now, couldn't this be done on codition 1 too?}
+
 \begin{enumerate}
-  \item Run the pipeline in conditions 2; after running each process 
+  \item Run the pipeline in Condition 2; after running each process 
   label it as \emph{transparent}, \emph{create differences} or
@@
   copies the results produced by the same process in condition 1 to the 
-  pipeline output in condition 2.
+  pipeline output in Condition 2.
 
@@
 This algorithm converges to a process tree without any 
-\emph{unknown} process at the end of the pipeline execution.
+\emph{undetermined} process at the end of the pipeline execution.
+
+\tristan{Now that you're doing this, do you still need subject clustering? It looks like this isn't
+much longer than regular execution.}
 

### 2019-05-28 paper.tex
More edits and comments

@@
   \State \texttt{\# Create a cluster for each tree in input list;}
-  \State \texttt{C = [t${_1}$, t${_2}$,...,t${_n}$]}
+  \State \texttt{C = [ [t${_1}$], [t${_2}$], \ldots , [t${_n}$] ]}
   \State \texttt{\# Define a 2 $\times$ 2 array to store distances}
@@
 Figure~\ref{fig:iterations} illustrates our incremental labelling 
-process for the example in Figure~\ref{fig:simple_script}. At every 
+process for the example in Figure~\ref{fig:simple_script} \tristan{This is outdated, you should
+update the figure and revise the text accordingly.}. At every 
 step, processes that created differences are shown in red, processes that 
@@
 
+\tristan{Add section introduction. Talk about OS}
+
 \subsection{HCP pipelines and datasets}
 
-The Human Connectome Project (HCP)~\cite{glasser2013minimal} developed 
-a set of pipelines to help extraction of structural, functional or 
-diffusion MRI data across a large set of high resolution MR images. 
-These pre-processing pipelines create results that are available in 
-standard volume and combined surface and volume spaces which makes it 
-easier for researchers to compare the images across the neuroimaging 
-spectrum.
+\tristan{Add introduction on HCP project and data}
 
-We focused on HCP structural pre-processing pipelines to (1)quantify 
-the effect of OS on HCP pre-processing pipelines (2) identify the 
-process(es) in the pipelines that are responsible for such effects.
+The Human Connectome Project (\href{https://www.humanconnectome.org}{HCP})
+developed a set of pre-processing pipelines to process structural,
+functional and diffusion MRI data acquired in the project. We focus on HCP
+pre-processing pipelines for structural data, in particular PreFreesurfer
+and Freesurfer. A detailed description of the analyses done in the
+pipelines is available in~\cite{glasser2013minimal}.
+% IF NEEDED, ADD MORE DETAILS ABOUT WHAT THE PIPELINES DO
+% The structural pre-processing pipelines consists of PreFreeSurfer, 
+% FreeSurfer and PostFreeSurfer. In this paper, 
+% we analyzed PreFreeSurfer pipeline which consist of various steps 
+% including correction of MR 
+% gradient nonlinearity distortions, align the T1w and T2w images using 
+% FSL FLIRT, Align native space to MNI template using ACPC and FSL FLIRT, 
+% brain extraction usin FSL FLIRT and FNIRT to MNI template, register T2w 
+% to T1w using FLIRT's BBR, perform a bias field correction, and register 
+% the subject's native structural space to MNI space.
 
-The structural pre-processing pipelines consists of PreFreeSurfer, 
-FreeSurfer and PostFreeSurfer. In this paper, 
-we analyzed PreFreeSurfer pipeline which consist of various steps 
-including correction of MR 
-gradient nonlinearity distortions, align the T1w and T2w images using 
-FSL FLIRT, Align native space to MNI template using ACPC and FSL FLIRT, 
-brain extraction usin FSL FLIRT and FNIRT to MNI template, register T2w 
-to T1w using FLIRT's BBR, perform a bias field correction, and register 
-the subject's native structural space to MNI space.
+In these experiments, we first quantify the effect of Linux version on
+these pipelines. Then we use our NURM tool to identify the processes in the
+pipelines that are responsible for such effects.
 

### 2019-06-03 paper.tex
Edit figures and change the context

@@
 
-Figure~\ref{fig:overview} shows an overview of our method \tristan{add reprozip
-tags to 1.b too. Add something to match ``(fixed)'' in 1.b -- cond 2}. For
+Figure~\ref{fig:overview} shows an overview of our method. For
 each tested condition, the pipeline is containerized with Docker, and its
 interface is described with Boutiques~\cite{glatard2017boutiques}.
-\reprozip~\cite{Chirigati2016} is added to the container images for
-provenance tracking. Our tool first clusters the subjects based on the
-similarity of their process trees produced by \reprozip
-(Fig.~\ref{fig:overview-cluster}) \tristan{``Hclustering'' isn't clear, just
-mention ``clustering''}. This results in a reduced set of representative
-process trees to evaluate. The processes in each representative process
-tree are then labeled from the differences found in their outputs in the
-two tested conditions (Fig.~\ref{fig:overview-classify}). To do that, a
-difference file is first produced that contains all the differences between
-conditions obtained by comparing file checksums
-(Fig.~\ref{fig:overview-classify} - (A)). Using the difference file and the
-process tree in Condition 1, we first identify temporary files and files
-written by more than a single process -- these files result in undetermined
-process labels \tristan{Add label in the Figure. Here and for any other step}.
-We then label processes in 4 categories (Fig.~\ref{fig:overview-classify} -
-(B)) depending whether they create differences (represented in red), remove
-differences (blue), neither create nor remove differences (green), or are
-undetermined (yellow). Process labelling is done in an iterative way,
-through breakpoints and incremental updates in Condition 2, until no
-undetermined process remains.
+\reprozip~\cite{Chirigati2016} is added to one of the container images for
+provenance tracking. 
+Our tool starts by capturing temporary and multi write files in condition 1 
+(Fig.~\ref{fig:overview-capturing}). 
+To do that, the tool first executes pipeline on condition 1 to obtain process graph 
+(Fig.~\ref{fig:overview-capturing} - (P)) and 
+result files. Subsequently, it generates the list of processes that create these intermediary files. 
+Using this list, the tool modifies docker image in condition 1 and exectutes the pipeline 
+again through it, which turns to produce results including temporary and overwritten
+files (Fig.~\ref{fig:overview-capturing} - (C1)).
+In the second step, the processes are labeled from 
+the differences found in their outputs in the
+two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order, 
+the tool builds docker image for condition 2 by modifying pipeline processes that 
+recognised from 
+the process graph. Then, the modified docker image is executed in condition 2 and 
+each process is labeled right 
+after the running by comparing checksum of the results obtained in condition 1 
+and output files of the process in condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
+We then label processes in 3 categories (Fig.~\ref{fig:overview-labelling} -
+(L)) depending whether they create differences (represented in red) or not (green). 
+Process labelling is done through hooks which are defined during the docker image modifying procedure, 
+and incremental updates in condition 2, 
+until the end of the pipeline execution.
 
@@
     \centering
-    \includegraphics[width=.7\columnwidth]{images/Slide1}
-    \caption{Subject clustering.}
-    \label{fig:overview-cluster}
+    \includegraphics[width=1\columnwidth]{images/Slide2}
+    \caption{Capturing of temp and multi write files.}
+    \label{fig:overview-capturing}
   \end{subfigure}
@@
     \centering
-     \includegraphics[width=1\columnwidth]{images/Slide2}
+     \includegraphics[width=1\columnwidth]{images/Slide1}
      \caption{Process labelling.}
-     \label{fig:overview-classify}
+     \label{fig:overview-labelling}
    \end{subfigure}
@@
 \subsection{Provenance capture and representation}
-\tristan{All these sections should refer to labels on Figure 1.}
 
-We use \reprozip to record: (1) the set of processes created by the
+We use \reprozip to create process graph (Fig.~\ref{fig:overview-capturing} - (P)), 
+which records: (1) the set of processes created by the
 pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
@@
   was captured by \reprozip while it did not appear in
-  Listing~\ref{listing:sample-script}. \tristan{The script has only 1 call to fslstats, this should be fixed.}
+  Listing~\ref{listing:sample-script}. 
 }
@@
 \end{figure}
+
+\subsection{Graph analysis}
+
+We label in two categories the processes of the graph of each subject:
+\begin{enumerate}
+\item Processes that read files that do not have differences and write files that do not have differences 
+are labeled \emph{transparent} (Figure~\ref{fig:processes}.a).
+\item Processes that read files 
+that do not have differences but write files that have differences 
+\emph{create} differences (Figure~\ref{fig:processes}.b).
+\end{enumerate}
+
+In addition, there are two more situations in labelling process as:
+\begin{enumerate}
+\item Processes that read files 
+that have differences and write files that do not have differences \emph{remove} 
+differences (Figure~\ref{fig:processes}.c).
+\item Processes that read files that have differences and write files that 
+also have differences are called
+\emph{undetermined} (Figure~\ref{fig:processes}.d).
+\end{enumerate}
+
+\begin{figure}%\centering
+\centering
+    \begin{subfigure}{0.2\linewidth}
+        \includegraphics[scale=0.3]{images/green.png}
+        \caption{Transparent}
+        \label{fig:green}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.2\linewidth}
+    \includegraphics[scale=0.3]{images/red.png}
+    \caption{Creates differences}
+    \label{fig:red}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.2\linewidth}
+    \includegraphics[scale=0.3]{images/blue.png}
+    \caption{Removes differences}
+    \label{fig:blue}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.2\linewidth}
+    \includegraphics[scale=0.3]{images/yellow.png}
+    \caption{Undetermined}
+    \label{fig:yellow}
+\end{subfigure}
+    \caption{Different type of labeled process based on the input/output files.
+  Dashed edges refer to the file dependencies between processes A and B 
+  if a file written by A is read by B. Solid black edges refer to the 
+  relationship between parent and child processes.}
+    \label{fig:processes}
+\end{figure}
+
+Undetermined processes are due to temporary files, and files that were
+written by more than a single process. In both situations, hooks must
+be added to the pipeline to capture intermediary file states and label
+undetermined processes.
+
+\subsubsection{Capturing temporary files}
+
+The labelling of processes that read or 
+write temporary files is undetermined because these files are deleted during 
+the execution. To address this issue, we replace in Condition 1 every process P that 
+writes temporary files with a modification of P that first calls P and then
+backs up all its output files to a read-only directory. This replacement is
+done by modifying the PATH environment variable to point to a directory
+containing a modified version of the executable called by the process. The
+modified executable will run the original executable and then backup its
+output files. \tristan{Add an example modified script (listing), and replace input checks on file names by checksums.}
+This 
+solution does not cover the temporary files that are removed by P 
+itself, but this is not a problem since these files do not play any role in 
+the subsequent steps of the pipeline.
+
+\subsubsection{Capturing multiple write files}
+
+Files written by multiple processes also lead 
+to undetermined labels. For a file F 
+written by all processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
+(1) check that processes in \textbf{P} do not write concurrently to F, 
+(2) we establish an order on \textbf{P} based on the creation timestamp 
+of the processes, (3) we replace every process $P_{i}$ in \textbf{P} by 
+a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
+directory \tristan{add example of wrapper}. Thus, multiple versions of F are saved and used in the 
+analysis.
+
+Furthermore, cycles may be present in the process graph in case a file 
+was written by more than one process. We remove such cycles by removing 
+file edges between processes A and B when A's process creation 
+timestamp is posterior to B's or when A=B. Indeed, such edges cannot 
+happen in practice unless A and B were running concurrently, which we 
+assume is not the case (the tool checks for that).
+
+After that all these modifications are done, we re-execute pipeline in condition 1.
+Figure~\ref{fig:overview-capturing} - (C1) shows pipeline results including intermediary files 
+captured as the outputs in condition 1.
+
+\subsubsection{Labelling undetermined processes}
+
+Since all the result files are captured in condition 1, the tool 
+is ready to label the piepeline processes based on the results in condition 2.
+However, another challenge is to label undetermined processes that read files with 
+differences and write files with differences too. In such 
+situations, it is not possible to determine from the pipeline results 
+whether the process created differences, removed some differences, or 
+was transparent.
+To address this issue, we set hooks in the pipeline 
+in Condition 2. Hooks are set by replacing We identify dall the pipeline 
+processes with a custom script. This custom script labels each process 
+as \emph{transparent} or \emph{create differences} immediately after running 
+the original executable. 
+The replacement is done through the PATH variable, as before. Moreover, 
+this custom script copies the results obtained from the same process in 
+Condition 1 to the pipeline output in Condition 2 if 
+it is invoked with 
+the arguments of a process that created differences. 
+Differences are identified by comparing process name and checksum of the output files.
+We developed an incremental approach that consists of the following steps:
+
+\begin{enumerate}
+  \item Start the pipeline execution in Condition 2; 
+        after hooking all the processes involved in the pipeline.
+  \item Label each process as \emph{transparent} or \emph{create differences} 
+        right after its running (Fig.~\ref{fig:overview-labelling} - (L)).
+  \item Check if process is \emph{create differences} then: 
+        copies the results produced by the same process in Condition 1 to the 
+        pipeline output in Condition 2 (Fig.~\ref{fig:overview-labelling} - (M)). 
+  \item Continue steps 2 and 3 until the end of the pipeline execution.
+\end{enumerate}
+
+This algorithm create a labeled process graph at the end of pipeline execution 
+(Fig.~\ref{fig:overview-labelling} - (LG)).
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{images/iterative_modif}
+  \caption{A simple example of steps of the proposed approach.}
+  \label{fig:iterations}
+\end{figure}
+
+Figure~\ref{fig:iterations} illustrates our incremental labelling 
+process for the example in Figure~\ref{fig:simple_script}. At every 
+step, processes that created differences are shown in red and other processes 
+(transparent processes) are in green. As in 
+Figure~\ref{fig:simple_script}, plain black edges represent the process 
+tree and dashed edges represent file dependencies: green edges 
+represent files with no differences, while red edges represent files with 
+differences. Temporary files are not represented because they have been 
+backed up as previously described.
+
+The four steps in Figure~\ref{fig:iterations} correspond to the 
+steps of the labelling algorithm. 
+At step one, process \texttt{1\#bet} is executed and then labeled 
+as transparent (green) as it produced files without differences.
+At step 2, \texttt{2\#bet2} 
+is labeled as difference creator (red) as it produced files with differences 
+from files without differences. Therefore, the files produced by \texttt{2\#bet2} in  
+condition 2 are replaced with the files produced by \texttt{2\#bet2} in 
+condition 1.
+At step 3, after running process \texttt{3\#fslstats}, it labeled as 
+transparent as it write files without differences.
+At step 4, process \texttt{4\#fslmaths} is labeled as difference creator 
+as the last process of the pipeline.
+As a result of these 4 steps, the final process labelling is: 
+\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red) 
+and the other processes are transparent.
+It should be noted that process that remove differrences are not recognised 
+by this method.
 
@@
 \tristan{how do you choose the representative process graph in each cluster?}
-
-\subsection{Graph analysis}
-
-We label in four categories the processes of the representative graph of each subject:
-\begin{enumerate}
-\item Processes that read files that do not have differences and write files that do not have differences 
-are labeled \emph{transparent} (Figure~\ref{fig:processes}.a).
-\item Processes that read files 
-that do not have differences but write files that have differences 
-\emph{create} differences (Figure~\ref{fig:processes}.b).
-\item Processes that read files 
-that have differences and write files that do not have differences \emph{remove} 
-differences (Figure~\ref{fig:processes}.c).
-\item Processes that read files that have differences and write files that 
-also have differences are called
-\emph{undetermined} (Figure~\ref{fig:processes}.d).
-\end{enumerate}
-
-\begin{figure}%\centering
-\centering
-    \begin{subfigure}{0.4\linewidth}
-        \includegraphics[scale=0.34]{images/green.png}
-        \caption{Transparent}
-        \label{fig:green}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-    \includegraphics[scale=0.34]{images/red.png}
-    \caption{Creates differences}
-    \label{fig:red}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-    \includegraphics[scale=0.34]{images/blue.png}
-    \caption{Removes differences}
-    \label{fig:blue}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.4\linewidth}
-    \includegraphics[scale=0.34]{images/yellow.png}
-    \caption{Undetermined}
-    \label{fig:yellow}
-\end{subfigure}
-    \caption{Different type of classified process based on the input/output files.
-  Dashed edges refer to the file dependencies between processes A and B 
-  if a file written by A is read by B. Solid black edges refer to the 
-  relationship between parent and child processes.}
-    \label{fig:processes}
-\end{figure}
-
-Undetermined processes are due to temporary files, and files that were
-written by more than a single process. In both situations, breakpoints must
-be added to the pipeline to capture intermediary file states and label
-undetermined processes.
-
-\subsubsection{Capturing temporary files}
-
-The labelling of processes that read or 
-write temporary files is undetermined because these files are deleted during 
-the execution. To address this issue, we replace in Condition 1 every process P that 
-writes temporary files with a modification of P that first calls P and then
-backs up all its output files to a read-only directory. This replacement is
-done by modifying the PATH environment variable to point to a directory
-containing a modified version of the executable called by the process. The
-modified executable will run the original executable and then backup its
-output files. \tristan{Add an example modified script (listing), and replace input checks on file names by checksums.}
-This 
-solution does not cover the temporary files that are removed by P 
-itself, but this is not a problem since these files do not play any role in 
-the subsequent steps of the pipeline.
-
-\subsubsection{Capturing multiple write files}
-
-Files written by multiple processes also lead 
-to undetermined labels. For a file F 
-written by all processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
-(1) check that processes in \textbf{P} do not write concurrently to F, 
-(2) we establish an order on \textbf{P} based on the creation timestamp 
-of the processes, (3) we replace every process $P_{i}$ in \textbf{P} by 
-a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
-directory \tristan{add example of wrapper}. Thus, multiple versions of F are saved and used in the 
-analysis.
-
-Furthermore, cycles may be present in the process graph in case a file 
-was written by more than one process. We remove such cycles by removing 
-file edges between processes A and B when A's process creation 
-timestamp is posterior to B's or when A=B. Indeed, such edges cannot 
-happen in practice unless A and B were running concurrently, which we 
-assume is not the case (the tool checks for that).
-
-we first capture temporary and overwritten files in condition 1
-we set various 
-breakpoints in the pipeline in condition 2 to check the output files 
-produced by each process after running. We then label processes in 4 
-categories. When processes that create differences are 
-found, their written files in condition 2 are replaced (as explained 
-later) and then re-label processes. This procedure (stop/resume) 
-continues until the end of the pipeline execution when all processes 
-are labeled correctly. \tristan{this looks completely out of place}
-
-\tristan{Explain that condition 1 is re-executed after all these modifications are done?}
-
-\subsubsection{Labelling undetermined processes}
-
-Another challenge is to label undetermined processes that read files with 
-differences and write files with differences too. In such 
-situations, it is not possible to determine from the pipeline results 
-whether the process created differences, removed some differences, or 
-was transparent.
-To address this issue, we set breakpoints in the pipeline 
-in Condition 2. Breakpoints are set by replacing all the pipeline 
-processes with a custom script. This custom script labels each process 
-as \emph{transparent}, \emph{create differences}, \emph{remove 
-differences} or \emph{undetermined} immediately after running the original executable. 
-The replacement is done through the PATH variable, as before. Moreover, 
-this custom script copies the results obtained from the same process in 
-Condition 1 to the pipeline output in Condition 2 if it is invoked with 
-the arguments of a process that created differences. 
-We developed an incremental approach that consists of the following steps:
-
-\tristan{Not sure that we should call that breakpoints now.}
-
-\tristan{Why do you need to modify condition 2 now, couldn't this be done on codition 1 too?}
-
-\begin{enumerate}
-  \item Run the pipeline in Condition 2; after running each process 
-  label it as \emph{transparent}, \emph{create differences} or
-  \emph{remove differences}.
-
-  \item If there is \emph{create differences} processes then: 
-  copies the results produced by the same process in condition 1 to the 
-  pipeline output in Condition 2.
-
-  \item Continue steps 1 and 2 until the end of the pipeline execution.
-\end{enumerate}
-
-This algorithm converges to a process tree without any 
-\emph{undetermined} process at the end of the pipeline execution.
-
-\tristan{Now that you're doing this, do you still need subject clustering? It looks like this isn't
-much longer than regular execution.}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{images/iterative_modif}
-  \caption{A simple example of steps of the proposed approach.}
-  \label{fig:iterations}
-\end{figure}
-
-Figure~\ref{fig:iterations} illustrates our incremental labelling 
-process for the example in Figure~\ref{fig:simple_script} \tristan{This is outdated, you should
-update the figure and revise the text accordingly.}. At every 
-step, processes that created differences are shown in red, processes that 
-removed differences are in blue, processes that propagated differences are in 
-yellow, and other processes (transparent processes) are in green. As in 
-Figure~\ref{fig:simple_script}, plain black edges represent the process 
-tree and dashed edges represent file dependencies: green edges 
-represent files with no differences, while red edges represent files with 
-differences. Temporary files are not represented because they have been 
-backed up as previously described.
-
-The three steps in Figure~\ref{fig:iterations} correspond to the 
-steps of the labelling algorithm. At step one, \texttt{2\#bet2} 
-is classified as difference creator (red) as it produced files with differences 
-from files without differences. \texttt{3\#fslstats} is classified as difference  
-remover (blue) as it produced files without differences from files with 
-differences. \texttt{4\#fslmaths} and \texttt{5\#fslstats} are classified as 
-unknown (yellow) as they produce files with differences from files with 
-differences.
-
-At step 2, the files produced by \texttt{2\#bet2} in the tested 
-condition are replaced with the files produced by \texttt{2\#bet2} in 
-the other condition. \texttt{3\#fslstats} is now classified as 
-transparent, \texttt{4\#fslmaths} is now classified as difference creator 
-and \texttt{5\#fslstats} is still unknown.
-
-At step 3, the files produced by \texttt{4\#fslmaths} in the tested 
-condition are replaced with the files produced by \texttt{4\#fslmaths} 
-in the other condition. \texttt{5\#fslstats} is now transparent.
-
-As a result of those 3 steps, the final process labelling is: 
-\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red), 
-\texttt{3\#fslstats} is difference remover (blue) and the other processes 
-are transparent.
 

### 2019-06-03 paper.tex
add Listing 2,3 for captured process

@@
 provenance tracking. 
-Our tool starts by capturing temporary and multi write files in condition 1 
+Our tool starts by capturing temporary and multi write files in Condition 1 
 (Fig.~\ref{fig:overview-capturing}). 
-To do that, the tool first executes pipeline on condition 1 to obtain process graph 
+To do that, the tool first executes pipeline on Condition 1 to obtain process graph 
 (Fig.~\ref{fig:overview-capturing} - (P)) and 
 result files. Subsequently, it generates the list of processes that create these intermediary files. 
-Using this list, the tool modifies docker image in condition 1 and exectutes the pipeline 
+Using this list, the tool modifies docker image in Condition 1 and exectutes the pipeline 
 again through it, which turns to produce results including temporary and overwritten
@@
 two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order, 
-the tool builds docker image for condition 2 by modifying pipeline processes that 
+the tool builds docker image for Condition 2 by modifying pipeline processes that 
 recognised from 
-the process graph. Then, the modified docker image is executed in condition 2 and 
+the process graph. Then, the modified docker image is executed in Condition 2 and 
 each process is labeled right 
-after the running by comparing checksum of the results obtained in condition 1 
-and output files of the process in condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
+after the running by comparing checksum of the results obtained in Condition 1 
+and output files of the process in Condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
 We then label processes in 3 categories (Fig.~\ref{fig:overview-labelling} -
@@
 Process labelling is done through hooks which are defined during the docker image modifying procedure, 
-and incremental updates in condition 2, 
+and incremental updates in Condition 2, 
 until the end of the pipeline execution.
@@
 modified executable will run the original executable and then backup its
-output files. \tristan{Add an example modified script (listing), and replace input checks on file names by checksums.}
+output files (see Listing~\ref{listing:modify-script}). 
 This 
@@
 a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
-directory \tristan{add example of wrapper}. Thus, multiple versions of F are saved and used in the 
+directory. Thus, multiple versions of F are saved and used in the 
 analysis.
@@
 
-After that all these modifications are done, we re-execute pipeline in condition 1.
+For example, in the sample pipeline in 
+Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} write a temporary file and 
+also file \emph{T1w\_brain\_bin.nii.gz} is first written by process \texttt{2\#bet2} and 
+then overwritten by process \texttt{3\#fslmaths}. The list of these process are captured in 
+Listing~\ref{listing:json_process}.
+Using this list, we modify relevant processes in Condition 1 and then re-execute pipeline 
+to store them in a backup folder.
 Figure~\ref{fig:overview-capturing} - (C1) shows pipeline results including intermediary files 
-captured as the outputs in condition 1.
+captured as the outputs in Condition 1.
+
+\begin{listing}
+  \begin{minted}[frame=single,
+    framesep=3mm,
+    linenos=false,
+    xleftmargin=0pt,
+    tabsize=4]{JSON}
+{
+  "multi_write_process": {
+    "fslmaths\00T1w_brain.nii.gz\00T1w_brain_bin.nii.gz":
+    {
+      "files":[
+        "/cond1/subj1/T1w_brain_bin.nii.gz"
+      ],
+      "pid": 78
+    },
+    "bet2\00/cond1/subj1/T1w.nii.gz\00T1w_brain.nii.gz":
+    {
+      "files":[
+        "/cond1/subj1/T1w_brain_bin.nii.gz"
+      ],
+      "pid": 75
+    }
+  },
+  "temp_process": {
+    "bet2\00/cond1/subj1/T1w.nii.gz\00T1w_brain.nii.gz": 
+    {
+      "files": [
+        "/cond1/subj1/temp.tmp"
+      ],
+      "pid": 75
+    } 
+  } 
+}
+
+\end{minted}
+\caption{List of process and their files from the example pipeline in 
+Listing~\ref{listing:sample-script} that should be captured.}
+\label{listing:json_process}
+\end{listing}
+
+\begin{listing}
+\begin{minted}[frame=single,
+  framesep=3mm,
+  linenos=false,
+  xleftmargin=0pt,
+  tabsize=4]{python}
+#!/bin/python
+
+# Execute process in cond2 from backup folder
+p_name = sys.argv[0]
+p_exec_cond2 = op.join('backup_scripts', p_name)
+i = 1
+while i < len(sys.argv):
+    p_exec_cond2 += " " + sys.argv[i]
+    i += 1
+subprocess.Popen(p_exec_cond2, shell=True).communicate()
+
+# Load list of process to be checked
+with open('path/to/p_list/', 'r') as p_list:
+    process_list = json.load(p_list)
+for p_exec_cond1, p_files in process_list.items():
+    # Comparing process exec in two conditions
+    if p_exec_cond1 == p_exec_cond2:
+        # Make backup copy
+        capture_file(p_files, from_cond1, to_backup)
+
+\end{minted}
+  \caption{Modify script.}
+  \label{listing:modify-script}
+\end{listing}
 
@@
 
-Since all the result files are captured in condition 1, the tool 
-is ready to label the piepeline processes based on the results in condition 2.
+Since all the result files are captured in Condition 1, the tool 
+is ready to label the piepeline processes based on the results in Condition 2.
 However, another challenge is to label undetermined processes that read files with 
@@
 the arguments of a process that created differences. 
-Differences are identified by comparing process name and checksum of the output files.
+Differences are identified by comparing checksum of the output files for 
+the same process in both conditions.
 We developed an incremental approach that consists of the following steps:
@@
 from files without differences. Therefore, the files produced by \texttt{2\#bet2} in  
-condition 2 are replaced with the files produced by \texttt{2\#bet2} in 
-condition 1.
+Condition 2 are replaced with the files produced by \texttt{2\#bet2} in 
+Condition 1.
 At step 3, after running process \texttt{3\#fslstats}, it labeled as 

### 2019-06-04 paper.tex
edit some typos

@@
 and output files of the process in Condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
-We then label processes in 3 categories (Fig.~\ref{fig:overview-labelling} -
+We then label processes in 2 categories (Fig.~\ref{fig:overview-labelling} -
 (L)) depending whether they create differences (represented in red) or not (green). 
@@
 \centering
-  \includegraphics[scale=0.25]{images/simple_graph}
+  \includegraphics[width=0.3\columnwidth]{images/simple_graph}
   \caption{Process tree and graph
@@
 also file \emph{T1w\_brain\_bin.nii.gz} is first written by process \texttt{2\#bet2} and 
-then overwritten by process \texttt{3\#fslmaths}. The list of these process are captured in 
+then overwritten by process \texttt{3\#fslmaths}. The list of these processes are captured in 
 Listing~\ref{listing:json_process}.
@@
 To address this issue, we set hooks in the pipeline 
-in Condition 2. Hooks are set by replacing We identify dall the pipeline 
+in Condition 2. Hooks are set by replacing the pipeline 
 processes with a custom script. This custom script labels each process 
@@
 and the other processes are transparent.
-It should be noted that process that remove differrences are not recognised 
+It should be noted that processes that remove differrences are not recognised 
 by this method.
@@
   \For{\texttt{j=i+1} to \texttt{len(C)}}
-  \State \texttt{D[i][j]} = \texttt{edit\_distance(C[i], C[j])} \tristan{how do you define the distance between two clusters?}
+  \State \texttt{D[i][j]} = $\min \left\{ \texttt{edit\_dist(a, b)}: \ \texttt{a} \in \texttt{C[i]}, \texttt{b} \in \texttt{C[j]} \right\}$
     \EndFor

### 2019-06-04 paper.tex
Updates

@@
 
-Figure~\ref{fig:overview} shows an overview of our method. For
-each tested condition, the pipeline is containerized with Docker, and its
-interface is described with Boutiques~\cite{glatard2017boutiques}.
-\reprozip~\cite{Chirigati2016} is added to one of the container images for
-provenance tracking. 
-Our tool starts by capturing temporary and multi write files in Condition 1 
-(Fig.~\ref{fig:overview-capturing}). 
-To do that, the tool first executes pipeline on Condition 1 to obtain process graph 
-(Fig.~\ref{fig:overview-capturing} - (P)) and 
-result files. Subsequently, it generates the list of processes that create these intermediary files. 
-Using this list, the tool modifies docker image in Condition 1 and exectutes the pipeline 
-again through it, which turns to produce results including temporary and overwritten
-files (Fig.~\ref{fig:overview-capturing} - (C1)).
-In the second step, the processes are labeled from 
-the differences found in their outputs in the
+Figure~\ref{fig:overview} shows an overview of our method. For each tested
+condition, Condition 1 and Condition 2, the pipeline is containerized with
+Docker, and its interface is described with
+Boutiques~\cite{glatard2017boutiques}. \reprozip~\cite{Chirigati2016} is
+added to Condition 1 for provenance tracking. Our tool starts by
+identifying temporary files and ``multi-write" files, i.e., files written
+by more than one process (Fig.~\ref{fig:overview-capturing}) \tristan{check if concurrent writes are discussed}. To do that,
+the tool first executes the pipeline in Condition 1, to obtain a process
+graph (Fig.~\ref{fig:overview-capturing} - (P)) and a list of result files.
+The list of processes that create intermediary or multi-write
+files is then generated. Using this list, the tool modifies Docker image in Condition 1 and
+exectutes the pipeline again through it, which turns to produce results
+including temporary and overwritten files
+(Fig.~\ref{fig:overview-capturing} - (C1)). In the second step, the
+processes are labeled from the differences found in their outputs in the
 two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order,
-the tool builds docker image for Condition 2 by modifying pipeline processes that 
-recognised from 
-the process graph. Then, the modified docker image is executed in Condition 2 and 
-each process is labeled right 
-after the running by comparing checksum of the results obtained in Condition 1 
-and output files of the process in Condition 2 (Fig.~\ref{fig:overview-labelling} - (D)). 
-We then label processes in 2 categories (Fig.~\ref{fig:overview-labelling} -
-(L)) depending whether they create differences (represented in red) or not (green). 
-Process labelling is done through hooks which are defined during the docker image modifying procedure, 
-and incremental updates in Condition 2, 
-until the end of the pipeline execution.
+the tool builds docker image for Condition 2 by modifying pipeline
+processes that recognised from the process graph. Then, the modified docker
+image is executed in Condition 2 and each process is labeled right after
+the running by comparing checksum of the results obtained in Condition 1
+and output files of the process in Condition 2
+(Fig.~\ref{fig:overview-labelling} - (D)). We then label processes in 2
+categories (Fig.~\ref{fig:overview-labelling} - (L)) depending whether they
+create differences (represented in red) or not (green). Process labelling
+is done through hooks which are defined during the docker image modifying
+procedure, and incremental updates in Condition 2, until the end of the
+pipeline execution.
+
+\tristan{Fig 1-a isn't in the git repo while I had asked it before and it is 
+written in the getting started guide anyway. It is annoying to have to ask for these things
+multiple times while I would like to focus on the content. Also, ppt is among the worst formats
+to put in a git repo as it doesn't play well with diffs.}
+
+\tristan{In Fig 1-a, change ``overwritten'' to ``multi-write''.}
+
+\tristan{We should also discuss the cardinality of ``input subject''. Are we talking
+about 1 subject or all subjects?}
 
@@
     \includegraphics[width=1\columnwidth]{images/Slide2}
-    \caption{Capturing of temp and multi write files.}
+    \caption{Identification of temporary and multi-write files.}
     \label{fig:overview-capturing}
@@
 
-We use \reprozip to create process graph (Fig.~\ref{fig:overview-capturing} - (P)), 
-which records: (1) the set of processes created by the
+We use \reprozip (Fig.~\ref{fig:overview-capturing} - (P))
+to capture: (1) the set of processes created by the
 pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
@@
 created by the pipeline and traversing processes through \texttt{clone()}
-and \texttt{fork()}. It also creates a process \emph{graph} from this tree
+and \texttt{fork()} dependencies. Our tool also creates a process \emph{graph} from this tree
 by adding edges corresponding to file dependencies between processes. A
@@
 #!/bin/bash
-if [ #$ !=1 ]
+
+# This script extracts the brain from an input image,
+# measures the volume of the extracted brain,
+# and creates a binarized brain mask.
+
+if [ $# != 1 ]
 then
-    echo "usage: 0 <inputimage.nii.gz>"
+    echo "usage: $0 <inputimage.nii.gz>"
     exit 1
@@
 input_image=$1
-bet_output="$(basename ${input_image} 
-                       .nii.gz)_brain.nii.gz"
-bet_output_binarized="$(basename ${input_image} 
-                       .nii.gz)_brain_bin.nii.gz"
+bet_output=$(basename "${input_image}" \
+                       .nii.gz)_brain.nii.gz
+bet_output_binarized=$(basename "${input_image}" \
+                       .nii.gz)_brain_bin.nii.gz
 
 # Run FSL bet, put result ${bet_output}
-bet ${input_image} ${bet_output} > bet_temp.out
+bet "${input_image}" "${bet_output}" > bet_temp.out
 echo "Voxels / Volume in brain mask:"
 # Run FSL stats on ${bet_output}
-fslstats ${bet_output} -V
+fslstats "${bet_output}" -V
 # Run FSL maths on ${bet_output}
-fslmaths ${bet_output} ${bet_output_binarized}
-echo "Voxels / Volume in binarized brain mask:"
-
+fslmaths "${bet_output}" "${bet_output_binarized}"
+# Remove temporary file
 \rm bet_temp.out
@@
   reconstruction, (2) the name of the executable run by the process.
-  Process 0 is the initial call to the \texttt{bash} interpreter, and
+  Process 0 is the initial call to the \texttt{sh} interpreter, and
   processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
@@
 
-We label in two categories the processes of the graph of each subject:
-\begin{enumerate}
-\item Processes that read files that do not have differences and write files that do not have differences 
-are labeled \emph{transparent} (Figure~\ref{fig:processes}.a).
-\item Processes that read files 
-that do not have differences but write files that have differences 
-\emph{create} differences (Figure~\ref{fig:processes}.b).
-\end{enumerate}
-
-In addition, there are two more situations in labelling process as:
-\begin{enumerate}
-\item Processes that read files 
-that have differences and write files that do not have differences \emph{remove} 
-differences (Figure~\ref{fig:processes}.c).
-\item Processes that read files that have differences and write files that 
-also have differences are called
-\emph{undetermined} (Figure~\ref{fig:processes}.d).
-\end{enumerate}
+We label in two categories the processes of the graph of each subject. First,
+processes that read files that do not have differences and write files that
+do not have differences are labeled \emph{transparent}
+(Figure~\ref{fig:green}). Second, processes that read files that do not
+have differences but write files that have differences \emph{create}
+differences (Figure~\ref{fig:red}). In addition, processes that read files
+that have differences and write files that also have differences are called
+\emph{undetermined} (Figure~\ref{fig:yellow}). Processes are also labeled
+undetermined when they read or write temporary files, or produce multi-write
+files. To resolve undetermined processes, hooks must be added to the pipeline to 
+execute these processes on files without differences, and to capture
+intermediary file states.
 
@@
     \begin{subfigure}{0.2\linewidth}
-    \includegraphics[scale=0.3]{images/blue.png}
-    \caption{Removes differences}
-    \label{fig:blue}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.2\linewidth}
     \includegraphics[scale=0.3]{images/yellow.png}
@@
 
-Undetermined processes are due to temporary files, and files that were
-written by more than a single process. In both situations, hooks must
-be added to the pipeline to capture intermediary file states and label
-undetermined processes.
+\tristan{You should use Fig. or Figure consistently}
 
@@
 
-The labelling of processes that read or 
+The label of processes that read or 
 write temporary files is undetermined because these files are deleted during 
@@
 writes temporary files with a modification of P that first calls P and then
-backs up all its output files to a read-only directory. This replacement is
+saves all its output files to a read-only directory. This replacement is
 done by modifying the PATH environment variable to point to a directory
 containing a modified version of the executable called by the process. The
-modified executable will run the original executable and then backup its
+modified executable will run the original executable and then save its
 output files (see Listing~\ref{listing:modify-script}). 
+\tristan{Add more comments, in particular to explain what the variables represent.}
 This 
@@
 the subsequent steps of the pipeline. 
+\begin{listing}
+  \begin{minted}[frame=single,
+    framesep=3mm,
+    linenos=false,
+    xleftmargin=0pt,
+    tabsize=4]{python}
+  #!/usr/bin/env python
+  import os.path as op
+
+  # Execute process in cond2 from backup folder
+  p_name = sys.argv[0]
+  p_exec_cond2 = op.join('backup_scripts', p_name)
+  p_exec_cond2 = p_exec_cond2 + " ".join(sys.argv[1:])
+  subprocess.Popen(p_exec_cond2, shell=True).communicate()
+
+  # Load list of process to be checked
+  with open('path/to/p_list/', 'r') as p_list:
+      process_list = json.load(p_list)
+
+  if p_exec_cond1 in process_list.keys():
+      # Make backup copy
+      capture_file(p_files, from_cond1, to_backup)
+
+  \end{minted}
+    \caption{Modified version of the script in Listing~\ref{listing:sample-script} to save temporary files.}
+    \label{listing:modify-script}
+  \end{listing}
 
@@
 of the processes, (3) we replace every process $P_{i}$ in \textbf{P} by 
-a wrapper that first calls $P_{i}$ and then backs up F to a read-only 
+a wrapper that first calls $P_{i}$ and then saves F to a read-only 
 directory. Thus, multiple versions of F are saved and used in the 
@@
 \label{listing:json_process}
-\end{listing}
-
-\begin{listing}
-\begin{minted}[frame=single,
-  framesep=3mm,
-  linenos=false,
-  xleftmargin=0pt,
-  tabsize=4]{python}
-#!/bin/python
-
-# Execute process in cond2 from backup folder
-p_name = sys.argv[0]
-p_exec_cond2 = op.join('backup_scripts', p_name)
-i = 1
-while i < len(sys.argv):
-    p_exec_cond2 += " " + sys.argv[i]
-    i += 1
-subprocess.Popen(p_exec_cond2, shell=True).communicate()
-
-# Load list of process to be checked
-with open('path/to/p_list/', 'r') as p_list:
-    process_list = json.load(p_list)
-for p_exec_cond1, p_files in process_list.items():
-    # Comparing process exec in two conditions
-    if p_exec_cond1 == p_exec_cond2:
-        # Make backup copy
-        capture_file(p_files, from_cond1, to_backup)
-
-\end{minted}
-  \caption{Modify script.}
-  \label{listing:modify-script}
 \end{listing}

### 2019-06-14 paper.tex
Edit Fig1 and revise corresponding text

@@
 identifying temporary files and ``multi-write" files, i.e., files written
-by more than one process (Fig.~\ref{fig:overview-capturing}) \tristan{check if concurrent writes are discussed}. To do that,
+by more than one process (Figure~\ref{fig:overview-capturing}). To do that,
 the tool first executes the pipeline in Condition 1, to obtain a process
-graph (Fig.~\ref{fig:overview-capturing} - (P)) and a list of result files.
+graph (Figure~\ref{fig:overview-capturing} - (1)) and a list of result files.
 The list of processes that create intermediary or multi-write
-files is then generated. Using this list, the tool modifies Docker image in Condition 1 and
+files is then generated (Figure~\ref{fig:overview-capturing} - (2)). 
+Using this list, the tool modifies Docker image in Condition 1 and
 exectutes the pipeline again through it, which turns to produce results
 including temporary and overwritten files
-(Fig.~\ref{fig:overview-capturing} - (C1)). In the second step, the
+(Figure~\ref{fig:overview-capturing} - (3,4)). In the second step, the
 processes are labeled from the differences found in their outputs in the
-two tested conditions (Fig.~\ref{fig:overview-labelling}). To this order,
+two tested conditions (Figure~\ref{fig:overview-labelling}). To this order,
 the tool builds docker image for Condition 2 by modifying pipeline
-processes that recognised from the process graph. Then, the modified docker
+processes (Figure~\ref{fig:overview-labelling} - (1)) that recognised from 
+the process graph. Then, the modified docker
 image is executed in Condition 2 and each process is labeled right after
@@
 and output files of the process in Condition 2
-(Fig.~\ref{fig:overview-labelling} - (D)). We then label processes in 2
-categories (Fig.~\ref{fig:overview-labelling} - (L)) depending whether they
-create differences (represented in red) or not (green). Process labelling
+(Figure~\ref{fig:overview-labelling} - (2)). We then label processes in 2
+categories depending whether they
+create differences (represented in red) or not (green). 
+The tool replaces output file with the same file obtained in Condition 1 
+if process creates differences (Figure~~\ref{fig:overview-labelling} - (3)).
+Process labelling
 is done through hooks which are defined during the docker image modifying
@@
 pipeline execution.
-
-\tristan{Fig 1-a isn't in the git repo while I had asked it before and it is 
-written in the getting started guide anyway. It is annoying to have to ask for these things
-multiple times while I would like to focus on the content. Also, ppt is among the worst formats
-to put in a git repo as it doesn't play well with diffs.}
-
-\tristan{In Fig 1-a, change ``overwritten'' to ``multi-write''.}
-
-\tristan{We should also discuss the cardinality of ``input subject''. Are we talking
-about 1 subject or all subjects?}
 
@@
     \centering
-    \includegraphics[width=1\columnwidth]{images/Slide2}
-    \caption{Identification of temporary and multi-write files.}
+    \includegraphics[width=1\columnwidth]{images/fig1-a}
+    \caption{Identification of temporary and multi-write files.
+    (\href{https://docs.google.com/drawings/d/1rsbFwPjNPNMRhXUl4RSK3G0OaiJC7K-buiHN0hPD8eQ/edit?usp=sharing}{Source})}
     \label{fig:overview-capturing}
@@
     \centering
-     \includegraphics[width=1\columnwidth]{images/Slide1}
-     \caption{Process labelling.}
+     \includegraphics[width=1\columnwidth]{images/fig1-b}
+     \caption{Labelling process graphs including intermediary files.
+     (\href{https://docs.google.com/drawings/d/1yblYuKWAD18aJe5JxBu2h1EysyoQA3YrsQj-1T3s0l8/edit?usp=sharing}{Source})}
      \label{fig:overview-labelling}
    \end{subfigure}
-   \caption{Method overview}
+   \caption{An overview of NURM tool in two steps. First capturing intermediary files (a); then
+            labelling process graphs (b).}
    \label{fig:overview}
@@
 
-We use \reprozip (Fig.~\ref{fig:overview-capturing} - (P))
+We use \reprozip 
 to capture: (1) the set of processes created by the
@@
 created by the pipeline and traversing processes through \texttt{clone()}
-and \texttt{fork()} dependencies. Our tool also creates a process \emph{graph} from this tree
+and \texttt{fork()} dependencies. Our tool also creates a process \emph{graph} 
+(Figure~\ref{fig:overview-capturing} - (1)) from this tree
 by adding edges corresponding to file dependencies between processes. A
@@
 
-\tristan{You should use Fig. or Figure consistently}
-
 \subsubsection{Capturing temporary files}
@@
 output files (see Listing~\ref{listing:modify-script}). 
-\tristan{Add more comments, in particular to explain what the variables represent.}
 This 
@@
 the subsequent steps of the pipeline.
+
 \begin{listing}
@@
   import os.path as op
+  import shutil
 
-  # Execute process in cond2 from backup folder
-  p_name = sys.argv[0]
-  p_exec_cond2 = op.join('backup_scripts', p_name)
-  p_exec_cond2 = p_exec_cond2 + " ".join(sys.argv[1:])
-  subprocess.Popen(p_exec_cond2, shell=True).communicate()
+  # Run the original executable using the same arguments
+  executed_file = sys.argv[0]
+  exec_argv = op.join('original_scripts', executed_file)
+  exec_argv = exec_argv + " ".join(sys.argv[1:])
+  subprocess.Popen(exec_argv, shell=True).communicate()
 
-  # Load list of process to be checked
+  # Load list of captured process to be checked
   with open('path/to/p_list/', 'r') as p_list:
@@
 
-  if p_exec_cond1 in process_list.keys():
-      # Make backup copy
-      capture_file(p_files, from_cond1, to_backup)
+  if exec_argv in process_list.keys():
+      # Save output files to read-only directory
+      p_files = process_list[exec_argv]
+      shutil.copy(p_files, '/read-only/dir')
 
   \end{minted}
-    \caption{Modified version of the script in Listing~\ref{listing:sample-script} to save temporary files.}
+    \caption{Modified version of the script in Listing~\ref{listing:sample-script} to save intermediary files.}
     \label{listing:modify-script}
@@
 
-\subsubsection{Capturing multiple write files}
+\subsubsection{Capturing multi-write files}
 
@@
 
-For example, in the sample pipeline in 
-Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} write a temporary file and 
+As an example of this capturing process, in the sample pipeline in 
+Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} writes a temporary file and 
 also file \emph{T1w\_brain\_bin.nii.gz} is first written by process \texttt{2\#bet2} and 
-then overwritten by process \texttt{3\#fslmaths}. The list of these processes are captured in 
+then overwritten by process \texttt{3\#fslmaths}. The list of these processes is captured in 
 Listing~\ref{listing:json_process}.
 Using this list, we modify relevant processes in Condition 1 and then re-execute pipeline 
-to store them in a backup folder.
-Figure~\ref{fig:overview-capturing} - (C1) shows pipeline results including intermediary files 
+to save them in a backup folder.
+Figure~\ref{fig:overview-capturing} - (4) shows pipeline results including intermediary files 
 captured as the outputs in Condition 1.
@@
 However, another challenge is to label undetermined processes that read files with 
-differences and write files with differences too. In such 
-situations, it is not possible to determine from the pipeline results 
-whether the process created differences, removed some differences, or 
+differences and write files with differences too. In this
+situation, it is not possible to determine from the pipeline results 
+whether the process created differences or 
 was transparent.
@@
 Condition 1 to the pipeline output in Condition 2 if 
-it is invoked with 
-the arguments of a process that created differences. 
+it is invoked with the arguments of a process that created differences 
+(Figure~\ref{fig:overview-labelling} - (3)). 
 Differences are identified by comparing checksum of the output files for 
-the same process in both conditions.
+the same process in both conditions (Figure~\ref{fig:overview-labelling} - (2)).
 We developed an incremental approach that consists of the following steps:
@@
   \item Label each process as \emph{transparent} or \emph{create differences} 
-        right after its running (Fig.~\ref{fig:overview-labelling} - (L)).
+        right after its running.
   \item Check if process is \emph{create differences} then: 
-        copies the results produced by the same process in Condition 1 to the 
-        pipeline output in Condition 2 (Fig.~\ref{fig:overview-labelling} - (M)). 
+        copy the results produced by the same process in Condition 1 to the 
+        pipeline output in Condition 2. 
   \item Continue steps 2 and 3 until the end of the pipeline execution.
@@
 
-This algorithm create a labeled process graph at the end of pipeline execution 
-(Fig.~\ref{fig:overview-labelling} - (LG)).
+This algorithm creates a \emph{labeled process tree} at the end of pipeline execution 
+which highlights processes that create differences.
+(Figure~\ref{fig:overview-labelling} - (4)).
 
@@
 differences. Temporary files are not represented because they have been 
-backed up as previously described.
+saved as previously described.
 
@@
 and the other processes are transparent.
-It should be noted that processes that remove differrences are not recognised 
-by this method.
 
@@
 
-Different subjects may lead to different process trees and graphs, due to
+Different subjects may lead to different process trees, due to
 variations that may be coming from differences in data types or
@@
 
-To address this issue, we use the tree edit distance~\cite{zhang1989simple}
-to measure the distance between labeled trees \tristan{how about graphs?
-also, why between \emph{labeled} trees and not just trees?}, defined as the
-minimum weighted \tristan{weighted by what?} number of edit operations
-required to transform one tree into the other one. Three edit operations
-are considered: node label modification, node removal, and node insertion.
-Each operation has an associated cost: 1 for insertion and removal, and
-either 1 or 2 for label modification depending on the type of modification. \tristan{Please check my edits.}
-
+We cluster subjects into different types to specify how many different 
+\emph{labeled process tree} are generated. This implies the effect of different 
+subjects on pipeline results. First, we cluster subjects based on their process 
+trees that are labeled by the process name. Then, we cluster subjects in each 
+category after re-labeling the processes that create differences. In this order, 
+the tree edit distance~\cite{zhang1989simple} is used as similarity measure for 
+representing the distance between labeled trees. Tree edit distance  is defined  
+as the minimum number of edit operations which is needed to transform one tree into 
+the other. Three edit operations are considered: node label modification, 
+node removal, and node insertion. Each operation has an associated cost of 1.
 With this distance, we cluster subjects using agglomerative hierarchical
@@
 
-\tristan{how do you choose the representative process graph in each cluster?}
-
 \section{Experiments}
@@
 and Freesurfer. A detailed description of the analyses done in the
-pipelines is available in~\cite{glasser2013minimal}.
+pipelines is available in~\cite{glasser2013}.
 % IF NEEDED, ADD MORE DETAILS ABOUT WHAT THE PIPELINES DO
@@
 In the previous study~\cite{Scaria2017}, we showed that pre-processing 
-pipelines of the Human Connectome Project~\cite{Glasser2013} were 
+pipelines of the Human Connectome Project~\cite{glasser2013} were 
 sensitive to operating system variation so that Figure 
@@
 The processes that introduce errors in PreFreeSurfer along with the 
-number of occurrences are depicted in Figures~\ref{fig:pfs_table, 
-fig:pfs_chart} including linear registration with “\emph{FLIRT}” (in 
+number of occurrences are depicted in Figures~\ref{fig:pfs_table} and 
+~\ref{fig:pfs_chart} including linear registration with “\emph{FLIRT}” (in 
 ACPC Alignment, BrainExtraction, DistortionCorrection, 
@@
 instead of masking the problem. These processes need to be reviewed to 
-understand and correct the cause of instabilities. }
+understand and correct the cause of instabilities.
+
+It should be noted that processes that remove differrences, read files with differences and 
+write files without differences, are not recognised 
+by this method.
+}
 

### 2019-06-21 paper.tex
Add results of PFS pipeline

@@
 in~\cite{Glatard2015, Gronenschild2012} specially on some of the main 
-neuroimaging pipelines including FSL, CIVET and Freesurfer pipelines. 
+neuroimaging pipelines including FSL, CIVET and FreeSurfer pipelines. 
 The binary differences are measured by the checksum method between 
@@
 
-Our tool reconstructs a process tree starting from the first process
+Our tool reconstructs a \emph{process tree} starting from the first process
 created by the pipeline and traversing processes through \texttt{clone()}
@@
 
-Different subjects may lead to different process trees, due to
+Different subjects may lead to different \emph{process tree}, due to
 variations that may be coming from differences in data types or
@@
 
+Different \emph{process tree} shows that pipeline execution followed by a different procedure. 
+This implies the possible impact of different subjects on pipeline results.
 We cluster subjects into different types to specify how many different 
-\emph{labeled process tree} are generated. This implies the effect of different 
-subjects on pipeline results. First, we cluster subjects based on their process 
-trees that are labeled by the process name. Then, we cluster subjects in each 
-category after re-labeling the processes that create differences. In this order, 
-the tree edit distance~\cite{zhang1989simple} is used as similarity measure for 
+\emph{process trees} are generated. Clustering is performed based on the \emph{process trees} 
+which nodes are labeled by the process name. 
+In this order, 
+the tree edit distance~\cite{zhang1989simple} is used as a similarity measure for 
 representing the distance between labeled trees. Tree edit distance  is defined  
@@
 any two subjects in a cluster is lower than a threshold parameter.
+
+Furthermore, we will show that even subjects in the same cluster, 
+having the same topological \emph{process tree} structure, 
+may have different labels as processes that create differences.
 
@@
 
-\tristan{Add section introduction. Talk about OS}
+The variety of computing infrastructures, particularly operating system, are 
+known to influence the reproducibility of computational pipelines because of the
+creation of small numerical differences. 
+These differences mainly correspond to the mathematical functions implemented 
+in different operating system libraries.
+For instance, changing the mathematical functions like \emph{expf()} and 
+\emph{cosf()} which manipulate the precision of floating-point representations, 
+between \emph{glibc} libraries in different operating systems can produce 
+small numerical differences.
+Furthermore, a similar issue is expected for any operating system which is 
+based on \emph{glibc}, the GNU C library.
+
+Indeed, the propagation and amplification of these tiny perturbations by analysis 
+pipelines may cause reproducibility issues. We focus on the differences 
+created by different version of operating systems in computational pipelines. 
+In particular, NURM-tool is tested on the neuroimaging applications which are 
+predominantly using mathematical libraries.
 
@@
 
-\tristan{Add introduction on HCP project and data}
+The Human Connectome Project (\href{https://www.humanconnectome.org}{HCP}) 
+is a significant effort that will acquire and analyze brain connectivity 
+data from 1200 healthy adults.
+It enables the neuroscience 
+research community to discover relationships between brain circuits and 
+indivisual behaviors. This helps to understand a wide range of brain disorders.
+The HCP project provides (1) database services (ConnectomeDB) for storing and 
+sharing primary and processed data freely, and (2) data analysis pipelines that 
+are available under an open source license.
 
-The Human Connectome Project (\href{https://www.humanconnectome.org}{HCP})
-developed a set of pre-processing pipelines to process structural,
+The HCP developed a set of pre-processing pipelines to process structural,
 functional and diffusion MRI data acquired in the project. We focus on HCP
-pre-processing pipelines for structural data, in particular PreFreesurfer
-and Freesurfer. A detailed description of the analyses done in the
+pre-processing pipelines for structural data, particularly PreFreeSurfer
+and FreeSurfer. A detailed description of the analyses done in the
 pipelines is available in~\cite{glasser2013}.
@@
 % the subject's native structural space to MNI space.
-
-In these experiments, we first quantify the effect of Linux version on
-these pipelines. Then we use our NURM tool to identify the processes in the
-pipelines that are responsible for such effects.
-
-We randomly selected 100 subjects from the HCP data release S500 and then 
-using the NURM-tool we cluster them into different groups. 
-Figure~\ref{fig:subj-clusters} shows the result of data clustering. 
-Among our dataset, we identified 4 types of subjects with 
-different topology of process trees. In addition, we verified that 
-the process trees were identical for all subjects of the same type in 
-different versions of CentOS operating system. The remainder of the 
-analysis is done separately for each subject type.
+In this paper, we randomly selected 100 unprocessed subjects from the HCP data release S500.
+We first quantify the effect of Linux version on
+these pipelines. Then we identify the processes in the
+pipelines that are responsible for such effects using NURM-tool.
 
@@
 deployment of different operating system versions on execution 
-platforms. The Docker images were built for the HCP PreFreeSurfer 
-pipeline v3.19.0 in 
-CentOS 6.8 and CentOS 7.2. Docker images are available on DockerHub for 
-reuse \url{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}.
+platforms. The Docker images were built for the HCP pre-processing 
+pipelines v3.19.0 (PreFreeSurfer and FreeSurfer) in 
+CentOS 6.8 and CentOS 7.2. Docker images are available on 
+\href{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}{DockerHub}
+for reuse.
 
@@
 occurs as a result of the pseudo-random processes used in the 
-pipelines. An example of a pseudo-random process function is, a random 
-number generator that would get initialized using a seed state. the 
-proposed method can be used to identify both kind of errors. The files 
-that are common to all the subjects only are taken into consideration 
-for comparison. The first step is identification of files with 
-differences in their checksums. This is identified using the checksums 
-that are recorded after the processing. Intra-OS errors are identified 
-using the run-number added as the suffix for the conditions. For 
-example, the two batches of subjects processed under the same condition 
-(CentOS6) are stored as run-1 and run-2. The files belonging to the 
-subjects stored under the above mentioned conditions are treated as 
-intra-OS runs.
+pipelines. 
+We have initialized random number generators using a seed state 
+to avoid intra-OS errors in our experiments. 
+Therefore, NURM-tool is able to identify the process in the pipelines 
+that create differences caused by the inter-OS errors.
 
@@
 
-In the previous study~\cite{Scaria2017}, we showed that pre-processing 
-pipelines of the Human Connectome Project~\cite{glasser2013} were 
+In the previous study~\cite{Scaria2017}, we showed that the HCP pre-processing 
+pipelines~\cite{glasser2013} were 
 sensitive to operating system variation so that Figure 
-\ref{fig:tissue_class} illustrates the binarized errors after brain 
+\ref{fig:tissue_class} illustrates the binarized differences after brain 
 tissue segmentation process in FreeSurfer. In addition, 
-Figure~\ref{fig:fnirt_result} shows the same errors occurred in 
+Figure~\ref{fig:fnirt_result} shows the same differences occurred in 
 non-linear registration process between CentOS6 and CentOS7 in PreFreeSurfer.
@@
   \includegraphics[width=\columnwidth]{images/brain_classification.png} 
-  \caption{Binarized error between brain segmentation results from 
-  FreeSurfer, subject 105216 (CentOS6 vs. CentOS7)~\cite{Scaria2017}.
-    } 
+  \caption{Binarized differences between brain segmentation results from 
+           FreeSurfer, subject 105216 (CentOS6 vs. CentOS7)~\cite{Scaria2017}.} 
   \label{fig:tissue_class}
@@
   \includegraphics[width=\columnwidth]{images/fnirt_result.png} 
-  \caption{Absolute errors between FNIRT results from PreFreeSurfer 
+  \caption{Absolute differences between FNIRT results from PreFreeSurfer 
   (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (CentOS6 vs. 
@@
 
-We identified processes that introduce errors in PreFreeSurfer 
-pipeline. Among the 117 data files produced by PreFreeSurfer, 21 did 
-not have any error for any subject, 92 had errors for all subjects and 
-4 had errors for 3 subjects only.
+We clustered subjects into different groups using the NURM-tool, 
+after that PreFreeSurfer pipeline has executed on input data.
+Figure~\ref{fig:subj-clusters} shows the result of data clustering. 
+Among our dataset, we identified 4 types of subjects with 
+different topology of \emph{process trees}. In addition, we verified that 
+the \emph{process trees} were identical for all subjects of the same type in 
+different versions of CentOS operating system.
 
-Figure \ref{fig:complete-graph} shows the annotated provenance graph 
-of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
-node in the graph represent an executed process in the pipeline. 
-Processes that created errors are shown in red, processes that removed 
-errors are in blue, and other processes are in green.  Squares denote 
-processes for which the labelling is uncertain, due to temporary 
-files that were removed during the execution. Black edges link 
-sub-processes to their parents while dashed edges denote file 
-dependencies between processes (green edges: files with no errors; red 
-edges: files with errors; yellow edges: temporary files).
-
-The processes that introduce errors in PreFreeSurfer along with the 
-number of occurrences are depicted in Figures~\ref{fig:pfs_table} and 
-~\ref{fig:pfs_chart} including linear registration with “\emph{FLIRT}” (in 
+We identified processes that introduce differences in PreFreeSurfer 
+pipeline. 
+The processes that introduce differences in PreFreeSurfer along with the 
+number of occurrences are depicted in Figure~\ref{fig:pfs_chart} 
+including linear registration with “\emph{FLIRT}” (in 
 ACPC Alignment, BrainExtraction, DistortionCorrection, 
@@
 BrainExtraction and AltasRegistration), image warping with 
-“\emph{new\_invwarp}” (in BrainExtraction and AtlasRegistration).  In 
-addition, errors were observed in image mean and standard-deviation 
+“\emph{new\_invwarp}” (in BrainExtraction and AtlasRegistration) and 
+“\emph{applywarp}” (in DistortionCorrection, BrainExtraction, AtlasRegistration).  
+In addition, differences were observed in image mean and standard-deviation 
 computations with “\emph{fslstats}” (in BiasFieldCorrection), and in 
-masked image extrapolation with “\emph{fslmaths}” (in 
-BiasFieldCorrection).  Besides, transformation format conversion with 
-“\emph{convertwarp}” (in DistortionCorrection) was able to remove 
-errors.
+masked image extrapolation with “\emph{fslmaths}” (in BiasFieldCorrection). 
+Besides, transformation format conversion with 
+“\emph{convertwarp}” (in DistortionCorrection) was able to create differences.
+It can be seen that \emph{fslmaths}, \emph{applywarp}, and \emph{flirt} 
+are the most frequent processes that create differences in the 
+PreFreeSurfer pipeline.
+\note{add: \emph{wb\_command}}
+
+Furthermore, we determined the frequency of each process that creates differences in 
+each pipeline element separately in Figure~\ref{fig:pfs_freq}. It is clear that 
+DistortionCorrection element in PreFreeSurfer has the most calls of the processes that 
+create differences.
+% Among the 117 data files produced by PreFreeSurfer, 21 did 
+% not have any error for any subject, 92 had errors for all subjects and 
+% 4 had errors for 3 subjects only.
+
+% Figure \ref{fig:complete-graph} shows the annotated provenance graph 
+% of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
+% node in the graph represent an executed process in the pipeline. 
+% Processes that created errors are shown in red, processes that removed 
+% errors are in blue, and other processes are in green.  Squares denote 
+% processes for which the labelling is uncertain, due to temporary 
+% files that were removed during the execution. Black edges link 
+% sub-processes to their parents while dashed edges denote file 
+% dependencies between processes (green edges: files with no errors; red 
+% edges: files with errors; yellow edges: temporary files).
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{images/pfs_table.png}
-  \caption{Process name along with the pipeline elements that introduce 
-  errors.}
-  \label{fig:pfs_table}
+    \includegraphics[width=\columnwidth]{images/pfs_barchart.png} \caption{Number 
+    of occurrences of processes that create differences in PreFreeSurfer 
+    among 100 subjects.} 
+    \label{fig:pfs_chart}
   \end{figure}
@@
 \centering
-  \includegraphics[width=\columnwidth]{images/pfs_chart.png} \caption{Number 
-  of occurrences of errors in PreFreeSurfer on subject 103515, red and 
-  blue bars indicate the processes which create and remove errors 
-  respectively.} 
-  \label{fig:pfs_chart}
+  \includegraphics[width=\columnwidth]{images/pfs_frequency.png}
+  \caption{Colorized processes along with the pipeline elements that introduce 
+           differences. The range between red and green colors show the most and 
+           least frequency of processes respectively.}
+  \label{fig:pfs_freq}
 \end{figure}
+
+\subsubsection{FreeSurfer pipeline analysis} 
+Result of FreeSurfer\dots
 
@@
 components automatically. The numerical instability in the 
-PreFreesurfer HCP pipeline arises mainly from linear and non-linear 
+PreFreeSurfer HCP pipeline arises mainly from linear and non-linear 
 registration processes implemented in FSL FLIRT and FNIRT.
@@
   \includegraphics[width=.9\textwidth]{images/graph}
-  \caption{A complete process graph from the PreFreesurfer pipeline.
+  \caption{A complete process graph from the PreFreeSurfer pipeline.
 Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}

### 2019-06-21 paper.tex
Comments on results

@@
 and FreeSurfer. A detailed description of the analyses done in the
-pipelines is available in~\cite{glasser2013}.
+pipelines is available in~\cite{glasser2013}. \TG{Introduce and present all the
+pipeline components mentioned in the results, such as BrainExtraction, ACPC alignment, \ldots
+The presentation should be at the right level of detail.}
 % IF NEEDED, ADD MORE DETAILS ABOUT WHAT THE PIPELINES DO
@@
 
-\subsection{Results}
+\section{Results}
 
@@
 \href{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}{DockerHub}
-for reuse.
+for reuse. \TG{Should go to experiments}
 
@@
 We have initialized random number generators using a seed state 
-to avoid intra-OS errors in our experiments. 
+to avoid intra-OS errors in our experiments \TG{how? where?}. 
 Therefore, NURM-tool is able to identify the process in the pipelines 
-that create differences caused by the inter-OS errors.
+that create differences caused by the inter-OS errors. \TG{Change errors to differences.}
+\TG{Check that again, on all subjects.}
 
@@
 
-In the previous study~\cite{Scaria2017}, we showed that the HCP pre-processing 
+In a previous study~\cite{Scaria2017}, we showed that the HCP pre-processing 
 pipelines~\cite{glasser2013} were 
-sensitive to operating system variation so that Figure 
+sensitive to operating system variation. Figure 
 \ref{fig:tissue_class} illustrates the binarized differences after brain 
-tissue segmentation process in FreeSurfer. In addition, 
-Figure~\ref{fig:fnirt_result} shows the same differences occurred in 
-non-linear registration process between CentOS6 and CentOS7 in PreFreeSurfer.
+tissue segmentation in FreeSurfer. In addition, 
+Figure~\ref{fig:fnirt_result} shows the same differences after 
+non-linear in PreFreeSurfer.
 
@@
 after that PreFreeSurfer pipeline has executed on input data.
-Figure~\ref{fig:subj-clusters} shows the result of data clustering. 
+Figure~\ref{fig:subj-clusters} shows the result of data clustering. \TG{Make a table instead of a figure. 
+Mention the threshold and any other clustering parameter.}
 Among our dataset, we identified 4 types of subjects with 
@@
 the \emph{process trees} were identical for all subjects of the same type in 
-different versions of CentOS operating system.
+different versions of CentOS operating system. \TG{Check what these 4 types correspond to}
 

### 2019-09-11 paper.tex
add 'Experiment' section and PFS resutls

--- 2019-06-21 paper.tex
@@
 
-Recently, scientists began to realize that the results of many 
-scientific experiments were not reproducible. According to the previous 
+\note{Recently, scientists began to realize that the results of many 
+scientific experiments were not reproducible.
+
+In particular, validation of the reproducibility has been widely investigated 
+in the field of neuroimaging in which uses optimized 
+processing methods with the purpose of functional and/or structural 
+assessments of the human brain.
+
+According to the previous 
 studies, variety of computational infrastructures including workstation 
@@
 
-In particular, the effect of operating system is quantified 
+In particular, the effect of the operating system is quantified 
 in~\cite{Glatard2015, Gronenschild2012} specially on some of the main 
@@
 indicate a significant disparity. 
+}
 
-This irreproducibility issues are reported as the result of the 
+These irreproducibility issues are reported as the result of the 
 creation, propagation, and amplification of small numerical 
@@
 Reproducibility Measurement (NURM) tool to identify the processes in 
-pipeline that create numerical differences across several runs. In this 
+the pipeline that create numerical differences across several runs. In this 
 paper, the execution results of pipelines across different operating 
 systems are considered to find out the cause of the irreproducibility 
-using the interposition techniques.
+using interposition techniques.
 
@@
 image is executed in Condition 2 and each process is labeled right after
-the running by comparing checksum of the results obtained in Condition 1
+the running by comparing the results obtained in Condition 1
 and output files of the process in Condition 2
@@
 the execution. To address this issue, we replace in Condition 1 every process P that 
-writes temporary files with a modification of P that first calls P and then
+writes temporary files by a wrapper that first calls P and then
 saves all its output files to a read-only directory. This replacement is
@@
   import shutil
+  import pipes
 
@@
   exec_argv = op.join('original_scripts', executed_file)
-  exec_argv = exec_argv + " ".join(sys.argv[1:])
+  exec_argv = exec_argv + " ".join(map(pipes.quote, sys.argv[1:]))
   subprocess.Popen(exec_argv, shell=True).communicate()
@@
 
-The variety of computing infrastructures, particularly operating system, are 
+The variety of computing infrastructures, particularly the operating system, are 
 known to influence the reproducibility of computational pipelines because of the
@@
 Indeed, the propagation and amplification of these tiny perturbations by analysis 
-pipelines may cause reproducibility issues. We focus on the differences 
+may cause reproducibility issues. We focus on the differences 
 created by different version of operating systems in computational pipelines. 
@@
 
-The Human Connectome Project (\href{https://www.humanconnectome.org}{HCP}) 
-is a significant effort that will acquire and analyze brain connectivity 
-data from 1200 healthy adults.
+We used our tool to evaluate the effect of the operating system on results 
+produced by the pipelines from the Human Connectome Project (\href{https://www.humanconnectome.org}{HCP}).
+The HCP initiative is a significant effort that will acquire and analyze 
+brain connectivity data from 1200 healthy adults.
 It enables the neuroscience 
 research community to discover relationships between brain circuits and 
-indivisual behaviors. This helps to understand a wide range of brain disorders.
+individual behaviors. This helps to understand a wide range of brain disorders.
 The HCP project provides (1) database services (ConnectomeDB) for storing and 
@@
 pre-processing pipelines for structural data, particularly PreFreeSurfer
-and FreeSurfer. A detailed description of the analyses done in the
+and FreeSurfer.
+
+According to~\cite{glasser2013}, main steps of the HCP PreFreeSurfer pipeline are:
+(1) Distortion Correction: 
+is the correction of geometric distortions for routine MRI scans 
+to produce an undistorted native structural volume space for each subject.
+(2) Anatomical Average: average same modality images (if more than one is available)
+(3) ACPC (Anterior Commissure, Posterior Commissure) Alignment: 
+is aligning images to the MNI space to have the same orientation as 
+the reference template.
+(4) Brain Extraction: includes a robust initial brain extraction, using FSL's linear (FLIRT) 
+and non-linear (FNIRT) registration. 
+(5) Bias Field Correction: applies intensity inhomogenity correction on the registered images.
+Calculates bias field using square root of the product of T1w and T2w images. 
+(6) Atlas-Registration: register the subject's native structural volume space to MNI space as the last step.
+
+Also, the main goals of HCP FreeSurfer pipeline are:
+(1) improve the robustness of brain extraction,
+(2) fine tune image registration,
+(3) accurately place the surface images with high resolution data,
+(4) perform FreeSurfer's standard folding-based surface registration 
+to their surface atlas.
+
+A detailed description of the analyses done in the
 pipelines is available in~\cite{glasser2013}.
-% IF NEEDED, ADD MORE DETAILS ABOUT WHAT THE PIPELINES DO
-% The structural pre-processing pipelines consists of PreFreeSurfer, 
-% FreeSurfer and PostFreeSurfer. In this paper, 
-% we analyzed PreFreeSurfer pipeline which consist of various steps 
-% including correction of MR 
-% gradient nonlinearity distortions, align the T1w and T2w images using 
-% FSL FLIRT, Align native space to MNI template using ACPC and FSL FLIRT, 
-% brain extraction usin FSL FLIRT and FNIRT to MNI template, register T2w 
-% to T1w using FLIRT's BBR, perform a bias field correction, and register 
-% the subject's native structural space to MNI space.
-In this paper, we randomly selected 100 unprocessed subjects from the HCP data release S500.
-We first quantify the effect of Linux version on
-these pipelines. Then we identify the processes in the
-pipelines that are responsible for such effects using NURM-tool.
 
-%\begin{figure}
-%\centering
-%  \includegraphics[width=\columnwidth]{images/overview.png}
-%  \caption{An overview of the proposed technique.}
-%  \label{fig:overview}
-%\end{figure}
+HCP data consists of brain images collected from an individual healthy person, 
+in the age range from 22-35 years~\cite{van2013wu}. Data from 1200 subjects 
+is available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
+Such a directory containing the HCP data is denoted by the term ``subject".
+In this paper, we randomly selected 100 unprocessed subjects from 
+the HCP data release S500. All the subjects were scanned using the 3-tesla 
+type of MRI scanners.
 
-%\begin{table}
-%\centering
-%\begin{threeparttable}
-%\caption{An overview of the HCP data (Humman Connectome DB).}
-%
-%\begin{tabular}{@{}llllll@{}}
-%\toprule
-%Subject & Release & Acquisition & Gender & Age      & Subj\_Type   \\ \midrule
-%103515  & Q1      & Q02         & F      & 26-30    & type1         \\
-%105216  & Q3      & Q03         & M      & 26-30    & type4         \\
-%103414  & Q2      & Q02         & F      & 22-25    & type1         \\
-%125525  & Q1      & Q01         & F      & 31-35    & type1         \\
-%142828  & Q1      & Q01         & M      & 31-35    & type1         \\
-%129533  & Q3      & Q04         & F      & 31-35    & type1         \\
-%103818  & Q1      & Q01         & F      & 31-35    & type3         \\
-%133928  & Q2      & Q03         & M      & 26-30    & type3         \\
-%148032  & Q3      & Q03         & F      & 31-35    & type2         \\
-%139637  & Q2      & Q03         & F      & 31-35    & type1         \\ \bottomrule
-%\end{tabular}
-%\begin{tablenotes}
-%      \small
-%      \item *Subj\_Type refers to the tree topology type of the subjects.
-%\end{tablenotes}
-%\end{threeparttable}
-%\label{table:data}
-%\end{table}
+\subsection{Processing of data}
+
+We executed the pipelines using Docker containers to simplify the 
+deployment of different operating system versions on execution 
+platforms. 
+The Docker images were built for the HCP pre-processing 
+pipelines v3.19.0 (PreFreeSurfer and FreeSurfer) in 
+CentOS 6.9 (Final) and CentOS 7.4 (Core). 
+CentOS is a widely used operating system among the Neuroscience community because of 
+the appropriate community support, network functions, safety, portability, and openness
+~\cite{hanke2011neuroscience}.
+These containers were installed with all the necessary software and libraries 
+required to run the HCP pipelines including FSL (version 5.0.6), 
+FreeSurfer (version 5.3.0-HCP, CentOS4 build), and Connectome Workbench (version 1.0).
+Docker images are available on 
+\href{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}{DockerHub}
+for reuse.
+
+The data were processed in the order PreFreeSurfer and FreeSurfer. 
+Unprocessed subject data used as the input data for PreFreeSurfer and its output 
+results used as the input data for FreeSurfer analysis in each operating system separately.
+Through all the subjects, first, two runs were made on the same operating systems 
+to be sure that identified differences are not coming from the operations with randomness.
+Then we identified the processes in the pipelines that are responsible for 
+the creation of differences using the NURM-tool.
+
+It should be noted that some files may have the same data but 
+different metadata in headers, including creation time, modification time and date, 
+which turns to non-identical checksums. 
+So, we added some other functions for the file comparison to be sure that differences 
+are in part of the file data. 
+Depending on the file type, we used functions from FreeSurfer software package to determine 
+whether two files differ, including  
+\emph{mri\_diff}, \emph{mris\_diff}, and \emph{lta\_diff} for comparing volumes, surface-based files, 
+and lta files (transforms) respectively. Also, text files are compared using a custom function.
 
@@
 
-We executed the pipelines using Docker containers to simplify the 
-deployment of different operating system versions on execution 
-platforms. The Docker images were built for the HCP pre-processing 
-pipelines v3.19.0 (PreFreeSurfer and FreeSurfer) in 
-CentOS 6.8 and CentOS 7.2. Docker images are available on 
-\href{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}{DockerHub}
-for reuse.
+Two types of differences can occur in the subjects due to the 
+differences in the operating systems. One is inter-OS differences caused by the 
+operating system library updates and the other type, intra-OS differences 
+occur as a result of the pseudo-random processes used in the 
+pipelines.
 
-Two types of errors can occur in the subjects due to the 
-errors in the operating systems. One is inter-OS error caused by the 
-operating system library updates and the other type, intra-OS errors 
-occurs as a result of the pseudo-random processes used in the 
-pipelines. 
-We have initialized random number generators using a seed state 
-to avoid intra-OS errors in our experiments. 
-Therefore, NURM-tool is able to identify the process in the pipelines 
-that create differences caused by the inter-OS errors.
+In our experiments, we identified the processes that create differences due to 
+using random variables in the FreeSurfer pipeline. 
+For example, process \emph{mris\_curvature}
+computes mean and Gaussian curvatures of a cortical surface using the random variable.
+To avoid such randomness we 
+initialized random number generators using a seed state in 
+the command line in which pipeline script calls the program.
+This ensures that results are the same from run to run.
+Therefore, NURM-tool can identify the process in the pipelines 
+that create differences caused by the inter-OS differences.
 
@@
 
-In the previous study~\cite{Scaria2017}, we showed that the HCP pre-processing 
+In a previous study~\cite{Scaria2017}, we showed that the HCP pre-processing 
 pipelines~\cite{glasser2013} were 
-sensitive to operating system variation so that Figure 
-\ref{fig:tissue_class} illustrates the binarized differences after brain 
-tissue segmentation process in FreeSurfer. In addition, 
-Figure~\ref{fig:fnirt_result} shows the same differences occurred in 
-non-linear registration process between CentOS6 and CentOS7 in PreFreeSurfer.
+sensitive to operating system variation.
+
+Figure~\ref{fig:fnirt_result} shows differences in the \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz} files. 
+These files are created as a part of ACPC-Alignment (Step 3 of PreFreeSurfer).
+The RMSE (root mean squared error) value of results is 0.029 and the dice coefficient is 0.125.
+The square pixels in the brain in this checkerboard images show the differences between results in two OSes. 
+In the next subsection, we will show that such misalignment is created due to 
+non-linear registration in \emph{FNIRT} process in PreFreeSurfer pipeline.
+
+Furthermore, Figure 
+\ref{fig:tissue_class} illustrates the binarized differences in the \emph{aseg.hires.mgz} files.
+This is the result of brain tissue segmentation across OSes in FreeSurfer pipeline. 
+Each color represents a different structure in the brain.
+The bright spots show relevant pixels in the image which are different.
+The RMSE and the dice coefficient values of results are 0.009 and 0.991 respectively.
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\columnwidth]{images/segmentation.png} 
+    \caption{Differences between FNIRT results from PreFreeSurfer 
+    (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
+    CentOS7). Right images show the zoomed version part of the images marked 
+    by the red rectangle.} 
+    \label{fig:fnirt_result}
+\end{figure}
 
@@
   \caption{Binarized differences between brain segmentation results from 
-           FreeSurfer, subject 105216 (CentOS6 vs. CentOS7)~\cite{Scaria2017}.} 
+           FreeSurfer (CentOS6 vs. CentOS7).} 
   \label{fig:tissue_class}
-\end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=\columnwidth]{images/fnirt_result.png} 
-  \caption{Absolute differences between FNIRT results from PreFreeSurfer 
-  (T2w\_acpc\_to\_MNI\_nonli.nii.gz), subject 104820 (CentOS6 vs. 
-  CentOS7)~\cite{Scaria2017}. } 
-  \label{fig:fnirt_result}
 \end{figure}
@@
 
-We clustered subjects into different groups using the NURM-tool, 
-after that PreFreeSurfer pipeline has executed on input data.
-Figure~\ref{fig:subj-clusters} shows the result of data clustering. 
+First, subjects are clustered into different groups using the clustering method as described before.
+We choosed the threshold value of zero for the clustering method. 
+This turns to have the subjects with the same topology in each cluster.
+Also, we used the nearest neighbor algorithm to calculate the distance between clusters.
+Table~\ref{table:data-clusters} shows the PreFreeSurfer result of subject clustering. 
 Among our dataset, we identified 4 types of subjects with 
-different topology of \emph{process trees}. In addition, we verified that 
+different topology of \emph{process trees}. 
+Differences correspond to the data processing that leads to producing disparate output file(s).
+For example, subject type 3 missed output file \emph{T1w2\_gdc.nii.gz} after processing 
+in comparison with subject type 1.
+Besides, we verified that 
 the \emph{process trees} were identical for all subjects of the same type in 
-different versions of CentOS operating system.
+different versions of CentOS.
 
-We identified processes that introduce differences in PreFreeSurfer 
-pipeline. 
-The processes that introduce differences in PreFreeSurfer along with the 
-number of occurrences are depicted in Figure~\ref{fig:pfs_chart} 
+\begin{table}
+\centering
+\begin{threeparttable}
+\caption{Clustering of 20 input subjects.}
+
+\begin{tabular}{@{}llllll@{}}
+\toprule
+Type   & Number of Subjects  & Missing File(s)                                    \\ \midrule
+type1  & 9                   & None                                               \\
+type2  & 7                   & [``T1w/T1w2\_gdc.nii.gz",                          \\
+       &                     &  ``T2w/T2w2\_gdc.nii.gz"]                          \\
+type3  & 2                   & [``T1w/T1w2\_gdc.nii.gz"]                          \\
+type4  & 2                   & [``T2w/T2w2\_gdc.nii.gz"]                          \\ \bottomrule
+\end{tabular}
+\begin{tablenotes}
+     \small
+     \item *Type refers to the topology type of the \emph{process tree} of subjects.
+\end{tablenotes}
+\end{threeparttable}
+\label{table:data-clusters}
+\end{table}
+
+Then, we determined the frequency of the processes that create differences in 
+the pipeline in Figure~\ref{fig:pfs_freq}. \emph{FLIRT} is the most frequent 
+process that produces differences.
+Furthermore, the percentage of processes that introduce differences 
+in PreFreeSurfer along with the 
+pipeline elements are depicted in Figure~\ref{fig:pfs_freq} 
 including linear registration with “\emph{FLIRT}” (in 
-ACPC Alignment, BrainExtraction, DistortionCorrection, 
+ACPC-Alignment, BrainExtraction, DistortionCorrection, 
 AtlasRegistration), non-linear registration with “\emph{FNIRT}” (in 
 BrainExtraction and AltasRegistration), image warping with 
-“\emph{new\_invwarp}” (in BrainExtraction and AtlasRegistration) and 
-“\emph{applywarp}” (in DistortionCorrection, BrainExtraction, AtlasRegistration).  
-In addition, differences were observed in image mean and standard-deviation 
-computations with “\emph{fslstats}” (in BiasFieldCorrection), and in 
-masked image extrapolation with “\emph{fslmaths}” (in BiasFieldCorrection). 
-Besides, transformation format conversion with 
-“\emph{convertwarp}” (in DistortionCorrection) was able to create differences.
-It can be seen that \emph{fslmaths}, \emph{applywarp}, and \emph{flirt} 
-are the most frequent processes that create differences in the 
-PreFreeSurfer pipeline.
-\note{add: \emph{wb\_command}}
+“\emph{new\_invwarp}” (in BrainExtraction and AtlasRegistration). 
+Also, differences were observed in image mean and standard-deviation 
+computations in image interpolation with “\emph{fslmaths}” (in AnatomicalAverage). 
+It can be seen that \emph{fslmaths} and \emph{FLIRT} 
+are the processes that at least have one subprocess to produce differences across all the subjects.
 
-Furthermore, we determined the frequency of each process that creates differences in 
-each pipeline element separately in Figure~\ref{fig:pfs_freq}. It is clear that 
-DistortionCorrection element in PreFreeSurfer has the most calls of the processes that 
-create differences.
-% Among the 117 data files produced by PreFreeSurfer, 21 did 
-% not have any error for any subject, 92 had errors for all subjects and 
-% 4 had errors for 3 subjects only.
-
-% Figure \ref{fig:complete-graph} shows the annotated provenance graph 
-% of the PreFreeSurfer pipeline executed on CentOS6 and CentOS7. Each 
-% node in the graph represent an executed process in the pipeline. 
-% Processes that created errors are shown in red, processes that removed 
-% errors are in blue, and other processes are in green.  Squares denote 
-% processes for which the labelling is uncertain, due to temporary 
-% files that were removed during the execution. Black edges link 
-% sub-processes to their parents while dashed edges denote file 
-% dependencies between processes (green edges: files with no errors; red 
-% edges: files with errors; yellow edges: temporary files).
+% \begin{figure}
+%   \centering
+%     \includegraphics[width=\columnwidth]{images/pfs_barchart.png} \caption{Number 
+%     of occurrences of processes that create differences in PreFreeSurfer 
+%     among 20 subjects.} 
+%     \label{fig:pfs_chart}
+%   \end{figure}
 
@@
 \centering
-    \includegraphics[width=\columnwidth]{images/pfs_barchart.png} \caption{Number 
-    of occurrences of processes that create differences in PreFreeSurfer 
-    among 100 subjects.} 
-    \label{fig:pfs_chart}
-  \end{figure}
-
-\begin{figure}
-\centering
-  \includegraphics[width=\columnwidth]{images/pfs_frequency.png}
+  \includegraphics[width=\columnwidth]{images/pfs_heatmap.png}
   \caption{Colorized processes along with the pipeline elements that introduce 
-           differences. The range between red and green colors show the most and 
-           least frequency of processes respectively.}
+           differences among 20 subjects. The range between red and green colors show the most and 
+           least frequency of processes respectively. Besides, the number 
+           of occurrences of each processes with differences is represented in the line below the heatmap.}
   \label{fig:pfs_freq}
@@
 \subsubsection{FreeSurfer pipeline analysis}
-Result of FreeSurfer\dots
+
+We found that inter-OS differences occur if pre-processed data are different in FreeSurfer. 
+Figure~\ref{fig:tissue_class} shows differences that coming from PreFreeSurfer analysis 
+as a result of propagation of differences. 
+NURM tool identified that there are no processs in the pipeline that create differences.
+This is because of using static builds of FreeSurfer which statically linked libraries (embeded libraries in the pipeline).
 
@@
 
-\note{Pipeline amplify small numerical differences because they are numerically 
+\note{Pipelines amplify small numerical differences because they are numerically 
 unstable. Furthermore, math libraries evolve over time, leading to 
@@
 understand and correct the cause of instabilities. 
-
-It should be noted that processes that remove differrences, read files with differences and 
-write files without differences, are not recognised 
-by this method.
 }
@@
 
-\begin{figure*}
-\centering
-  \includegraphics[width=.9\textwidth]{images/graph}
-  \caption{A complete process graph from the PreFreeSurfer pipeline.
-Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
-  \label{fig:complete-graph}
-\end{figure*}
+% \begin{figure*}
+% \centering
+%   \includegraphics[width=.9\textwidth]{images/graph}
+%   \caption{A complete process graph from the PreFreeSurfer pipeline.
+% Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
+%   \label{fig:complete-graph}
+% \end{figure*}
 
-\begin{figure*}
-\centering
-  \includegraphics[width=.8\textwidth]{images/hclusters}
-  \caption{Different data types clustered among 100 subjects.}
-  \label{fig:subj-clusters}
-\end{figure*}
+% \begin{figure*}
+% \centering
+%   \includegraphics[width=.8\textwidth]{images/hclusters}
+%   \caption{Different data types clustered among 100 subjects.}
+%   \label{fig:subj-clusters}
+% \end{figure*}
 

### 2019-09-19 paper.tex
Add Results and Discussion section

@@
 
-In particular, the effect of the operating system is quantified 
-in~\cite{Glatard2015, Gronenschild2012} specially on some of the main 
-neuroimaging pipelines including FSL, CIVET and FreeSurfer pipelines. 
-The binary differences are measured by the checksum method between 
-results of the same analyses on different operating systems, which 
-indicate a significant disparity. 
+In particular, the effect of 
+the operating system across some of the main 
+neuroimaging pipelines including FSL, CIVET and FreeSurfer pipelines 
+is studied in~\cite{Glatard2015, Gronenschild2012} that indicates a significant disparity. 
 }
@@
 
-The variety of computing infrastructures, particularly the operating system, are 
+The variety of computing infrastructures, particularly the operating system, is 
 known to influence the reproducibility of computational pipelines because of the
@@
 small numerical differences.
-Furthermore, a similar issue is expected for any operating system which is 
+Besides, a similar issue is expected for any operating system which is 
 based on \emph{glibc}, the GNU C library.
@@
 Indeed, the propagation and amplification of these tiny perturbations by analysis 
-may cause reproducibility issues. We focus on the differences 
+may cause a significant reproducibility issue. We focus on the differences 
 created by different version of operating systems in computational pipelines. 
@@
 
-\subsection{HCP pipelines and datasets}
+\subsection{HCP pipelines and dataset}
 
@@
 
-According to~\cite{glasser2013}, main steps of the HCP PreFreeSurfer pipeline are:
-(1) Distortion Correction: 
-is the correction of geometric distortions for routine MRI scans 
-to produce an undistorted native structural volume space for each subject.
-(2) Anatomical Average: average same modality images (if more than one is available)
-(3) ACPC (Anterior Commissure, Posterior Commissure) Alignment: 
+According to~\cite{glasser2013}, main sub-pipelines of the HCP PreFreeSurfer are:
+(1) Distortion Correction (DC) 
+to correct the geometric distortions for routine MRI scans 
+and produce an undistorted native structural volume space for each subject,
+(2) Anatomical Average (AAve) to average same modality images (if more than one is available),
+(3) ACPC (Anterior Commissure, Posterior Commissure) Alignment (ACPC-A) 
 is aligning images to the MNI space to have the same orientation as 
-the reference template.
-(4) Brain Extraction: includes a robust initial brain extraction, using FSL's linear (FLIRT) 
-and non-linear (FNIRT) registration. 
-(5) Bias Field Correction: applies intensity inhomogenity correction on the registered images.
-Calculates bias field using square root of the product of T1w and T2w images. 
-(6) Atlas-Registration: register the subject's native structural volume space to MNI space as the last step.
+the reference template,
+(4) Brain Extraction (BExt) includes a robust initial brain extraction, using FSL's linear (FLIRT) 
+and non-linear (FNIRT) registration, 
+(5) Bias Field Correction (BFC) applies intensity inhomogenity correction on the registered images, 
+%Calculates bias field using square root of the product of T1w and T2w images. 
+(6) Atlas-Registration (AR) to register the subject's native structural volume space to MNI space as the last step.
 
@@
 
-A detailed description of the analyses done in the
+A detailed description of the analyses done in these
 pipelines is available in~\cite{glasser2013}.
@@
 HCP data consists of brain images collected from an individual healthy person, 
-in the age range from 22-35 years~\cite{van2013wu}. Data from 1200 subjects 
+in the age range between 22-35 years~\cite{van2013wu}. Data from 1200 subjects 
 is available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
 Such a directory containing the HCP data is denoted by the term ``subject".
-In this paper, we randomly selected 100 unprocessed subjects from 
+In this paper, we randomly selected 20 unprocessed subjects from 
 the HCP data release S500. All the subjects were scanned using the 3-tesla 
@@
 \emph{mri\_diff}, \emph{mris\_diff}, and \emph{lta\_diff} for comparing volumes, surface-based files, 
-and lta files (transforms) respectively. Also, text files are compared using a custom function.
+and lta files (transforms) respectively. 
+These functions check data based on the volume dimentions, 
+resolutions, pixel intensities,  
+the number of surface vertices and neighbors, and geometry information. 
+It then returns files differ if there be differences in any of these parameters.
+Also, text files are compared using a custom function that compares the content line by line.
 
@@
 In our experiments, we identified the processes that create differences due to 
-using random variables in the FreeSurfer pipeline. 
+intra-OS differences.  
 For example, process \emph{mris\_curvature}
-computes mean and Gaussian curvatures of a cortical surface using the random variable.
+computes mean and Gaussian curvatures of a cortical surface using 
+the random variable in FreeSurfer.
 To avoid such randomness we 
@@
 This ensures that results are the same from run to run.
-Therefore, NURM-tool can identify the process in the pipelines 
-that create differences caused by the inter-OS differences.
+Therefore, NURM-tool can identify the processes in the pipelines 
+that create inter-OS differences.
 
@@
 
-Figure~\ref{fig:fnirt_result} shows differences in the \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz} files. 
-These files are created as a part of ACPC-Alignment (Step 3 of PreFreeSurfer).
+Figure~\ref{fig:fnirt_result} shows differences in the files 
+that are created as a part of ACPC-Alignment in PreFreeSurfer.
+The square pixels in the brain in the checkerboard images show the differences between results in two OSes. 
 The RMSE (root mean squared error) value of results is 0.029 and the dice coefficient is 0.125.
-The square pixels in the brain in this checkerboard images show the differences between results in two OSes. 
-In the next subsection, we will show that such misalignment is created due to 
+In the next subsection, it is shown that such misalignment is created due to 
 non-linear registration in \emph{FNIRT} process in PreFreeSurfer pipeline.
@@
 Furthermore, Figure 
-\ref{fig:tissue_class} illustrates the binarized differences in the \emph{aseg.hires.mgz} files.
-This is the result of brain tissue segmentation across OSes in FreeSurfer pipeline. 
-Each color represents a different structure in the brain.
-The bright spots show relevant pixels in the image which are different.
-The RMSE and the dice coefficient values of results are 0.009 and 0.991 respectively.
+\ref{fig:tissue_class} illustrates the heatmap binarized differences 
+of the brain tissue segmentation across OSes in FreeSurfer pipeline.
+These differences are accumulated from 20 subject results.
+So, the magnitude of differences is in the range of 1-20 as the total number of subjects.
+The darker blue pixels show differences among less subjests and 
+the brighter blue pixels are related to the differences among more subjects.
+
+%The brighter spots show relevant pixels in the segmented image which are different among more subjects.
+%The RMSE and the dice coefficient values of results are 0.009 and 0.991 respectively.
+%Each color represents a different structure in the brain.
 
@@
     \includegraphics[width=\columnwidth]{images/segmentation.png} 
-    \caption{Differences between FNIRT results from PreFreeSurfer 
+    \caption{Differences between FNIRT results of one subject from PreFreeSurfer 
     (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
@@
   \includegraphics[width=\columnwidth]{images/brain_classification.png} 
-  \caption{Binarized differences between brain segmentation results from 
-           FreeSurfer (CentOS6 vs. CentOS7).} 
+  \caption{Heatmap binarized differences between brain segmentation results (file \emph{aseg.hires.mgz}) from 
+          20 subjects in FreeSurfer (CentOS6 vs. CentOS7).} 
   \label{fig:tissue_class}
@@
 First, subjects are clustered into different groups using the clustering method as described before.
-We choosed the threshold value of zero for the clustering method. 
+We set the threshold value to zero for the clustering method. 
 This turns to have the subjects with the same topology in each cluster.
@@
 Furthermore, the percentage of processes that introduce differences 
-in PreFreeSurfer along with the 
-pipeline elements are depicted in Figure~\ref{fig:pfs_freq} 
+along with the 
+sub-pipelines are depicted in this figure 
 including linear registration with “\emph{FLIRT}” (in 
-ACPC-Alignment, BrainExtraction, DistortionCorrection, 
-AtlasRegistration), non-linear registration with “\emph{FNIRT}” (in 
-BrainExtraction and AltasRegistration), image warping with 
-“\emph{new\_invwarp}” (in BrainExtraction and AtlasRegistration). 
+ACPC-Alignment, Brain Extraction, Distortion Correction, 
+Atlas Registration), non-linear registration with “\emph{FNIRT}” (in 
+Brain Extraction and Altas Registration), image warping with 
+“\emph{new\_invwarp}” (in Brain Extraction and Atlas Registration). 
 Also, differences were observed in image mean and standard-deviation 
-computations in image interpolation with “\emph{fslmaths}” (in AnatomicalAverage). 
+computations in image interpolation with “\emph{fslmaths}” (in Anatomical Average). 
 It can be seen that \emph{fslmaths} and \emph{FLIRT} 
@@
   \includegraphics[width=\columnwidth]{images/pfs_heatmap.png}
-  \caption{Colorized processes along with the pipeline elements that introduce 
+  \caption{Heatmap of difference frequency of processes along with the sub-pipelines that introduce 
            differences among 20 subjects. The range between red and green colors show the most and 
-           least frequency of processes respectively. Besides, the number 
-           of occurrences of each processes with differences is represented in the line below the heatmap.}
+           least frequency of processes respectively. Besides, the total number 
+           of occurrences of each processes is represented in the bracket in each cell.}
   \label{fig:pfs_freq}
@@
 
-We found that inter-OS differences occur if pre-processed data are different in FreeSurfer. 
-Figure~\ref{fig:tissue_class} shows differences that coming from PreFreeSurfer analysis 
-as a result of propagation of differences. 
-NURM tool identified that there are no processs in the pipeline that create differences.
-This is because of using static builds of FreeSurfer which statically linked libraries (embeded libraries in the pipeline).
+We executed FreeSurfer pipeline on both OSes using a fixed input data, result of PreFreeSurfer on Centos 6, 
+to determine the creation of differences. 
+NURM tool identidied \emph{mris\_make\_surfaces} as the only process in the pipeline that create differences.
+This process generates surface files for cortical and white matter surfaces, curvature file for cortical thickness and 
+surface file estimate for layer IV of cortical sheet. 
+In this analysis, the number of process that create differences are dropped becasue of 
+using the static builds of FreeSurfer in which libaries are embeded in the pipeline package.
+
+However, we found that inter-OS differences occur when pre-processed data are different in FreeSurfer. 
+Using the result data from PreFreeSurfer, we executed FreeSurfer pipeline in each operating system separately.
+FreeSurfer pipeline is unstable as it propagates and amplifies differences 
+through the pipeline analysis.
+Figure~\ref{fig:tissue_class} shows inter-OS differences  
+as a result of propagation and amplification of differences. % that coming from PreFreeSurfer outputs.
 
@@
 
-\note{Pipelines amplify small numerical differences because they are numerically 
-unstable. Furthermore, math libraries evolve over time, leading to 
-different numerical errors. we listed some of the irreproducibility 
-causes of the pipelines as we mentioned in the previous section 
-overally along with exact command arguments.
+We indentified 4 processes that create differences 
+among all 28 distinct processes that are involved in the PreFreeSurfer execution. 
+The numerical instability in the 
+PreFreeSurfer HCP pipeline arises mainly from linear 
+registration processes implemented in FSL \emph{FLIRT}. 
+We can see that \emph{FLIRT} creates differences in almost all the sub-pipelines. 
+Therefore, it is important to more investigate about the registration procedure and find 
+a way to stabilize this process.
 
-Mention that this was only possible because the unprocessed data was 
-shared in the first place. DICOM to Nifti conversion was out of scope 
-and may introduce other issues.}
+In the second experiment, 
+the analysis results obtained by the pipeline show that inter-OS differences are visually substabtial.
+We see that statically builds of the pipeline improves reproducibility, but 
+it would not improve the numerical stability of the pipeline. 
+FreeSurfer pipeline propagates and amplifies the numerical differences that are created by the 
+differences in the underlying operating system libraries.
+
+In addition to variations of the operating system, there are other factors that may affect 
+the results such as software versions, compilers, hardwares, etc.
+Generally, we can use the Monte Carlo Arithmatic (MCA) to determine the numerical instability of the pipelines.
+This argorithm perturbs every floating-point operation 
+in the analysis based on the random rounding in each run. 
+By this algorithm, NURM tool can easily locates which parts of the analysed source code are to be responsible 
+for floating-point related instabilities.
+
+Although we could identify the processes in the pipeline that create differeces, the functions 
+that amplify these differences should be quantified in the future works.
+For this purpose, we can evaluate the final result after perturbing the identified processes in the pipeline iterratively.
+As a result, the processes that amplify differences would have a significant effect on the results independently.
 
@@
 
-\note{Our technique is able to characterize the stability of a pipeline's 
-components automatically. The numerical instability in the 
-PreFreeSurfer HCP pipeline arises mainly from linear and non-linear 
-registration processes implemented in FSL FLIRT and FNIRT.
+Our technique is able to characterize the stability of a pipeline
+automatically. We identified the main causes of the inter-OS differences as (1) the evolution 
+of the math libraries over the time and, (2) the instability of the pipelines. 
+There are two ways to tackle this problem. The easy but less preferred solution is masking the instabilities.
+This can be done by using, (1) single operating system for the processing of subjects, 
+(2) containerizing the pipelines so that the 
+processing is done on a more controlled environment, 
+(3) increasing the numerical precision of the arithmetics, 
+(4) building static executables by removing the host operating system library dependencies. 
+These solutions only makes the problem invisible but the instability still remains.
 
-There are a few ways to impede such instabilities:
-\begin{itemize}
-\item Use a single operating system
-\item Containerize pipelines
-\item Increase numerical precision
-\item Be stricter on truncation and rounding standards (IEEE 754)
-\item Build static executable
-\end{itemize}
-
-The results still suffer from small perturbations literally because of 
-the fact that pipeline are not numerically stable. The preferred 
-solution is to detect and fix numerical instability of the pipeline 
-instead of masking the problem. These processes need to be reviewed to 
-understand and correct the cause of instabilities. 
-}
+The preferred solution is to find the origin of amplification of the differences and 
+then fixing the particular functions/modules in the processing pipelines.
 
 \section{Acknowledgments}
+
+\note{
+CBRAIN team. Compute Canada(Calcul Quebec).
 
@@
 
-\note{CBRAIN team. Compute Canada(Calcul Quebec).}
+}
 

### 2019-09-25 paper.tex
Comments on Experiments and results

@@
 \usepackage{amsmath}
+\usepackage{ulem}
 \DeclareMathOperator*{\argmin}{argmin}
@@
 \newcommand{\tristan}[1]{\color{blue}\textbf{From Tristan:}#1\color{black}}
+\newcommand{\english}[1]{\uwave{#1}}
 
@@
 Indeed, the propagation and amplification of these tiny perturbations by analysis 
-may cause a significant reproducibility issue. We focus on the differences 
-created by different version of operating systems in computational pipelines. 
+may cause a significant reproducibility issue. \tristan{All the above is introduction and should be moved there.
+The introduction of this experiment section should explain:
+\begin{enumerate}
+  \item What is the goal of the experiment
+  \item What are the expected results 
+  \item Any required context that is specific to the experiment and not described before
+  \item The main components of the section: ``This section describes 
+  the datasets, pipelines, etc used in the experiments."
+\end{enumerate}
+}
+We focus on the differences 
+created by different versions of operating systems in computational pipelines. 
 In particular, NURM-tool is tested on the neuroimaging applications which are 
@@
 produced by the pipelines from the Human Connectome Project (\href{https://www.humanconnectome.org}{HCP}).
-The HCP initiative is a significant effort that will acquire and analyze 
+The HCP initiative is a significant effort that will acquire \tristan{check future tense} and analyze 
 brain connectivity data from 1200 healthy adults.
@@
 to their surface atlas.
+\tristan{The two previous paragraphs should focus on a technical description of the pipelines, 
+possibly with a workflow figure, 
+and refer to ref [12] for any image processing detail.}
 
@@
 
-HCP data consists of brain images collected from an individual healthy person, 
+HCP data consists of brain images collected from an individual healthy human subjects, 
 in the age range between 22-35 years~\cite{van2013wu}. Data from 1200 subjects 
@@
 type of MRI scanners.
+\tristan{remove details about age range and directory. Instead, explain the type of
+imaging that is available: T1 (multiple instances), DWI, functional.}
 
@@
 
-The data were processed in the order PreFreeSurfer and FreeSurfer. 
-Unprocessed subject data used as the input data for PreFreeSurfer and its output 
-results used as the input data for FreeSurfer analysis in each operating system separately.
-Through all the subjects, first, two runs were made on the same operating systems 
-to be sure that identified differences are not coming from the operations with randomness.
-Then we identified the processes in the pipelines that are responsible for 
-the creation of differences using the NURM-tool.
+The data were processed with PreFreeSurfer and FreeSurfer. 
+\english{Unprocessed subject data used as the input data for PreFreeSurfer and its output 
+results used as the input data for FreeSurfer analysis in each operating system separately}.
+Each subject was processed twice on the same operating system to detect
+variability coming from pseudo-random operations. 
+Then we used NURM-tool to identify the processes in the pipelines that
+created differences.
 
-It should be noted that some files may have the same data but 
-different metadata in headers, including creation time, modification time and date, 
-which turns to non-identical checksums. 
-So, we added some other functions for the file comparison to be sure that differences 
-are in part of the file data. 
+\tristan{This paragraph should be merged with the sentence on pseudo-randomness. 
+Prose should also be simplified and made more direct.}
+Within-OS variability may also be coming from other factors than pseudo-randomness. For instance, 
+metadata may differ due to the presence of timestamps, file paths and other 
+execution-specific information. \tristan{in the tool description, talk about the possibility to add distance functions 
+that are specific to file types and are computed in addition to checksums.}
+To address this issue, we created distance functions \english{for the file comparison to be sure that differences 
+are in part of the file data}. 
+\tristan{summarize the following sentences, keep it high level, and provide a link to the
+metrics file that you used.}
 Depending on the file type, we used functions from FreeSurfer software package to determine 
@@
 
-Two types of differences can occur in the subjects due to the 
-differences in the operating systems. One is inter-OS differences caused by the 
+Two types of differences can occur in the subjects due to the differences
+in the operating systems. One is inter-OS differences caused by the
 operating system library updates and the other type, intra-OS differences
-occur as a result of the pseudo-random processes used in the 
-pipelines.
+occur as a result of the pseudo-random processes used in the pipelines.
+\tristan{Move this up to intro and/or experiments, replace "inter-" by
+"between-" and "intra-" by "within-"}
+
+\subsubsection{Within-OS binary differences}
 
@@
 intra-OS differences.  
-For example, process \emph{mris\_curvature}
+For example \tristan{list the processes exhaustively}, process \emph{mris\_curvature}
 computes mean and Gaussian curvatures of a cortical surface using 
@@
 Therefore, NURM-tool can identify the processes in the pipelines 
-that create inter-OS differences.
+that create inter-OS differences. \tristan{sharpen a bit the writing}
 
-\subsubsection{Inter-OS binary differences}
+\subsubsection{Between-OS binary differences}
+
+\tristan{Before showing a particular case, you should describe 
+an overview of the differences:
+\begin{enumerate}
+  \item What is the frequency of binary differences and where do they happen in the pipelines? Do they happen for all subjects?
+  \item What is the magnitude of the differences? Explain the metrics, report them across subjects for different files.
+  \item Give visual examples.
+\end{enumerate}
 
@@
 pipelines~\cite{glasser2013} were 
-sensitive to operating system variation.
+sensitive to operating system variation \tristan{this is intro}.
 
 Figure~\ref{fig:fnirt_result} shows differences in the files 
-that are created as a part of ACPC-Alignment in PreFreeSurfer.
+that are created as a part of ACPC-Alignment in PreFreeSurfer. \tristan{add link to gif animation}
 The square pixels in the brain in the checkerboard images show the differences between results in two OSes. 
 The RMSE (root mean squared error) value of results is 0.029 and the dice coefficient is 0.125.
-In the next subsection, it is shown that such misalignment is created due to 
-non-linear registration in \emph{FNIRT} process in PreFreeSurfer pipeline.
 
 Furthermore, Figure 
-\ref{fig:tissue_class} illustrates the heatmap binarized differences 
-of the brain tissue segmentation across OSes in FreeSurfer pipeline.
+\ref{fig:tissue_class} illustrates the sum of binarized differences 
+of the brain tissue segmentation across OSes in FreeSurfer pipeline, for the 20 subjects \tristan{add a color 
+scale to the figure}.
 These differences are accumulated from 20 subject results.

### 2019-10-02 paper.tex
Edited the text of the experiment section

--- 2019-09-19 paper.tex
@@
    \end{subfigure}
-   \caption{An overview of NURM tool in two steps. First capturing intermediary files (a); then
+   \caption{An overview of NURM-tool in two steps. First capturing intermediary files (a); then
             labelling process graphs (b).}
@@
 (Figure~\ref{fig:overview-labelling} - (3)).
+
 Differences are identified by comparing checksum of the output files for 
 the same process in both conditions (Figure~\ref{fig:overview-labelling} - (2)).
+In addition, the distance functions that are specific to file types can be added 
+to our tool to ensure that differences belong to the file data part.
 We developed an incremental approach that consists of the following steps:
@@
 
-The variety of computing infrastructures, particularly the operating system, is 
-known to influence the reproducibility of computational pipelines because of the
-creation of small numerical differences. 
-These differences mainly correspond to the mathematical functions implemented 
-in different operating system libraries.
-For instance, changing the mathematical functions like \emph{expf()} and 
-\emph{cosf()} which manipulate the precision of floating-point representations, 
-between \emph{glibc} libraries in different operating systems can produce 
-small numerical differences.
-Besides, a similar issue is expected for any operating system which is 
-based on \emph{glibc}, the GNU C library.
+In this experiment, we analyse the numerical reproducibility of the computational pipelines 
+and identify the origin of differences across the operating systems. 
+Two types of differences can occur in the subjects due to the differences
+in the operating systems. One is between-OS differences caused by the
+operating system library updates and the other type, within-OS differences
+occur as a result of the pseudo-random processes used in the pipelines.
 
-Indeed, the propagation and amplification of these tiny perturbations by analysis 
-may cause a significant reproducibility issue. We focus on the differences 
-created by different version of operating systems in computational pipelines. 
 In particular, NURM-tool is tested on the neuroimaging applications which are 
-predominantly using mathematical libraries.
+predominantly using mathematical libraries. Therefore, we expect to find 
+differences as a result of changing the mathematical functions between operating system libraries.
+
+This section describes datasets and pipelines used for the analyses, 
+how data processed through the pipelines, and the results.
 
@@
 
-We used our tool to evaluate the effect of the operating system on results 
-produced by the pipelines from the Human Connectome Project (\href{https://www.humanconnectome.org}{HCP}).
-The HCP initiative is a significant effort that will acquire and analyze 
+We used our tool to evaluate the reproducibility of the pipelines from the Human Connectome 
+Project (\href{https://www.humanconnectome.org}{HCP}).
+The HCP initiative is a significant effort to acquire and analyze 
 brain connectivity data from 1200 healthy adults.
@@
 sharing primary and processed data freely, and (2) data analysis pipelines that 
-are available under an open source license.
+are available under an open-source license.
 
 The HCP developed a set of pre-processing pipelines to process structural,
-functional and diffusion MRI data acquired in the project. We focus on HCP
+functional, and diffusion MRI data acquired in the project. We focus on HCP
 pre-processing pipelines for structural data, particularly PreFreeSurfer
@@
 
-According to~\cite{glasser2013}, main sub-pipelines of the HCP PreFreeSurfer are:
-(1) Distortion Correction (DC) 
-to correct the geometric distortions for routine MRI scans 
-and produce an undistorted native structural volume space for each subject,
-(2) Anatomical Average (AAve) to average same modality images (if more than one is available),
-(3) ACPC (Anterior Commissure, Posterior Commissure) Alignment (ACPC-A) 
-is aligning images to the MNI space to have the same orientation as 
-the reference template,
-(4) Brain Extraction (BExt) includes a robust initial brain extraction, using FSL's linear (FLIRT) 
-and non-linear (FNIRT) registration, 
-(5) Bias Field Correction (BFC) applies intensity inhomogenity correction on the registered images, 
-%Calculates bias field using square root of the product of T1w and T2w images. 
-(6) Atlas-Registration (AR) to register the subject's native structural volume space to MNI space as the last step.
+According to~\cite{glasser2013}, the sequence of the 
+sub-pipelines in PreFreeSurfer includes 
+(1) Distortion Correction (DC), 
+(2) Anatomical Average (AAve), 
+(3) ACPC (Anterior Commissure, Posterior Commissure) Alignment (ACPC-A), 
+(4) Brain Extraction (BExt), 
+(5) Bias Field Correction (BFC), 
+(6) Atlas-Registration (AR).
 
-Also, the main goals of HCP FreeSurfer pipeline are:
-(1) improve the robustness of brain extraction,
-(2) fine tune image registration,
-(3) accurately place the surface images with high resolution data,
-(4) perform FreeSurfer's standard folding-based surface registration 
-to their surface atlas.
-
+Also, the sequence of the main sub-pipelines in FreeSurfer includes 
+(1) Downsampling images, 
+(2) T1w image registration, 
+(3) Placement of surface images, 
+(4) Folding-based surface registration.
 A detailed description of the analyses done in these
@@
 
-HCP data consists of brain images collected from an individual healthy person, 
-in the age range between 22-35 years~\cite{van2013wu}. Data from 1200 subjects 
+Furthermore, the average time per subject taken for PreFreeSurfer processing is 2.2 hours, 
+and FreeSurfer processing is 8.3 hours (20 runs), 
+and the average output file size of PreFreeSurfer processing is 2.7 GB, 
+and FreeSurfer processing is 4.1 GB (20 subjects).
+
+HCP data collected from different types of imaging techniques, including 
+structural imaging (sMRI), functional imaging (fMRI), and diffusion imaging (dMRI).
+The structural images include T1-weighted (T1w) and T2-weighted (T2w) images which 
+help in the diagnosis of the brain injury.
+The functional images, including task-based fMRI and resting-state scans, 
+enable the measurement of functional activations within brain areas. 
+Diffusion imaging is another kind of MRI technique, which measures 
+the anatomical connectivity between regions.
+See~\cite{van2013wu} for a discussion of these various terms.
+
+Data from 1200 subjects 
 is available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
-Such a directory containing the HCP data is denoted by the term ``subject".
 In this paper, we randomly selected 20 unprocessed subjects from 
@@
 
-The data were processed in the order PreFreeSurfer and FreeSurfer. 
-Unprocessed subject data used as the input data for PreFreeSurfer and its output 
-results used as the input data for FreeSurfer analysis in each operating system separately.
-Through all the subjects, first, two runs were made on the same operating systems 
-to be sure that identified differences are not coming from the operations with randomness.
-Then we identified the processes in the pipelines that are responsible for 
-the creation of differences using the NURM-tool.
+The data were processed with PreFreeSurfer and FreeSurfer. 
+First, PreFreeSurfer was utilized to analyse the unprocessed data, and then its 
+results were used as the input of FreeSurfer.
 
-It should be noted that some files may have the same data but 
-different metadata in headers, including creation time, modification time and date, 
-which turns to non-identical checksums. 
-So, we added some other functions for the file comparison to be sure that differences 
-are in part of the file data. 
-Depending on the file type, we used functions from FreeSurfer software package to determine 
-whether two files differ, including  
-\emph{mri\_diff}, \emph{mris\_diff}, and \emph{lta\_diff} for comparing volumes, surface-based files, 
+Each subject was processed twice on the same operating system to detect
+variability coming from pseudo-random operations. 
+Within-OS variability may also be coming from other factors than pseudo-randomness. 
+For instance, metadata may differ due to the presence of timestamps, 
+file paths, and other execution-specific information. 
+To ensure that differences belong to the file data part, 
+we added distance functions specific to the file types.
+These functions are \emph{mri\_diff}, \emph{mris\_diff}, and \emph{lta\_diff} 
+for comparing volumes, surface-based files, 
 and lta files (transforms) respectively. 
-These functions check data based on the volume dimentions, 
-resolutions, pixel intensities,  
-the number of surface vertices and neighbors, and geometry information. 
-It then returns files differ if there be differences in any of these parameters.
-Also, text files are compared using a custom function that compares the content line by line.
+These functions compare files based on the image dimensions, 
+pixel intensities, the number of vertices, and geometry information
+(see~\cite{fischl2012freesurfer} for more details).
+
+Furthermore, we clustered subjects into different type using 
+the clustering method as described before.
+We set the threshold value to zero for the clustering method. 
+This turns to cluster the subjects with the same topology into the same type.
+Also, we used the nearest neighbor algorithm to calculate the distance between clusters.
 
@@
 
-Two types of differences can occur in the subjects due to the 
-differences in the operating systems. One is inter-OS differences caused by the 
-operating system library updates and the other type, intra-OS differences 
-occur as a result of the pseudo-random processes used in the 
-pipelines.
+\subsubsection{Within-OS differences}
 
-In our experiments, we identified the processes that create differences due to 
-intra-OS differences.  
-For example, process \emph{mris\_curvature}
+In our experiments, we identified two processes that create differences due to 
+within-OS differences.  
+Registration process with \emph{mri\_segreg} and process \emph{mris\_curvature} that 
 computes mean and Gaussian curvatures of a cortical surface using 
 the random variable in FreeSurfer.
-To avoid such randomness we 
+To avoid such pseudo-randomness, we 
 initialized random number generators using a seed state in 
-the command line in which pipeline script calls the program.
-This ensures that results are the same from run to run.
-Therefore, NURM-tool can identify the processes in the pipelines 
-that create inter-OS differences.
+the command line in which pipeline script calls the process.
+In this way, NURM-tool identifies the processes in the pipelines 
+that create between-OS differences.
 
-\subsubsection{Inter-OS binary differences}
+\subsubsection{Between-OS differences in PreFreeSurfer}
+%\subsubsection{PreFreeSurfer pipeline analysis}
 
-In a previous study~\cite{Scaria2017}, we showed that the HCP pre-processing 
-pipelines~\cite{glasser2013} were 
-sensitive to operating system variation.
-
-Figure~\ref{fig:fnirt_result} shows differences in the files 
-that are created as a part of ACPC-Alignment in PreFreeSurfer.
-The square pixels in the brain in the checkerboard images show the differences between results in two OSes. 
-The RMSE (root mean squared error) value of results is 0.029 and the dice coefficient is 0.125.
-In the next subsection, it is shown that such misalignment is created due to 
-non-linear registration in \emph{FNIRT} process in PreFreeSurfer pipeline.
-
-Furthermore, Figure 
-\ref{fig:tissue_class} illustrates the heatmap binarized differences 
-of the brain tissue segmentation across OSes in FreeSurfer pipeline.
-These differences are accumulated from 20 subject results.
-So, the magnitude of differences is in the range of 1-20 as the total number of subjects.
-The darker blue pixels show differences among less subjests and 
-the brighter blue pixels are related to the differences among more subjects.
-
-%The brighter spots show relevant pixels in the segmented image which are different among more subjects.
-%The RMSE and the dice coefficient values of results are 0.009 and 0.991 respectively.
-%Each color represents a different structure in the brain.
-
-\begin{figure}
-  \centering
-    \includegraphics[width=\columnwidth]{images/segmentation.png} 
-    \caption{Differences between FNIRT results of one subject from PreFreeSurfer 
-    (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
-    CentOS7). Right images show the zoomed version part of the images marked 
-    by the red rectangle.} 
-    \label{fig:fnirt_result}
-\end{figure}
-
-\begin{figure}
-%  \includegraphics{brain\_classification}
-\centering
-  \includegraphics[width=\columnwidth]{images/brain_classification.png} 
-  \caption{Heatmap binarized differences between brain segmentation results (file \emph{aseg.hires.mgz}) from 
-          20 subjects in FreeSurfer (CentOS6 vs. CentOS7).} 
-  \label{fig:tissue_class}
-\end{figure}
-
-\subsubsection{PreFreeSurfer pipeline analysis}
-
-First, subjects are clustered into different groups using the clustering method as described before.
-We set the threshold value to zero for the clustering method. 
-This turns to have the subjects with the same topology in each cluster.
-Also, we used the nearest neighbor algorithm to calculate the distance between clusters.
 Table~\ref{table:data-clusters} shows the PreFreeSurfer result of subject clustering. 
-Among our dataset, we identified 4 types of subjects with 
-different topology of \emph{process trees}. 
-Differences correspond to the data processing that leads to producing disparate output file(s).
+Among our dataset, we identified four types of subjects with 
+the different topology of \emph{process trees}. 
+Differences correspond to the data processing that leads to producing a disparate output file(s).
 For example, subject type 3 missed output file \emph{T1w2\_gdc.nii.gz} after processing 
@@
 
-Then, we determined the frequency of the processes that create differences in 
-the pipeline in Figure~\ref{fig:pfs_freq}. \emph{FLIRT} is the most frequent 
-process that produces differences.
+The frequency of the processes that create differences in 
+the pipeline among 20 subjects is determined in Figure~\ref{fig:pfs_freq}. 
+\emph{FLIRT} is the most frequent process that produces differences.
 Furthermore, the percentage of processes that introduce differences 
@@
 “\emph{new\_invwarp}” (in Brain Extraction and Atlas Registration). 
-Also, differences were observed in image mean and standard-deviation 
+Besides, differences were observed in image mean and standard-deviation 
 computations in image interpolation with “\emph{fslmaths}” (in Anatomical Average). 
-It can be seen that \emph{fslmaths} and \emph{FLIRT} 
+Results show that \emph{fslmaths} and \emph{FLIRT} 
 are the processes that at least have one subprocess to produce differences across all the subjects.
@@
 
-\subsubsection{FreeSurfer pipeline analysis}
+We created checkerboard images from the image results of one subject in two OSes.
+Alternating rectangular regions between two images, creating a characteristic 
+checkerboard-like pattern of varying magnitudes.
+Figure~\ref{fig:fnirt_result} shows differences in the files 
+that are created as a part of ACPC-Alignment in PreFreeSurfer.
+Each rectangle includes the pixels with different intensity values between images.
 
-We executed FreeSurfer pipeline on both OSes using a fixed input data, result of PreFreeSurfer on Centos 6, 
-to determine the creation of differences. 
-NURM tool identidied \emph{mris\_make\_surfaces} as the only process in the pipeline that create differences.
-This process generates surface files for cortical and white matter surfaces, curvature file for cortical thickness and 
-surface file estimate for layer IV of cortical sheet. 
-In this analysis, the number of process that create differences are dropped becasue of 
-using the static builds of FreeSurfer in which libaries are embeded in the pipeline package.
+\begin{figure}
+  \centering
+    \includegraphics[width=\columnwidth]{images/segmentation.png} 
+    \caption{Differences between FNIRT results of one subject from PreFreeSurfer 
+    (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
+    CentOS7). Right images show the zoomed version part of the images marked 
+    by the red rectangle.
+    Here is 
+    \href{https://raw.githubusercontent.com/ali4006/HCP-reproducibility-paper/master/images/brain_classification.gif}{the link}
+    to the gif animation of the same images.
+} 
+    \label{fig:fnirt_result}
+\end{figure}
 
-However, we found that inter-OS differences occur when pre-processed data are different in FreeSurfer. 
-Using the result data from PreFreeSurfer, we executed FreeSurfer pipeline in each operating system separately.
-FreeSurfer pipeline is unstable as it propagates and amplifies differences 
-through the pipeline analysis.
-Figure~\ref{fig:tissue_class} shows inter-OS differences  
-as a result of propagation and amplification of differences. % that coming from PreFreeSurfer outputs.
+\subsubsection{Between-OS differences in FreeSurfer}
+
+In the first test, the results of PreFreeSurfer on CentOS 6
+feed to the FreeSurfer pipeline on both OSes.
+We fixed input data to determine whether the pipeline creates differences.
+We processed 20 subjects using FreeSurfer.
+NURM-tool identified \emph{mris\_make\_surfaces} as the only process in 
+the pipeline that creates differences in 10 subjects.
+This process generates surface files for cortical and white matter surfaces (see~\cite{} for more details).
+Also, FreeSurfer did not introduce any differences in 10 subjects.
+
+Results show that the static builds of FreeSurfer in which uses the embedded libraries, 
+significantly avoids the creation of differences.
+However, we found that between-OS differences occur when pre-processed data are different in FreeSurfer.
+
+In the second test, the results of PreFreeSurfer feed to the FreeSurfer 
+pipeline on each operating system separately.
+Figure~\ref{fig:tissue_class} shows between-OS differences  
+as a result of propagation and amplification of differences in FreeSurfer. % that coming from PreFreeSurfer outputs. 
+This figure illustrates the sum of binarized differences 
+from 20 subject results of the brain tissue segmentation across OSes.
+Therefore, the magnitude of differences is in the range of 1-20 as the total number of subjects.
+
+\begin{figure}
+%  \includegraphics{brain\_classification}
+\centering
+  \includegraphics[width=\columnwidth]{images/brain_classification.png} 
+  \caption{Heatmap binarized differences between brain segmentation results (file \emph{aseg.hires.mgz}) from 
+          20 subjects in FreeSurfer (CentOS6 vs. CentOS7).} 
+  \label{fig:tissue_class}
+\end{figure}
 
@@
 
-We indentified 4 processes that create differences 
+We identified four processes that create differences 
 among all 28 distinct processes that are involved in the PreFreeSurfer execution. 
@@
 We can see that \emph{FLIRT} creates differences in almost all the sub-pipelines. 
-Therefore, it is important to more investigate about the registration procedure and find 
+Therefore, it is important to investigate the registration procedure and find 
 a way to stabilize this process.
@@
 In the second experiment, 
-the analysis results obtained by the pipeline show that inter-OS differences are visually substabtial.
+the analysis results obtained by the pipeline show that between-OS differences are visually substantial.
 We see that statically builds of the pipeline improves reproducibility, but 
@@
 
-In addition to variations of the operating system, there are other factors that may affect 
-the results such as software versions, compilers, hardwares, etc.
-Generally, we can use the Monte Carlo Arithmatic (MCA) to determine the numerical instability of the pipelines.
-This argorithm perturbs every floating-point operation 
+In addition to variations of the operating system, other factors exist which may influence 
+the reproducibility of the analyses such as software versions, compilers, hardware, etc.
+Generally, we can use the Monte Carlo Arithmetic (MCA) to determine the numerical stability of the pipelines.
+This algorithm perturbs every floating-point operation 
 in the analysis based on the random rounding in each run. 
-By this algorithm, NURM tool can easily locates which parts of the analysed source code are to be responsible 
-for floating-point related instabilities.
+By this algorithm, NURM-tool can easily locate which parts of the analyzed source code are responsible 
+for the floating-point related instabilities.
 
-Although we could identify the processes in the pipeline that create differeces, the functions 
-that amplify these differences should be quantified in the future works.
-For this purpose, we can evaluate the final result after perturbing the identified processes in the pipeline iterratively.
-As a result, the processes that amplify differences would have a significant effect on the results independently.
+Although we identify the processes in the pipeline that create differences, 
+the amplification of these differences should be quantified in the future works.
+For this purpose, we can evaluate the final result after perturbing the identified processes in the pipeline iteratively.
+As a result, the processes that amplify differences would have a significant effect on the results.
 
@@
 
-Our technique is able to characterize the stability of a pipeline
-automatically. We identified the main causes of the inter-OS differences as (1) the evolution 
-of the math libraries over the time and, (2) the instability of the pipelines. 
+Our technique can characterize the stability of the pipelines
+automatically. We identified the main causes of the between-OS differences as (1) the evolution 
+of the math libraries over time and, (2) the instability of the pipelines. 
 There are two ways to tackle this problem. The easy but less preferred solution is masking the instabilities.
@@
 (4) building static executables by removing the host operating system library dependencies. 
-These solutions only makes the problem invisible but the instability still remains.
+These solutions only make the problem invisible, but the pipelines are stiil unstable.
 

### 2019-10-03 paper.tex
Updates

@@
 (Figure~\ref{fig:overview-labelling} - (3)). 
-
 Differences are identified by comparing checksum of the output files for 
@@
 
-\subsection{Processing of data}
+\subsection{Data processing}
 
@@
 
-In our experiments, we identified two processes that create differences due to 
-within-OS differences.  
-Registration process with \emph{mri\_segreg} and process \emph{mris\_curvature} that 
-computes mean and Gaussian curvatures of a cortical surface using 
-the random variable in FreeSurfer.
-To avoid such pseudo-randomness, we 
-initialized random number generators using a seed state in 
-the command line in which pipeline script calls the process.
-In this way, NURM-tool identifies the processes in the pipelines 
-that create between-OS differences.
+We didn't observe any within-OS difference in PreFreeSurfer, however, we
+ identified two FreeSurfer processes leading to within-OS differences due
+ to the use of pseudo-random numbers: image registration with
+ \emph{mri\_segreg}, and cortical surface curvature
+ estimations with \emph{mris\_curvature}. To address this issue, we fixed
+ the seed used by the pseudo-random number generator, through an option of
+ the FreeSurfer tools. In this way, our analysis can focus on between-OS
+ differences.
 
@@
 
-Table~\ref{table:data-clusters} shows the PreFreeSurfer result of subject clustering. 
-Among our dataset, we identified four types of subjects with 
-the different topology of \emph{process trees}. 
-Differences correspond to the data processing that leads to producing a disparate output file(s).
-For example, subject type 3 missed output file \emph{T1w2\_gdc.nii.gz} after processing 
-in comparison with subject type 1.
-Besides, we verified that 
-the \emph{process trees} were identical for all subjects of the same type in 
-different versions of CentOS.
+We identified four types of subjects with different PreFreeSurfer process
+trees (Table~\ref{table:data-clusters}). Differences between subject types
+are coming from different  numbers of T1 and T2 images in the raw data:
+subjects of type 1 have 2 T1 images and 2 T2 images, while subjects of type
+4 only have 1 T1 image and 1 T2 image. \tristan{check that} This leads to
+different sets of output files. We verified that the process trees were
+identical for all subjects of the same type in different versions of
+CentOS.
 
 \begin{table}
+\label{table:data-clusters}
 \centering
@@
 \end{threeparttable}
-\label{table:data-clusters}
 \end{table}
@@
 The frequency of the processes that create differences in 
-the pipeline among 20 subjects is determined in Figure~\ref{fig:pfs_freq}. 
+the pipeline is shown in Figure~\ref{fig:pfs_freq}. 
 \emph{FLIRT} is the most frequent process that produces differences.

### 2019-10-09 paper.tex

@@
 Although new techniques such as virtualization, version control 
-systems, and provenance management tools provide more reliable 
+systems, and provenance management tools provide a more reliable 
 environment and significantly improve the reproducibility of scientific 
@@
 present a technique to identify the processes in the pipeline that 
-create numerical differences along the execution, and we apply this 
+create numerical differences along with the execution, and we apply this 
 technique to the HCP structural pre-processing pipelines.
@@
 
-A reproducible study provides a context in which one is able to get 
-results that are consistent with the original 
-work~\cite{plesser2018reproducibility}. Research findings are expected 
-to be reproducible so that their authenticity and reliability can be 
-evaluated.
-
+Reproducibility is a crucial element of the scientific works, 
+as it enables researchers to evaluate authenticity and reliability 
+of the findings~\cite{plesser2018reproducibility}.
 Reproducibility is defined as the ability to regenerate the same 
 results as the original findings when the experiment is reanalyzed by 
-exactly the same analytic methods, software package, parameters, and 
+the same analytic methods, software package, parameters, and 
 data~\cite{peng2011reproducible}. 
-
 In addition, numerical reproducibility is defined as the ability to 
 regenerate bit for bit identical results from multiple 
-runs~\cite{hill2017numerical}. This Numerical reproducibility can be 
-checked by comparing the binary content of the results which is 
-computed by the checksum method.
-Regarding these definitions, it must be pointed that a
-computation might be reproducible, but not 
-be numerically reproducible. For instance, small numerical differences 
-created during the pipeline execution may hamper numerical 
-reproducibility, but be negligible in the final results.
+runs~\cite{hill2017numerical}. 
+We considered the numerical reproducibility in our experiments by comparing 
+the binary content of the results using the checksum method.
 
-\note{Recently, scientists began to realize that the results of many 
-scientific experiments were not reproducible.
-
-In particular, validation of the reproducibility has been widely investigated 
+Recently, validation of the reproducibility has been widely investigated 
 in the field of neuroimaging in which uses optimized 
-processing methods with the purpose of functional and/or structural 
+processing methods with the purpose of functional and structural 
 assessments of the human brain.
-
 According to the previous 
-studies, variety of computational infrastructures including workstation 
+studies, the variety of computing infrastructures including workstation 
 types, parallelization methods, operating systems, and analysis 
-packages are known to influence reproducibility~\cite{Gronenschild2012, 
+packages are known to influence the reproducibility of the analyses~\cite{Gronenschild2012, 
 diethelm2012limits, Glatard2015, bowring2019exploring}.
 
-In particular, the effect of 
-the operating system across some of the main 
-neuroimaging pipelines including FSL, CIVET and FreeSurfer pipelines 
-is studied in~\cite{Glatard2015, Gronenschild2012} that indicates a significant disparity. 
-}
+In particular, the studies on the effect of the operating system on the results show that 
+computational pipelines create small numerical differences~\cite{Glatard2015, Scaria2017}.
+These differences mainly correspond to the mathematical functions implemented 
+in different operating system libraries.
+For instance, changing the mathematical functions like \emph{expf()} and 
+\emph{cosf()} which manipulate the precision of floating-point representations, 
+between \emph{glibc} libraries in different operating systems can produce 
+small numerical differences.
+Besides, a similar issue is expected for any operating system which is 
+based on \emph{glibc}, the GNU C library.
 
@@
 pipeline and magnitude of the differences. In many cases, numerical 
-instability is an important issue for reproducibility.
+instability is an essential issue for reproducibility.
 
-There are different solutions to make reproducible analysis including 
-containerization techniques that encapsulate software/hardware 
+There are different solutions to improve the reproducibility of the analysis, 
+including containerization techniques that encapsulate software/hardware 
 dependencies, provenance capturing tools, and version control systems. 
 However, a comprehensive solution requires to fix the numerical 
-instabilities. For this purpose, we introduce a Numerical 
-Reproducibility Measurement (NURM) tool to identify the processes in 
-the pipeline that create numerical differences across several runs. In this 
-paper, the execution results of pipelines across different operating 
-systems are considered to find out the cause of the irreproducibility 
-using interposition techniques.
+instabilities instead of masking the issue. For this purpose, we introduce 
+the Numerical Reproducibility Measurement (NURM) tool to identify the processes in 
+the pipeline that create numerical differences across several runs. 
+Furthermore, we can fix the instability of the identified processes 
+in the next steps.
 
@@
 
-\tristan{define subjects, explain neuroimaging, bring context.}
 \section{Tool description}
@@
 Docker, and its interface is described with
-Boutiques~\cite{glatard2017boutiques}. \reprozip~\cite{Chirigati2016} is
+Boutiques~\cite{glatard2017boutiques}. \reprozip~\cite{chirigati2016reprozip} is
 added to Condition 1 for provenance tracking. Our tool starts by
@@
 The tool replaces output file with the same file obtained in Condition 1 
-if process creates differences (Figure~~\ref{fig:overview-labelling} - (3)).
+if process creates differences (Figure~\ref{fig:overview-labelling} - (3)).
 Process labelling
@@
 
-Different subjects may lead to different \emph{process tree}, due to
+Different subjects may lead to different process tree, due to
 variations that may be coming from differences in data types or
@@
 
-Different \emph{process tree} shows that pipeline execution followed by a different procedure. 
+Different process tree shows that pipeline execution followed by a different procedure. 
 This implies the possible impact of different subjects on pipeline results.
 We cluster subjects into different types to specify how many different 
-\emph{process trees} are generated. Clustering is performed based on the \emph{process trees} 
+process trees are generated. Clustering is performed based on the process trees 
 which nodes are labeled by the process name. 
@@
 Furthermore, we will show that even subjects in the same cluster, 
-having the same topological \emph{process tree} structure, 
+having the same topological process tree structure, 
 may have different labels as processes that create differences.
@@
 subjects of type 1 have 2 T1 images and 2 T2 images, while subjects of type
-4 only have 1 T1 image and 1 T2 image. \tristan{check that} This leads to
+2 only have 1 T1 image and 1 T2 image. This leads to
 different sets of output files. We verified that the process trees were
@@
 \begin{table}
-\label{table:data-clusters}
 \centering
@@
 \caption{Clustering of 20 input subjects.}
+\label{table:data-clusters}
 
@@
 type1  & 9                   & None                                               \\
-type2  & 7                   & [``T1w/T1w2\_gdc.nii.gz",                          \\
-       &                     &  ``T2w/T2w2\_gdc.nii.gz"]                          \\
-type3  & 2                   & [``T1w/T1w2\_gdc.nii.gz"]                          \\
-type4  & 2                   & [``T2w/T2w2\_gdc.nii.gz"]                          \\ \bottomrule
+type2  & 7                   & [``T1w", ``T2w"]                                   \\
+type3  & 2                   & [``T1w"]                                           \\
+type4  & 2                   & [``T2w"]                                           \\ \bottomrule
 \end{tabular}
@@
      \small
-     \item *Type refers to the topology type of the \emph{process tree} of subjects.
+     \item *Type refers to the topology type of the process tree of subjects.
 \end{tablenotes}
@@
 
-The frequency of the processes that create differences in 
-the pipeline is shown in Figure~\ref{fig:pfs_freq}. 
-\emph{FLIRT} is the most frequent process that produces differences.
-Furthermore, the percentage of processes that introduce differences 
-along with the 
-sub-pipelines are depicted in this figure 
-including linear registration with “\emph{FLIRT}” (in 
-ACPC-Alignment, Brain Extraction, Distortion Correction, 
-Atlas Registration), non-linear registration with “\emph{FNIRT}” (in 
-Brain Extraction and Altas Registration), image warping with 
-“\emph{new\_invwarp}” (in Brain Extraction and Atlas Registration). 
+Figure~\ref{fig:pfs_freq} shows the frequency of the processes that create differences 
+in the pipeline (the number in the bracket).
+Furthermore, this figure represents the percentage of the processes that introduce 
+differences along with the sub-pipelines, including linear registration 
+with “\emph{FLIRT}” (in ACPC-Alignment, Brain Extraction, Distortion Correction, 
+Atlas Registration), non-linear registration with “\emph{FNIRT}” (in Brain Extraction 
+and Altas Registration), image warping with “\emph{new\_invwarp}” (in Brain Extraction 
+and Atlas Registration). 
 Besides, differences were observed in image mean and standard-deviation 
 computations in image interpolation with “\emph{fslmaths}” (in Anatomical Average).
-Results show that \emph{fslmaths} and \emph{FLIRT} 
-are the processes that at least have one subprocess to produce differences across all the subjects.
 
-% \begin{figure}
-%   \centering
-%     \includegraphics[width=\columnwidth]{images/pfs_barchart.png} \caption{Number 
-%     of occurrences of processes that create differences in PreFreeSurfer 
-%     among 20 subjects.} 
-%     \label{fig:pfs_chart}
-%   \end{figure}
+Results show that \emph{FLIRT} is the most frequent process that produces 
+differences in ACPC-Alignment across all the subjects, Distortion Correction 
+in 95\% of the subjects, and Atlas Registration in 85\% of the subjects. 
+In addition, there is a \emph{fslmaths} process in Anatomical Average that 
+creates differences in all 11 subjects in which it has appeared.
 
@@
 
-We created checkerboard images from the image results of one subject in two OSes.
-Alternating rectangular regions between two images, creating a characteristic 
-checkerboard-like pattern of varying magnitudes.
-Figure~\ref{fig:fnirt_result} shows differences in the files 
-that are created as a part of ACPC-Alignment in PreFreeSurfer.
-Each rectangle includes the pixels with different intensity values between images.
+To visually represent the magnitude of variations, we created checkerboard 
+images from the image results of one subject in two OSes. Figure~\ref{fig:fnirt_result} 
+shows the checkerboard images from the results of ACPC-Alignment in PreFreeSurfer. 
+Each rectangle includes the replaced pixels with different intensity values between images.
+
+%Alternating rectangular regions between two images, creating a characteristic 
+%checkerboard-like pattern of varying magnitudes.
 
@@
 NURM-tool identified \emph{mris\_make\_surfaces} as the only process in 
-the pipeline that creates differences in 10 subjects.
-This process generates surface files for cortical and white matter surfaces (see~\cite{} for more details).
+the pipeline that creates differences among 10 subjects.
+This process generates surface files for cortical and white matter surfaces 
+(see~\cite{fischl2012freesurfer} for more detail of the process). 
 Also, FreeSurfer did not introduce any differences in 10 subjects.
@@
 However, we found that between-OS differences occur when pre-processed data are different in FreeSurfer. 
-
-In the second test, the results of PreFreeSurfer feed to the FreeSurfer 
+For this purpose, in the second test, the results of PreFreeSurfer feed to the FreeSurfer 
 pipeline on each operating system separately.
@@
 from 20 subject results of the brain tissue segmentation across OSes.
-Therefore, the magnitude of differences is in the range of 1-20 as the total number of subjects.
+The magnitude of differences is in the range of 1-20 as the total number of subjects.
 
@@
 the amplification of these differences should be quantified in the future works.
-For this purpose, we can evaluate the final result after perturbing the identified processes in the pipeline iteratively.
+For this purpose, we can evaluate the final result after perturbing the identified 
+processes in the pipeline iteratively.
 As a result, the processes that amplify differences would have a significant effect on the results.

### 2019-10-09 paper.tex
A few fixes in results

@@
 different sets of output files. We verified that the process trees were
-identical for all subjects of the same type in different versions of
+identical for all subjects of the same type, for both versions of
 CentOS.
@@
 
-Figure~\ref{fig:pfs_freq} shows the frequency of the processes that create differences 
-in the pipeline (the number in the bracket).
-Furthermore, this figure represents the percentage of the processes that introduce 
-differences along with the sub-pipelines, including linear registration 
-with “\emph{FLIRT}” (in ACPC-Alignment, Brain Extraction, Distortion Correction, 
-Atlas Registration), non-linear registration with “\emph{FNIRT}” (in Brain Extraction 
-and Altas Registration), image warping with “\emph{new\_invwarp}” (in Brain Extraction 
+Figure~\ref{fig:pfs_freq} shows the frequency at which  processes create differences 
+in the pipeline, and how differences are distribute among sub-pipelines.
+Differences were observed in linear registration 
+with FSL FLIRT (in ACPC-Alignment, Brain Extraction, Distortion Correction, 
+Atlas Registration), in non-linear registration with FSL FNIRT (in Brain Extraction 
+and Altas Registration), and in image warping with FSL \emph{new\_invwarp} (in Brain Extraction 
 and Atlas Registration). 
-Besides, differences were observed in image mean and standard-deviation 
-computations in image interpolation with “\emph{fslmaths}” (in Anatomical Average).
+Differences were also observed in image mean 
+computations with FSL maths  (in Anatomical Average).
 

### 2019-10-16 paper.tex

@@
 \toprule
-Type   & Number of Subjects  & Missing File(s)                                    \\ \midrule
-type1  & 9                   & None                                               \\
-type2  & 7                   & [``T1w", ``T2w"]                                   \\
-type3  & 2                   & [``T1w"]                                           \\
-type4  & 2                   & [``T2w"]                                           \\ \bottomrule
+Type   & Number of Subjects  & Available Modalities                                         \\ \midrule
+type1  & 9                   & [``T1w\_MPR1", ``T1w\_MPR2",                     \\
+&  &                            ``T2w\_SPC1", ``T2w\_SPC2"]     \\
+type2  & 8                   & [``T1w\_MPR1", ``T2w\_SPC1"]                                 \\
+type3  & 1                   & [``T1w\_MPR1", ``T2w\_SPC1",   \\
+&  &                            ``T2w\_SPC2"]                   \\
+type4  & 2                   & [``T1w\_MPR1", ``T1w\_MPR2",   \\
+&  &                            ``T2w\_SPC1"]                   \\ \bottomrule
 \end{tabular}
@@
   \includegraphics[width=\columnwidth]{images/pfs_heatmap.png}
-  \caption{Heatmap of difference frequency of processes along with the sub-pipelines that introduce 
-           differences among 20 subjects. The range between red and green colors show the most and 
-           least frequency of processes respectively. Besides, the total number 
-           of occurrences of each processes is represented in the bracket in each cell.}
+  \caption{Heatmap frequency of processes along with the sub-pipelines that introduce 
+           differences among 20 subjects. The range between red and green colors shows the most and 
+           least frequency of processes, respectively. 
+           Each cell describes the frequency of the specific process in the sub-pipelines.        
+           Each cell may split into multiple sub-cells when the specific process 
+           called by different options (distinct process).
+           Annotations represent `N [M] x Y" in which N is the percentage of the distinct process 
+           that creates differences out of all M times that is occurred in that sub-pipeline. 
+           Also, times Y shows the number of the distinct process with the same distribution.
+           For example, the cell of process FLIRT in sub-pipeline AAve shows 10 distinct processes  
+           with the frequency of 10 that introduce no differences.}
   \label{fig:pfs_freq}
@@
     Here is 
-    \href{https://raw.githubusercontent.com/ali4006/HCP-reproducibility-paper/master/images/brain_classification.gif}{the link}
-    to the gif animation of the same images.
+    \href{https://raw.githubusercontent.com/ali4006/HCP-reproducibility-paper/master/images/brain_classification.gif}
+    {the link} to the gif animation of the same images.
 } 
@@
 \centering
-  \includegraphics[width=\columnwidth]{images/brain_classification.png} 
+  \includegraphics[width=\columnwidth]{images/brain_classification_mni.png} 
   \caption{Heatmap binarized differences between brain segmentation results (file \emph{aseg.hires.mgz}) from 
-          20 subjects in FreeSurfer (CentOS6 vs. CentOS7).} 
+          20 subjects in FreeSurfer (CentOS6 vs. CentOS7).
+          We calculated some statistics of the differences based on the dice value of the segmented regions.
+          Here is 
+          \href{https://encsconcordiaca-my.sharepoint.com/:x:/g/personal/m_alari_encs_concordia_ca/EfNB-QMcivhIv6j0r9weffwBCHujf8RbaKjiYalIFa8cbQ?e=hyvQSU}
+          {the link} to the excel file.} 
   \label{fig:tissue_class}
@@
 the amplification of these differences should be quantified in the future works.
-For this purpose, we can evaluate the final result after perturbing the identified 
-processes in the pipeline iteratively.
-As a result, the processes that amplify differences would have a significant effect on the results.
 

### 2019-10-30 paper.tex
edit images and captions

--- 2019-10-09 paper.tex
@@
 \usepackage{ulem}
+\usepackage{pifont}
 \DeclareMathOperator*{\argmin}{argmin}
@@
 diethelm2012limits, Glatard2015, bowring2019exploring}.
+These irreproducibility issues are reported as the result of the 
+creation, propagation, and amplification of small numerical 
+differences.
 
-In particular, the studies on the effect of the operating system on the results show that 
-computational pipelines create small numerical differences~\cite{Glatard2015, Scaria2017}.
+In particular, the effect of the operating systems on  
+computational pipelines shows the creation of small numerical differences~\cite{Glatard2015, Scaria2017}.
 These differences mainly correspond to the mathematical functions implemented 
@@
 
-These irreproducibility issues are reported as the result of the 
-creation, propagation, and amplification of small numerical 
-differences~\cite{Gronenschild2012, diethelm2012limits, Glatard2015, 
-bowring2019exploring}. In this case, the analysis pipelines are said to 
-be numerically unstable. Numerical instability is a characteristic of 
-the pipelines which amplify small numerical differences and then hamper 
-the reproducibility of the analyses depending on the length of the 
-pipeline and magnitude of the differences. In many cases, numerical 
-instability is an essential issue for reproducibility.
-
-There are different solutions to improve the reproducibility of the analysis, 
-including containerization techniques that encapsulate software/hardware 
-dependencies, provenance capturing tools, and version control systems. 
-However, a comprehensive solution requires to fix the numerical 
-instabilities instead of masking the issue. For this purpose, we introduce 
-the Numerical Reproducibility Measurement (NURM) tool to identify the processes in 
-the pipeline that create numerical differences across several runs. 
-Furthermore, we can fix the instability of the identified processes 
-in the next steps.
+There are different approaches to improve the reproducibility of the analysis, 
+including containerization techniques that encapsulate software/hardware dependencies, 
+provenance capturing tools, and version control systems. 
+However, a comprehensive solution requires to fix the numerical instabilities instead of 
+masking the problem. For this purpose, bagging is one solution that has been used to 
+stabilize motion estimation analysis in fMRI~\cite{Glatard2018hbm}. 
+Bootstrap aggregation reduces the instabilities, but it needs expensive computations.
+As the second solution, a debugging tool can help to identify and 
+fix the instabilities in the pipeline. Therefore, we introduce 
+the Numerical Reproducibility Measurement (NURM) tool to identify 
+the processes in the pipeline that create numerical differences across different conditions. 
+Furthermore, we can stabilize the identified processes in the next steps.
 
@@
 process labelling to characterize differences. We describe the HCP 
-preprocessing pipelines as a popular neuroimaging project to test the 
+pre-processing pipelines as a popular neuroimaging project to test the 
 NURM-tool and evaluate its functionality. Finally, we reported the 
@@
   system calls). Dashed edges represent file dependencies: temporary 
-  files are in yellow and result files are in green.
+  files are in purple and result files are in black.
   Every node in the tree is labeled using (1) a process id created by our
@@
 that have differences and write files that also have differences are called
-\emph{undetermined} (Figure~\ref{fig:yellow}). Processes are also labeled
+\emph{undetermined} (Figure~\ref{fig:grey}). Processes are also labeled
 undetermined when they read or write temporary files, or produce multi-write
@@
 \centering
-    \begin{subfigure}{0.2\linewidth}
-        \includegraphics[scale=0.3]{images/green.png}
+    \begin{subfigure}{0.3\linewidth}
+      \centering
+        \includegraphics[scale=0.36]{images/green.png}
         \caption{Transparent}
@@
     \hfill
-    \begin{subfigure}{0.2\linewidth}
-    \includegraphics[scale=0.3]{images/red.png}
+    \begin{subfigure}{0.35\linewidth}
+      \centering
+    \includegraphics[scale=0.36]{images/red.png}
     \caption{Creates differences}
@@
     \hfill
-    \begin{subfigure}{0.2\linewidth}
-    \includegraphics[scale=0.3]{images/yellow.png}
+    \begin{subfigure}{0.3\linewidth}
+      \centering
+    \includegraphics[scale=0.36]{images/grey.png}
     \caption{Undetermined}
-    \label{fig:yellow}
+    \label{fig:grey}
 \end{subfigure}
@@
   Dashed edges refer to the file dependencies between processes A and B 
-  if a file written by A is read by B. Solid black edges refer to the 
-  relationship between parent and child processes.}
+  if a file written by A is read by B.}
     \label{fig:processes}
@@
   executed_file = sys.argv[0]
-  exec_argv = op.join('original_scripts', executed_file)
-  exec_argv = exec_argv + " ".join(map(pipes.quote, sys.argv[1:]))
-  subprocess.Popen(exec_argv, shell=True).communicate()
+  exec_proc = op.join('original_scripts', executed_file)
+  exec_argv = " ".join(map(pipes.quote, sys.argv[1:]))
+  exec_cmd = exec_proc + exec_argv
+  subprocess.Popen(exec_cmd, shell=True).communicate()
 
@@
 
-  if exec_argv in process_list.keys():
+  if exec_cmd in process_list.keys():
       # Save output files to read-only directory
-      p_files = process_list[exec_argv]
+      p_files = process_list[exec_cmd]
       shutil.copy(p_files, '/read-only/dir')
@@
 In addition, the distance functions that are specific to file types can be added 
-to our tool to ensure that differences belong to the file data part.
+to our tool to measure differences more precisely. 
 We developed an incremental approach that consists of the following steps:
@@
 
-Furthermore, the average time per subject taken for PreFreeSurfer processing is 2.2 hours, 
-and FreeSurfer processing is 8.3 hours (20 runs), 
-and the average output file size of PreFreeSurfer processing is 2.7 GB, 
-and FreeSurfer processing is 4.1 GB (20 subjects).
+Furthermore, the average time per subject taken for PreFreeSurfer processing is $\approx$ 2.2 hours, 
+and FreeSurfer processing is $\approx$ 8.3 hours (20 runs), 
+and the average output file size of 20 subjects in PreFreeSurfer is $\approx$ 2.7 GB, 
+and FreeSurfer is $\approx$ 4.1 GB.
 
@@
 help in the diagnosis of the brain injury.
+The images are indexed with a suffix if several scans of the same modality were acquired.
+For example, some data may include two T1w images with low and high resolutions.
 The functional images, including task-based fMRI and resting-state scans, 
@@
 the anatomical connectivity between regions.
-See~\cite{van2013wu} for a discussion of these various terms.
+See~\cite{van2013wu} for a discussion of these various modalities.
 
@@
 is available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
-In this paper, we randomly selected 20 unprocessed subjects from 
-the HCP data release S500. All the subjects were scanned using the 3-tesla 
-type of MRI scanners.
+In this paper, we randomly selected 20 unprocessed subjects 
+containing the 3-tesla type of MRI scanners from the HCP data release S500.
 
@@
 estimations with \emph{mris\_curvature}. To address this issue, we fixed
-the seed used by the pseudo-random number generator, through an option of
-the FreeSurfer tools. In this way, our analysis can focus on between-OS
-differences.
+the seed used by the pseudo-random number generator, through defining 
+option ``--seed'' in the command-line of the processes.
+Our analysis can focus on between-OS differences.
 
@@
 
-\begin{tabular}{@{}llllll@{}}
+\begin{tabular}{cccccc}
 \toprule
-Type   & Number of Subjects  & Missing File(s)                                    \\ \midrule
-type1  & 9                   & None                                               \\
-type2  & 7                   & [``T1w", ``T2w"]                                   \\
-type3  & 2                   & [``T1w"]                                           \\
-type4  & 2                   & [``T2w"]                                           \\ \bottomrule
+       &                        &  \multicolumn{4}{c}{Available Modalities}    \\ 
+\cmidrule(lr){3-6}       
+Type   &   \makecell{Number of \\ Subjects}   &  T1w\_1          & T1w\_2      & T2w\_1          & T2w\_2        \\ \midrule
+1      &               9                      &   \ding{51}      &   \ding{51} &   \ding{51}     &   \ding{51}   \\ 
+2      &               8                      &   \ding{51}      &             &   \ding{51}     &               \\ 
+3      &               1                      &   \ding{51}      &             &   \ding{51}     &   \ding{51}   \\ 
+4      &               2                      &   \ding{51}      &   \ding{51} &   \ding{51}     &               \\ 
+\bottomrule
 \end{tabular}
@@
 Figure~\ref{fig:pfs_freq} shows the frequency at which  processes create differences 
-in the pipeline, and how differences are distribute among sub-pipelines.
+in the pipeline, and how differences are distributed among sub-pipelines.
 Differences were observed in linear registration 
@@
 and Altas Registration), and in image warping with FSL \emph{new\_invwarp} (in Brain Extraction 
-and Atlas Registration). 
-Differences were also observed in image mean 
+and Atlas Registration). Differences were also observed in image mean 
 computations with FSL maths  (in Anatomical Average).
 
-Results show that \emph{FLIRT} is the most frequent process that produces 
-differences in ACPC-Alignment across all the subjects, Distortion Correction 
-in 95\% of the subjects, and Atlas Registration in 85\% of the subjects. 
-In addition, there is a \emph{fslmaths} process in Anatomical Average that 
-creates differences in all 11 subjects in which it has appeared.
+Results show the variability between processes when 
+the same tool behaves differently in different process instances. For example, 
+FLIRT tool always produces files without differences in \emph{AAve} sub-pipeline but 
+mostly creates different files in \emph{AAli} sub-pipeline.
 
+Furthermore, different frequencies show the variability between subjects.
+Results show that the same process can be reproducible for some subjects but not for others. 
+Between subject differences can be a result of different subject types. 
+The similar findings in the other studies show that the subject 
+is an important factor in reproducibility~\cite{bowring2019exploring}.
+
+% make figure in two columns use {figure*}
 \begin{figure}
 \centering
-  \includegraphics[width=\columnwidth]{images/pfs_heatmap.png}
-  \caption{Heatmap of difference frequency of processes along with the sub-pipelines that introduce 
-           differences among 20 subjects. The range between red and green colors show the most and 
-           least frequency of processes respectively. Besides, the total number 
-           of occurrences of each processes is represented in the bracket in each cell.}
+  \includegraphics[width=\columnwidth,height=4cm]{images/pfs_heatmap.png}
+  \caption{Heatmap frequency of processes along with the sub-pipelines that introduce 
+           differences among 20 subjects. 
+           Each column represents the pipeline tool, and each row shows the sub-pipeline.
+           Each cell may contain several process instances of a tool in a sub-pipeline.
+           Annotations describe the frequency distribution of the instances as ``N [M] x Y" 
+           in which N is the percentage of the process 
+           that creates differences out of all M times that is occurred in the sub-pipeline. 
+           Also, times Y shows the number of instances with the same distribution.
+           For example, the cell of FLIRT tool in sub-pipeline AAve shows 10 process instances  
+           with the frequency of 10 that introduce no differences.
+           The range between red and green colors shows the highest and lowest 
+           proportion of occurrences of the processes with differences.
+           Also, grey cells indicate that the pipeline tool is not executed in the sub-pipeline.}
   \label{fig:pfs_freq}
@@
 
-To visually represent the magnitude of variations, we created checkerboard 
-images from the image results of one subject in two OSes. Figure~\ref{fig:fnirt_result} 
+Figure~\ref{fig:fnirt_result} 
 shows the checkerboard images from the results of ACPC-Alignment in PreFreeSurfer. 
 Each rectangle includes the replaced pixels with different intensity values between images.
+The misaligned regions visually represent significant differences in the alignment 
+procedure.
 
@@
   \centering
-    \includegraphics[width=\columnwidth]{images/segmentation.png} 
+    \includegraphics[width=\columnwidth]{images/t2w_alignment.png} 
     \caption{Differences between FNIRT results of one subject from PreFreeSurfer 
     (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
-    CentOS7). Right images show the zoomed version part of the images marked 
-    by the red rectangle.
+    CentOS7). Sagittal (top) and axial (bottom) views of the brain in dimension of X=130, Y=155, Z=130.
+    Right images show the zoomed version part of the images marked by the red rectangle.
     Here is 
-    \href{https://raw.githubusercontent.com/ali4006/HCP-reproducibility-paper/master/images/brain_classification.gif}{the link}
-    to the gif animation of the same images.
+    \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
+    {the link} to the gif animation of the same images.
 } 
@@
 
-In the first test, the results of PreFreeSurfer on CentOS 6
-feed to the FreeSurfer pipeline on both OSes. 
-We fixed input data to determine whether the pipeline creates differences.
-We processed 20 subjects using FreeSurfer.
 NURM-tool identified \emph{mris\_make\_surfaces} as the only process in 
-the pipeline that creates differences among 10 subjects.
+the pipeline that creates differences in 10 subjects among 20 subjects.
 This process generates surface files for cortical and white matter surfaces 
-(see~\cite{fischl2012freesurfer} for more detail of the process). 
+(see~\cite{fischl2012freesurfer} for more detail). 
 Also, FreeSurfer did not introduce any differences in 10 subjects.
 
-Results show that the static builds of FreeSurfer in which uses the embedded libraries, 
+Results show that the static builds of FreeSurfer in which uses the embedded libraries 
 significantly avoids the creation of differences.
 However, we found that between-OS differences occur when pre-processed data are different in FreeSurfer. 
-For this purpose, in the second test, the results of PreFreeSurfer feed to the FreeSurfer 
-pipeline on each operating system separately.
-Figure~\ref{fig:tissue_class} shows between-OS differences  
-as a result of propagation and amplification of differences in FreeSurfer. % that coming from PreFreeSurfer outputs. 
-This figure illustrates the sum of binarized differences 
+FreeSurfer may not introduce differences, but it propagates and amplifies 
+the current differences that appeared in PreFreeSurfer results across the operating systems.
+
+Figure~\ref{fig:tissue_class} illustrates the sum of binarized differences 
 from 20 subject results of the brain tissue segmentation across OSes.
-The magnitude of differences is in the range of 1-20 as the total number of subjects.
+This figure shows between-OS differences  
+as a result of the propagation and amplification of differences in FreeSurfer. 
+Differences are mostly localized at the interfaces between tissues.
+
+Furthermore, Figure~\ref{fig:scatter_plot} plots the mean of Dice coefficients 
+for the 45 segmented regions obtained on two operating systems.
+High Dice coefficients (more than 0.95) in most of the regions indicate minor differences.
+Dice values can be very low , down to 0.5, in particular in small structures 
+such as the vessels and lateral ventricles.
 
@@
 \centering
-  \includegraphics[width=\columnwidth]{images/brain_classification.png} 
+  \includegraphics[width=\columnwidth]{images/brain_segmentation_mni.png} 
   \caption{Heatmap binarized differences between brain segmentation results (file \emph{aseg.hires.mgz}) from 
-          20 subjects in FreeSurfer (CentOS6 vs. CentOS7).} 
+          20 subjects in FreeSurfer (CentOS6 vs. CentOS7). 
+          All binarized differences were resampled to the default MNI152 volume template.
+          Here is 
+          \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
+          {the link} to the gif animation from a single subject.} 
   \label{fig:tissue_class}
+\end{figure}
+
+\begin{figure}
+  \centering
+    \includegraphics[width=\columnwidth]{images/scatter_plot.png} 
+    \caption{The scatter plot of the mean of Dice coefficients between segmented regions 
+            from 20 subjetcs in FreeSurfer (CentOS6 vs. CentOS7). 
+            The complete Dice coefficient values are 
+            \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/data/dice_values.csv}
+            {available}.} 
+    \label{fig:scatter_plot}
   \end{figure}
@@
 the amplification of these differences should be quantified in the future works.
-For this purpose, we can evaluate the final result after perturbing the identified 
-processes in the pipeline iteratively.
-As a result, the processes that amplify differences would have a significant effect on the results.
+
+\note{FREESURFER RESULTS: 
+These are consistant results to the paper~\cite{Glatard2015}, as it is shown the propaggation of differences in cortical thickness extraction analysis.
+The magnitude of differences from the operating system changes is out of our expectaion}
 

### 2019-11-06 paper.tex
update result images and add discussion

@@
 Between subject differences can be a result of different subject types. 
-The similar findings in the other studies show that the subject 
-is an important factor in reproducibility~\cite{bowring2019exploring}.
+For instance, the subjects contain two T1 images need more calculations 
+to produce an average image between them.
+The similar findings in another study~\cite{bowring2019exploring} were reported 
+the effect of subject type on reproducibility.
 
 % make figure in two columns use {figure*}
-\begin{figure}
+\begin{figure*}
 \centering
-  \includegraphics[width=\columnwidth,height=4cm]{images/pfs_heatmap.png}
+  \includegraphics[width=\textwidth]{images/pfs_heatmap.png}
   \caption{Heatmap frequency of processes along with the sub-pipelines that introduce 
@@
   \label{fig:pfs_freq}
-\end{figure}
+\end{figure*}
 
@@
     (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
-    CentOS7). Sagittal (top) and axial (bottom) views of the brain in dimension of X=130, Y=155, Z=130.
-    Right images show the zoomed version part of the images marked by the red rectangle.
-    Here is 
+    CentOS7). Sagittal (left) and coronal (right) views of the brain are in dimension of X=130, Y=155, Z=130.
     \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
-    {the link} to the gif animation of the same images.
+    {the link} to the animated version of alignment results shows the better visualization of differences.
 } 
@@
 Differences are mostly localized at the interfaces between tissues.
+In our dataset, we found subjects with the min and max Dice coefficient 
+values of 0.89 and 0.96 respectively.
 
-Furthermore, Figure~\ref{fig:scatter_plot} plots the mean of Dice coefficients 
-for the 45 segmented regions obtained on two operating systems.
+Furthermore, Figure~\ref{fig:scatter_plot} plots the Dice coefficients 
+for the 44 segmented regions obtained on the two operating systems.
 High Dice coefficients (more than 0.95) in most of the regions indicate minor differences.
-Dice values can be very low , down to 0.5, in particular in small structures 
+Dice values can be very low, down to 0.5, in particular in small structures 
 such as the vessels and lateral ventricles.
@@
           All binarized differences were resampled to the default MNI152 volume template.
-          Here is 
           \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
-          {the link} to the gif animation from a single subject.} 
+          {The link} to the animated version of segmentation results of one subject represents the significance of differences.} 
   \label{fig:tissue_class}
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
-    \includegraphics[width=\columnwidth]{images/scatter_plot.png} 
-    \caption{The scatter plot of the mean of Dice coefficients between segmented regions 
-            from 20 subjetcs in FreeSurfer (CentOS6 vs. CentOS7). 
-            The complete Dice coefficient values are 
+    \includegraphics[width=\textwidth]{images/scatter_plot.png} 
+    \caption{The scatter plots of the Dice coefficients between segmented regions 
+            from 20 subjects in FreeSurfer (CentOS6 vs. CentOS7). 
+            The x-axes are the average number of voxels in each region.
+            Regions are divided into four sub-plots from the smallest regions in the top-left plot 
+            to the largest ones in the bottom-right plot.
+            The complete table of Dice coefficient values is 
             \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/data/dice_values.csv}
@@
     \label{fig:scatter_plot}
-  \end{figure}
+  \end{figure*}
 
@@
 We can see that \emph{FLIRT} creates differences in almost all the sub-pipelines. 
-Therefore, it is important to investigate the registration procedure and find 
-a way to stabilize this process.
+We monitored the library system calls using \emph{ltrace} to check that registration processes
+use the same libraries. There were no differences in the library dependencies used by 
+the tool in two operating systems, but we found a different number of calls to the \emph{glibc}.
+More investigations need to stabilize registration processes.
 
@@
 the analysis results obtained by the pipeline show that between-OS differences are visually substantial.
-We see that statically builds of the pipeline improves reproducibility, but 
-it would not improve the numerical stability of the pipeline. 
+The statically builds of the pipeline would not improve the numerical stability of the pipeline. 
 FreeSurfer pipeline propagates and amplifies the numerical differences that are created by the 
 differences in the underlying operating system libraries.
+These results are consistent with the findings in the paper~\cite{Glatard2015}, 
+as it has confirmed the propagation of differences in cortical thickness extraction analysis using FreeSurfer.
 
-In addition to variations of the operating system, other factors exist which may influence 
-the reproducibility of the analyses such as software versions, compilers, hardware, etc.
-Generally, we can use the Monte Carlo Arithmetic (MCA) to determine the numerical stability of the pipelines.
-This algorithm perturbs every floating-point operation 
-in the analysis based on the random rounding in each run. 
-By this algorithm, NURM-tool can easily locate which parts of the analyzed source code are responsible 
-for the floating-point related instabilities.
+We obtained results between-OS versions; further investigations could be done over the OS families (e.g., Linux VS. Windows).
+In addition to variations of the operating system, numerical instability of the pipelines have been 
+explored when varying software, hardware and implementation details, or by adding negligible perturbations.
+In the field of numerical analysis, we can investigate these issue using the Monte Carlo Arithmetic (MCA), 
+a method instrumenting floating-point operations based on the random rounding algorithms at a target precision.
+By this algorithm, we can locate the instabilities for the given pipeline, tool, or operations.
 
@@
 the amplification of these differences should be quantified in the future works.
-
-\note{FREESURFER RESULTS: 
-These are consistant results to the paper~\cite{Glatard2015}, as it is shown the propaggation of differences in cortical thickness extraction analysis.
-The magnitude of differences from the operating system changes is out of our expectaion}
+We can use distance-preserving hashes rather than MD5 checksums or metrics specific to the file type like Dice coefficient 
+for the registration processes to measure the magnitude of differences.
 
@@
 
-The preferred solution is to find the origin of amplification of the differences and 
+\note{The preferred solution is to find the origin of amplification of the differences and 
 then fixing the particular functions/modules in the processing pipelines. 
+Bagging is a fundamental technique toward reducing the instability of the analysis.
+Therefore, we can use bagging for the parts in the pipeline that amplify differences, 
+but it still is computationaly expensive. }
 

### 2019-11-29 paper.tex
notes on intro

@@
 
-\title{NURM: a tool to 
-locate numerical differences in pipelines}
+\title{NURM: file-based localization of numerical /differences/ without instrumentation}
 
@@
 \section{Introduction}
+
+% Reproducibility: crisis, numerical stability
+% Containerization isn't the solution
+
+% Complex neuroimaging pipelines
+In neuroimaging, our primary application field, data analyses often consist
+of computational pipelines aggregating heterogeneous codes to implement a
+specific function. For instance ...
+
+% Verificarlo and MCA, require recompilation
+Numerical instabilities originate in the limited \todo{precision} of floating-point representations. 
+Traditionnaly, it has been evaluated through
+
+% -> File-based analysis of pipelines without instrumentation
+% Reprozip
 

### 2019-11-29 paper.tex

@@
 These functions are \emph{mri\_diff}, \emph{mris\_diff}, and \emph{lta\_diff} 
-for comparing volumes, surface-based files, 
+from FreeSurfer tool for comparing volumes, surface-based files, 
 and lta files (transforms) respectively. 
@@
 (see~\cite{fischl2012freesurfer} for more details). 
+Moreover, we used the Dice coefficient metric to evaluate the similarity of 
+the image segmentation results, which is defined by
+\[DICE=\frac{2(|X|\cap|Y|)}{|X|+|Y|}\]
+The coefficient between 0 to 1, 1 means that two images are totally identical, while if 
+X and Y have no voxels in common, it is equal to 0.
 
@@
     \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
-    {the link} to the animated version of alignment results shows the better visualization of differences.
+    {The link} to the animated version of alignment results shows the better visualization of differences.
 } 
@@
 In our dataset, we found subjects with the min and max Dice coefficient 
-values of 0.89 and 0.96 respectively.
+values of 0.89 and 0.96, respectively.
 
@@
 such as the vessels and lateral ventricles.
+
+We formulated the linear regression model for the Dice and region size variables as:
+\[D=1 + T\epsilon ;    T=\frac{-1}{\overline{R}}\]
+It is easier to obtain high overlaps with large regions, 
+while smaller regions are harder to be similarly segmented.
+We obtained the epsilon value of $\approx$ 8.3 voxels ($\approx$ 3mm) which indicate the minimum number
+of voxels in each region that needs to be segmented.
 
@@
             from 20 subjects in FreeSurfer (CentOS6 vs. CentOS7). 
-            The x-axes are the average number of voxels in each region.
-            Regions are divided into four sub-plots from the smallest regions in the top-left plot 
-            to the largest ones in the bottom-right plot.
+            The x-axes are the number of voxels in each region.
+            Right plot shows an overview of similarities between all regions and 
+            left plot represents a closer view of differences between 
+            regions smaller than $10^4$ voxels.
             The complete table of Dice coefficient values is 
@@
 In the second experiment, 
-the analysis results obtained by the pipeline show that between-OS differences are visually substantial.
-The statically builds of the pipeline would not improve the numerical stability of the pipeline. 
+the analyses obtained by the FreeSurfer show substantial between-OS differences.
+These results show the sensitivity of the segmentation process to the 
+size of regions, however, it might be also sensitive to the location and shape of 
+the segmented regions.
+We demonstrate that statically builds of the pipeline would not improve the numerical stability of the pipeline. 
 FreeSurfer pipeline propagates and amplifies the numerical differences that are created by the 
@@
 
-We obtained results between-OS versions; further investigations could be done over the OS families (e.g., Linux VS. Windows).
+We obtained results between OS versions; further investigations could be done over the OS families (e.g., Linux VS. Windows).
 In addition to variations of the operating system, numerical instability of the pipelines have been 
-explored when varying software, hardware and implementation details, or by adding negligible perturbations.
+explored when varying software, hardware and implementation details, or even by adding negligible 
+perturbations like one voxel noise in the input dataset.
 In the field of numerical analysis, we can investigate these issue using the Monte Carlo Arithmetic (MCA), 
 a method instrumenting floating-point operations based on the random rounding algorithms at a target precision.
-By this algorithm, we can locate the instabilities for the given pipeline, tool, or operations.
+By this algorithm, we can locate the instabilities for the given pipelines, tools, or operations.
 
+We demonstrate that computational models in neuroimaging are sensitive to numerical instabilities. 
 Although we identify the processes in the pipeline that create differences, 
 the amplification of these differences should be quantified in the future works.
-We can use distance-preserving hashes rather than MD5 checksums or metrics specific to the file type like Dice coefficient 
-for the registration processes to measure the magnitude of differences.
+We can use distance-preserving hashes rather than MD5 checksums or metrics specific to the file type like the Dice coefficient 
+for the registration and segmentation processes to measure the magnitude of differences.
 
@@
 Our technique can characterize the stability of the pipelines
-automatically. We identified the main causes of the between-OS differences as (1) the evolution 
-of the math libraries over time and, (2) the instability of the pipelines. 
+automatically. We identified the leading causes of the between-OS differences as (1) the evolution 
+of the math libraries over time and (2) the instability of the pipelines. 
 There are two ways to tackle this problem. The easy but less preferred solution is masking the instabilities.
@@
 (4) building static executables by removing the host operating system library dependencies. 
-These solutions only make the problem invisible, but the pipelines are stiil unstable.
+These solutions only make the problem invisible, but the pipelines are still unstable.
 
-\note{The preferred solution is to find the origin of amplification of the differences and 
+\note{The preferred solution is to find the origin of amplification of differences and 
 then fixing the particular functions/modules in the processing pipelines. 
@@
 Therefore, we can use bagging for the parts in the pipeline that amplify differences, 
-but it still is computationaly expensive. }
+but it still is computationally expensive. }
 
@@
 
-\note{
-CBRAIN team. Compute Canada(Calcul Quebec).
-
-Data were provided by the Human Connectome Project, WU-Minn 
+This research was enabled in part by support provided by 
+Calcul Quebec (http://www.calculquebec.ca) and 
+Compute Canada (http://www.computecanada.ca).
+Also, data were provided by the Human Connectome Project, WU-Minn 
 Consortium (Principal Investigators: David Van Essen and Kamil Ugurbil; 
@@
 Center for Systems Neuroscience at Washington University.
-
-}
 

### 2019-11-29 paper.tex

@@
 
-\title{NURM: file-based localization of numerical /differences/ without instrumentation}
+\title{NURM: a tool to 
+locate numerical differences in pipelines}
 
@@
 \section{Introduction}
-
-% Reproducibility: crisis, numerical stability
-% Containerization isn't the solution
-
-% Complex neuroimaging pipelines
-In neuroimaging, our primary application field, data analyses often consist
-of computational pipelines aggregating heterogeneous codes to implement a
-specific function. For instance ...
-
-% Verificarlo and MCA, require recompilation
-Numerical instabilities originate in the limited \todo{precision} of floating-point representations. 
-Traditionnaly, it has been evaluated through
-
-% -> File-based analysis of pipelines without instrumentation
-% Reprozip
 

### 2019-12-16 paper.tex

@@
 
-\title{NURM: a tool to 
+\title{Spot: a tool to 
 locate numerical differences in pipelines}
@@
 fix the instabilities in the pipeline. Therefore, we introduce 
-the Numerical Reproducibility Measurement (NURM) tool to identify 
+the numerical reproducibility measurement tool named \emph{Spot} to identify 
 the processes in the pipeline that create numerical differences across different conditions. 
@@
 
-The remainder of this paper describes the NURM-tool in different parts 
+The remainder of this paper describes the Spot tool in different parts 
 including the way of capturing and representation of the provenance 
@@
 pre-processing pipelines as a popular neuroimaging project to test the 
-NURM-tool and evaluate its functionality. Finally, we reported the 
+Spot tool and evaluate its functionality. Finally, we reported the 
 experimental results which describe the processes that are responsible 
@@
    \end{subfigure}
-   \caption{An overview of NURM-tool in two steps. First capturing intermediary files (a); then
+   \caption{An overview of Spot tool in two steps. First capturing intermediary files (a); then
             labelling process graphs (b).}
@@
 
-In particular, NURM-tool is tested on the neuroimaging applications which are 
+In particular, Spot tool is tested on the neuroimaging applications which are 
 predominantly using mathematical libraries. Therefore, we expect to find 
@@
 
-NURM-tool identified \emph{mris\_make\_surfaces} as the only process in 
+Spot tool identified \emph{mris\_make\_surfaces} as the only process in 
 the pipeline that creates differences in 10 subjects among 20 subjects.

### 2019-12-16 paper.tex

@@
 
-\title{Spot: a tool to 
-locate numerical differences in pipelines}
-
+\title{File-based localization of numerical perturbations in data analysis pipelines}
 \begin{document}
@@
 \author[1]{Ali Salari}
-\author[1]{Lalet Scaria}
 \author[2,3]{Gregory Kiar}
@@
 
-Many experiments show that computational analysis results are still not 
-completely reproducible. Computational environments including different 
-operating systems, hardware categories, and software versions are known 
-to have effects on the results produced by analysis pipelines. These 
-effects are presumably due to the creation, propagation and 
-amplification of small numerical differences across the pipelines. 
-Although new techniques such as virtualization, version control 
-systems, and provenance management tools provide a more reliable 
-environment and significantly improve the reproducibility of scientific 
-findings. However, the precise causes of such instabilities and the 
-path along which they propagate in the pipelines are unclear.  We 
-present a technique to identify the processes in the pipeline that 
-create numerical differences along with the execution, and we apply this 
-technique to the HCP structural pre-processing pipelines.
+Operating systems and other computational factors are known to have an
+effect on the results produced by neuroimaging pipelines, presumably due to
+the creation, propagation and amplification of numerical errors along data
+analysis pipelines. However, the precise causes of such instabilities and
+the path along which they propagate in the pipelines are unclear, while
+this process could play a major role in the current reproducibility crisis.
+We present a technique to identify the computational processes in a
+pipeline that create numerical differences when executed in different
+computational conditions. Our technique leverages system-call interception to reconstruct
+processing and file graphs without pipeline instrumentation. Applying our tool to
+the structural pre-processing pipelines released by the Human Connectome
+Project, we find that linear and non-linear registration are the cause of
+most numerical instabilities in such pipelines, which confirms previous
+findings.
 
@@
 \begin{keywords}
-Reproducibility; Numerical Instability; Neuroimaging.
+Computational reproducibility; Numerical instability; Neuroimaging.
 \end{keywords}
@@
 
-Reproducibility is a crucial element of the scientific works, 
-as it enables researchers to evaluate authenticity and reliability 
-of the findings~\cite{plesser2018reproducibility}.
-Reproducibility is defined as the ability to regenerate the same 
-results as the original findings when the experiment is reanalyzed by 
-the same analytic methods, software package, parameters, and 
-data~\cite{peng2011reproducible}. 
-In addition, numerical reproducibility is defined as the ability to 
-regenerate bit for bit identical results from multiple 
-runs~\cite{hill2017numerical}. 
-We considered the numerical reproducibility in our experiments by comparing 
-the binary content of the results using the checksum method.
+% Reproducibility: crisis, numerical stability
+% Containerization isn't the solution
 
-Recently, validation of the reproducibility has been widely investigated 
-in the field of neuroimaging in which uses optimized 
-processing methods with the purpose of functional and structural 
-assessments of the human brain.
-According to the previous 
-studies, the variety of computing infrastructures including workstation 
-types, parallelization methods, operating systems, and analysis 
-packages are known to influence the reproducibility of the analyses~\cite{Gronenschild2012, 
-diethelm2012limits, Glatard2015, bowring2019exploring}.
-These irreproducibility issues are reported as the result of the 
-creation, propagation, and amplification of small numerical 
-differences.
+Numerical stability is believed to play a central role in the differences
+observed in data analysis pipelines across computational infrastructures.
+Variability resulting from operating system upgrades, parallelization, or
+even minor data perturbations may originate in errors that accumulate in
+numerical computations, with important consequences on the reproducibility
+of data analysis results. However, it remains challenging to identify the
+exact elements in a complex pipeline that are prone to numerical instability,
+leaving pipeline developers and users with the daunting task to manually
+compare results obtained in different conditions.
 
-In particular, the effect of the operating systems on  
-computational pipelines shows the creation of small numerical differences~\cite{Glatard2015, Scaria2017}.
-These differences mainly correspond to the mathematical functions implemented 
-in different operating system libraries.
-For instance, changing the mathematical functions like \emph{expf()} and 
-\emph{cosf()} which manipulate the precision of floating-point representations, 
-between \emph{glibc} libraries in different operating systems can produce 
-small numerical differences.
-Besides, a similar issue is expected for any operating system which is 
-based on \emph{glibc}, the GNU C library.
+In neuroimaging, our primary application field, data analyses often consist
+of hundreds of computational processes, sometimes coming from multiple
+toolboxes, that are aggregated to perform a specific function. For
+instance, the fMRIprep pipeline~\cite{esteban2019fmriprep} assembles software blocks
+from FSL, AFNI, Freesurfer and ANTs to provide a state-of-the art
+functional MRI processing tool with minimal user input. Another example are the pipelines of the Human
+Connectome Project~\cite{glasser2013} that combine tools from FSL and Freesurfer to pre-process
+structural, functional and diffusion data from their
+open dataset. In both cases, the pipelines leverage toolboxes that are
+widely trusted in the community. At the same time, substantial results
+variations have been observed in these toolboxes resulting from minor data
+or infrastructure perturbations~\cite{Glatard2015, Gronenschild2012, Lewis2017-ll}, which suggests that further investigation of their
+numerical conditioning is required. For such complex pipelines, a
+lightweight solution has to be found to perform such evaluations with
+limited code instrumentation.
 
-There are different approaches to improve the reproducibility of the analysis, 
-including containerization techniques that encapsulate software/hardware dependencies, 
-provenance capturing tools, and version control systems. 
-However, a comprehensive solution requires to fix the numerical instabilities instead of 
-masking the problem. For this purpose, bagging is one solution that has been used to 
-stabilize motion estimation analysis in fMRI~\cite{Glatard2018hbm}. 
-Bootstrap aggregation reduces the instabilities, but it needs expensive computations.
-As the second solution, a debugging tool can help to identify and 
-fix the instabilities in the pipeline. Therefore, we introduce 
-the numerical reproducibility measurement tool named \emph{Spot} to identify 
-the processes in the pipeline that create numerical differences across different conditions. 
-Furthermore, we can stabilize the identified processes in the next steps.
+Numerical evaluations are traditionnally performed using techniques such as
+interval arithmetics~\cite{hickey2001interval} that require complete code re-writes and are
+therefore barely applicable to complex pipelines. Recently, Monte-Carlo
+Arithmetic~\cite{Parker1997-qq, Denis2016-wo} provided a practical way to
+evaluate the uncertainty of numerical results without the need to rewrite
+the application in a different paradigm. By perturbating floating-point
+computations, it introduces a controllable amount of noise in the
+pipelines, effectively sampling results from a random distribution. While
+this technique is very appealing, it suffers from two main issues that make
+it impractical at the scale of a complete pipeline. First, it requires that
+all software components be recompiled for MCA instrumentation, which is not
+always feasible. Second, it multiplies the execution time by a factor of 10
+to 100, which is impractical when executions already take a few hours to
+complete.
 
-The remainder of this paper describes the Spot tool in different parts 
-including the way of capturing and representation of the provenance 
-information, the method of clustering of different subject types, and 
-process labelling to characterize differences. We describe the HCP 
-pre-processing pipelines as a popular neuroimaging project to test the 
-Spot tool and evaluate its functionality. Finally, we reported the 
-experimental results which describe the processes that are responsible 
-for the differences in the pipelines.
+We present Spot, a tool to identify the source of numerical differences in
+complex pipelines without instrumentation. Using system-call interception
+through the Reprozip tool~\cite{chirigati2016reprozip}, Spot traverses graphs of processes and
+intermediary files to pinpoint the pipeline components that are unstable
+across execution conditions. When differences start accumulating,
+effectively masking any further instability, it restores clean data copies
+through a set of wrapper scripts. Wrapper scripts are also used to restore
+temporary data that might have been deleted during the execution, and to
+disambiguate files that have been written by multiple processes. The remainder of this paper 
+presents the design of our tool, and its 
+application to pre-processing pipelines of the HCP project.
+
+% % Verificarlo and MCA, require recompilation
+% Numerical instabilities originate in the limited \todo{precision} of floating-point representations. 
+% Traditionnaly, it has been evaluated through
+
+% % -> File-based analysis of pipelines without instrumentation
+% % Reprozip
+
+% Reproducibility is a crucial element of the scientific works, 
+% as it enables researchers to evaluate authenticity and reliability 
+% of the findings~\cite{plesser2018reproducibility}.
+% Reproducibility is defined as the ability to regenerate the same 
+% results as the original findings when the experiment is reanalyzed by 
+% the same analytic methods, software package, parameters, and 
+% data~\cite{peng2011reproducible}. 
+% In addition, numerical reproducibility is defined as the ability to 
+% regenerate bit for bit identical results from multiple 
+% runs~\cite{hill2017numerical}. 
+% We considered the numerical reproducibility in our experiments by comparing 
+% the binary content of the results using the checksum method.
+
+% Recently, validation of the reproducibility has been widely investigated 
+% in the field of neuroimaging in which uses optimized 
+% processing methods with the purpose of functional and structural 
+% assessments of the human brain.
+% According to the previous 
+% studies, the variety of computing infrastructures including workstation 
+% types, parallelization methods, operating systems, and analysis 
+% packages are known to influence the reproducibility of the analyses~\cite{Gronenschild2012, 
+% diethelm2012limits, Glatard2015, bowring2019exploring}.
+% These irreproducibility issues are reported as the result of the 
+% creation, propagation, and amplification of small numerical 
+% differences.
+
+% In particular, the effect of the operating systems on  
+% computational pipelines shows the creation of small numerical differences~\cite{Glatard2015, Scaria2017}.
+% These differences mainly correspond to the mathematical functions implemented 
+% in different operating system libraries.
+% For instance, changing the mathematical functions like \emph{expf()} and 
+% \emph{cosf()} which manipulate the precision of floating-point representations, 
+% between \emph{glibc} libraries in different operating systems can produce 
+% small numerical differences.
+% Besides, a similar issue is expected for any operating system which is 
+% based on \emph{glibc}, the GNU C library.
+
+% There are different approaches to improve the reproducibility of the analysis, 
+% including containerization techniques that encapsulate software/hardware dependencies, 
+% provenance capturing tools, and version control systems. 
+% However, a comprehensive solution requires to fix the numerical instabilities instead of 
+% masking the problem. For this purpose, bagging is one solution that has been used to 
+% stabilize motion estimation analysis in fMRI~\cite{Glatard2018hbm}. 
+% Bootstrap aggregation reduces the instabilities, but it needs expensive computations.
+% As the second solution, a debugging tool can help to identify and 
+% fix the instabilities in the pipeline. Therefore, we introduce 
+% the numerical reproducibility measurement tool named \emph{Spot} to identify 
+% the processes in the pipeline that create numerical differences across different conditions. 
+% Furthermore, we can stabilize the identified processes in the next steps.
 
@@
 
+% mention fuzzy libmath as a way to do this across distributions.
+
 \section{Conclusion}
@@
 but it still is computationally expensive. }
+
+% importance of data-driven evaluations, subject clustering, etc
+
+% future work with MCA
+
+% link registration to our motion estimation study
 

### 2020-01-06 paper.tex

@@
 computational conditions. Our technique leverages system-call interception to reconstruct
-processing and file graphs without pipeline instrumentation. Applying our tool to
-the structural pre-processing pipelines released by the Human Connectome
-Project, we find that linear and non-linear registration are the cause of
+processing and file graphs without pipeline instrumentation. By applying our tool to
+the HCP pre-processing pipelines, 
+we find that linear and non-linear registration are the cause of
 most numerical instabilities in such pipelines, which confirms previous
@@
 structural, functional and diffusion data from their
-open dataset. In both cases, the pipelines leverage toolboxes that are
+open dataset. In both cases, pipelines leverage toolboxes that are
 widely trusted in the community. At the same time, substantial results
@@
 
-Figure~\ref{fig:overview} shows an overview of our method. For each tested
+Spot identifies processes that create differences from a process graph of the pipeline execution.
+This tool compares file dependencies through all the processes based on the result files in two conditions.
+In this paper, different conditions refer to various operating systems.
+This tool is applicable on the complex pipelines in which tracing a large number of processes is difficult.
+
+Figure~\ref{fig:overview-tool} shows an overview of our method. For each tested
 condition, Condition 1 and Condition 2, the pipeline is containerized with
 Docker, and its interface is described with
-Boutiques~\cite{glatard2017boutiques}. \reprozip~\cite{chirigati2016reprozip} is
-added to Condition 1 for provenance tracking. Our tool starts by
-identifying temporary files and ``multi-write" files, i.e., files written
-by more than one process (Figure~\ref{fig:overview-capturing}). To do that,
-the tool first executes the pipeline in Condition 1, to obtain a process
-graph (Figure~\ref{fig:overview-capturing} - (1)) and a list of result files.
-The list of processes that create intermediary or multi-write
-files is then generated (Figure~\ref{fig:overview-capturing} - (2)). 
-Using this list, the tool modifies Docker image in Condition 1 and
-exectutes the pipeline again through it, which turns to produce results
-including temporary and overwritten files
-(Figure~\ref{fig:overview-capturing} - (3,4)). In the second step, the
+Boutiques~\cite{glatard2017boutiques}. 
+\reprozip is added to Condition 1 for provenance tracking. 
+This captures process dependencies through system call interceptions, 
+then Spot tool creates related process trees (Figure~\ref{fig:overview-tool} - (1)).
+Different subjects produce different process trees. 
+Hierachical clustering used to group subjects into different types (Figure~\ref{fig:overview-tool} - (2)).
+The rest of the procedure is done separately on each subject type.
+Our tool continues by
+restoring temporary files and ``multi-write" files, i.e., files written
+by more than one process (Figure~\ref{fig:overview-tool} - (3)). 
+To do so, the tool creates a list of processes 
+that introduce intermediary files and replaces their corresponding executables with a wrapper script. 
+The tool re-executes the pipeline in a modified Docker image in Condition 1.
+This copies files in the list to a read-only directory.
+In the next step, the
 processes are labeled from the differences found in their outputs in the
-two tested conditions (Figure~\ref{fig:overview-labelling}). To this order,
-the tool builds docker image for Condition 2 by modifying pipeline
-processes (Figure~\ref{fig:overview-labelling} - (1)) that recognised from 
-the process graph. Then, the modified docker
-image is executed in Condition 2 and each process is labeled right after
-the running by comparing the results obtained in Condition 1
-and output files of the process in Condition 2
-(Figure~\ref{fig:overview-labelling} - (2)). We then label processes in 2
-categories depending whether they
+two tested conditions (Figure~\ref{fig:overview-tool} - (4)). 
+The tool builds the Docker image for Condition 2 by modifying all
+the processes in the pipeline. Then, the modified Docker
+image is executed in Condition 2 and each process is labeled after
+running by comparing the results obtained in Condition 1
+and output files of the process in Condition 2. We label processes in 2
+categories depending on whether they
 create differences (represented in red) or not (green). 
 The tool replaces output file with the same file obtained in Condition 1 
-if process creates differences (Figure~\ref{fig:overview-labelling} - (3)).
+if the process creates differences.
 Process labelling
-is done through hooks which are defined during the docker image modifying
-procedure, and incremental updates in Condition 2, until the end of the
-pipeline execution.
+is done through hooks which are defined during the Docker image modifying
+procedure in Condition 2.
+Spot tool is available at https://github.com/big-data-lab-team/repro-tools.
 
@@
   \centering
-  \begin{subfigure}{\columnwidth}
-    \centering
-    \includegraphics[width=1\columnwidth]{images/fig1-a}
-    \caption{Identification of temporary and multi-write files.
-    (\href{https://docs.google.com/drawings/d/1rsbFwPjNPNMRhXUl4RSK3G0OaiJC7K-buiHN0hPD8eQ/edit?usp=sharing}{Source})}
-    \label{fig:overview-capturing}
-  \end{subfigure}
-   \begin{subfigure}{\columnwidth}
-    \centering
-     \includegraphics[width=1\columnwidth]{images/fig1-b}
-     \caption{Labelling process graphs including intermediary files.
-     (\href{https://docs.google.com/drawings/d/1yblYuKWAD18aJe5JxBu2h1EysyoQA3YrsQj-1T3s0l8/edit?usp=sharing}{Source})}
-     \label{fig:overview-labelling}
-   \end{subfigure}
-   \caption{An overview of Spot tool in two steps. First capturing intermediary files (a); then
-            labelling process graphs (b).}
-   \label{fig:overview}
+    \includegraphics[width=1\columnwidth]{images/overview-fig}
+    \caption{An overview of Spot tool in four steps.
+    (\href{https://encsconcordiaca-my.sharepoint.com/:p:/g/personal/m_alari_encs_concordia_ca/EYx_aIZgYKlDuuXKPKi5ckIBo_WmJeE1Czr9rqG4gLHOOA?e=H8gack}{Source})}
+    \label{fig:overview-tool}
   \end{figure}
@@
 and \texttt{fork()} dependencies. Our tool also creates a process \emph{graph} 
-(Figure~\ref{fig:overview-capturing} - (1)) from this tree
+from this tree
 by adding edges corresponding to file dependencies between processes. A
@@
 Listing~\ref{listing:sample-script}.
+
+Different subjects may lead to different process trees, due to
+variations that may be coming from differences in data types or
+cardinality. For instance, some images may have been acquired multiple
+times in some subjects. Some of these differences can be neglected, for
+instance, when a data decompression step is present at the beginning of the
+execution for some subjects only, and other ones cannot, when different
+processing paths are used in different subjects.
+Accordingly, we cluster subjects into different types.
+
 \begin{listing}
@@
 
-# Run FSL bet, put result ${bet_output}
+# Run FSL bet, put result in ${bet_output}
 bet "${input_image}" "${bet_output}" > bet_temp.out
@@
 
+\subsection{Subject clustering}
+
+We cluster subjects into different types to specify how many different 
+process trees are generated. 
+Nodes in the process trees are labeled by the process name.
+The tree edit distance~\cite{zhang1989simple} is used as a similarity measure for 
+representing the distance between labeled trees. 
+Tree edit distance  is defined  
+as the minimum number of edit operations which is needed to transform one tree into 
+the other. Three edit operations are considered: node label modification, 
+node removal, and node insertion. Each operation has an associated cost of 1.
+With this distance, we cluster subjects using agglomerative hierarchical
+clustering as implemented in SciPy~\cite{oliphant2007scipy}.
+As a result, the distance between
+any two subjects in a cluster is lower than a threshold parameter.
+
+We will show that different types of subjects certainly produce different results, 
+while subjects in the same type can be either reproducible or not.
+
 \subsection{Graph analysis}
 
-We label in two categories the processes of the graph of each subject. First,
+We clustered subjects into different types that each one refers to a specific graph.
+The processes of the graph of each subject are labeled into two categories. First,
 processes that read files that do not have differences and write files that
@@
 (1) check that processes in \textbf{P} do not write concurrently to F, 
-(2) we establish an order on \textbf{P} based on the creation timestamp 
-of the processes, (3) we replace every process $P_{i}$ in \textbf{P} by 
+(2) establish an order on \textbf{P} based on the creation timestamp 
+of the processes, (3) replace every process $P_{i}$ in \textbf{P} by 
 a wrapper that first calls $P_{i}$ and then saves F to a read-only 
@@
 happen in practice unless A and B were running concurrently, which we 
-assume is not the case (the tool checks for that).
+assume is not the case.
 
-As an example of this capturing process, in the sample pipeline in 
-Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} writes a temporary file and 
-also file \emph{T1w\_brain\_bin.nii.gz} is first written by process \texttt{2\#bet2} and 
-then overwritten by process \texttt{3\#fslmaths}. The list of these processes is captured in 
+As an example, in the sample pipeline in Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} writes a temporary file.
+Also, file \emph{T1w\_brain\_bin.nii.gz} is first written by \texttt{2\#bet2} and then overwritten by process \texttt{3\#fslmaths}.
+The complete list of the processes that write intermediary files is captured in 
 Listing~\ref{listing:json_process}.
-Using this list, we modify relevant processes in Condition 1 and then re-execute pipeline 
+Using this list, we modify processes in Condition 1 and then re-execute pipeline 
 to save them in a backup folder.
-Figure~\ref{fig:overview-capturing} - (4) shows pipeline results including intermediary files 
-captured as the outputs in Condition 1.
 
@@
 
-Since all the result files are captured in Condition 1, the tool 
-is ready to label the piepeline processes based on the results in Condition 2.
+When all the result files are captured in Condition 1, the tool 
+is ready to label the pipeline processes in Condition 2.
 However, another challenge is to label undetermined processes that read files with 
-differences and write files with differences too. In this
-situation, it is not possible to determine from the pipeline results 
-whether the process created differences or 
-was transparent.
+differences and write files with differences too. 
+In this case, it is not possible to determine from the pipeline results 
+whether the process creates differences or 
+is transparent.
 To address this issue, we set hooks in the pipeline 
 in Condition 2. Hooks are set by replacing the pipeline 
-processes with a custom script. This custom script labels each process 
-as \emph{transparent} or \emph{create differences} immediately after running 
+processes with a wrapper script.
+The replacement is done by modifying the PATH environment variable, as explained before. 
+The wrapper script labels each process 
+as \emph{transparent} or \emph{create differences} after running 
 the original executable. 
-The replacement is done through the PATH variable, as before. Moreover, 
-this custom script copies the results obtained from the same process in 
+Then, it copies the results obtained from the same process in 
 Condition 1 to the pipeline output in Condition 2 if 
-it is invoked with the arguments of a process that created differences 
-(Figure~\ref{fig:overview-labelling} - (3)). 
-Differences are identified by comparing checksum of the output files for 
-the same process in both conditions (Figure~\ref{fig:overview-labelling} - (2)).
-In addition, the distance functions that are specific to file types can be added 
-to our tool to measure differences more precisely. 
+it is invoked with the arguments of a process that created differences.
+Differences are identified by comparing the checksum of the output files for 
+the same process in both conditions.
+Besides, to measure the magnitude of the differences, we can add distance functions specific to file types.
 We developed an incremental approach that consists of the following steps:
@@
   \item Label each process as \emph{transparent} or \emph{create differences} 
-        right after its running.
+        after running.
   \item Check if process is \emph{create differences} then: 
@@
 which highlights processes that create differences.
-(Figure~\ref{fig:overview-labelling} - (4)).
 
@@
   \includegraphics[width=\columnwidth]{images/iterative_modif}
-  \caption{A simple example of steps of the proposed approach.}
+  \caption{Illustration of the labelling approach.}
   \label{fig:iterations}
@@
 step, processes that created differences are shown in red and other processes 
-(transparent processes) are in green. As in 
-Figure~\ref{fig:simple_script}, plain black edges represent the process 
-tree and dashed edges represent file dependencies: green edges 
-represent files with no differences, while red edges represent files with 
-differences. Temporary files are not represented because they have been 
-saved as previously described.
+(transparent processes) are in green. Temporary files are not represented 
+because they have been saved as previously described.
 
@@
 At step 3, after running process \texttt{3\#fslstats}, it labeled as 
-transparent as it write files without differences.
+transparent as it writes files without differences.
 At step 4, process \texttt{4\#fslmaths} is labeled as difference creator 
@@
 
-\subsection{Subject clustering}
-
-Different subjects may lead to different process tree, due to
-variations that may be coming from differences in data types or
-cardinality. For instance, some images may have been acquired multiple
-times in some subjects. Some of these differences can be neglected, for
-instance when a data decompression step is present at the beginning of the
-execution for some subjects only, and other ones cannot, when different
-processing paths are used in different subjects.
-
-Different process tree shows that pipeline execution followed by a different procedure. 
-This implies the possible impact of different subjects on pipeline results.
-We cluster subjects into different types to specify how many different 
-process trees are generated. Clustering is performed based on the process trees 
-which nodes are labeled by the process name. 
-In this order, 
-the tree edit distance~\cite{zhang1989simple} is used as a similarity measure for 
-representing the distance between labeled trees. Tree edit distance  is defined  
-as the minimum number of edit operations which is needed to transform one tree into 
-the other. Three edit operations are considered: node label modification, 
-node removal, and node insertion. Each operation has an associated cost of 1.
-With this distance, we cluster subjects using agglomerative hierarchical
-clustering as implemented in SciPy~\cite{oliphant2007scipy}
-(Algorithm~\ref{algo:hclustering}). As a result, the distance between
-any two subjects in a cluster is lower than a threshold parameter.
-
-Furthermore, we will show that even subjects in the same cluster, 
-having the same topological process tree structure, 
-may have different labels as processes that create differences.
-
-\begin{algorithm}[h!]
-\caption{Hierarchical clustering algorithm from SciPy}
-\label{algo:hclustering}
-\begin{algorithmic}
-
-  \State /* Input: - a list of n trees; \texttt{T = [t${_1}$, t${_2}$,...,t${_n}$]}
-  \State /*\quad \quad \quad \quad - a distance threshold value; \texttt{threshold}
-  \State /* Output: a list of clusters; \texttt{C = [c${_1}$,c${_2}$,...,c${_k}$]}
-  \State \texttt{\# Create a cluster for each tree in input list;}
-  \State \texttt{C = [ [t${_1}$], [t${_2}$], \ldots , [t${_n}$] ]}
-  \State \texttt{\# Define a 2 $\times$ 2 array to store distances}
-  \State \texttt{D} = \texttt{array(n, n)}
-  \While{\texttt{len(C)} > \texttt{1}}  
-  \State \texttt{\# Select the two nearest clusters}
-  \For{\texttt{i=1} to \texttt{n}}
-  \For{\texttt{j=i+1} to \texttt{len(C)}}
-  \State \texttt{D[i][j]} = $\min \left\{ \texttt{edit\_dist(a, b)}: \ \texttt{a} \in \texttt{C[i]}, \texttt{b} \in \texttt{C[j]} \right\}$
-    \EndFor
-  \EndFor
-  \State \texttt{i, j} = $\argmin \left\{ \texttt{D[i][j]}, \ \texttt{i, j} \in \llbracket 1, \texttt{len(C)}\rrbracket^2 \right\}$
-  \If{\texttt{D[i][j]} > \texttt{threshold}}
-  \State \Return \texttt{C}
-  \Else
-  \State \texttt{\# Merge C[j] into C[i]}
-  \State \texttt{C[i]} = \texttt{merge}(\texttt{C[i] and C[j]})
-  \State \texttt{\# Remove C[j] from C}
-  \State \texttt{remove(C[j], C)}
-  \EndIf
-  \EndWhile
-%  \State Function edit\_distance (C${_1}$, C${_2}$)
-%  \State EndFunction
-\end{algorithmic}
-\end{algorithm}
-
 \section{Experiments}
 
-In this experiment, we analyse the numerical reproducibility of the computational pipelines 
+In this experiment, we analyse the numerical reproducibility of computational pipelines 
 and identify the origin of differences across the operating systems. 
@@
 
-This section describes datasets and pipelines used for the analyses, 
-how data processed through the pipelines, and the results.
+This section describes datasets and pipelines used for the analyses, and the way of data processing.
 
@@
 Project (\href{https://www.humanconnectome.org}{HCP}).
-The HCP initiative is a significant effort to acquire and analyze 
+The HCP initiative is an effort to acquire and analyse 
 brain connectivity data from 1200 healthy adults.
@@
 individual behaviors. This helps to understand a wide range of brain disorders.
-The HCP project provides (1) database services (ConnectomeDB) for storing and 
-sharing primary and processed data freely, and (2) data analysis pipelines that 
+The HCP project provides database services (ConnectomeDB) for storing and 
+sharing primary and processed data freely, and data analysis pipelines that 
 are available under an open-source license.
@@
 structural imaging (sMRI), functional imaging (fMRI), and diffusion imaging (dMRI).
-The structural images include T1-weighted (T1w) and T2-weighted (T2w) images which 
-help in the diagnosis of the brain injury.
+The structural images include T1-weighted (T1w) and T2-weighted (T2w) images, which 
+help in the diagnosis of brain injury.
 The images are indexed with a suffix if several scans of the same modality were acquired.
@@
 
-Data from 1200 subjects 
-is available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
 In this paper, we randomly selected 20 unprocessed subjects 
-containing the 3-tesla type of MRI scanners from the HCP data release S500.
+containing 3-tesla type of MRI scanners from the HCP data release S500 
+available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}.
 
@@
 results were used as the input of FreeSurfer.
-
 Each subject was processed twice on the same operating system to detect
@@
 Within-OS variability may also be coming from other factors than pseudo-randomness. 
-For instance, metadata may differ due to the presence of timestamps, 
+For instance, the same files may have different metadata such as timestamps, 
 file paths, and other execution-specific information. 
-To ensure that differences belong to the file data part, 
-we added distance functions specific to the file types.
-These functions are \emph{mri\_diff}, \emph{mris\_diff}, and \emph{lta\_diff} 
-from FreeSurfer tool for comparing volumes, surface-based files, 
-and lta files (transforms) respectively. 
+To ensure that the differences are not related to the metadata, 
+we used specific functions from FreeSurfer tool including 
+\emph{mri\_diff}, \emph{mris\_diff}, and \emph{lta\_diff}.
 These functions compare files based on the image dimensions, 
-pixel intensities, the number of vertices, and geometry information
+voxel intensities, the number of vertices, and geometry information
 (see~\cite{fischl2012freesurfer} for more details). 
-Moreover, we used the Dice coefficient metric to evaluate the similarity of 
+In addition, we used the Dice coefficient metric to evaluate the similarity of 
 the image segmentation results, which is defined by
 \[DICE=\frac{2(|X|\cap|Y|)}{|X|+|Y|}\]
-The coefficient between 0 to 1, 1 means that two images are totally identical, while if 
-X and Y have no voxels in common, it is equal to 0.
+The Dice coefficient ranges from 0, indicating X and Y have no voxels in common, to 1, if 
+two images are identical.
 
-Furthermore, we clustered subjects into different type using 
-the clustering method as described before.
-We set the threshold value to zero for the clustering method. 
-This turns to cluster the subjects with the same topology into the same type.
+Finally, to cluster the subjects, the threshold value is set to zero in the clustering method. 
 Also, we used the nearest neighbor algorithm to calculate the distance between clusters.
@@
 
-We didn't observe any within-OS difference in PreFreeSurfer, however, we
+We did not observe any within-OS difference in PreFreeSurfer, however, we
 identified two FreeSurfer processes leading to within-OS differences due
@@
 
-Results show the variability between processes when 
-the same tool behaves differently in different process instances. For example, 
+The results show two types of differences, between-process and between-subject.
+Between-process differences occur when the same tool behaves differently in different process instances.
+For example, 
 FLIRT tool always produces files without differences in \emph{AAve} sub-pipeline but 
 mostly creates different files in \emph{AAli} sub-pipeline.
-
-Furthermore, different frequencies show the variability between subjects.
-Results show that the same process can be reproducible for some subjects but not for others. 
-Between subject differences can be a result of different subject types. 
-For instance, the subjects contain two T1 images need more calculations 
-to produce an average image between them.
+Also, different frequencies show between-subjects variability.
+Accordingly, the same process can be reproducible for some subjects but not for others. 
 The similar findings in another study~\cite{bowring2019exploring} were reported 
-the effect of subject type on reproducibility.
+the effect of between-subject differences on reproducibility.
+% Between subject differences can be a result of different subject types. 
+% For instance, the subjects contain two T1 images need more calculations 
+% to produce an average image between them.
 
@@
   \includegraphics[width=\textwidth]{images/pfs_heatmap.png}
-  \caption{Heatmap frequency of processes along with the sub-pipelines that introduce 
-           differences among 20 subjects. 
+  \caption{Frequency heatmap of processes in the tools that introduce differences among 20 subjects. 
            Each column represents the pipeline tool, and each row shows the sub-pipeline.
-           Each cell may contain several process instances of a tool in a sub-pipeline.
+           Each cell may contain several instances of the process of a tool in a sub-pipeline.
            Annotations describe the frequency distribution of the instances as ``N [M] x Y" 
-           in which N is the percentage of the process 
-           that creates differences out of all M times that is occurred in the sub-pipeline. 
-           Also, times Y shows the number of instances with the same distribution.
-           For example, the cell of FLIRT tool in sub-pipeline AAve shows 10 process instances  
-           with the frequency of 10 that introduce no differences.
+           in which N is the proportion of the process 
+           that creates differences out of all M times that is occurred. 
+           Also, times Y shows the number of processes with the same distribution.
+           For example, the cell of FLIRT tool in the sub-pipeline AAve shows 10 instances of the process 
+           with the frequency of 11 that introduce no differences.
            The range between red and green colors shows the highest and lowest 
            proportion of occurrences of the processes with differences.
-           Also, grey cells indicate that the pipeline tool is not executed in the sub-pipeline.}
+           Also, grey cells indicate that the tool is not executed in the sub-pipeline.}
   \label{fig:pfs_freq}
@@
 Figure~\ref{fig:fnirt_result} 
-shows the checkerboard images from the results of ACPC-Alignment in PreFreeSurfer. 
-Each rectangle includes the replaced pixels with different intensity values between images.
+shows checkerboard image from the results of ACPC-Alignment in PreFreeSurfer. 
+Each rectangle includes the replaced voxels with different intensity values between images.
 The misaligned regions visually represent significant differences in the alignment 
@@
     (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
-    CentOS7). Sagittal (left) and coronal (right) views of the brain are in dimension of X=130, Y=155, Z=130.
+    CentOS7).
     \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
@@
 
+We processed 20 subjects using FreeSurfer and 
 Spot tool identified \emph{mris\_make\_surfaces} as the only process in 
-the pipeline that creates differences in 10 subjects among 20 subjects.
+the pipeline that creates differences in 10 subjects.
 This process generates surface files for cortical and white matter surfaces 
@@
 Differences are mostly localized at the interfaces between tissues.
-In our dataset, we found subjects with the min and max Dice coefficient 
-values of 0.89 and 0.96, respectively.
+In our dataset, we found subjects with min Dice value of 0.89 and max Dice  
+value of 0.96.
 
-Furthermore, Figure~\ref{fig:scatter_plot} plots the Dice coefficients 
-for the 44 segmented regions obtained on the two operating systems.
-High Dice coefficients (more than 0.95) in most of the regions indicate minor differences.
-Dice values can be very low, down to 0.5, in particular in small structures 
-such as the vessels and lateral ventricles.
-
-We formulated the linear regression model for the Dice and region size variables as:
-\[D=1 + T\epsilon ;    T=\frac{-1}{\overline{R}}\]
-It is easier to obtain high overlaps with large regions, 
-while smaller regions are harder to be similarly segmented.
-We obtained the epsilon value of $\approx$ 8.3 voxels ($\approx$ 3mm) which indicate the minimum number
-of voxels in each region that needs to be segmented.
+Furthermore, the scatter plots of the Dice coefficients of the 44 segmented regions obtained 
+from two operating systems are shown in Figure~\ref{fig:scatter_plot}.
+This figure shows high Dice coefficients (more than 0.95) in large regions and very low 
+Dice coefficients (below 0.5) in small structures particularly in the vessels and lateral ventricles.
 
@@
           \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
-          {The link} to the animated version of segmentation results of one subject represents the significance of differences.} 
+          {The link} to the animated version of segmentation results of one subject presents more clear illustration of differences.} 
   \label{fig:tissue_class}
@@
             from 20 subjects in FreeSurfer (CentOS6 vs. CentOS7). 
-            The x-axes are the number of voxels in each region.
-            Right plot shows an overview of similarities between all regions and 
-            left plot represents a closer view of differences between 
+            The x-axes are the number of voxels in each region and 
+            y-axes are Dice values that are in different scales in two plots.
+            The right plot shows the similarity between regions larger than $10^4$ voxels. 
+            The left plot represents the differences between 
             regions smaller than $10^4$ voxels.
@@
   \end{figure*}
+
+  % We formulated the linear regression model for the Dice and region size variables as:
+  % \[D=1 + T\epsilon ;    T=\frac{-1}{\overline{R}}\]
+  % It is easier to obtain high overlaps with large regions, 
+  % while smaller regions are harder to be similarly segmented.
+  % We obtained the epsilon value of $\approx$ 8.3 voxels ($\approx$ 3mm) which indicate the minimum number
+  % of voxels in each region that needs to be segmented.
 
@@
 the tool in two operating systems, but we found a different number of calls to the \emph{glibc}.
-More investigations need to stabilize registration processes.
 
@@
 the segmented regions.
-We demonstrate that statically builds of the pipeline would not improve the numerical stability of the pipeline. 
+We demonstrate that static build of the pipeline would not improve the numerical stability of the pipeline. 
 FreeSurfer pipeline propagates and amplifies the numerical differences that are created by the 

### 2020-02-16 paper.tex
Updates

--- 2019-12-16 paper.tex
@@
 
-Operating systems and other computational factors are known to have an
-effect on the results produced by neuroimaging pipelines, presumably due to
-the creation, propagation and amplification of numerical errors along data
-analysis pipelines. However, the precise causes of such instabilities and
-the path along which they propagate in the pipelines are unclear, while
-this process could play a major role in the current reproducibility crisis.
-We present a technique to identify the computational processes in a
-pipeline that create numerical differences when executed in different
-computational conditions. Our technique leverages system-call interception to reconstruct
-processing and file graphs without pipeline instrumentation. Applying our tool to
-the structural pre-processing pipelines released by the Human Connectome
+Data analysis pipelines are known to be impacted by operating system
+updates and other infrastructural factors, presumably due to the creation,
+propagation and amplification of numerical errors. However, the precise
+causes of such instabilities and the path along which they propagate in the
+pipelines are unclear, while this process could play a major role in the
+current reproducibility crisis. We present a technique to identify the
+computational processes in a pipeline that create numerical differences
+when executed in different computational conditions. Our technique
+leverages system-call interception to reconstruct processing and file
+graphs without pipeline instrumentation. By applying our tool to the
+structural pre-processing pipelines released by the Human Connectome
 Project, we find that linear and non-linear registration are the cause of
@@
 instance, the fMRIprep pipeline~\cite{esteban2019fmriprep} assembles software blocks
-from FSL, AFNI, Freesurfer and ANTs to provide a state-of-the art
+from FSL~\cite{jenkinson2012fsl}, AFNI~\cite{cox2012afni}, Freesurfer~\cite{fischl2012freesurfer} and ANTs~\cite{avants2009advanced} to provide a state-of-the art
 functional MRI processing tool with minimal user input. Another example are the pipelines of the Human

### 2020-02-24 paper.tex

--- 2020-01-06 paper.tex
@@
 
-Operating systems and other computational factors are known to have an
-effect on the results produced by neuroimaging pipelines, presumably due to
-the creation, propagation and amplification of numerical errors along data
-analysis pipelines. However, the precise causes of such instabilities and
-the path along which they propagate in the pipelines are unclear, while
-this process could play a major role in the current reproducibility crisis.
-We present a technique to identify the computational processes in a
-pipeline that create numerical differences when executed in different
-computational conditions. Our technique leverages system-call interception to reconstruct
-processing and file graphs without pipeline instrumentation. By applying our tool to
-the HCP pre-processing pipelines, 
-we find that linear and non-linear registration are the cause of
+Data analysis pipelines are known to be impacted by operating system
+updates and other infrastructural factors, presumably due to the creation,
+propagation and amplification of numerical errors. However, the precise
+causes of such instabilities and the path along which they propagate in the
+pipelines are unclear, while this process could play a major role in the
+current reproducibility crisis. We present a technique to identify the
+computational processes in a pipeline that create numerical differences
+when executed in different computational conditions. Our technique
+leverages system-call interception to reconstruct processing and file
+graphs without pipeline instrumentation. By applying our tool to the
+structural pre-processing pipelines released by the Human Connectome
+Project, we find that linear and non-linear registration are the cause of
 most numerical instabilities in such pipelines, which confirms previous

### 2020-02-24 paper.tex
Working on the method

@@
 \newcommand{\english}[1]{\uwave{#1}}
+\newcommand{\toolname}[0]{Spot\xspace}
 
@@
 
-Spot identifies processes that create differences from a process graph of the pipeline execution.
-This tool compares file dependencies through all the processes based on the result files in two conditions.
-In this paper, different conditions refer to various operating systems.
-This tool is applicable on the complex pipelines in which tracing a large number of processes is difficult.
-
-Figure~\ref{fig:overview-tool} shows an overview of our method. For each tested
-condition, Condition 1 and Condition 2, the pipeline is containerized with
-Docker, and its interface is described with
-Boutiques~\cite{glatard2017boutiques}. 
-\reprozip is added to Condition 1 for provenance tracking. 
-This captures process dependencies through system call interceptions, 
-then Spot tool creates related process trees (Figure~\ref{fig:overview-tool} - (1)).
-Different subjects produce different process trees. 
-Hierachical clustering used to group subjects into different types (Figure~\ref{fig:overview-tool} - (2)).
-The rest of the procedure is done separately on each subject type.
-Our tool continues by
-restoring temporary files and ``multi-write" files, i.e., files written
-by more than one process (Figure~\ref{fig:overview-tool} - (3)). 
-To do so, the tool creates a list of processes 
-that introduce intermediary files and replaces their corresponding executables with a wrapper script. 
-The tool re-executes the pipeline in a modified Docker image in Condition 1.
-This copies files in the list to a read-only directory.
-In the next step, the
-processes are labeled from the differences found in their outputs in the
-two tested conditions (Figure~\ref{fig:overview-tool} - (4)). 
-The tool builds the Docker image for Condition 2 by modifying all
-the processes in the pipeline. Then, the modified Docker
-image is executed in Condition 2 and each process is labeled after
-running by comparing the results obtained in Condition 1
-and output files of the process in Condition 2. We label processes in 2
-categories depending on whether they
-create differences (represented in red) or not (green). 
-The tool replaces output file with the same file obtained in Condition 1 
-if the process creates differences.
-Process labelling
-is done through hooks which are defined during the Docker image modifying
-procedure in Condition 2.
-Spot tool is available at https://github.com/big-data-lab-team/repro-tools.
+Figure~\ref{fig:overview-tool} shows an overview of our method. The output
+of 2 computational conditions is compared through 3 main steps. At step 1,
+we build a low-level provenance graph for the processing of each subject in
+Condition 1, where nodes are OS processes created by the application and
+edges are file dependencies. Such provenance graphs reconstruct a model of
+the application workflow, regardless of the original programming language
+and dependencies. They may, however, slightly differ across the processed
+subjects since different execution paths might be followed due to
+variations in the data. To address this issue, we cluster subjects based on
+their provenance graph (step 2), resulting in a reduced set of $k$
+representative provenance graphs. These $k$ graphs and their corresponding
+subjects /TODO: ???/ are then used for process labeling (step 3), where
+Condition 2 is executed step by step, comparing the files produced by each
+process to their counterparts in Condition 1, making sure that temporary
+files and files written by more than a single process are correctly saved.
 
@@
 
-We use \reprozip 
+We use the \reprozip tool~\cite{rampin2016reprozip}
 to capture: (1) the set of processes created by the
@@
 
-Our tool reconstructs a \emph{process tree} starting from the first process
+\toolname reconstructs a \emph{process tree} starting from the first process
 created by the pipeline and traversing processes through \texttt{clone()}
-and \texttt{fork()} dependencies. Our tool also creates a process \emph{graph} 
-from this tree
-by adding edges corresponding to file dependencies between processes. A
+and \texttt{fork()} dependencies. It also creates a process \emph{graph} 
+from this tree by adding edges corresponding to file dependencies between processes. A
 file dependency is defined between processes A and B if a file written by A
@@
 
-Different subjects may lead to different process trees, due to
-variations that may be coming from differences in data types or
-cardinality. For instance, some images may have been acquired multiple
-times in some subjects. Some of these differences can be neglected, for
-instance, when a data decompression step is present at the beginning of the
-execution for some subjects only, and other ones cannot, when different
-processing paths are used in different subjects.
-Accordingly, we cluster subjects into different types.
-
 \begin{listing}
-\begin{minted}[frame=single,
-  framesep=3mm,
-  linenos=false,
-  xleftmargin=0pt,
-  tabsize=4]{bash}
-#!/bin/bash
-
-# This script extracts the brain from an input image,
-# measures the volume of the extracted brain,
-# and creates a binarized brain mask.
-
-if [ $# != 1 ]
-then
-    echo "usage: $0 <inputimage.nii.gz>"
-    exit 1
-fi
-
-# Parse argument, set file names
-input_image=$1
-bet_output=$(basename "${input_image}" \
-                       .nii.gz)_brain.nii.gz
-bet_output_binarized=$(basename "${input_image}" \
-                       .nii.gz)_brain_bin.nii.gz
-
-# Run FSL bet, put result in ${bet_output}
-bet "${input_image}" "${bet_output}" > bet_temp.out
-echo "Voxels / Volume in brain mask:"
-# Run FSL stats on ${bet_output}
-fslstats "${bet_output}" -V
-# Run FSL maths on ${bet_output}
-fslmaths "${bet_output}" "${bet_output_binarized}"
-# Remove temporary file
-\rm bet_temp.out
-\end{minted}
+  \inputminted{bash}{"bin/example.sh"}
   \caption{Example pipeline}
@@
 
-We cluster subjects into different types to specify how many different 
-process trees are generated. 
-Nodes in the process trees are labeled by the process name.
-The tree edit distance~\cite{zhang1989simple} is used as a similarity measure for 
-representing the distance between labeled trees. 
-Tree edit distance  is defined  
-as the minimum number of edit operations which is needed to transform one tree into 
-the other. Three edit operations are considered: node label modification, 
-node removal, and node insertion. Each operation has an associated cost of 1.
-With this distance, we cluster subjects using agglomerative hierarchical
-clustering as implemented in SciPy~\cite{oliphant2007scipy}.
-As a result, the distance between
-any two subjects in a cluster is lower than a threshold parameter.
+Different subjects may lead to different provenance graphs, due to
+variations that may be coming from differences in data types or
+cardinality. For instance, some images may have been acquired multiple
+times in some subjects, for noise-reduction purposes. Some of these
+differences can be neglected, for instance, when a data decompression step
+is present at the beginning of the execution for some subjects only, but
+other differences cannot be ignored, for instance when complete different
+processing paths are used in different subjects.
 
-We will show that different types of subjects certainly produce different results, 
-while subjects in the same type can be either reproducible or not.
+To reduce the number of provenance graphs to be examined, we cluster
+process trees using agglomerative hierarchical clustering, as implemented
+in SciPy~\cite{oliphant2007scipy}. We use the tree edit
+distance~\cite{zhang1989simple} between process trees, as implemented in
+the zss Python package. This distance is defined as the minimum number of
+edit operations to transform one tree into the other. Three edit operations
+are considered: node label modification, node removal, and node insertion.
+Each operation has an associated cost of 1. 
+//If threshold is 0, why do we even need the edit distance?//
 

### 2020-02-28 paper.tex
Updates in methodo

@@
 
+\todo{explain the issue with propagating differences and with transient files}
+
 % % Verificarlo and MCA, require recompilation
@@
 
-Figure~\ref{fig:overview-tool} shows an overview of our method. The output
-of 2 computational conditions is compared through 3 main steps. At step 1,
-we build a low-level provenance graph for the processing of each subject in
-Condition 1, where nodes are OS processes created by the application and
-edges are file dependencies. Such provenance graphs reconstruct a model of
-the application workflow, regardless of the original programming language
-and dependencies. They may, however, slightly differ across the processed
-subjects since different execution paths might be followed due to
-variations in the data. To address this issue, we cluster subjects based on
-their provenance graph (step 2), resulting in a reduced set of $k$
-representative provenance graphs. These $k$ graphs and their corresponding
-subjects /TODO: ???/ are then used for process labeling (step 3), where
-Condition 2 is executed step by step, comparing the files produced by each
-process to their counterparts in Condition 1, making sure that temporary
-files and files written by more than a single process are correctly saved.
+The steps implemented in \toolname to compare pipeline executions in
+different conditions are outlined in Figure~\ref{fig:overview-tool}. First,
+in Condition 1, a directed, rooted, bipartite provenance graph is built for
+each subject, where nodes represent application processes or files, and
+edges represent read or write file accesses (Step 1 in Figure
+Figure~\ref{fig:overview-tool}). \emph{Transient} files are then identified as
+ files that are either deleted during the execution, or modified by
+multiple processes (Step 2). \toolname re-runs the pipeline to save these files
+right after they are produced, resulting in a provenance \emph{DAG}
+(Directed Acyclic Graph) in which file nodes have a single parent (in-degree
+of 1) (Step 3). This DAG is then compared to the one produced from
+Condition 2, in a step-by-step process that prevents the propagation of differences.
+\toolname finally produces a labeled graph that identifies the
+non-reproducible processes of the pipeline.
 
@@
 
-\subsection{Provenance capture and representation}
+  \todo{review and summarize assumptions:
+  * No concurrent writes from multiple processes
+  * A given process cannot modify the bytes it wrote to a file
+  But we support non-concurrent writes by different processes. This is common in neuroimaging.
+  }
+
+\subsection{Recording of provenance graphs}
 
@@
 temporary files. \reprozip collects this information through the
-\texttt{ptrace()} system call and stores it in a SQLite database.
+\texttt{ptrace()} system call, with no required instrumentation of the pipeline.
+\toolname then reconstructs a provenance graph by creating process and file
+nodes from the \reprozip trace, and by adding directed edges corresponding
+to file reads and file writes (Figure~\ref{fig:simple_script}).
 
-\toolname reconstructs a \emph{process tree} starting from the first process
-created by the pipeline and traversing processes through \texttt{clone()}
-and \texttt{fork()} dependencies. It also creates a process \emph{graph} 
-from this tree by adding edges corresponding to file dependencies between processes. A
-file dependency is defined between processes A and B if a file written by A
-is read by B. Figure~\ref{fig:simple_script} shows an example of a process
-tree and graph constructed from the example pipeline in
-Listing~\ref{listing:sample-script}.
+\todo{Is every process linked to an executable?}
+
+Provenance graphs may be subject-dependent, which justifies iterating
+\toolname on multiple subjects. For instance, data acquisitions may be
+repeated a different number of times in each subject, for noise-reduction
+purposes. Some of these differences can be neglected, for instance when a
+data decompression step is present at the beginning of the execution for
+some subjects only, but other differences cannot be ignored, for instance
+when complete different processing paths are used in different subjects.
 
@@
   \includegraphics[width=0.3\columnwidth]{images/simple_graph}
-  \caption{Process tree and graph
+  \caption{Process graph
   constructed from the example pipeline in
@@
   was captured by \reprozip while it did not appear in
-  Listing~\ref{listing:sample-script}. 
+  Listing~\ref{listing:sample-script}. \todo{Tristan: check that, this is not a bipartite graph now}. 
 }
@@
 
-\subsection{Subject clustering}
+\subsection{Capture of transient files}
 
-Different subjects may lead to different provenance graphs, due to
-variations that may be coming from differences in data types or
-cardinality. For instance, some images may have been acquired multiple
-times in some subjects, for noise-reduction purposes. Some of these
-differences can be neglected, for instance, when a data decompression step
-is present at the beginning of the execution for some subjects only, but
-other differences cannot be ignored, for instance when complete different
-processing paths are used in different subjects.
+\toolname captures temporary files in Condition 1 by replacing every
+process P by a wrapper that first calls P and then saves the produced
+temporary files to a read-only directory. This process replacement is done
+without pipeline instrumentation, by pre-pending to the PATH environment
+variable a directory containing a wrapper script named as the executable
+called by the process (see Listing~\ref{listing:modify-script}).
 
-To reduce the number of provenance graphs to be examined, we cluster
-process trees using agglomerative hierarchical clustering, as implemented
-in SciPy~\cite{oliphant2007scipy}. We use the tree edit
-distance~\cite{zhang1989simple} between process trees, as implemented in
-the zss Python package. This distance is defined as the minimum number of
-edit operations to transform one tree into the other. Three edit operations
-are considered: node label modification, node removal, and node insertion.
-Each operation has an associated cost of 1. 
-//If threshold is 0, why do we even need the edit distance?//
-
-\subsection{Graph analysis}
-
-We clustered subjects into different types that each one refers to a specific graph.
-The processes of the graph of each subject are labeled into two categories. First,
-processes that read files that do not have differences and write files that
-do not have differences are labeled \emph{transparent}
-(Figure~\ref{fig:green}). Second, processes that read files that do not
-have differences but write files that have differences \emph{create}
-differences (Figure~\ref{fig:red}). In addition, processes that read files
-that have differences and write files that also have differences are called
-\emph{undetermined} (Figure~\ref{fig:grey}). Processes are also labeled
-undetermined when they read or write temporary files, or produce multi-write
-files. To resolve undetermined processes, hooks must be added to the pipeline to 
-execute these processes on files without differences, and to capture
-intermediary file states.
-
-\begin{figure}%\centering
-\centering
-    \begin{subfigure}{0.3\linewidth}
-      \centering
-        \includegraphics[scale=0.36]{images/green.png}
-        \caption{Transparent}
-        \label{fig:green}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.35\linewidth}
-      \centering
-    \includegraphics[scale=0.36]{images/red.png}
-    \caption{Creates differences}
-    \label{fig:red}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.3\linewidth}
-      \centering
-    \includegraphics[scale=0.36]{images/grey.png}
-    \caption{Undetermined}
-    \label{fig:grey}
-\end{subfigure}
-    \caption{Different type of labeled process based on the input/output files.
-  Dashed edges refer to the file dependencies between processes A and B 
-  if a file written by A is read by B.}
-    \label{fig:processes}
-\end{figure}
-
-\subsubsection{Capturing temporary files}
-
-The label of processes that read or 
-write temporary files is undetermined because these files are deleted during 
-the execution. To address this issue, we replace in Condition 1 every process P that 
-writes temporary files by a wrapper that first calls P and then
-saves all its output files to a read-only directory. This replacement is
-done by modifying the PATH environment variable to point to a directory
-containing a modified version of the executable called by the process. The
-modified executable will run the original executable and then save its
-output files (see Listing~\ref{listing:modify-script}). 
-This 
-solution does not cover the temporary files that are removed by P 
-itself, but this is not a problem since these files do not play any role in 
-the subsequent steps of the pipeline.
+% The temporary files that are removed by P 
+% itself won't be saved, but this is not a problem since these files do not play any role in 
+% the subsequent steps of the pipeline.
 
@@
   \end{minted}
-    \caption{Modified version of the script in Listing~\ref{listing:sample-script} to save intermediary files.}
+    \caption{Wrapper script used to save temporary and modified files.}
     \label{listing:modify-script}
@@
 
-\subsubsection{Capturing multi-write files}
+Files written by multiple processes are disambiguated as follows. For a
+ file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
+ $P_{n}$\}, \toolname first checks that processes in \textbf{P} do not
+ write concurrently to F, which would violate our assumptions. Then, it
+ replaces every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
+ calls $P_{i}$ and then saves F to a read-only directory. In this way, the
+ multiple versions of F are preserved for comparison. \toolname finally
+ updates the provenance graph accordingly, so that all files in the graph
+ have an in-degree of 1. This operation also makes the provenance graph
+ acyclic, since we assumed that a process cannot modify a file once it was
+ written \todo{check/prove that}.
 
-Files written by multiple processes also lead 
-to undetermined labels. For a file F 
-written by all processes in \textbf{P} = \{$P_{1}$, \ldots $P_{n}$\}, we 
-(1) check that processes in \textbf{P} do not write concurrently to F, 
-(2) establish an order on \textbf{P} based on the creation timestamp 
-of the processes, (3) replace every process $P_{i}$ in \textbf{P} by 
-a wrapper that first calls $P_{i}$ and then saves F to a read-only 
-directory. Thus, multiple versions of F are saved and used in the 
-analysis.
-
-Furthermore, cycles may be present in the process graph in case a file 
-was written by more than one process. We remove such cycles by removing 
-file edges between processes A and B when A's process creation 
-timestamp is posterior to B's or when A=B. Indeed, such edges cannot 
-happen in practice unless A and B were running concurrently, which we 
-assume is not the case.
-
-As an example, in the sample pipeline in Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} writes a temporary file.
-Also, file \emph{T1w\_brain\_bin.nii.gz} is first written by \texttt{2\#bet2} and then overwritten by process \texttt{3\#fslmaths}.
-The complete list of the processes that write intermediary files is captured in 
-Listing~\ref{listing:json_process}.
-Using this list, we modify processes in Condition 1 and then re-execute pipeline 
-to save them in a backup folder.
+As an example, in the sample pipeline in
+Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} writes a
+temporary file. Also, file \emph{T1w\_brain\_bin.nii.gz} is first written
+by \texttt{2\#bet2} and then overwritten by process \texttt{3\#fslmaths}
+\todo{it's not the case, check}. The complete list of the processes that
+write intermediary files is captured in Listing~\ref{listing:json_process}.
+Using this list, we modify processes in Condition 1 and then re-execute
+pipeline to save them in a backup folder.
 
@@
 
-\subsubsection{Labelling undetermined processes}
+\subsection{Process labelling}
 
-When all the result files are captured in Condition 1, the tool 
-is ready to label the pipeline processes in Condition 2.
-However, another challenge is to label undetermined processes that read files with 
-differences and write files with differences too. 
-In this case, it is not possible to determine from the pipeline results 
-whether the process creates differences or 
-is transparent.
-To address this issue, we set hooks in the pipeline 
-in Condition 2. Hooks are set by replacing the pipeline 
-processes with a wrapper script.
-The replacement is done by modifying the PATH environment variable, as explained before. 
-The wrapper script labels each process 
-as \emph{transparent} or \emph{create differences} after running 
-the original executable. 
-Then, it copies the results obtained from the same process in 
-Condition 1 to the pipeline output in Condition 2 if 
-it is invoked with the arguments of a process that created differences.
-Differences are identified by comparing the checksum of the output files for 
-the same process in both conditions.
-Besides, to measure the magnitude of the differences, we can add distance functions specific to file types.
-We developed an incremental approach that consists of the following steps:
+Once all the files have been captured in Condition 1, including transient
+files, \toolname labels pipeline processes through a step-by-step execution
+in Condition 2. Each process in Condition 2 is run and its output files are
+compared to the ones produced in Condition 1. If no differences are found,
+the process is marked as reproducible. Otherwise, the process is marked as
+irrreproducible and the output files produced in Condition 1 are copied to
+Condition 2, to ensure that differences do not propagate further in the
+pipeline. Processes are instrumented transparently, through a modification
+of the \texttt{PATH} variable similar to the one described previously.
 
-\begin{enumerate}
-  \item Start the pipeline execution in Condition 2; 
-        after hooking all the processes involved in the pipeline.
-  \item Label each process as \emph{transparent} or \emph{create differences} 
-        after running.
-  \item Check if process is \emph{create differences} then: 
-        copy the results produced by the same process in Condition 1 to the 
-        pipeline output in Condition 2. 
-  \item Continue steps 2 and 3 until the end of the pipeline execution.
-\end{enumerate}
+By default, differences in output files are identified by comparing file
+checksums. Other comparison functions can also be defined for specific file
+types, for instance to ignore sections of the file containing timestamps or
+other execution-specific information.
 
-This algorithm creates a \emph{labeled process tree} at the end of pipeline execution 
-which highlights processes that create differences.
+Finally, \toolname creates a labeled provenance graph, 
+highlighting non-reproducible processes.
 
@@
   \includegraphics[width=\columnwidth]{images/iterative_modif}
-  \caption{Illustration of the labelling approach.}
+  \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red. \todo{update with bi-partite graph}}
   \label{fig:iterations}
@@
 Figure~\ref{fig:iterations} illustrates our incremental labelling 
-process for the example in Figure~\ref{fig:simple_script}. At every 
-step, processes that created differences are shown in red and other processes 
-(transparent processes) are in green. Temporary files are not represented 
-because they have been saved as previously described.
-
-The four steps in Figure~\ref{fig:iterations} correspond to the 
-steps of the labelling algorithm. 
+process for the example in Figure~\ref{fig:simple_script}. 
 At step one, process \texttt{1\#bet} is executed and then labeled 
-as transparent (green) as it produced files without differences.
+as reproducible (green) as it produced files without differences.
 At step 2, \texttt{2\#bet2} 
-is labeled as difference creator (red) as it produced files with differences 
+is labeled as non-reproducible (red) as it produced files with differences 
 from files without differences. Therefore, the files produced by \texttt{2\#bet2} in  
@@
 Condition 1.
-At step 3, after running process \texttt{3\#fslstats}, it labeled as 
-transparent as it writes files without differences.
-At step 4, process \texttt{4\#fslmaths} is labeled as difference creator 
+At step 3, \texttt{3\#fslstats} is labeled as 
+reproducible as it writes files without differences.
+At step 4, \texttt{4\#fslmaths} is labeled as non-reproducible 
 as the last process of the pipeline.
-As a result of these 4 steps, the final process labelling is: 
-\texttt{2\#bet2} and \texttt{4\#fslmath} are difference creators (red) 
-and the other processes are transparent.
 
@@
 \subsection{Results}
+
+% \subsection{Subject clustering}
+
+% To reduce the number of provenance graphs to be examined, we cluster
+% process trees using agglomerative hierarchical clustering, as implemented
+% in SciPy~\cite{oliphant2007scipy}. We use the tree edit
+% distance~\cite{zhang1989simple} between process trees, as implemented in
+% the zss Python package. This distance is defined as the minimum number of
+% edit operations to transform one tree into the other. Three edit operations
+% are considered: node label modification, node removal, and node insertion.
+% Each operation has an associated cost of 1. 
+% //If threshold is 0, why do we even need the edit distance?//
 
@@
 
+Method:
+* we assume no concurrent file accesses.
+
+* we assume that the provenance graph is oriented, i.e., a process cannot read or write in the same file. Reasonnable in general but wouldn't work 
+for, say, out of core computations.
+
+Comment on the overhead, number of executions (3, 2 in condition 1, 1 in condition 2)
+
+Applicability: no instrumentation required, fast.
+But not able to capture dependencies in memory.
+
 We identified four processes that create differences 

### 2020-02-28 paper.tex

@@
 The steps implemented in \toolname to compare pipeline executions in
-different conditions are outlined in Figure~\ref{fig:overview-tool}. First,
+2 conditions are outlined in Figure~\ref{fig:overview-tool}. First,
 in Condition 1, a directed, rooted, bipartite provenance graph is built for
 each subject, where nodes represent application processes or files, and
-edges represent read or write file accesses (Step 1 in Figure
-Figure~\ref{fig:overview-tool}). \emph{Transient} files are then identified as
+edges represent read or write file accesses (Step 1 in Figure~\ref{fig:overview-tool}). \emph{Transient} files are then identified as
  files that are either deleted during the execution, or modified by
@@
 of 1) (Step 3). This DAG is then compared to the one produced from
-Condition 2, in a step-by-step process that prevents the propagation of differences.
+Condition 2, in a step-by-step process that prevents the propagation of differences (step 4).
 \toolname finally produces a labeled graph that identifies the
@@
   * A given process cannot modify the bytes it wrote to a file
+  * Process are command lines
   But we support non-concurrent writes by different processes. This is common in neuroimaging.
@@
 
-\subsection{Recording of provenance graphs}
+\subsection{Recording provenance graphs}
 
@@
 to capture: (1) the set of processes created by the
-pipeline, using the \texttt{clone()} or \texttt{fork()} system call, and
-(2) the set of files read, written and executed by each process, including
+pipeline, and
+(2) the set of files read and written by each process, including
 temporary files. \reprozip collects this information through the
@@
 
-Provenance graphs may be subject-dependent, which justifies iterating
-\toolname on multiple subjects. For instance, data acquisitions may be
+Provenance graphs may be subject-dependent, which justifies the processing
+of multiple subjects. For instance, data acquisitions may be
 repeated a different number of times in each subject, for noise-reduction
@@
 data decompression step is present at the beginning of the execution for
-some subjects only, but other differences cannot be ignored, for instance
+some subjects only, but other differences cannot, for instance
 when complete different processing paths are used in different subjects.
@@
 
-\subsection{Capture of transient files}
+\subsection{Capturing transient files}
 
@@
 temporary files to a read-only directory. This process replacement is done
-without pipeline instrumentation, by pre-pending to the PATH environment
+without pipeline instrumentation, by pre-pending to the \texttt{PATH} environment
 variable a directory containing a wrapper script named as the executable
-called by the process (see Listing~\ref{listing:modify-script}).
-
-% The temporary files that are removed by P 
-% itself won't be saved, but this is not a problem since these files do not play any role in 
-% the subsequent steps of the pipeline.
-
-\begin{listing}
-  \begin{minted}[frame=single,
-    framesep=3mm,
-    linenos=false,
-    xleftmargin=0pt,
-    tabsize=4]{python}
-  #!/usr/bin/env python
-  import os.path as op
-  import shutil
-  import pipes
-
-  # Run the original executable using the same arguments
-  executed_file = sys.argv[0]
-  exec_proc = op.join('original_scripts', executed_file)
-  exec_argv = " ".join(map(pipes.quote, sys.argv[1:]))
-  exec_cmd = exec_proc + exec_argv
-  subprocess.Popen(exec_cmd, shell=True).communicate()
-
-  # Load list of captured process to be checked
-  with open('path/to/p_list/', 'r') as p_list:
-      process_list = json.load(p_list)
-
-  if exec_cmd in process_list.keys():
-      # Save output files to read-only directory
-      p_files = process_list[exec_cmd]
-      shutil.copy(p_files, '/read-only/dir')
-
-  \end{minted}
-    \caption{Wrapper script used to save temporary and modified files.}
-    \label{listing:modify-script}
-  \end{listing}
+called by P.
 
@@
  replaces every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
- calls $P_{i}$ and then saves F to a read-only directory. In this way, the
- multiple versions of F are preserved for comparison. \toolname finally
+ calls $P_{i}$ and then saves F to a read-only directory. In this way,
+ successive versions of F are preserved for comparison. \toolname finally
  updates the provenance graph accordingly, so that all files in the graph
  have an in-degree of 1. This operation also makes the provenance graph
- acyclic, since we assumed that a process cannot modify a file once it was
- written \todo{check/prove that}.
+ acyclic, since we assumed that a process can only write a single time to a file.
 
@@
 
-\subsection{Process labelling}
+\subsection{Labelling processes}
 
 Once all the files have been captured in Condition 1, including transient
-files, \toolname labels pipeline processes through a step-by-step execution
+ones, \toolname labels pipeline processes through a step-by-step execution
 in Condition 2. Each process in Condition 2 is run and its output files are
@@
 the process is marked as reproducible. Otherwise, the process is marked as
-irrreproducible and the output files produced in Condition 1 are copied to
+non-reproducible and the output files produced in Condition 1 are copied to
 Condition 2, to ensure that differences do not propagate further in the
 pipeline. Processes are instrumented transparently, through a modification
-of the \texttt{PATH} variable similar to the one described previously.
-
-By default, differences in output files are identified by comparing file
+of the \texttt{PATH} variable similar to the one described previously. By
+default, differences in output files are identified by comparing file
 checksums. Other comparison functions can also be defined for specific file
 types, for instance to ignore sections of the file containing timestamps or
-other execution-specific information.
-
-Finally, \toolname creates a labeled provenance graph, 
-highlighting non-reproducible processes.
+other execution-specific information. Finally, \toolname creates a labeled
+provenance graph, highlighting non-reproducible processes.
 

### 2020-03-05 paper.tex
Update Figure-1
update example script, add new pg figures, add bin data samples

@@
 
-\begin{figure}
+\begin{figure*}
   \centering
-    \includegraphics[width=1\columnwidth]{images/overview-fig}
+    \includegraphics[width=.8\textwidth]{images/spot-diagram}
     \caption{An overview of Spot tool in four steps.
-    (\href{https://encsconcordiaca-my.sharepoint.com/:p:/g/personal/m_alari_encs_concordia_ca/EYx_aIZgYKlDuuXKPKi5ckIBo_WmJeE1Czr9rqG4gLHOOA?e=H8gack}{Source})}
+    (\href{https://docs.google.com/drawings/d/1OB3sB8kkK17Q516-TcmXKh3TwXSLDToJB11IVfUH5AU/edit?usp=sharing}{Source})}
     \label{fig:overview-tool}
-  \end{figure}
+  \end{figure*}
 
@@
 nodes from the \reprozip trace, and by adding directed edges corresponding
-to file reads and file writes (Figure~\ref{fig:simple_script}).
+to file reads and file writes (Figure~\ref{fig:simple_script}-left).
 
@@
 \centering
-  \includegraphics[width=0.3\columnwidth]{images/simple_graph}
-  \caption{Process graph
+  \includegraphics[width=1\columnwidth]{images/provenance-graphs}
+  \caption{Provenance graphs
   constructed from the example pipeline in
   Listing~\ref{listing:sample-script}.
-  Processes that read or write
-  temporary files are 
-  represented with squares. Plain edges 
+  Nodes are devided into two sets including processes that are shaped 
+  in the circles and files that are shaped in the boxes.
+  Transient files are represented with grey color boxes. 
+  Plain edges 
   represent the process tree (\texttt{fork()} or \texttt{clone()} 
-  system calls). Dashed edges represent file dependencies: temporary 
-  files are in purple and result files are in black.
-  Every node in the tree is labeled using (1) a process id created by our
-  reconstruction, (2) the name of the executable run by the process.
+  system calls). Dashed edges represent file dependencies.
+  Every node in the graph is labeled using (1) a process id created by our
+  reconstruction, (2) the name of the executable run by the process or the file name.
   Process 0 is the initial call to the \texttt{sh} interpreter, and
-  processes 1, 3 and 4 are the calls to FSL bet, stats and maths made in
-  Listing~\ref{listing:sample-script}. Process 2 is forked by process 1: it
-  was captured by \reprozip while it did not appear in
-  Listing~\ref{listing:sample-script}. \todo{Tristan: check that, this is not a bipartite graph now}. 
+  the rest of processes are the calls to FSL robustfov, bet, stats and maths made in
+  Listing~\ref{listing:sample-script}. Processes 75 and 149 are forked by processes 4 and 78, respectively: 
+  they were captured by \reprozip while they did not appear in
+  Listing~\ref{listing:sample-script}.  
 }
@@
  updates the provenance graph accordingly, so that all files in the graph
- have an in-degree of 1. This operation also makes the provenance graph
+ have an in-degree of 1 (Figure~\ref{fig:simple_script}-right). This operation also makes the provenance graph
  acyclic, since we assumed that a process can only write a single time to a file.
@@
 As an example, in the sample pipeline in
-Listing~\ref{listing:sample-script}, process \texttt{2\#bet2} writes a
-temporary file. Also, file \emph{T1w\_brain\_bin.nii.gz} is first written
-by \texttt{2\#bet2} and then overwritten by process \texttt{3\#fslmaths}
-\todo{it's not the case, check}. The complete list of the processes that
-write intermediary files is captured in Listing~\ref{listing:json_process}.
+Listing~\ref{listing:sample-script}, process \texttt{3\#robustfov} writes a
+temporary file. Also, file \emph{T1w\_brain.nii.gz} is first written
+by \texttt{75\#bet2} and then overwritten by process \texttt{149\#bet2}. 
+The complete list of the processes that
+write transient files is captured in Listing~\ref{listing:json_process}.
 Using this list, we modify processes in Condition 1 and then re-execute
@@
   "multi_write_process": {
-    "fslmaths\00T1w_brain.nii.gz\00T1w_brain_bin.nii.gz":
+    "bet2\00T1w\00T1w_brain":
     {
       "files":[
-        "/cond1/subj1/T1w_brain_bin.nii.gz"
+        "/cond1/subj1/T1w_brain.nii.gz"
       ],
-      "pid": 78
+      "pid": 75
     },
-    "bet2\00/cond1/subj1/T1w.nii.gz\00T1w_brain.nii.gz":
+    "bet2\00T1w_roi\00T1w_brain":
     {
       "files":[
-        "/cond1/subj1/T1w_brain_bin.nii.gz"
+        "/cond1/subj1/T1w_brain.nii.gz"
       ],
-      "pid": 75
+      "pid": 149
     }
@@
   "temp_process": {
-    "bet2\00/cond1/subj1/T1w.nii.gz\00T1w_brain.nii.gz": 
+    "robustfov\00-i\00T1w.nii.gz\00-r\00T1w_roi.nii.gz": 
     {
       "files": [
-        "/cond1/subj1/temp.tmp"
+        "/cond1/subj1/T1w_roi.nii.gz"
       ],
-      "pid": 75
+      "pid": 3
     } 
@@
 \end{minted}
-\caption{List of process and their files from the example pipeline in 
-Listing~\ref{listing:sample-script} that should be captured.}
+\caption{List of processes and transient files from the example pipeline in 
+Listing~\ref{listing:sample-script}.}
 \label{listing:json_process}
@@
 
-\begin{figure}
+\begin{figure*}
   \centering
-  \includegraphics[width=\columnwidth]{images/iterative_modif}
-  \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red. \todo{update with bi-partite graph}}
+  \includegraphics[width=.8\textwidth]{images/labelling-process}
+  \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red.}
   \label{fig:iterations}
-\end{figure}
+\end{figure*}
 
@@
 process for the example in Figure~\ref{fig:simple_script}. 
-At step one, process \texttt{1\#bet} is executed and then labeled 
+At step one, process \texttt{3\#robustfov} is executed and then labeled 
 as reproducible (green) as it produced files without differences.
-At step 2, \texttt{2\#bet2} 
-is labeled as non-reproducible (red) as it produced files with differences 
-from files without differences. Therefore, the files produced by \texttt{2\#bet2} in  
-Condition 2 are replaced with the files produced by \texttt{2\#bet2} in 
+At step 2, \texttt{75\#bet2} 
+is also labeled as reproducible as it produced files without differences. 
+At step 3 \texttt{149\#bet2} is labeled as non-reproducible (red) as it produced files with differences 
+from files without differences. Therefore, the files produced by \texttt{149\#bet2} in  
+Condition 2 are replaced with the files produced by \texttt{149\#bet2} in 
 Condition 1.
-At step 3, \texttt{3\#fslstats} is labeled as 
-reproducible as it writes files without differences.
-At step 4, \texttt{4\#fslmaths} is labeled as non-reproducible 
-as the last process of the pipeline.
+At step 4, \texttt{152\#fslmaths} is labeled as non-reproducible as it writes files with differences.
+At step 5, \texttt{153\#fslstats} is labeled as 
+reproducible as the last process of the pipeline.
 

### 2020-03-05 bin/example.sh

@@ -11,9 +11,16 @@ input_image=$1
 base_name=$(basename ${input_image} .nii.gz)
 bet_output="${base_name}_brain.nii.gz"
 bet_output_binarized="${base_name}_brain_bin.nii.gz"
+robustfov_output="${base_name}_roi.nii.gz"
+
+# Cropping to remove neck and lower head
+robustfov -i ${input_image} -r ${robustfov_output}
 
 # Run FSL bet, put result in ${bet_output}
 bet ${input_image} ${bet_output}
+echo "Voxels / volume in brain mask:"
+fslstats ${bet_output} -V
+bet ${robustfov_output} ${bet_output}
 echo "Voxels / volume in brain mask:"
 fslstats ${bet_output} -V
 
@@ -23,4 +30,4 @@ echo "Voxels / volume in binarized brain mask:"
 fslstats ${bet_output_binarized} -V
 
 # Remove temporary file
-\rm ${bet_output}
+\rm ${robustfov_output}

### 2020-03-23 paper.tex
Updates and comments in tool description

@@
 % Furthermore, we can stabilize the identified processes in the next steps.
+\todo{explain creation vs propagation}
+\todo{define pipeline}
 
@@
 
-The steps implemented in \toolname to compare pipeline executions in
-2 conditions are outlined in Figure~\ref{fig:overview-tool}. First,
-in Condition 1, a directed, rooted, bipartite provenance graph is built for
-each subject, where nodes represent application processes or files, and
-edges represent read or write file accesses (Step 1 in Figure~\ref{fig:overview-tool}). \emph{Transient} files are then identified as
+\toolname identifies the processes in a pipeline that produce different results in two different conditions.  
+In Condition 1, a directed, rooted, bipartite provenance graph is recorded for
+each subject, where nodes represent application processes and files, and
+edges represent read and write file accesses (Step 1 in Figure~\ref{fig:overview-tool}).
+\emph{Transient} files are then identified as
  files that are either deleted during the execution, or modified by
@@
 of 1) (Step 3). This DAG is then compared to the one produced from
-Condition 2, in a step-by-step process that prevents the propagation of differences (step 4).
+Condition 2, in a step-by-step execution to prevent the propagation of differences (step 4).
 \toolname finally produces a labeled graph that identifies the
 non-reproducible processes of the pipeline.
+
+\toolname makes the following assumptions on the pipeline, meant to ensure that 
+a file can be unambiguously associated with the process that created it:
+\begin{enumerate}
+  \item Files are not written concurrently by different processes, although sequential writes by multiple processes are supported;
+  \item In case a process writes multiple versions of the same file, only a single version is made available to other processes.
+\end{enumerate} 
+In addition, \toolname also requires that processes are associated to a
+command line (executable and arguments), to facilitate process instrumentation.
+
+\tristan{Ali: please review these assumptions and check if any other one is required}
 
@@
     \includegraphics[width=.8\textwidth]{images/spot-diagram}
-    \caption{An overview of Spot tool in four steps.
-    (\href{https://docs.google.com/drawings/d/1OB3sB8kkK17Q516-TcmXKh3TwXSLDToJB11IVfUH5AU/edit?usp=sharing}{Source})}
+    \caption{\toolname overview.}
     \label{fig:overview-tool}
@@
 
-  \todo{review and summarize assumptions:
-  * No concurrent writes from multiple processes
-  * A given process cannot modify the bytes it wrote to a file
-  * Process are command lines
-  But we support non-concurrent writes by different processes. This is common in neuroimaging.
-  }
-
 \subsection{Recording provenance graphs}
 
-We use the \reprozip tool~\cite{rampin2016reprozip}
+\toolname uses the \reprozip tool~\cite{rampin2016reprozip}
 to capture: (1) the set of processes created by the
@@
 nodes from the \reprozip trace, and by adding directed edges corresponding
-to file reads and file writes (Figure~\ref{fig:simple_script}-left).
-
-\todo{Is every process linked to an executable?}
+to file reads and file writes (Figure~\ref{fig:simple_script}-left). \tristan{the new example is 
+confusing: it has too many processes and its purpose is confusing. This should be simplified. 
+See suggestion in example.sh: it has only a few processes, and it has a temporary and a multiple-write file. 
+If you agree that this example is enough, could you update the figure?}
 
@@
   Listing~\ref{listing:sample-script}.
-  Nodes are devided into two sets including processes that are shaped 
-  in the circles and files that are shaped in the boxes.
-  Transient files are represented with grey color boxes. 
+  Processes are represented with circles and files with rectangles.
+  Transient files are shaded in gray. 
   Plain edges 
   represent the process tree (\texttt{fork()} or \texttt{clone()} 
-  system calls). Dashed edges represent file dependencies.
+  system calls) \tristan{Does spot make any use of the process tree at any stage? If not, I think we may want to remove these edges.}. Dashed edges represent file dependencies.
   Every node in the graph is labeled using (1) a process id created by our
@@
 process P by a wrapper that first calls P and then saves the produced
-temporary files to a read-only directory. This process replacement is done
-without pipeline instrumentation, by pre-pending to the \texttt{PATH} environment
-variable a directory containing a wrapper script named as the executable
+temporary files to a read-only directory. This process replacement does not require 
+modifying the pipeline: it only pre-pends to the \texttt{PATH} environment
+variable a directory containing a wrapper script named after the executable
 called by P.
@@
  have an in-degree of 1 (Figure~\ref{fig:simple_script}-right). This operation also makes the provenance graph
- acyclic, since we assumed that a process can only write a single time to a file.
+ acyclic, since we assumed that a process could only release a single version of a file.
 
@@
 by \texttt{75\#bet2} and then overwritten by process \texttt{149\#bet2}.
-The complete list of the processes that
-write transient files is captured in Listing~\ref{listing:json_process}.
-Using this list, we modify processes in Condition 1 and then re-execute
-pipeline to save them in a backup folder.
-
-\begin{listing}
-  \begin{minted}[frame=single,
-    framesep=3mm,
-    linenos=false,
-    xleftmargin=0pt,
-    tabsize=4]{JSON}
-{
-  "multi_write_process": {
-    "bet2\00T1w\00T1w_brain":
-    {
-      "files":[
-        "/cond1/subj1/T1w_brain.nii.gz"
-      ],
-      "pid": 75
-    },
-    "bet2\00T1w_roi\00T1w_brain":
-    {
-      "files":[
-        "/cond1/subj1/T1w_brain.nii.gz"
-      ],
-      "pid": 149
-    }
-  },
-  "temp_process": {
-    "robustfov\00-i\00T1w.nii.gz\00-r\00T1w_roi.nii.gz": 
-    {
-      "files": [
-        "/cond1/subj1/T1w_roi.nii.gz"
-      ],
-      "pid": 3
-    } 
-  } 
-}
-
-\end{minted}
-\caption{List of processes and transient files from the example pipeline in 
-Listing~\ref{listing:sample-script}.}
-\label{listing:json_process}
-\end{listing}
 
@@
 
-Once all the files have been captured in Condition 1, including transient
-ones, \toolname labels pipeline processes through a step-by-step execution
-in Condition 2. Each process in Condition 2 is run and its output files are
-compared to the ones produced in Condition 1. If no differences are found,
-the process is marked as reproducible. Otherwise, the process is marked as
-non-reproducible and the output files produced in Condition 1 are copied to
-Condition 2, to ensure that differences do not propagate further in the
-pipeline. Processes are instrumented transparently, through a modification
-of the \texttt{PATH} variable similar to the one described previously. By
-default, differences in output files are identified by comparing file
-checksums. Other comparison functions can also be defined for specific file
-types, for instance to ignore sections of the file containing timestamps or
-other execution-specific information. Finally, \toolname creates a labeled
+Once transient files are captured, \toolname labels pipeline processes
+through a step-by-step execution in Condition 2. Each process in Condition
+2 is run and its output files are compared to the ones produced in
+Condition 1. If no differences are found, the process is marked as
+reproducible. Otherwise, the process is marked as non-reproducible and the
+output files produced in Condition 1 are copied to Condition 2, to ensure
+that differences do not propagate further in the pipeline. Processes are
+instrumented transparently, through a modification of the \texttt{PATH}
+variable similar to the one described previously. By default, differences
+in output files are identified by comparing file checksums. Other
+comparison functions can also be defined for specific file types, for
+instance to ignore sections of the file containing timestamps or other
+execution-specific information. Finally, \toolname creates a labeled
 provenance graph, highlighting non-reproducible processes.
@@
 
+* granularity is that of processes, no in-memory checks.
+
 Method:

### 2020-03-23 bin/example.sh

@@ -8,26 +8,15 @@ fi
 
 # Parse argument, set output file names
 input_image=$1
-base_name=$(basename ${input_image} .nii.gz)
-bet_output="${base_name}_brain.nii.gz"
-bet_output_binarized="${base_name}_brain_bin.nii.gz"
-robustfov_output="${base_name}_roi.nii.gz"
-
-# Cropping to remove neck and lower head
-robustfov -i ${input_image} -r ${robustfov_output}
 
 # Run FSL bet, put result in ${bet_output}
-bet ${input_image} ${bet_output}
-echo "Voxels / volume in brain mask:"
-fslstats ${bet_output} -V
-bet ${robustfov_output} ${bet_output}
-echo "Voxels / volume in brain mask:"
-fslstats ${bet_output} -V
+bet ${input_image} output.nii.gz
 
 # Create binary mask
-fslmaths ${bet_output} -bin ${bet_output_binarized}
+fslmaths output.nii.gz -bin output.nii.gz
+
 echo "Voxels / volume in binarized brain mask:"
-fslstats ${bet_output_binarized} -V
+fslstats output.nii.gz -V
 
 # Remove temporary file
-\rm ${robustfov_output}
+\rm output.nii.gz

### 2020-03-24 paper.tex
Some comments on results

@@
 \usepackage{pifont}
+
+\hypersetup{
+colorlinks=true,% hyperlinks will be coloured
+linkcolor=blue,% hyperlink text will be blue
+linkbordercolor=blue,% hyperlink border will be blue
+}
+\makeatletter
+\Hy@AtBeginDocument{%
+  \def\@pdfborder{0 0 1}% Overrides border definition set with colorlinks=true
+  \def\@pdfborderstyle{/S/U/W 1}% Overrides border style set with colorlinks=true
+                                % Hyperlink border style will be underline of width 1pt
+}
+\makeatother
+
 \DeclareMathOperator*{\argmin}{argmin}
@@
 
-In this experiment, we analyse the numerical reproducibility of computational pipelines 
-and identify the origin of differences across the operating systems. 
-Two types of differences can occur in the subjects due to the differences
-in the operating systems. One is between-OS differences caused by the
-operating system library updates and the other type, within-OS differences
-occur as a result of the pseudo-random processes used in the pipelines.
+We applied \toolname to find the origin of between-OS differences that
+occur in many neuroimaging pipelines. In particular, we studied the minimal
+pre-processing pipelines released by the Human Connectome Project
+(\href{https://www.humanconnectome.org}{HCP}), a leading initiative in
+neurosciences.
 
-In particular, Spot tool is tested on the neuroimaging applications which are 
-predominantly using mathematical libraries. Therefore, we expect to find 
-differences as a result of changing the mathematical functions between operating system libraries.
+% In this experiment, we analyse the numerical reproducibility of computational pipelines 
+% and identify the origin of differences across the operating systems. 
+% Two types of differences can occur in the subjects due to the differences
+% in the operating systems. One is between-OS differences caused by the
+% operating system library updates and the other type, within-OS differences
+% occur as a result of the pseudo-random processes used in the pipelines.
 
-This section describes datasets and pipelines used for the analyses, and the way of data processing.
+% In particular, Spot tool is tested on the neuroimaging applications which are 
+% predominantly using mathematical libraries. Therefore, we expect to find 
+% differences as a result of changing the mathematical functions between operating system libraries.
+
+% This section describes datasets and pipelines used for the analyses, and the way of data processing.
 
@@
 
-We used our tool to evaluate the reproducibility of the pipelines from the Human Connectome 
-Project (\href{https://www.humanconnectome.org}{HCP}).
-The HCP initiative is an effort to acquire and analyse 
-brain connectivity data from 1200 healthy adults.
-It enables the neuroscience 
-research community to discover relationships between brain circuits and 
-individual behaviors. This helps to understand a wide range of brain disorders.
-The HCP project provides database services (ConnectomeDB) for storing and 
-sharing primary and processed data freely, and data analysis pipelines that 
-are available under an open-source license.
+% We used our tool to evaluate the reproducibility of the pipelines from the Human Connectome 
+% Project .
+% The HCP initiative is an effort to acquire and analyse 
+% brain connectivity data from 1200 healthy adults.
+% It enables the neuroscience 
+% research community to discover relationships between brain circuits and 
+% individual behaviors. This helps to understand a wide range of brain disorders.
+% The HCP project provides database services (ConnectomeDB) for storing and 
+% sharing primary and processed data freely, and data analysis pipelines that 
+% are available under an open-source license.
 
@@
 functional, and diffusion MRI data acquired in the project. We focus on HCP
-pre-processing pipelines for structural data, particularly PreFreeSurfer
-and FreeSurfer.
-
-According to~\cite{glasser2013}, the sequence of the 
-sub-pipelines in PreFreeSurfer includes 
-(1) Distortion Correction (DC), 
-(2) Anatomical Average (AAve), 
-(3) ACPC (Anterior Commissure, Posterior Commissure) Alignment (ACPC-A), 
-(4) Brain Extraction (BExt), 
-(5) Bias Field Correction (BFC), 
-(6) Atlas-Registration (AR).
-
-Also, the sequence of the main sub-pipelines in FreeSurfer includes 
-(1) Downsampling images, 
-(2) T1w image registration, 
-(3) Placement of surface images, 
-(4) Folding-based surface registration.
+pre-processing pipelines for structural data, and particularly
+PreFreeSurfer and FreeSurfer. 
 A detailed description of the analyses done in these
 pipelines is available in~\cite{glasser2013}. 
+In summary, PreFreeSurfer consists of the following steps: 
+\begin{itemize}
+\item Distortion Correction (DC), 
+\item Anatomical Average (AAve), 
+\item ACPC (Anterior Commissure, Posterior Commissure) Alignment (ACPC-A), 
+\item Brain Extraction (BExt), 
+\item Bias Field Correction (BFC), 
+\item Atlas-Registration (AR).
+\end{itemize}
+And FreeSurfer consists of the following ones:
+\begin{itemize}
+\item Downsampling images, 
+\item T1w image registration, 
+\item Placement of surface images, 
+\item Folding-based surface registration.
+\end{itemize}
+The average processing time per subject is approximately 2 hours for
+PreFreeSurfer and 8 hours for FreeSurfer. The average output file size is
+approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
+\tristan{Ali, are these numbers per subject or for 20 subjects?}
 
-Furthermore, the average time per subject taken for PreFreeSurfer processing is $\approx$ 2.2 hours, 
-and FreeSurfer processing is $\approx$ 8.3 hours (20 runs), 
-and the average output file size of 20 subjects in PreFreeSurfer is $\approx$ 2.7 GB, 
-and FreeSurfer is $\approx$ 4.1 GB.
-
-HCP data collected from different types of imaging techniques, including 
-structural imaging (sMRI), functional imaging (fMRI), and diffusion imaging (dMRI).
-The structural images include T1-weighted (T1w) and T2-weighted (T2w) images, which 
-help in the diagnosis of brain injury.
-The images are indexed with a suffix if several scans of the same modality were acquired.
-For example, some data may include two T1w images with low and high resolutions.
-The functional images, including task-based fMRI and resting-state scans, 
-enable the measurement of functional activations within brain areas. 
-Diffusion imaging is another kind of MRI technique, which measures 
-the anatomical connectivity between regions.
+We randomly selected 20 unprocessed subjects 
+from the HCP data release S500 
+available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
+\tristan{Can you let me know where the data is located, so that I could describe it here?}
 See~\cite{van2013wu} for a discussion of these various modalities.
 
-In this paper, we randomly selected 20 unprocessed subjects 
-containing 3-tesla type of MRI scanners from the HCP data release S500 
-available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}.
+% HCP data collected from different types of imaging techniques, including 
+% structural imaging (sMRI), functional imaging (fMRI), and diffusion imaging (dMRI).
+% The structural images include T1-weighted (T1w) and T2-weighted (T2w) images, which 
+% help in the diagnosis of brain injury.
+% The images are indexed with a suffix if several scans of the same modality were acquired.
+% For example, some data may include two T1w images with low and high resolutions.
+% The functional images, including task-based fMRI and resting-state scans, 
+% enable the measurement of functional activations within brain areas. 
+% Diffusion imaging is another kind of MRI technique, which measures 
+% the anatomical connectivity between regions.
 
@@
 
-We executed the pipelines using Docker containers to simplify the 
-deployment of different operating system versions on execution 
-platforms. 
-The Docker images were built for the HCP pre-processing 
-pipelines v3.19.0 (PreFreeSurfer and FreeSurfer) in 
-CentOS 6.9 (Final) and CentOS 7.4 (Core). 
-CentOS is a widely used operating system among the Neuroscience community because of 
-the appropriate community support, network functions, safety, portability, and openness
-~\cite{hanke2011neuroscience}.
-These containers were installed with all the necessary software and libraries 
-required to run the HCP pipelines including FSL (version 5.0.6), 
-FreeSurfer (version 5.3.0-HCP, CentOS4 build), and Connectome Workbench (version 1.0).
-Docker images are available on 
-\href{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}{DockerHub}
-for reuse.
+We built Docker images for the HCP pre-processing pipelines v3.19.0
+(PreFreeSurfer and FreeSurfer) in CentOS 6.9 (Final) and CentOS 7.4 (Core),
+a widely-used operating system in neurosciences
+~\cite{hanke2011neuroscience}. Container images are available on
+\href{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}{DockerHub}.
+They contain the software dependencies of the HCP pipelines, including FSL
+(version 5.0.6), FreeSurfer (version 5.3.0-HCP, CentOS4 build), and
+Connectome Workbench (version 1.0).
 
-The data were processed with PreFreeSurfer and FreeSurfer. 
-First, PreFreeSurfer was utilized to analyse the unprocessed data, and then its 
-results were used as the input of FreeSurfer.
-Each subject was processed twice on the same operating system to detect
-variability coming from pseudo-random operations. 
-Within-OS variability may also be coming from other factors than pseudo-randomness. 
-For instance, the same files may have different metadata such as timestamps, 
-file paths, and other execution-specific information. 
-To ensure that the differences are not related to the metadata, 
-we used specific functions from FreeSurfer tool including 
-\emph{mri\_diff}, \emph{mris\_diff}, and \emph{lta\_diff}.
-These functions compare files based on the image dimensions, 
-voxel intensities, the number of vertices, and geometry information
-(see~\cite{fischl2012freesurfer} for more details). 
-In addition, we used the Dice coefficient metric to evaluate the similarity of 
-the image segmentation results, which is defined by
-\[DICE=\frac{2(|X|\cap|Y|)}{|X|+|Y|}\]
-The Dice coefficient ranges from 0, indicating X and Y have no voxels in common, to 1, if 
-two images are identical.
+We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2
+CentOS versions. Each subject was processed twice on the same operating
+system to detect potential within-OS variability, for instance coming from
+pseudo-random operations.
 
-Finally, to cluster the subjects, the threshold value is set to zero in the clustering method. 
-Also, we used the nearest neighbor algorithm to calculate the distance between clusters.
+\tristan{See what to do with this paragraph}
+Within-OS variability may also be coming from other factors than
+pseudo-randomness. For instance, the same files may have different metadata
+such as timestamps, file paths, and other execution-specific information.
+To ensure that the differences are not related to the metadata, we used
+specific functions from FreeSurfer tool including \emph{mri\_diff},
+\emph{mris\_diff}, and \emph{lta\_diff}. These functions compare files
+based on the image dimensions, voxel intensities, the number of vertices,
+and geometry information (see~\cite{fischl2012freesurfer} for more
+details) \tristan{Ali, are these functions used by \toolname to compare results?}.
+
+In addition, we used the Dice coefficient metric to evaluate the
+similarity of image segmentation results, defined by:
+\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\].
+
+% Finally, to cluster the subjects, the threshold value is set to zero in the clustering method. 
+% Also, we used the nearest neighbor algorithm to calculate the distance between clusters.
 
@@
 
-We did not observe any within-OS difference in PreFreeSurfer, however, we
-identified two FreeSurfer processes leading to within-OS differences due
+We did not observe any within-OS difference in PreFreeSurfer. In
+FreeSurfer, we identified 2 processes leading to within-OS differences due
 to the use of pseudo-random numbers: image registration with
-\emph{mri\_segreg}, and cortical surface curvature
-estimations with \emph{mris\_curvature}. To address this issue, we fixed
-the seed used by the pseudo-random number generator, through defining 
-option ``--seed'' in the command-line of the processes.
-Our analysis can focus on between-OS differences.
+\texttt{mri\_segreg}, and cortical surface curvature estimations with
+\texttt{mris\_curvature}. To remove these differences, we fixed the random
+seed used by FreeSurfer using the \texttt{--seed} command-line option.
 
@@
 
-We identified four types of subjects with different PreFreeSurfer process
-trees (Table~\ref{table:data-clusters}). Differences between subject types
-are coming from different  numbers of T1 and T2 images in the raw data:
-subjects of type 1 have 2 T1 images and 2 T2 images, while subjects of type
-2 only have 1 T1 image and 1 T2 image. This leads to
-different sets of output files. We verified that the process trees were
-identical for all subjects of the same type, for both versions of
-CentOS.
+We identified four types of subjects with different PreFreeSurfer
+provenance graphs (Table~\ref{table:data-clusters}). Differences between
+subject types are coming from different  numbers of T1 and T2 images in the
+raw data: subjects of type 1 have 2 T1 images and 2 T2 images, while
+subjects of type 2 only have 1 T1 image and 1 T2 image. We verified that
+the provenance graphs were identical for all subjects of the same type, for
+both versions of CentOS.
 
@@
 \begin{threeparttable}
-\caption{Clustering of 20 input subjects.}
+\caption{Types of provenance graphs in PreFreeSurfer.}
 \label{table:data-clusters}
@@
 \toprule
-       &                        &  \multicolumn{4}{c}{Available Modalities}    \\ 
+       &                        &  \multicolumn{4}{c}{Image modalities in subject data}    \\ 
 \cmidrule(lr){3-6}       
@@
 \end{tabular}
-\begin{tablenotes}
-     \small
-     \item *Type refers to the topology type of the process tree of subjects.
-\end{tablenotes}
 \end{threeparttable}
@@
 
-Figure~\ref{fig:pfs_freq} shows the frequency at which  processes create differences 
-in the pipeline, and how differences are distributed among sub-pipelines.
+Figure~\ref{fig:pfs_freq} shows the frequency of non-reproducible pipeline processes
+by PreFreeSurfer pipeline step.
 Differences were observed in linear registration 
@@
 Atlas Registration), in non-linear registration with FSL FNIRT (in Brain Extraction 
-and Altas Registration), and in image warping with FSL \emph{new\_invwarp} (in Brain Extraction 
+and Altas Registration), and in image warping with FSL \texttt{new\_invwarp} (in Brain Extraction 
 and Atlas Registration). Differences were also observed in image mean 
-computations with FSL maths  (in Anatomical Average).
+computations with FSL maths  (in Anatomical Average). \tristan{figure and caption to be fixed.}
 
-The results show two types of differences, between-process and between-subject.
+Two types of differences are observed: between-process and between-subject.
 Between-process differences occur when the same tool behaves differently in different process instances.
-For example, 
-FLIRT tool always produces files without differences in \emph{AAve} sub-pipeline but 
-mostly creates different files in \emph{AAli} sub-pipeline.
-Also, different frequencies show between-subjects variability.
-Accordingly, the same process can be reproducible for some subjects but not for others. 
-The similar findings in another study~\cite{bowring2019exploring} were reported 
-the effect of between-subject differences on reproducibility.
+For example, \texttt{flirt}
+always produces files without differences in the \texttt{AAve} sub-pipeline but 
+almost always creates different files in the \texttt{AAli} sub-pipeline.
+Conversely, the same process can be reproducible for some subjects but not for others.
+
 % Between subject differences can be a result of different subject types. 
@@
   \includegraphics[width=\textwidth]{images/pfs_heatmap.png}
-  \caption{Frequency heatmap of processes in the tools that introduce differences among 20 subjects. 
-           Each column represents the pipeline tool, and each row shows the sub-pipeline.
-           Each cell may contain several instances of the process of a tool in a sub-pipeline.
-           Annotations describe the frequency distribution of the instances as ``N [M] x Y" 
+  \caption{Frequency heatmap of non-reproducible PreFreeSurfer processes (N=20). 
+           Rows refer to PreFreeSurfer pipeline steps.
+           Cell labels describe the frequency distribution of the instances as ``N [M] x Y" 
            in which N is the proportion of the process 
@@
 
-Figure~\ref{fig:fnirt_result} 
-shows checkerboard image from the results of ACPC-Alignment in PreFreeSurfer. 
-Each rectangle includes the replaced voxels with different intensity values between images.
-The misaligned regions visually represent significant differences in the alignment 
-procedure.
-
-%Alternating rectangular regions between two images, creating a characteristic 
-%checkerboard-like pattern of varying magnitudes.
+To give an idea of the observed differences, Figure~\ref{fig:fnirt_result} 
+compares results of ACPC-Alignment in PreFreeSurfer. \tristan{Ali, there is no
+FNIRT process in ACPC-Alignment, please check.}
 
@@
     \includegraphics[width=\columnwidth]{images/t2w_alignment.png} 
-    \caption{Differences between FNIRT results of one subject from PreFreeSurfer 
+    \caption{Differences between FNIRT results from PreFreeSurfer 
     (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
-    CentOS7).
+    CentOS7). An animated version of the comparison is available 
     \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
-    {The link} to the animated version of alignment results shows the better visualization of differences.
+    {here}.
 } 
@@
 
-We processed 20 subjects using FreeSurfer and 
-Spot tool identified \emph{mris\_make\_surfaces} as the only process in 
-the pipeline that creates differences in 10 subjects.
-This process generates surface files for cortical and white matter surfaces 
-(see~\cite{fischl2012freesurfer} for more detail). 
-Also, FreeSurfer did not introduce any differences in 10 subjects.
+The only non-reproducible process identified in FreeSurfer was
+\texttt{mris\_make\_surfaces}, which was the case for 10 out of 20
+subjects. \tristan{do you know why this was the case: is this CLI
+dynamically linked?} This process generates the cortical and white matter
+surfaces. However, this does not mean that FreeSurfer results were
+identical for the remaining 10 subjects and for the other processes, due to the
+propagation of PreFreeSurfer differences.
 
-Results show that the static builds of FreeSurfer in which uses the embedded libraries 
-significantly avoids the creation of differences.
-However, we found that between-OS differences occur when pre-processed data are different in FreeSurfer. 
-FreeSurfer may not introduce differences, but it propagates and amplifies 
-the current differences that appeared in PreFreeSurfer results across the operating systems.
+For instance, Figure~\ref{fig:tissue_class} illustrates the sum of
+binarized between-OS differences among the 20 whole-brain segmentations. An
+animated comparison of the segmentations obtained in 1 subject is available
+\href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
+{here}.
 
-Figure~\ref{fig:tissue_class} illustrates the sum of binarized differences 
-from 20 subject results of the brain tissue segmentation across OSes.
-This figure shows between-OS differences  
-as a result of the propagation and amplification of differences in FreeSurfer. 
-Differences are mostly localized at the interfaces between tissues.
-In our dataset, we found subjects with min Dice value of 0.89 and max Dice  
-value of 0.96.
-
-Furthermore, the scatter plots of the Dice coefficients of the 44 segmented regions obtained 
-from two operating systems are shown in Figure~\ref{fig:scatter_plot}.
-This figure shows high Dice coefficients (more than 0.95) in large regions and very low 
-Dice coefficients (below 0.5) in small structures particularly in the vessels and lateral ventricles.
+The Dice coefficients associated with the 44 regions segmented by
+FreeSurfer are shown in Figure~\ref{fig:scatter_plot}. This figure shows
+high Dice coefficients (more than 0.95) in large regions and very low Dice
+coefficients (below 0.5) in small structures particularly in the vessels
+and lateral ventricles.
 
@@
   \includegraphics[width=\columnwidth]{images/brain_segmentation_mni.png} 
-  \caption{Heatmap binarized differences between brain segmentation results (file \emph{aseg.hires.mgz}) from 
-          20 subjects in FreeSurfer (CentOS6 vs. CentOS7). 
-          All binarized differences were resampled to the default MNI152 volume template.
-          \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
-          {The link} to the animated version of segmentation results of one subject presents more clear illustration of differences.} 
+  \caption{Sum of binarized between-OS differences between whole-brain FreeSurfer segmentations (N=20),
+           resampled and overlaid to the MNI152 volume template.
+} 
   \label{fig:tissue_class}
@@
 
-* granularity is that of processes, no in-memory checks.
+  The similar findings in another study~\cite{bowring2019exploring} were reported 
+  the effect of between-subject differences on reproducibility.
+
+* granularity is that
+ of processes, no in-memory checks.
 

### 2020-03-24 paper.tex

@@
   \item In case a process writes multiple versions of the same file, only a single version is made available to other processes.
+  \item Temporary files are not removed by the processes that create them.
 \end{enumerate} 
@@
 command line (executable and arguments), to facilitate process instrumentation.
-
-\tristan{Ali: please review these assumptions and check if any other one is required}
 
@@
 nodes from the \reprozip trace, and by adding directed edges corresponding
-to file reads and file writes (Figure~\ref{fig:simple_script}-left). \tristan{the new example is 
-confusing: it has too many processes and its purpose is confusing. This should be simplified. 
-See suggestion in example.sh: it has only a few processes, and it has a temporary and a multiple-write file. 
-If you agree that this example is enough, could you update the figure?}
+to file reads and file writes (Figure~\ref{fig:simple_script}-left).
 
@@
 \centering
-  \includegraphics[width=1\columnwidth]{images/provenance-graphs}
+  \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
   \caption{Provenance graphs
@@
   represent the process tree (\texttt{fork()} or \texttt{clone()} 
-  system calls) \tristan{Does spot make any use of the process tree at any stage? If not, I think we may want to remove these edges.}. Dashed edges represent file dependencies.
+  system calls) 
+  \tristan{Does spot make any use of the process tree at any stage? If not, 
+  I think we may want to remove these edges. 
+  Ali: Yes, it uses process tree structure to do some tree operations like search, filter, or remove.
+  Also using the process tree we do clustering later}. 
+  Dashed edges represent file dependencies.
   Every node in the graph is labeled using (1) a process id created by our
@@
   Process 0 is the initial call to the \texttt{sh} interpreter, and
-  the rest of processes are the calls to FSL robustfov, bet, stats and maths made in
-  Listing~\ref{listing:sample-script}. Processes 75 and 149 are forked by processes 4 and 78, respectively: 
-  they were captured by \reprozip while they did not appear in
+  the rest of processes are the calls to FSL bet, stats and maths made in
+  Listing~\ref{listing:sample-script}. Process 73 is forked by process 2: 
+  it was captured by \reprozip while it did not appear in
   Listing~\ref{listing:sample-script}.  
@@
 As an example, in the sample pipeline in
-Listing~\ref{listing:sample-script}, process \texttt{3\#robustfov} writes a
-temporary file. Also, file \emph{T1w\_brain.nii.gz} is first written
-by \texttt{75\#bet2} and then overwritten by process \texttt{149\#bet2}.
+Listing~\ref{listing:sample-script}, file \emph{output.nii.gz} is first written
+by \texttt{73\#bet2} and then overwritten by process \texttt{75\#fslmaths} as a temporary file.
 
@@
 
-\begin{figure*}
+\begin{figure}
   \centering
-  \includegraphics[width=.8\textwidth]{images/labelling-process}
+  \includegraphics[width=.8\columnwidth]{images/labelling-process}
   \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red.}
   \label{fig:iterations}
-\end{figure*}
+\end{figure}
 
@@
 process for the example in Figure~\ref{fig:simple_script}. 
-At step one, process \texttt{3\#robustfov} is executed and then labeled 
+At step one, process \texttt{73\#bet2} is labeled as non-reproducible (red) 
+as it produced files with differences 
+from files without differences. Therefore, the files produced by \texttt{73\#bet2} in  
+Condition 2 are replaced with the files produced by \texttt{73\#bet2} in 
+Condition 1.
+At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
 as reproducible (green) as it produced files without differences.
-At step 2, \texttt{75\#bet2} 
-is also labeled as reproducible as it produced files without differences. 
-At step 3 \texttt{149\#bet2} is labeled as non-reproducible (red) as it produced files with differences 
-from files without differences. Therefore, the files produced by \texttt{149\#bet2} in  
-Condition 2 are replaced with the files produced by \texttt{149\#bet2} in 
-Condition 1.
-At step 4, \texttt{152\#fslmaths} is labeled as non-reproducible as it writes files with differences.
-At step 5, \texttt{153\#fslstats} is labeled as 
+At step 3, \texttt{76\#fslstats} is also labeled as 
 reproducible as the last process of the pipeline.

### 2020-03-27 paper.tex
reply comments in results

--- 2020-03-24 paper.tex
@@
 approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
-\tristan{Ali, are these numbers per subject or for 20 subjects?}
+\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
 
@@
 available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
-\tristan{Can you let me know where the data is located, so that I could describe it here?}
+\tristan{Can you let me know where the data is located, so that I could describe it here?
+Subjects are located in consider:/data/asalari/ali-tests/PFS\_Centos6\_Traced
+and consider:/data/asalari/ali-tests/PFS\_Centos7\_Traced 
+and outputs are in consider:/data/asalari/ali-tests/nurm-out}
 See~\cite{van2013wu} for a discussion of these various modalities.
@@
 and geometry information (see~\cite{fischl2012freesurfer} for more
-details) \tristan{Ali, are these functions used by \toolname to compare results?}.
+details) \tristan{Ali, are these functions used by \toolname to compare results?
+Yes, they have been used based on the file format}.
 
@@
            Also, times Y shows the number of processes with the same distribution.
-           For example, the cell of FLIRT tool in the sub-pipeline AAve shows 10 instances of the process 
-           with the frequency of 11 that introduce no differences.
+           For example, the FNIRT process occurred at most 6 times in sub-pipeline DC in each subject. 
+           Also, this shows 3 different distributions among the subjects.
+           The left cell shows 2 instances of the FNIRT that introduce differences in 19 out of 20 subjects.
+           The middle cell indicates three instances of the FNIRT that create no differences for all 20 subjects.
+           The right cell shows one instance of the FNIRT that introduces differences in 17 out of 20 subjects.
            The range between red and green colors shows the highest and lowest 
@@
 To give an idea of the observed differences, Figure~\ref{fig:fnirt_result} 
-compares results of ACPC-Alignment in PreFreeSurfer. \tristan{Ali, there is no
-FNIRT process in ACPC-Alignment, please check.}
+compares results of Brain Extraction in PreFreeSurfer. \tristan{Ali, there is no
+FNIRT process in ACPC-Alignment, please check. Ali: yes, it was brain extraction.}
 
@@
 subjects. \tristan{do you know why this was the case: is this CLI
-dynamically linked?} This process generates the cortical and white matter
+dynamically linked? Ali: It includes dynamic linking and I guess it may be affected by this.} 
+This process generates the cortical and white matter
 surfaces. However, this does not mean that FreeSurfer results were

### 2020-03-29 paper.tex
Added a comment in methods

@@
 
-\toolname identifies the processes in a pipeline that produce different results in two different conditions.  
+\toolname identifies the processes in a pipeline that produce different results when executed in two different conditions.  
 In Condition 1, a directed, rooted, bipartite provenance graph is recorded for
@@
 (Directed Acyclic Graph) in which file nodes have a single parent (in-degree
-of 1) (Step 3). This DAG is then compared to the one produced from
+of 1) (Step 3) \tristan{Ali, this is not true in Figure 1: at step 3 there are still transient files}. This DAG is then compared to the one produced from
 Condition 2, in a step-by-step execution to prevent the propagation of differences (step 4).

### 2020-03-29 paper.tex
Some fixes in methods

@@
 of 1) (Step 3) \tristan{Ali, this is not true in Figure 1: at step 3 there are still transient files}. This DAG is then compared to the one produced from
-Condition 2, in a step-by-step execution to prevent the propagation of differences (step 4).
+Condition 2, in a step-by-step execution that prevents the propagation of differences (Step 4).
 \toolname finally produces a labeled graph that identifies the
@@
 \begin{enumerate}
-  \item Files are not written concurrently by different processes, although sequential writes by multiple processes are supported;
+  \item Files are not written concurrently by different processes, although sequential writes by multiple processes are supported.
   \item In case a process writes multiple versions of the same file, only a single version is made available to other processes.
-  \item Temporary files are not removed by the processes that create them.
+  \item Temporary files used by multiple processes are not removed by the process that created them. \tristan{explain}
 \end{enumerate} 
@@
   \end{figure*}
+
+\tristan{The text switches between active and passive voice, this should be fixed.}
 
@@
 
-Provenance graphs may be subject-dependent, which justifies the processing
-of multiple subjects. For instance, data acquisitions may be
-repeated a different number of times in each subject, for noise-reduction
-purposes. Some of these differences can be neglected, for instance when a
+Provenance graphs are often subject-dependent.
+%, which justifies the processing
+% of multiple subjects. For instance, data acquisitions may be
+% repeated a different number of times in each subject, for noise-reduction
+% purposes. 
+Some of these differences can be neglected, for instance when a
 data decompression step is present at the beginning of the execution for
@@
 
-\toolname captures temporary files in Condition 1 by replacing every
+\toolname captures temporary files by replacing every
 process P by a wrapper that first calls P and then saves the produced
@@
 by \texttt{73\#bet2} and then overwritten by process \texttt{75\#fslmaths} as a temporary file. 
+\tristan{add this comment to caption}
 

### 2020-03-31 paper.tex
Edits in method

@@
 
-\toolname identifies the processes in a pipeline that produce different results when executed in two different conditions.  
-In Condition 1, a directed, rooted, bipartite provenance graph is recorded for
-each subject, where nodes represent application processes and files, and
-edges represent read and write file accesses (Step 1 in Figure~\ref{fig:overview-tool}).
-\emph{Transient} files are then identified as
- files that are either deleted during the execution, or modified by
-multiple processes (Step 2). \toolname re-runs the pipeline to save these files
-right after they are produced, resulting in a provenance \emph{DAG}
-(Directed Acyclic Graph) in which file nodes have a single parent (in-degree
-of 1) (Step 3) \tristan{Ali, this is not true in Figure 1: at step 3 there are still transient files}. This DAG is then compared to the one produced from
-Condition 2, in a step-by-step execution that prevents the propagation of differences (Step 4).
-\toolname finally produces a labeled graph that identifies the
-non-reproducible processes of the pipeline.
+\toolname identifies the processes in a pipeline that produce different
+results when executed in two different conditions. First, a directed
+bipartite provenance graph is recorded for each subject, where nodes
+represent application processes and files, and edges represent read and
+write file accesses (Step 1 in Figure~\ref{fig:overview-tool}). Second,
+transient files, i.e., files that are either deleted during pipeline
+execution or modified by multiple processes, are identified (Step 2) and
+disambiguated, resulting in a provenance \emph{DAG} (Directed Acyclic
+Graph) in which file nodes have a single parent (in-degree of 1) (Step 3)
+\tristan{Ali, this is not true in Figure 1: at step 3 there are still
+transient files}. DAGs produced in the two conditions are then compared, in
+a step-by-step execution that prevents the propagation of differences in the pipeline (Step
+4). Finally, a labeled graph is produced that identifies the
+non-reproducible processes in the pipeline.
 
-\toolname makes the following assumptions on the pipeline, meant to ensure that 
-a file can be unambiguously associated with the process that created it:
+To ensure that a file can be unambiguously associated with the process that
+created it, we assume that pipeline processes can be re-ordered such that:
 \begin{enumerate}
-  \item Files are not written concurrently by different processes, although sequential writes by multiple processes are supported.
-  \item In case a process writes multiple versions of the same file, only a single version is made available to other processes.
-  \item Temporary files used by multiple processes are not removed by the process that created them. \tristan{explain}
+\item Processes don't run concurrently;
+\item Each process sequentially reads, computes, and writes. 
 \end{enumerate}
-In addition, \toolname also requires that processes are associated to a
-command line (executable and arguments), to facilitate process instrumentation.
+In practice, pipeline processes may still run concurrently provided that
+they don't write concurrently to the same files. A process may also
+interleave file writes with computing, for instance when different file
+blocks are processed sequentially. However, only a single version of the
+file must eventually be made available to the other processes. Similarly,
+in case processes write and delete temporary files throughout their
+execution, these temporary files must not be read by any other process in
+the pipeline.
+
+In addition, we also require that processes are
+associated to a command line (executable and arguments), to facilitate
+process instrumentation.
 
@@
 
-\tristan{The text switches between active and passive voice, this should be fixed.}
-
 \subsection{Recording provenance graphs}
 
-\toolname uses the \reprozip tool~\cite{rampin2016reprozip}
+We use the \reprozip tool~\cite{rampin2016reprozip}
 to capture: (1) the set of processes created by the
@@
 \texttt{ptrace()} system call, with no required instrumentation of the pipeline.
-\toolname then reconstructs a provenance graph by creating process and file
-nodes from the \reprozip trace, and by adding directed edges corresponding
-to file reads and file writes (Figure~\ref{fig:simple_script}-left).
+Using the \reprozip trace, \toolname reconstructs a provenance graph by creating process and file
+nodes and by adding directed edges corresponding
+to file reads and writes (Figure~\ref{fig:provenance-graph}).
 
-Provenance graphs are often subject-dependent.
+Provenance graphs are often subject-dependent, due to variations in input data among subjects.
 %, which justifies the processing
@@
 some subjects only, but other differences cannot, for instance
-when complete different processing paths are used in different subjects.
+when complete different processing paths are used in different subjects. \tristan{We need to conclude on that comment}
 
@@
 \begin{figure}
-\centering
-  \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
+\begin{subfigure}{0.4\linewidth}
+  [ FIGURE GOES HERE ]
+
+  \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
+  \label{fig:provenance-graph}
+\end{subfigure}
+\hfill
+\begin{subfigure}{0.4\linewidth}
+  [ FIGURE GOES HERE ]
+
+  \caption{Provenance DAG, with disambiguated transient files.}
+  \label{fig:provenance-dag}
+\end{subfigure}
+  % \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
   \caption{Provenance graphs
-  constructed from the example pipeline in
+  created from the example pipeline in
   Listing~\ref{listing:sample-script}.
-  Processes are represented with circles and files with rectangles.
-  Transient files are shaded in gray. 
-  Plain edges 
-  represent the process tree (\texttt{fork()} or \texttt{clone()} 
-  system calls) 
-  \tristan{Does spot make any use of the process tree at any stage? If not, 
-  I think we may want to remove these edges. 
-  Ali: Yes, it uses process tree structure to do some tree operations like search, filter, or remove.
-  Also using the process tree we do clustering later}. 
-  Dashed edges represent file dependencies.
-  Every node in the graph is labeled using (1) a process id created by our
-  reconstruction, (2) the name of the executable run by the process or the file name.
-  Process 0 is the initial call to the \texttt{sh} interpreter, and
-  the rest of processes are the calls to FSL bet, stats and maths made in
-  Listing~\ref{listing:sample-script}. Process 73 is forked by process 2: 
-  it was captured by \reprozip while it did not appear in
-  Listing~\ref{listing:sample-script}.  
+  Processes are represented with circles, files with rectangles, and read/write accesses 
+  with plain edges. 
+  For visualization purposes, the process tree is represented too, with
+  dashed edges. Every node in the graph is labeled using (1) a process id
+  created by our reconstruction, (2) the name of the executable run by the
+  process or the file name. Process 0 is the initial call to the
+  \texttt{sh} interpreter; the other processes are the calls to FSL bet,
+  stats and maths made in Listing~\ref{listing:sample-script}. Process 73
+  is forked by process 2: it was captured by \reprozip while it did not
+  appear in Listing~\ref{listing:sample-script}.
 }
@@
 
-\toolname captures temporary files by replacing every
+We capture temporary files by replacing every
 process P by a wrapper that first calls P and then saves the produced
-temporary files to a read-only directory. This process replacement does not require 
-modifying the pipeline: it only pre-pends to the \texttt{PATH} environment
-variable a directory containing a wrapper script named after the executable
+temporary files to a read-only directory. This process replacement is done by pre-pending
+ to the \texttt{PATH} environment
+variable a directory that contains a wrapper script named after the executable
 called by P.
 
-Files written by multiple processes are disambiguated as follows. For a
+Files written by multiple processes are disambiguated using a similar technique. For a
  file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
- $P_{n}$\}, \toolname first checks that processes in \textbf{P} do not
- write concurrently to F, which would violate our assumptions. Then, it
- replaces every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
+ $P_{n}$\}, we first check that processes in \textbf{P} do not
+ write concurrently to F, which would violate our assumptions. Then, we
+ replace every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
  calls $P_{i}$ and then saves F to a read-only directory. In this way,
- successive versions of F are preserved for comparison. \toolname finally
- updates the provenance graph accordingly, so that all files in the graph
- have an in-degree of 1 (Figure~\ref{fig:simple_script}-right). This operation also makes the provenance graph
+ successive versions of F are preserved for comparison. We finally
+ update the provenance graph accordingly, so that all files in the graph
+ have an in-degree of 1 (Figure~\ref{fig:provenance-dag}). This operation also makes the provenance graph
  acyclic, since we assumed that a process could only release a single version of a file.
-
-As an example, in the sample pipeline in
-Listing~\ref{listing:sample-script}, file \emph{output.nii.gz} is first written
-by \texttt{73\#bet2} and then overwritten by process \texttt{75\#fslmaths} as a temporary file. 
-\tristan{add this comment to caption}
 
@@
 
-Once transient files are captured, \toolname labels pipeline processes
-through a step-by-step execution in Condition 2. Each process in Condition
-2 is run and its output files are compared to the ones produced in
-Condition 1. If no differences are found, the process is marked as
+After capturing transient files, we re-run the pipeline
+step by step in Condition 2 to label processes. The output files
+produced by a process in Condition 2 are compared to the ones produced
+in Condition 1. If no differences are found, the process is marked as
 reproducible. Otherwise, the process is marked as non-reproducible and the
@@
 comparison functions can also be defined for specific file types, for
-instance to ignore sections of the file containing timestamps or other
-execution-specific information. Finally, \toolname creates a labeled
-provenance graph, highlighting non-reproducible processes.
+instance to ignore file headers or file sections containing timestamps.
+\toolname finally creates a labeled
+provenance graph highlighting non-reproducible processes.
+
+Figure~\ref{fig:iterations} illustrates the incremental labelling 
+process for the example in Figure~\ref{fig:simple_script}. 
+At step 1, process \texttt{73\#bet2} is labeled as non-reproducible (red) 
+as it produces files with differences. Subsequently, the files produced by \texttt{73\#bet2} in  
+Condition 2 are replaced with the files produced by \texttt{73\#bet2} in 
+Condition 1.
+At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
+as reproducible (green) as it produced files without differences.
+At step 3, \texttt{76\#fslstats} is also labeled as 
+reproducible as it didn't produce any file.
 
@@
 \end{figure}
-
-Figure~\ref{fig:iterations} illustrates our incremental labelling 
-process for the example in Figure~\ref{fig:simple_script}. 
-At step one, process \texttt{73\#bet2} is labeled as non-reproducible (red) 
-as it produced files with differences 
-from files without differences. Therefore, the files produced by \texttt{73\#bet2} in  
-Condition 2 are replaced with the files produced by \texttt{73\#bet2} in 
-Condition 1.
-At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
-as reproducible (green) as it produced files without differences.
-At step 3, \texttt{76\#fslstats} is also labeled as 
-reproducible as the last process of the pipeline.
 
@@
 
+Propagation is very important.
+
   The similar findings in another study~\cite{bowring2019exploring} were reported 

### 2020-03-31 paper.tex
Edits in results

@@
 some subjects only, but other differences cannot, for instance
-when complete different processing paths are used in different subjects. \tristan{We need to conclude on that comment}
+when complete different processing paths are used in different subjects. \toolname 
+includes helpers to identify different instances of provenance graphs \tristan{mention tree distance here}.
 
@@
 
-We applied \toolname to find the origin of between-OS differences that
-occur in many neuroimaging pipelines. In particular, we studied the minimal
+We applied the \toolname to find the origin of between-OS differences that
+frequently occur in neuroimaging pipelines. In particular, we studied the minimal
 pre-processing pipelines released by the Human Connectome Project
@@
 pre-processing pipelines for structural data, and particularly
-PreFreeSurfer and FreeSurfer. 
-A detailed description of the analyses done in these
+on PreFreeSurfer and FreeSurfer. 
+A detailed description of the analyses done by these
 pipelines is available in~\cite{glasser2013}. 
@@
 \item Anatomical Average (AAve), 
-\item ACPC (Anterior Commissure, Posterior Commissure) Alignment (ACPC-A), 
+\item Anterior/Posterior Commissure Alignment (ACPC-A), 
 \item Brain Extraction (BExt), 
@@
 \end{itemize}
-And FreeSurfer consists of the following ones:
+And FreeSurfer consists of the following ones: \tristan{Check these steps}
 \begin{itemize}
-\item Downsampling images, 
+\item Image downsampling, 
 \item T1w image registration, 
-\item Placement of surface images, 
-\item Folding-based surface registration.
+\item Surface placement, 
+\item Surface registration.
 \end{itemize}
-The average processing time per subject is approximately 2 hours for
-PreFreeSurfer and 8 hours for FreeSurfer. The average output file size is
+The average processing time per subject is approximately 2~hours for
+PreFreeSurfer and 8~hours for FreeSurfer. The average output file size is
 approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
-\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
+%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
 
@@
 available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
-\tristan{Can you let me know where the data is located, so that I could describe it here?
-Subjects are located in consider:/data/asalari/ali-tests/PFS\_Centos6\_Traced
-and consider:/data/asalari/ali-tests/PFS\_Centos7\_Traced 
-and outputs are in consider:/data/asalari/ali-tests/nurm-out}
-See~\cite{van2013wu} for a discussion of these various modalities.
+For each subject, available data consisted of 1 or 2 T1 images, and 1 or 2 T2 images. 
+% \tristan{Can you let me know where the data is located, so that I could describe it here?
+% Subjects are located in consider:/data/asalari/ali-tests/PFS\_Centos6\_Traced
+% and consider:/data/asalari/ali-tests/PFS\_Centos7\_Traced 
+% and outputs are in consider:/data/asalari/ali-tests/nurm-out}
+Acquisition protocols and parameters are detailed in~\cite{van2013wu}. \tristan{Add summary of matrix sizes.}
 
@@
 We built Docker images for the HCP pre-processing pipelines v3.19.0
-(PreFreeSurfer and FreeSurfer) in CentOS 6.9 (Final) and CentOS 7.4 (Core),
-a widely-used operating system in neurosciences
-~\cite{hanke2011neuroscience}. Container images are available on
+(PreFreeSurfer and FreeSurfer) in CentOS 6.9 (Final) and CentOS 7.4 (Core), available on
 \href{https://hub.docker.com/r/bigdatalabteam/hcp-prefreesurfer/}{DockerHub}.
-They contain the software dependencies of the HCP pipelines, including FSL
+Container images contain the HCP software dependencies, including FSL
 (version 5.0.6), FreeSurfer (version 5.3.0-HCP, CentOS4 build), and
@@
 CentOS versions. Each subject was processed twice on the same operating
-system to detect potential within-OS variability, for instance coming from
-pseudo-random operations.
-
-\tristan{See what to do with this paragraph}
-Within-OS variability may also be coming from other factors than
-pseudo-randomness. For instance, the same files may have different metadata
-such as timestamps, file paths, and other execution-specific information.
-To ensure that the differences are not related to the metadata, we used
-specific functions from FreeSurfer tool including \emph{mri\_diff},
-\emph{mris\_diff}, and \emph{lta\_diff}. These functions compare files
-based on the image dimensions, voxel intensities, the number of vertices,
-and geometry information (see~\cite{fischl2012freesurfer} for more
-details) \tristan{Ali, are these functions used by \toolname to compare results?
-Yes, they have been used based on the file format}.
-
-In addition, we used the Dice coefficient metric to evaluate the
-similarity of image segmentation results, defined by:
+system to detect within-OS variability coming from pseudo-random
+operations. We ignored execution-specific information such as file path or
+timestamps by using FreeSurfer tools \texttt{mri\_diff},
+\texttt{mris\_diff}, and \texttt{lta\_diff} that determine if MRI volumes
+or transformations differ. To compare segmentations, we used the Dice coefficient defined as follows:
 \[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\].
@@
 Figure~\ref{fig:pfs_freq} shows the frequency of non-reproducible pipeline processes
-by PreFreeSurfer pipeline step.
+in PreFreeSurfer. 
 Differences were observed in linear registration 
-with FSL FLIRT (in ACPC-Alignment, Brain Extraction, Distortion Correction, 
+with FSL FLIRT (in ACPC-Alignment, Brain Extraction, Distortion Correction, and
 Atlas Registration), in non-linear registration with FSL FNIRT (in Brain Extraction 
-and Altas Registration), and in image warping with FSL \texttt{new\_invwarp} (in Brain Extraction 
+and Atlas Registration), and in image warping with FSL \texttt{new\_invwarp} (in Brain Extraction 
 and Atlas Registration). Differences were also observed in image mean 
@@
   \includegraphics[width=\textwidth]{images/pfs_heatmap.png}
-  \caption{Frequency heatmap of non-reproducible PreFreeSurfer processes (N=20). 
-           Rows refer to PreFreeSurfer pipeline steps.
-           Cell labels describe the frequency distribution of the instances as ``N [M] x Y" 
-           in which N is the proportion of the process 
-           that creates differences out of all M times that is occurred. 
-           Also, times Y shows the number of processes with the same distribution.
-           For example, the FNIRT process occurred at most 6 times in sub-pipeline DC in each subject. 
-           Also, this shows 3 different distributions among the subjects.
-           The left cell shows 2 instances of the FNIRT that introduce differences in 19 out of 20 subjects.
-           The middle cell indicates three instances of the FNIRT that create no differences for all 20 subjects.
-           The right cell shows one instance of the FNIRT that introduces differences in 17 out of 20 subjects.
-           The range between red and green colors shows the highest and lowest 
-           proportion of occurrences of the processes with differences.
-           Also, grey cells indicate that the tool is not executed in the sub-pipeline.}
+  \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 
+  \tristan{Remove frequency values as they are shown by the color already. Remove multipliers, multiply cells instead. Cells should just contain subject numbers.}
+  Each cell represents the occurrence of a particular command line in a
+  pipeline step among Anatomical Average (AAve), Anterior/Posterior
+  Commissure Alignment (ACPC-A), Brain Extraction (BExt), Bias Field
+  Correction (BFC), or Atlas-Registration (AR). Cell labels indicate the
+  number of subjects where the corresponding process appears. For example,
+  the FLIRT tool was invoked 6 times in step DC for each of the 20
+  subjects: 2 instances weren't reproducible in 95\% of the subjects, 3
+  instances were always reproducible, and 1 instance wasn't reproducible in
+  85\% of the subjects.
+        }
   \label{fig:pfs_freq}
@@
 
-To give an idea of the observed differences, Figure~\ref{fig:fnirt_result} 
-compares results of Brain Extraction in PreFreeSurfer. \tristan{Ali, there is no
-FNIRT process in ACPC-Alignment, please check. Ali: yes, it was brain extraction.}
-
+To visually illustrate the observed differences,
+Figure~\ref{fig:fnirt_result} compares FNIRT results in Brain Extraction
+for a particular subject. Differences appear to be very substantial, in
+particular in the frontal and parietal lobes. 
 \begin{figure}
@@
     \includegraphics[width=\columnwidth]{images/t2w_alignment.png} 
-    \caption{Differences between FNIRT results from PreFreeSurfer 
-    (checkerboard image \emph{T2w\_acpc\_to\_MNI\_nonli.nii.gz}) (CentOS6 vs. 
-    CentOS7). An animated version of the comparison is available 
+    \caption{Differences between FNIRT results in PreFreeSurfer's Brain Extraction (CentOS6 vs. 
+    CentOS7) \tristan{Improve contrast}. An animated version of the comparison is available 
     \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
@@
 \section{Discussion}
+
+The reproducibility of a given tool may vary across subjects, and across parameters. 
+See FLIRT in DC in Figure 4.
 

### 2020-04-01 paper.tex
Edited results section

@@
 
-Two types of differences are observed: between-process and between-subject.
-Between-process differences occur when the same tool behaves differently in different process instances.
-For example, \texttt{flirt}
-always produces files without differences in the \texttt{AAve} sub-pipeline but 
-almost always creates different files in the \texttt{AAli} sub-pipeline.
-Conversely, the same process can be reproducible for some subjects but not for others.
-
 % Between subject differences can be a result of different subject types. 
@@
     \includegraphics[width=\columnwidth]{images/t2w_alignment.png} 
-    \caption{Differences between FNIRT results in PreFreeSurfer's Brain Extraction (CentOS6 vs. 
+    \caption{Differences between T2 FNIRT results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
     CentOS7) \tristan{Improve contrast}. An animated version of the comparison is available 
-    \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
+    \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
     {here}.
@@
 
-The only non-reproducible process identified in FreeSurfer was
-\texttt{mris\_make\_surfaces}, which was the case for 10 out of 20
-subjects. \tristan{do you know why this was the case: is this CLI
-dynamically linked? Ali: It includes dynamic linking and I guess it may be affected by this.} 
-This process generates the cortical and white matter
-surfaces. However, this does not mean that FreeSurfer results were
-identical for the remaining 10 subjects and for the other processes, due to the
-propagation of PreFreeSurfer differences.
+The only non-reproducible process identified by \toolname in FreeSurfer was
+\texttt{mris\_make\_surfaces} (cortical and white matter surfaces
+generation), which is also the only dynamically-linked executable in this pipeline
+\tristan{Ali, could you check that this is indeed the only dynamic
+executable in the pipeline?}. \texttt{mris\_make\_surfaces} produced different results for
+10 out of 20 subjects.
 
-For instance, Figure~\ref{fig:tissue_class} illustrates the sum of
-binarized between-OS differences among the 20 whole-brain segmentations. An
-animated comparison of the segmentations obtained in 1 subject is available
-\href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
-{here}.
-
-The Dice coefficients associated with the 44 regions segmented by
-FreeSurfer are shown in Figure~\ref{fig:scatter_plot}. This figure shows
-high Dice coefficients (more than 0.95) in large regions and very low Dice
-coefficients (below 0.5) in small structures particularly in the vessels
-and lateral ventricles.
+While by design \toolname only identifies processes that create
+differences between conditions, pipeline results may still differ due to
+the propagation and amplification of differences created at previous
+processing steps, PreFreeSurfer in our case. We observed the effect of this
+propagation in FreeSurfer results, as shown in
+Figure~\ref{fig:tissue_class} for whole-brain segmentations. Furthermore,
+the Dice coefficients associated with the 44 regions segmented by
+FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing low Dice
+values (below 0.5) in the smallest structures.
 
@@
   \includegraphics[width=\columnwidth]{images/brain_segmentation_mni.png} 
-  \caption{Sum of binarized between-OS differences between whole-brain FreeSurfer segmentations (N=20),
-           resampled and overlaid to the MNI152 volume template.
+  \caption{Sum of binarized differences between whole-brain FreeSurfer
+  segmentations obtained from PreFreeSurfer processings in CentOS6 vs CentOS7
+   (N=20). Segmentations were resampled and overlaid to the MNI152 volume
+  template. An animated comparison of segmentations obtained for a particular subject is available
+\href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
+{here}.
 } 
@@
   \centering
-    \includegraphics[width=\textwidth]{images/scatter_plot.png} 
-    \caption{The scatter plots of the Dice coefficients between segmented regions 
-            from 20 subjects in FreeSurfer (CentOS6 vs. CentOS7). 
-            The x-axes are the number of voxels in each region and 
-            y-axes are Dice values that are in different scales in two plots.
-            The right plot shows the similarity between regions larger than $10^4$ voxels. 
-            The left plot represents the differences between 
-            regions smaller than $10^4$ voxels.
-            The complete table of Dice coefficient values is 
-            \href{https://github.com/ali4006/HCP-reproducibility-paper/blob/master/data/dice_values.csv}
-            {available}.} 
+  \begin{subfigure}{0.45\linewidth}
+    [FIGURE GOES HERE]
+    \caption{Regions with \emph{less} than 10,000 voxels.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.45\linewidth}
+    [FIGURE GOES HERE]
+    \caption{Regions with \emph{more} than 10,000 voxels.}
+  \end{subfigure}
+    % \includegraphics[width=\textwidth]{images/scatter_plot.png} 
+    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).
+        To improve readability, the two graphs are on different y scales. \tristan{Ali, could you separate the figures in 2 and insert them in each subfigure?}} 
     \label{fig:scatter_plot}
@@
 
+Static vs dynamic executables
+
 The reproducibility of a given tool may vary across subjects, and across parameters. 
 See FLIRT in DC in Figure 4.
+Two types of differences are observed: between-process and between-subject.
+Between-process differences occur when the same tool behaves differently in different process instances.
+For example, \texttt{flirt}
+always produces files without differences in the \texttt{AAve} sub-pipeline but 
+almost always creates different files in the \texttt{AAli} sub-pipeline.
+Conversely, the same process can be reproducible for some subjects but not for others.
 

### 2020-04-01 paper.tex
First version of discussion

@@
 
+\tristan{Ali, it would be useful to show a complete PreFreesurfer graph here.}
+
 % Between subject differences can be a result of different subject types. 
@@
 
-Static vs dynamic executables
+\subsection{Key findings}
 
-The reproducibility of a given tool may vary across subjects, and across parameters. 
-See FLIRT in DC in Figure 4.
-Two types of differences are observed: between-process and between-subject.
-Between-process differences occur when the same tool behaves differently in different process instances.
-For example, \texttt{flirt}
-always produces files without differences in the \texttt{AAve} sub-pipeline but 
-almost always creates different files in the \texttt{AAli} sub-pipeline.
-Conversely, the same process can be reproducible for some subjects but not for others.
+Linear and non-linear registration with FSL were found to
+frequently lead to differences between results obtained with different
+operating systems. This does not come as a surprise given the instabilities
+associated with these processes. It also corroborates our previous findings
+in~\cite{Glatard2015} where fMRI pre-processing using FSL was found to be
+non-reproducible across operating systems starting from the motion
+correction step, a step that uses FSL's \texttt{flirt} tool internally. It
+would be relevant to investigate if the observed instability of
+registration processes generalizes to other toolkits, or if it is specific
+to FSL. In view of the effect of small data perturbations in a variety of
+toolboxes and processes, such as cortical surface extraction using
+FreeSurfer and CIVET~\cite{Lewis2017-ll} or connectome estimation using
+Dipy~\cite{kiar2019comparing}, it is possible that this observation
+generalizes widely across toolboxes and requires a deeper investigation of
+the stability of linear and non-linear registration.
 
-Propagation is very important.
+While only a handful or processes were found non-reproducible across the
+tested operating systems, including most notably linear and non-linear
+registration, the effect of such instabilities were found to propagate
+widely in the pipelines, resulting in Dice coefficients frequently below
+0.9 for the amygdala or ventricles. This illustrates the need to conduct such 
+reproducibility studies for entire pipelines rather than isolated processes.
 
-  The similar findings in another study~\cite{bowring2019exploring} were reported 
-  the effect of between-subject differences on reproducibility.
+Furthermore, as shown in Figure~\ref{fig:pfs_freq}, the reproducibility of
+a given tool may vary across subjects and across processing parameters. For
+instance, linear registration with FLIRT seems to be fully reproducible in
+the Anatomical Average sub-pipeline, while it is highly non-reproducible in
+ACPC Alignment. In Brain Extraction, the same tool was found reproducible
+for some subjects only. Therefore, reproducibility studies need to be
+performed on several representative subjects rather than only a couple of
+them. While this is common practice in neurosciences, it is common that
+software tests be executed only on a single dataset, to reduce the
+associated computational load. It is likely, however, that the observed
+instabilities stem from numerical edge cases that may not manifest in all
+datasets.
 
-* granularity is that
- of processes, no in-memory checks.
+While our findings give insight on the reasons why image analysis processes
+are unstable across operating systems, they do not provide a complete
+uncertainty quantification for neuroimaging pipelines across operating
+systems. Although CentOS is a popular distribution in
+neurosciences~\cite{hanke2011neuroscience}, different observations could be
+made on other operating systems. The numerical noise introduced by
+operating system updates is a realistic one, as such updates are likely to
+occur throughout the time span of a neuroscience study, but it is also
+uncontrolled, as it originates in updates of low-level libraries by
+third-party developers. A possible method to study this problem more
+comprehensively would be to introduce controlled numerical perturbations in
+pipelines, which could be done either using data noise, or through
+computational noise introduced by Monte-Carlo
+Arithmetic~\cite{Parker1997-qq}. The work in~\cite{kiar2019comparing}
+discusses and compares these two techniques.
 
-Method:
-* we assume no concurrent file accesses.
+\subsection{\toolname evaluation}
 
-* we assume that the provenance graph is oriented, i.e., a process cannot read or write in the same file. Reasonnable in general but wouldn't work 
-for, say, out of core computations.
+The processes identified by \toolname as non-reproducible were all
+associated with dynamically-linked executables \tristan{Ali, please check
+that}. This makes complete sense as statically-linked executables wouldn't
+be impacted by library updates. Moreover, the hypothetical effects of
+hardware or Linux kernel updates were not measured, as the different
+operating sytems were deployed in Docker containers on the same host, that
+is, using the same kernel and hardware.
 
-Comment on the overhead, number of executions (3, 2 in condition 1, 1 in condition 2)
+To evaluate the reproducibility of a pipeline, \toolname needs to execute
+it 3 times in order to (1) record a first \reprozip trace, (2) save
+transient files, and (3) compare results in the second condition. This is
+one more execution than the theoretical minimum of 2. It might be possible
+to further reduce this overhead, by executing at step (2) only the
+processes depending on transient files.
 
-Applicability: no instrumentation required, fast.
-But not able to capture dependencies in memory.
+We demonstrated the applicability of our approach by evaluating two of the
+arguably most complex pipelines in neuroimaging. Technically, these
+pipelines consist of a mix of tools assembled from different toolboxes
+through a variety of scripts written in different languages. Our file-based
+approach, notably enabled by \reprozip, was able to analyze these pipelines
+without requiring their instrumentation, which saved a very substantial
+technical effort. The assumptions made on the pipeline structure, related
+to the absence of concurrent writes, were not violated in our analysis, and
+are likely to not impede \toolname's applicability to the most common
+neuroimaging pipelines.
 
-We identified four processes that create differences 
-among all 28 distinct processes that are involved in the PreFreeSurfer execution. 
-The numerical instability in the 
-PreFreeSurfer HCP pipeline arises mainly from linear 
-registration processes implemented in FSL \emph{FLIRT}. 
-We can see that \emph{FLIRT} creates differences in almost all the sub-pipelines. 
-We monitored the library system calls using \emph{ltrace} to check that registration processes
-use the same libraries. There were no differences in the library dependencies used by 
-the tool in two operating systems, but we found a different number of calls to the \emph{glibc}.
-
-In the second experiment, 
-the analyses obtained by the FreeSurfer show substantial between-OS differences.
-These results show the sensitivity of the segmentation process to the 
-size of regions, however, it might be also sensitive to the location and shape of 
-the segmented regions.
-We demonstrate that static build of the pipeline would not improve the numerical stability of the pipeline. 
-FreeSurfer pipeline propagates and amplifies the numerical differences that are created by the 
-differences in the underlying operating system libraries.
-These results are consistent with the findings in the paper~\cite{Glatard2015}, 
-as it has confirmed the propagation of differences in cortical thickness extraction analysis using FreeSurfer.
-
-We obtained results between OS versions; further investigations could be done over the OS families (e.g., Linux VS. Windows).
-In addition to variations of the operating system, numerical instability of the pipelines have been 
-explored when varying software, hardware and implementation details, or even by adding negligible 
-perturbations like one voxel noise in the input dataset.
-In the field of numerical analysis, we can investigate these issue using the Monte Carlo Arithmetic (MCA), 
-a method instrumenting floating-point operations based on the random rounding algorithms at a target precision.
-By this algorithm, we can locate the instabilities for the given pipelines, tools, or operations.
-
-We demonstrate that computational models in neuroimaging are sensitive to numerical instabilities. 
-Although we identify the processes in the pipeline that create differences, 
-the amplification of these differences should be quantified in the future works.
-We can use distance-preserving hashes rather than MD5 checksums or metrics specific to the file type like the Dice coefficient 
-for the registration and segmentation processes to measure the magnitude of differences.
-
-% mention fuzzy libmath as a way to do this across distributions.
+However, file-based analyses also have limitations related to the
+granularity at which they operate. Indeed, differences can only be
+identified at the level of an entire operating-system process, which can
+correspond to arbitrary amounts of code. Narrowing down the analysis to
+particular libraries, functions, or even code sections would require
+another approach. Similarly, \toolname would not be able to detect
+differences in data not saved in files, and passed to subsequent processes
+in memory. A common scenario in neuroimaging pipelines is that tools return
+results in their standard output, which is parsed by the calling process
+and passed to subsequent ones through variables. Differences that occur in data 
+that is processed in this way would not be detected by \toolname.
 
@@
 
-Our technique can characterize the stability of the pipelines
-automatically. We identified the leading causes of the between-OS differences as (1) the evolution 
-of the math libraries over time and (2) the instability of the pipelines. 
-There are two ways to tackle this problem. The easy but less preferred solution is masking the instabilities.
-This can be done by using, (1) single operating system for the processing of subjects, 
-(2) containerizing the pipelines so that the 
-processing is done on a more controlled environment, 
-(3) increasing the numerical precision of the arithmetics, 
-(4) building static executables by removing the host operating system library dependencies. 
-These solutions only make the problem invisible, but the pipelines are still unstable.
+Summary of key findings.
 
-\note{The preferred solution is to find the origin of amplification of differences and 
-then fixing the particular functions/modules in the processing pipelines. 
-Bagging is a fundamental technique toward reducing the instability of the analysis.
-Therefore, we can use bagging for the parts in the pipeline that amplify differences, 
-but it still is computationally expensive. }
-
-% importance of data-driven evaluations, subject clustering, etc
-
-% future work with MCA
-
-% link registration to our motion estimation study
+Spot works well but is a first step toward deeper investigation of 
+particular processes.
 

### 2020-04-03 paper.tex
add figure script and update some comments

@@
 disambiguated, resulting in a provenance \emph{DAG} (Directed Acyclic
-Graph) in which file nodes have a single parent (in-degree of 1) (Step 3)
-\tristan{Ali, this is not true in Figure 1: at step 3 there are still
-transient files}. DAGs produced in the two conditions are then compared, in
+Graph) in which file nodes have a single parent (in-degree of 1) (Step 3). 
+DAGs produced in the two conditions are then compared, in
 a step-by-step execution that prevents the propagation of differences in the pipeline (Step
@@
 when complete different processing paths are used in different subjects. \toolname 
-includes helpers to identify different instances of provenance graphs \tristan{mention tree distance here}.
+includes helpers to identify different instances of provenance graphs.
+For this purpose, \toolname measures the distance between process trees using 
+the tree edit distance method~\cite{zhang1989simple}, as implemented in 
+the zss Python package. This distance is defined as the minimum number of
+edit operations to transform one tree into the other. Three edit operations
+are considered: node label modification, node removal, and node insertion.
+Each operation has an associated cost of 1.
 
@@
 \begin{figure}
-\begin{subfigure}{0.4\linewidth}
-  [ FIGURE GOES HERE ]
-
+\begin{subfigure}[t]{0.4\linewidth}
+  \centering
+  \includegraphics[width=.7\linewidth]{images/p-graph}
   \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
@@
 \hfill
-\begin{subfigure}{0.4\linewidth}
-  [ FIGURE GOES HERE ]
-
+\begin{subfigure}[t]{0.4\linewidth}
+  \centering
+  \includegraphics[width=.7\linewidth]{images/p-graph-dag}
   \caption{Provenance DAG, with disambiguated transient files.}
@@
 \end{itemize}
-And FreeSurfer consists of the following ones: \tristan{Check these steps}
+And FreeSurfer consists of the following ones:
 \begin{itemize}
@@
 \item T1w image registration, 
+\item T1w image segmentation,
 \item Surface placement, 
@@
 For each subject, available data consisted of 1 or 2 T1 images, and 1 or 2 T2 images. 
+The voxel size of the acquired T1 and T2 images was 0.7 × 0.7 × 0.7 mm, 
+and the dimension of the images was 256 x 320 x 320.
 % \tristan{Can you let me know where the data is located, so that I could describe it here?
@@
 % and outputs are in consider:/data/asalari/ali-tests/nurm-out}
-Acquisition protocols and parameters are detailed in~\cite{van2013wu}. \tristan{Add summary of matrix sizes.}
+Acquisition protocols and parameters are detailed in~\cite{van2013wu}.
 
@@
 and Atlas Registration). Differences were also observed in image mean 
-computations with FSL maths  (in Anatomical Average). \tristan{figure and caption to be fixed.}
-
-\tristan{Ali, it would be useful to show a complete PreFreesurfer graph here.}
-
+computations with FSL maths  (in Anatomical Average). 
+A complete provenance graph of the PreFreeSurfer is illustrated in Figure~\ref{fig:complete_pfs}
+\begin{figure}
+  \centering
+    \includegraphics[width=\columnwidth]{images/pfs-intransient} 
+    \caption{A complete provenance graph from the PreFreesurfer pipeline.
+     Full-resolution image is available 
+     \href{https://drive.google.com/file/d/1xonnMq4hLEe6GwMp4P7umikvrIAeflDe/view?usp=sharing}{here}.}
+} 
+    \label{fig:complete_pfs}
+\end{figure}
 % Between subject differences can be a result of different subject types. 
@@
   \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 
-  \tristan{Remove frequency values as they are shown by the color already. Remove multipliers, multiply cells instead. Cells should just contain subject numbers.}
   Each cell represents the occurrence of a particular command line in a
@@
   the FLIRT tool was invoked 6 times in step DC for each of the 20
-  subjects: 2 instances weren't reproducible in 95\% of the subjects, 3
+  subjects: 2 instances weren't reproducible in 19 subjects, 3
   instances were always reproducible, and 1 instance wasn't reproducible in
-  85\% of the subjects.
+  17 subjects.
         }
@@
     \caption{Differences between T2 FNIRT results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
-    CentOS7) \tristan{Improve contrast}. An animated version of the comparison is available 
+    CentOS7). An animated version of the comparison is available 
     \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
@@
 \tristan{Ali, could you check that this is indeed the only dynamic
-executable in the pipeline?}. \texttt{mris\_make\_surfaces} produced different results for
+executable in the pipeline?
+Ali: There is another dynamically-linked executable named mris\_curvature 
+that doesn't create difference when set a seed for that.
+Also, another executables are in directory freesurfer/mni/bin/ such as volume\_hist, 
+mincinfo, mincresample, \dots
+that are dynamically-linked but do not create differences.}. 
+\texttt{mris\_make\_surfaces} produced different results for
 10 out of 20 subjects.
@@
   \begin{subfigure}{0.45\linewidth}
-    [FIGURE GOES HERE]
+    \includegraphics[width=\linewidth]{images/scatter_plot_1.png}
     \caption{Regions with \emph{less} than 10,000 voxels.}
@@
   \begin{subfigure}{0.45\linewidth}
-    [FIGURE GOES HERE]
+    \includegraphics[width=\linewidth]{images/scatter_plot_2.png}
     \caption{Regions with \emph{more} than 10,000 voxels.}
@@
     \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).
-        To improve readability, the two graphs are on different y scales. \tristan{Ali, could you separate the figures in 2 and insert them in each subfigure?}} 
+        To improve readability, the two graphs are on different y scales.}
     \label{fig:scatter_plot}
@@
 The processes identified by \toolname as non-reproducible were all
-associated with dynamically-linked executables \tristan{Ali, please check
-that}. This makes complete sense as statically-linked executables wouldn't
+associated with dynamically-linked executables. 
+This makes complete sense as statically-linked executables wouldn't
 be impacted by library updates. Moreover, the hypothetical effects of

### 2020-04-03 bin/code/Dice_region_csv.py

@@ -1,0 +1,138 @@
+#!/usr/bin/env python
+
+import argparse
+import logging
+import nibabel
+import sys
+import os.path as op
+import numpy as np
+import numpy.ma as ma
+import seaborn as sns
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+import csv
+# from dask import delayed
+# import dask.array as da
+
+
+def log_error(message):
+    logging.error(message)
+    sys.exit(1)
+
+
+def remove_uncommons(regions2, regions, voxels_2):
+    uncommon2 = np.setdiff1d(regions2, regions).tolist()
+    for i in uncommon2:
+        ind = np.where(regions2 == i)[0].tolist()[0]
+        regions2 = np.delete(regions2, ind)
+        voxels_2 = np.delete(voxels_2, ind)
+    return regions2, voxels_2
+
+
+def main():
+    parser = argparse.ArgumentParser(description="Computes Dice coefficient \
+                                     of the regions between images.")
+    parser.add_argument("first_image",
+                        help='First input T1W segmented image like aseg.nii')
+    parser.add_argument("second_image",
+                        help='Second input T1W segmented image like aseg.nii')
+    parser.add_argument("output",
+                        help='Output CSV file')
+
+    args = parser.parse_args()
+    output = args.output
+    # Load images using nibabel
+    im1 = nibabel.load(args.first_image)
+    im2 = nibabel.load(args.second_image)
+
+    # Check that both images have the same dimensions
+    # shape1 = im1.header.get_data_shape()
+    # shape2 = im2.header.get_data_shape()
+    data1 = im1.get_data()
+    data2 = im2.get_data()
+    subj_info = {}
+    voxel_counts = {}
+
+    region_labels = {0: 'Background', 42: 'Right-Cerebral-Cortex',
+                     3: ' Left-Cerebral-Cortex',
+                     41: 'Right-Cerebral-White-Matter',
+                     2: 'Left-Cerebral-White-Matter',
+                     47: 'Right-Cerebellum-Cortex',
+                     8: 'Left-Cerebellum-Cortex',
+                     16: 'Brain-Stem', 46: 'Right-Cerebellum-White-Matter',
+                     7: 'Left-Cerebellum-White-Matter',
+                     10: 'Left-Thalamus-Proper*',
+                     49: 'Right-Thalamus-Proper*',
[…]

### 2020-04-03 bin/code/binarized_heatmap_img.py

@@ -1,0 +1,83 @@
+#!/usr/bin/env python
+
+import argparse
+import logging
+import nibabel
+import sys
+import os
+import numpy as np
+import numpy.ma as ma
+import seaborn as sns
+import matplotlib.pyplot as plt
+import matplotlib as mpl
+
+
+def log_error(message):
+    logging.error(message)
+    sys.exit(1)
+
+
+def main():
+    parser = argparse.ArgumentParser(description="Computes a \
+                                     binarized heatmap of the images.")
+    parser.add_argument("input_imgs",
+                        help='Input folder of binarized difference images')
+    parser.add_argument("output_image_file",
+                        help='output folder')
+    parser.add_argument("ref_image_file",
+                        help='reference image like MNI-152')
+
+    args = parser.parse_args()
+    # Load images using nibabel
+    img_sum = ''
+    for img in os.listdir(args.input_imgs):
+        if os.path.splitext(os.path.basename(img))[1] in ['.nii', '.gz']:
+            img1 = nibabel.load(os.path.join(args.input_imgs, img))
+            data1 = img1.get_data()
+            if img_sum == '':
+                img_sum = data1
+                continue
+            img_sum = img_sum + data1
+    im_ref = nibabel.load(args.ref_image_file)
+    im_data_ref = im_ref.get_data()
+
+    # Check that both images have the same dimensions
+    # shape1 = im1.header.get_data_shape()
+    # shape2 = im_ref.header.get_data_shape() 
+    hor_view = img_sum[129,:,:]
+    hor_view_ref = im_data_ref[98,:,:]
+    ver_view = img_sum[:,155,:]
+    ver_view_ref = im_data_ref[:,116,:]
+    axi_view = img_sum[:,:,130]
+    axi_view_ref = im_data_ref[:,:,94]
+
+    # Heatmap plots
+    startcolor = '#990033'
+    midcolor = '#ffff00'
+    endcolor = '#FFFFFF'
+    own_cmap1 = mpl.colors.LinearSegmentedColormap.from_list(
+                    'own2', [startcolor, midcolor, endcolor])
+    fig = plt.figure(figsize=(25,10), facecolor='black')
[…]

### 2020-04-03 bin/code/provenance_graph.py

@@ -1,0 +1,230 @@
+import zss
+import argparse
+import os
+import re
+import sqlite3
+import hashlib
+import json
+from sqlite3 import Error
+from graphviz import Digraph as Di
+
+
+# Returns the children of the process
+def get_the_child_processes(process_cursor, pid):
+    process_id_query = '''
+            SELECT id
+            FROM processes
+            WHERE parent = %s
+            '''
+    process_cursor.execute(process_id_query % pid)
+    child_list = process_cursor.fetchall()
+    chlst = []
+    for child2 in child_list:
+        chlst.append(child2[0])
+    return chlst
+
+
+# Returns the process name
+def get_the_processes_name(executed_cursor, pid):
+    process_name_query = '''
+                SELECT name, argv
+                FROM executed_files
+                WHERE process = %s
+                '''
+    executed_cursor.execute(process_name_query % pid)
+    process_name = executed_cursor.fetchall()
+    if process_name != []:
+        process_name = str(process_name[0][0]).split('/')[-1:][0]
+    else:
+        process_name = ""
+    return process_name
+
+
+# Returns all opened files (both W/R modes)
+def get_opened_files_list(openfile_cursor, pid):
+    opened_files_query = '''
+            SELECT process, name, mode, id
+            FROM opened_files
+            WHERE process = %s AND mode <= 2
+            '''
+    openfile_cursor.execute(opened_files_query % pid)
+    return openfile_cursor.fetchall()
+
+
+def create_provenance_graph(db_path, pid, graph_, pipe_list=[], multi_list={},
+                            tmp_list=[], total_proc_diff=[], dag=False):
+    try:
+        db = sqlite3.connect(db_path)
+    except Error as e:
+        print(e)
+    process_cursor = db.cursor()
+    executed_cursor = db.cursor()
+    openfile_cursor = db.cursor()
+    # Get the list of child process from pid
+    child_list = get_the_child_processes(process_cursor, pid)
+    # Get the process name
+    process_name = get_the_processes_name(executed_cursor, pid)
+    # Get the list of opened files (w/r) from pid
+    process_ofiles = get_opened_files_list(openfile_cursor, pid)
+
+    # graph_.attr('node', shape='circle', style='filled', fillcolor='white')
+    if pid in total_proc_diff:
+        graph_.node(str(pid), ''.join([str(pid), '#', process_name]),
+                    shape='circle', style='filled', fillcolor='red')
+    elif total_proc_diff != []:
[…]

### 2020-04-04 paper.tex
small fix
Fixed figure 8

@@
 or transformations differ. To compare segmentations, we used the Dice coefficient defined as follows:
-\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\].
+\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\]
 
@@
      \href{https://drive.google.com/file/d/1xonnMq4hLEe6GwMp4P7umikvrIAeflDe/view?usp=sharing}{here}.}
-} 
     \label{fig:complete_pfs}
@@
 
-While by design \toolname only identifies processes that create
-differences between conditions, pipeline results may still differ due to
-the propagation and amplification of differences created at previous
-processing steps, PreFreeSurfer in our case. We observed the effect of this
+While by design \toolname only identifies processes that create differences
+between conditions, pipeline results may still differ due to the
+propagation and amplification of differences created at previous processing
+steps, PreFreeSurfer in our case. We observed the effect of this
 propagation in FreeSurfer results, as shown in
@@
 the Dice coefficients associated with the 44 regions segmented by
-FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing low Dice
-values (below 0.5) in the smallest structures.
+FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing, for some subjects, Dice values
+below 0.9 in structures such as the corpus callosum, pallidum, amygdala,
+accumbens area, and ventricles.
 
@@
 
-\begin{figure*}
-  \centering
-  \begin{subfigure}{0.45\linewidth}
-    \includegraphics[width=\linewidth]{images/scatter_plot_1.png}
-    \caption{Regions with \emph{less} than 10,000 voxels.}
-  \end{subfigure}
-  \hfill
-  \begin{subfigure}{0.45\linewidth}
-    \includegraphics[width=\linewidth]{images/scatter_plot_2.png}
-    \caption{Regions with \emph{more} than 10,000 voxels.}
-  \end{subfigure}
-    % \includegraphics[width=\textwidth]{images/scatter_plot.png} 
-    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).
-        To improve readability, the two graphs are on different y scales.}
+\begin{figure}
+  \hspace*{-1cm}
+  \includegraphics[width=1.2\linewidth]{images/dice_regions.pdf}
+    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).}
     \label{fig:scatter_plot}
-  \end{figure*}
+  \end{figure}
 

### 2020-04-04 bin/code/regions.py

@@ -1,0 +1,89 @@
+from matplotlib import pyplot as plt
+import matplotlib._color_data as mcd
+import random
+
+
+plt.figure(figsize=(15,15))
+
+with open('./data/fs_seg_dice_accumulated_20sbj.csv') as f:
+    lines = f.readlines()
+
+
+dices = {}
+sizes = {}
+for i, line in enumerate(lines):
+    if i == 0:
+        regions = line.split(',')
+        regions = [ x.replace('-', ' ').replace('_', ' ').replace('\n', '') for x in regions]
+        continue
+    if i % 2 == 0:
+        pointer = sizes 
+    else:
+        pointer = dices
+    vals = [float(x) for x in line.split(',')]
+    for j, val in enumerate(vals):
+        region = regions[j]
+        if pointer.get(region) is None:
+            pointer[region] = []
+        pointer[region] += [ val ]
+
+
+palette_colors = [ "#000000", "#00FF00", "#0000FF", "#FF0000", "#01FFFE", "#FFA6FE", "#FFDB66", "#006401", "#010067", "#95003A", "#007DB5", "#FF00F6", "#FFEEE8", "#774D00", "#90FB92", "#0076FF", "#D5FF00", "#FF937E", "#6A826C", "#FF029D", "#FE8900", "#7A4782", "#7E2DD2", "#85A900", "#FF0056", "#A42400", "#00AE7E", "#683D3B", "#BDC6FF", "#263400", "#BDD393", "#00B917", "#9E008E", "#001544", "#C28C9F", "#FF74A3", "#01D0FF", "#004754", "#E56FFE", "#788231", "#0E4CA1", "#91D0CB", "#BE9970", "#968AE8", "#BB8800", "#43002C", "#DEFF74", "#00FFC6", "#FFE502",
+                   "#620E00", "#008F9C", "#98FF52", "#7544B1", "#B500FF", "#00FF78", "#FF6E41", "#005F39", "#6B6882", "#5FAD4E", "#A75740", "#A5FFD2" ]
+
+
+# pick a color or left/right region
+colors={}
+nc = 0
+for i, region in enumerate(regions):
+    c = region.replace('Left', '').replace('Right', '')
+    if colors.get(c) is None:
+        colors[c] = palette_colors[nc]
+        nc = nc + 1
+    colors[region] = colors[c]
+
[…]

### 2020-04-05 bin/code/regions.py
Improved Fig 8

@@ -27,63 +27,28 @@ for i, line in enumerate(lines):
             pointer[region] = []
         pointer[region] += [ val ]
 
+# Sort regions by median dice
+from statistics import mean, median
+sorted_regions = sorted(regions, key=lambda x: median(dices[x]))
 
-palette_colors = [ "#000000", "#00FF00", "#0000FF", "#FF0000", "#01FFFE", "#FFA6FE", "#FFDB66", "#006401", "#010067", "#95003A", "#007DB5", "#FF00F6", "#FFEEE8", "#774D00", "#90FB92", "#0076FF", "#D5FF00", "#FF937E", "#6A826C", "#FF029D", "#FE8900", "#7A4782", "#7E2DD2", "#85A900", "#FF0056", "#A42400", "#00AE7E", "#683D3B", "#BDC6FF", "#263400", "#BDD393", "#00B917", "#9E008E", "#001544", "#C28C9F", "#FF74A3", "#01D0FF", "#004754", "#E56FFE", "#788231", "#0E4CA1", "#91D0CB", "#BE9970", "#968AE8", "#BB8800", "#43002C", "#DEFF74", "#00FFC6", "#FFE502",
-                   "#620E00", "#008F9C", "#98FF52", "#7544B1", "#B500FF", "#00FF78", "#FF6E41", "#005F39", "#6B6882", "#5FAD4E", "#A75740", "#A5FFD2" ]
+data = [ dices[region] for region in sorted_regions ]
+mean_sizes = [ [ mean(sizes[region]) for x in sizes[region]] for region in sorted_regions ]
 
+from math import log
+plt.boxplot(data, showfliers=False,
+            labels = range(len(sorted_regions)))
+plt.gca().xaxis.grid(True)
 
-# pick a color or left/right region
-colors={}
-nc = 0
-for i, region in enumerate(regions):
-    c = region.replace('Left', '').replace('Right', '')
-    if colors.get(c) is None:
-        colors[c] = palette_colors[nc]
-        nc = nc + 1
-    colors[region] = colors[c]
+for i, region in enumerate(sorted_regions):
+    plt.scatter( [i+1 for x in dices[region]], dices[region],
+                 s=30,
+                 #c=colors[region], 
+                 label=f'{i} - {region}',
+                 marker='.')
 
-prefixes = [ 'left', 'right', 'inf', 'lat', 'lateral', '3rd', '4th', 'non' ]
-
[…]

### 2020-04-05 paper.tex
Fixed Fig 2

@@
 % Furthermore, we can stabilize the identified processes in the next steps.
-\todo{explain creation vs propagation}
+\todo{explain creation vs proSpagation}
 \todo{define pipeline}
+\todo{explain the need for reduced instrumentation}
 
@@
 \toolname identifies the processes in a pipeline that produce different
-results when executed in two different conditions. First, a directed
+results in different execution conditions. First, a directed
 bipartite provenance graph is recorded for each subject, where nodes
@@
 execution or modified by multiple processes, are identified (Step 2) and
-disambiguated, resulting in a provenance \emph{DAG} (Directed Acyclic
+disambiguated, resulting in a provenance DAG (Directed Acyclic
 Graph) in which file nodes have a single parent (in-degree of 1) (Step 3). 
-DAGs produced in the two conditions are then compared, in
+DAGs produced in different conditions are then compared, in
 a step-by-step execution that prevents the propagation of differences in the pipeline (Step
-4). Finally, a labeled graph is produced that identifies the
+4). The resulting labeled graph identifies the
 non-reproducible processes in the pipeline.
@@
 To ensure that a file can be unambiguously associated with the process that
-created it, we assume that pipeline processes can be re-ordered such that:
+created it, we assume that the pipeline can be transformed such that:
 \begin{enumerate}
@@
 blocks are processed sequentially. However, only a single version of the
-file must eventually be made available to the other processes. Similarly,
-in case processes write and delete temporary files throughout their
-execution, these temporary files must not be read by any other process in
-the pipeline.
-
-In addition, we also require that processes are
+file must eventually be made available to the other processes. In
+particular, in case a process deletes a file that it had created itself,
+this file must not be used by any other process. Finally, we also require that processes are
 associated to a command line (executable and arguments), to facilitate
@@
 % purposes. 
-Some of these differences can be neglected, for instance when a
-data decompression step is present at the beginning of the execution for
-some subjects only, but other differences cannot, for instance
-when complete different processing paths are used in different subjects. \toolname 
-includes helpers to identify different instances of provenance graphs.
-For this purpose, \toolname measures the distance between process trees using 
-the tree edit distance method~\cite{zhang1989simple}, as implemented in 
-the zss Python package. This distance is defined as the minimum number of
-edit operations to transform one tree into the other. Three edit operations
-are considered: node label modification, node removal, and node insertion.
-Each operation has an associated cost of 1.
+Some of these differences can be neglected, for instance when a data
+decompression step is present at the beginning of the execution for some
+subjects only, but other differences cannot, for instance when complete
+different processing paths are used in different subjects. \toolname
+includes helpers to identify different instances of provenance graphs. In
+particular, it can cluster process trees, where nodes are processes and
+edges are \texttt{fork()} or \texttt{clone()} system calls, using the tree
+edit distance~\cite{zhang1989simple} implemented in Python's \texttt{zss} package.
 
@@
   \centering
-  \includegraphics[width=.7\linewidth]{images/p-graph}
+  \includegraphics[width=.7\linewidth]{images/p-graph.pdf}
   \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
@@
   \centering
-  \includegraphics[width=.7\linewidth]{images/p-graph-dag}
+  \includegraphics[width=.7\linewidth]{images/p-graph-dag.pdf}
   \caption{Provenance DAG, with disambiguated transient files.}
@@
 
-\subsection{Labelling processes}
+\subsection{Labeling processes}
 
@@
 in Condition 1. If no differences are found, the process is marked as
-reproducible. Otherwise, the process is marked as non-reproducible and the
+reproducible. Otherwise, the process is marked as non-reproducible, and the
 output files produced in Condition 1 are copied to Condition 2, to ensure
@@
 
-Figure~\ref{fig:iterations} illustrates the incremental labelling 
+Figure~\ref{fig:iterations} illustrates a hypothetical incremental labeling 
 process for the example in Figure~\ref{fig:simple_script}. 
 At step 1, process \texttt{73\#bet2} is labeled as non-reproducible (red) 
-as it produces files with differences. Subsequently, the files produced by \texttt{73\#bet2} in  
+as it produces files with differences, and the files produced by \texttt{73\#bet2} in  
 Condition 2 are replaced with the files produced by \texttt{73\#bet2} in 
-Condition 1.
-At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
+Condition 1. At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
 as reproducible (green) as it produced files without differences.
@@
   \centering
-  \includegraphics[width=.8\columnwidth]{images/labelling-process}
-  \caption{Illustration of the labelling approach. Non-reproducible processes are shown in red.}
+  \includegraphics[width=.8\columnwidth]{images/p-graph-dag-labelled.pdf}
+  \caption{Illustration of the labeling approach. Non-reproducible processes are shown in red.}
   \label{fig:iterations}
@@
 
-We applied the \toolname to find the origin of between-OS differences that
-frequently occur in neuroimaging pipelines. In particular, we studied the minimal
+We applied the \toolname to the minimal
 pre-processing pipelines released by the Human Connectome Project
@@
 \centering
-  \includegraphics[width=\textwidth]{images/pfs_heatmap.png}
+  \includegraphics[width=\textwidth]{./data/pfs-20sbj-output/heatmap.png}
   \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 
@@
 
+Mention non-instrumentation / reprozip
+
 Spot works well but is a first step toward deeper investigation of 

### 2020-04-06 paper.tex
Cleaned up figures
Fixed fig 2 and 3 (now fig 1)

@@
 
-Data analysis pipelines are known to be impacted by operating system
-updates and other infrastructural factors, presumably due to the creation,
-propagation and amplification of numerical errors. However, the precise
-causes of such instabilities and the path along which they propagate in the
-pipelines are unclear, while this process could play a major role in the
-current reproducibility crisis. We present a technique to identify the
-computational processes in a pipeline that create numerical differences
-when executed in different computational conditions. Our technique
-leverages system-call interception to reconstruct processing and file
-graphs without pipeline instrumentation. By applying our tool to the
-structural pre-processing pipelines released by the Human Connectome
-Project, we find that linear and non-linear registration are the cause of
-most numerical instabilities in such pipelines, which confirms previous
-findings.
+% Data analysis pipelines are known to be impacted by operating system
+% updates and other infrastructural factors, presumably due to the creation,
+% propagation and amplification of numerical errors. However, the precise
+% causes of such instabilities and the path along which they propagate in the
+% pipelines are unclear, while this process could play a major role in the
+% current reproducibility crisis. We present a technique to identify the
+% computational processes in a pipeline that create numerical differences
+% when executed in different computational conditions. Our technique
+% leverages system-call interception to reconstruct processing and file
+% graphs without pipeline instrumentation. By applying our tool to the
+% structural pre-processing pipelines released by the Human Connectome
+% Project, we find that linear and non-linear registration are the cause of
+% most numerical instabilities in such pipelines, which confirms previous
+% findings.
 
@@
 
+\newpage
+
 \section{Tool description}
@@
 represent application processes and files, and edges represent read and
-write file accesses (Step 1 in Figure~\ref{fig:overview-tool}). Second,
+write file accesses (Figure~\ref{fig:provenance-graph}). Second,
 transient files, i.e., files that are either deleted during pipeline
-execution or modified by multiple processes, are identified (Step 2) and
+execution or modified by multiple processes, are identified and
 disambiguated, resulting in a provenance DAG (Directed Acyclic
-Graph) in which file nodes have a single parent (in-degree of 1) (Step 3). 
+Graph) in which file nodes have a single parent (in-degree of 1) (Figure~\ref{fig:provenance-dag}). 
 DAGs produced in different conditions are then compared, in
-a step-by-step execution that prevents the propagation of differences in the pipeline (Step
-4). The resulting labeled graph identifies the
+a step-by-step execution that prevents the propagation of differences in the pipeline (Figure~\ref{fig:labeled-dag}).
+ The resulting labeled graph identifies the
 non-reproducible processes in the pipeline.
+
+\begin{figure*}
+  \begin{subfigure}[t]{0.3\linewidth}
+    \centering
+    \includegraphics[width=.5\linewidth]{images/p-graph.pdf}
+    \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
+    \label{fig:provenance-graph}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.3\linewidth}
+    \centering
+    \includegraphics[width=.4\linewidth]{images/p-graph-dag.pdf}
+    \caption{Provenance DAG, with disambiguated transient files.}
+    \label{fig:provenance-dag}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}[t]{0.3\linewidth}
+      \centering
+      \includegraphics[width=.4\linewidth]{images/p-graph-dag-labelled.pdf}
+     \caption{Labeled DAG comparing 2 execution conditions, showing 1 non-reproducible process.}
+     \label{fig:labeled-dag} 
+  \end{subfigure}
+    % \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
+    \caption{Provenance graphs
+    created from the example pipeline in
+    Listing~\ref{listing:sample-script}.
+    Processes are represented with circles, files with rectangles, and read/write accesses 
+    with plain edges. 
+    For visualization purposes, the process tree is represented too, with
+    dashed edges. Every node in the graph is labeled using (1) a process id
+    created by our reconstruction, (2) the name of the executable run by the
+    process or the file name. Process 1 is the initial call to the
+    \texttt{sh} interpreter; the other processes are the calls to FSL bet,
+    stats and maths made in Listing~\ref{listing:sample-script}. Process 73
+    is forked by process 2: it was captured by \reprozip while it did not
+    appear in Listing~\ref{listing:sample-script}.
+  }
+    \label{fig:spot-example}
+  \end{figure*}
 
@@
 
-\begin{figure*}
-  \centering
-    \includegraphics[width=.8\textwidth]{images/spot-diagram}
-    \caption{\toolname overview.}
-    \label{fig:overview-tool}
-  \end{figure*}
+\begin{listing}
+  \inputminted{bash}{"bin/example.sh"}
+  \caption{Example pipeline}
+  \label{listing:sample-script}
+\end{listing}
+\tristan{Ali, I added a file in Listing 1, could you re-generate the trace?}
 
@@
 
-\begin{listing}
-  \inputminted{bash}{"bin/example.sh"}
-  \caption{Example pipeline}
-  \label{listing:sample-script}
-\end{listing}
-
-\begin{figure}
-\begin{subfigure}[t]{0.4\linewidth}
-  \centering
-  \includegraphics[width=.7\linewidth]{images/p-graph.pdf}
-  \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
-  \label{fig:provenance-graph}
-\end{subfigure}
-\hfill
-\begin{subfigure}[t]{0.4\linewidth}
-  \centering
-  \includegraphics[width=.7\linewidth]{images/p-graph-dag.pdf}
-  \caption{Provenance DAG, with disambiguated transient files.}
-  \label{fig:provenance-dag}
-\end{subfigure}
-  % \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
-  \caption{Provenance graphs
-  created from the example pipeline in
-  Listing~\ref{listing:sample-script}.
-  Processes are represented with circles, files with rectangles, and read/write accesses 
-  with plain edges. 
-  For visualization purposes, the process tree is represented too, with
-  dashed edges. Every node in the graph is labeled using (1) a process id
-  created by our reconstruction, (2) the name of the executable run by the
-  process or the file name. Process 0 is the initial call to the
-  \texttt{sh} interpreter; the other processes are the calls to FSL bet,
-  stats and maths made in Listing~\ref{listing:sample-script}. Process 73
-  is forked by process 2: it was captured by \reprozip while it did not
-  appear in Listing~\ref{listing:sample-script}.
-}
-  \label{fig:simple_script}
-\end{figure}
-
 \subsection{Capturing transient files}
@@
 
-After capturing transient files, we re-run the pipeline
-step by step in Condition 2 to label processes. The output files
-produced by a process in Condition 2 are compared to the ones produced
-in Condition 1. If no differences are found, the process is marked as
+After capturing transient files in the first condition, we re-run the pipeline
+step by step in the second one to label processes. The output files
+produced by a process in both conditions are compared.
+ If no differences are found, the process is marked as
 reproducible. Otherwise, the process is marked as non-reproducible, and the
-output files produced in Condition 1 are copied to Condition 2, to ensure
+output files produced in the first condition to the second one, to ensure
 that differences do not propagate further in the pipeline. Processes are
@@
 
-Figure~\ref{fig:iterations} illustrates a hypothetical incremental labeling 
-process for the example in Figure~\ref{fig:simple_script}. 
-At step 1, process \texttt{73\#bet2} is labeled as non-reproducible (red) 
-as it produces files with differences, and the files produced by \texttt{73\#bet2} in  
-Condition 2 are replaced with the files produced by \texttt{73\#bet2} in 
-Condition 1. At step 2, process \texttt{75\#fslmaths} is executed and then labeled 
-as reproducible (green) as it produced files without differences.
-At step 3, \texttt{76\#fslstats} is also labeled as 
-reproducible as it didn't produce any file.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=.8\columnwidth]{images/p-graph-dag-labelled.pdf}
-  \caption{Illustration of the labeling approach. Non-reproducible processes are shown in red.}
-  \label{fig:iterations}
-\end{figure}
+Figure~\ref{fig:labeled-dag} illustrates a hypothetical incremental
+labeling of the example in Listing~\ref{listing:sample-script}. Process
+\texttt{73\#bet2} is labeled as non-reproducible (red) as it produces files
+with differences. To prevent the propagation of these differences, the
+files produced by \texttt{73\#bet2} in Condition 2 are replaced with the
+files produced by \texttt{73\#bet2} in Condition 1. Processes
+\texttt{75\#fslmaths} and \texttt{76\#fslstats} are then executed and
+labeled as reproducible (green) as they produce files without differences.
 

### 2020-04-06 paper.tex
Update Fig4

--- 2020-04-04 paper.tex
@@
 or transformations differ. To compare segmentations, we used the Dice coefficient defined as follows:
-\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\]
+\[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\].
 
@@
   \centering
-    \includegraphics[width=\columnwidth]{images/pfs-intransient} 
+    \includegraphics[width=\columnwidth]{images/pfs-labeled.png} 
     \caption{A complete provenance graph from the PreFreesurfer pipeline.
      Full-resolution image is available 
-     \href{https://drive.google.com/file/d/1xonnMq4hLEe6GwMp4P7umikvrIAeflDe/view?usp=sharing}{here}.}
+     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
+} 
     \label{fig:complete_pfs}
@@
 
-While by design \toolname only identifies processes that create differences
-between conditions, pipeline results may still differ due to the
-propagation and amplification of differences created at previous processing
-steps, PreFreeSurfer in our case. We observed the effect of this
+While by design \toolname only identifies processes that create
+differences between conditions, pipeline results may still differ due to
+the propagation and amplification of differences created at previous
+processing steps, PreFreeSurfer in our case. We observed the effect of this
 propagation in FreeSurfer results, as shown in
@@
 the Dice coefficients associated with the 44 regions segmented by
-FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing, for some subjects, Dice values
-below 0.9 in structures such as the corpus callosum, pallidum, amygdala,
-accumbens area, and ventricles.
+FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing low Dice
+values (below 0.5) in the smallest structures.
 
@@
 
-\begin{figure}
-  \hspace*{-1cm}
-  \includegraphics[width=1.2\linewidth]{images/dice_regions.pdf}
-    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).}
+\begin{figure*}
+  \centering
+  \begin{subfigure}{0.45\linewidth}
+    \includegraphics[width=\linewidth]{images/scatter_plot_1.png}
+    \caption{Regions with \emph{less} than 10,000 voxels.}
+  \end{subfigure}
+  \hfill
+  \begin{subfigure}{0.45\linewidth}
+    \includegraphics[width=\linewidth]{images/scatter_plot_2.png}
+    \caption{Regions with \emph{more} than 10,000 voxels.}
+  \end{subfigure}
+    % \includegraphics[width=\textwidth]{images/scatter_plot.png} 
+    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).
+        To improve readability, the two graphs are on different y scales.}
     \label{fig:scatter_plot}
-  \end{figure}
+  \end{figure*}
 

### 2020-04-06 paper.tex

--- 2020-04-06 paper.tex
@@
 \end{listing}
-\tristan{Ali, I added a file in Listing 1, could you re-generate the trace?}
 
@@
 %\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
+The \reprozip traced files show that approximately PreFreeSurfer opened  
+83K files and created 7.7K processes, and FreeSurfer opened 62K files and 
+created 4K processes in total for a single subject.
 
@@
   \centering
-    \includegraphics[width=\columnwidth]{images/pfs-intransient} 
+    \includegraphics[width=\columnwidth]{images/pfs-labeled.png} 
     \caption{A complete provenance graph from the PreFreesurfer pipeline.
      Full-resolution image is available 
-     \href{https://drive.google.com/file/d/1xonnMq4hLEe6GwMp4P7umikvrIAeflDe/view?usp=sharing}{here}.}
+     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
     \label{fig:complete_pfs}
@@
 \centering
-  \includegraphics[width=\textwidth]{./data/pfs-20sbj-output/heatmap.png}
+  \includegraphics[width=\textwidth]{./data/pfs-20sbj-output/heatmap.pdf}
   \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 

### 2020-04-06 paper.tex
Fixes in method

--- 2020-04-06 paper.tex
@@
 \newcommand{\toolname}[0]{Spot\xspace}
+\newcommand{\flirt}[0]{\texttt{flirt}\xspace}
+\newcommand{\fnirt}[0]{\texttt{fnirt}\xspace}
 
@@
 step by step in the second one to label processes. The output files
-produced by a process in both conditions are compared.
- If no differences are found, the process is marked as
-reproducible. Otherwise, the process is marked as non-reproducible, and the
-output files produced in the first condition to the second one, to ensure
+produced by a process in both conditions are compared: if no differences are found, the process is marked as
+reproducible; otherwise, the process is marked as non-reproducible, and the
+output files produced in the first condition are copied to the second one, to ensure
 that differences do not propagate further in the pipeline. Processes are
@@
 
-We applied the \toolname to the minimal
+We applied \toolname to the minimal
 pre-processing pipelines released by the Human Connectome Project
@@
 \begin{itemize}
-\item Distortion Correction (DC), 
-\item Anatomical Average (AAve), 
+\item Gradient Distortion Correction (DC), 
+\item Alignment and Anatomical Average (AAve), T1w(s), T2w(s),  
 \item Anterior/Posterior Commissure Alignment (ACPC-A), 
 \item Brain Extraction (BExt), 
+\item \tristan{Ali, why isn't BBR cross-modal registration here and in the heatmap?}
 \item Bias Field Correction (BFC), 
@@
 available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
-For each subject, available data consisted of 1 or 2 T1 images, and 1 or 2 T2 images. 
-The voxel size of the acquired T1 and T2 images was 0.7 × 0.7 × 0.7 mm, 
-and the dimension of the images was 256 x 320 x 320.
-% \tristan{Can you let me know where the data is located, so that I could describe it here?
-% Subjects are located in consider:/data/asalari/ali-tests/PFS\_Centos6\_Traced
-% and consider:/data/asalari/ali-tests/PFS\_Centos7\_Traced 
-% and outputs are in consider:/data/asalari/ali-tests/nurm-out}
+For each subject, available data consisted of 1 or 2 T1 images, and 1 or 2 T2 images, with 
+256 x 320 x 320 voxels of size 0.7 × 0.7 × 0.7 mm.
 Acquisition protocols and parameters are detailed in~\cite{van2013wu}.
@@
 \texttt{mris\_diff}, and \texttt{lta\_diff} that determine if MRI volumes
-or transformations differ. To compare segmentations, we used the Dice coefficient defined as follows:
+or transformations differ. To compare segmentations $X$ and $Y$, we used the Dice coefficient defined as follows:
 \[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\]
@@
 \texttt{mri\_segreg}, and cortical surface curvature estimations with
-\texttt{mris\_curvature}. To remove these differences, we fixed the random
-seed used by FreeSurfer using the \texttt{--seed} command-line option.
+\texttt{mris\_curvature}. Fixing the random seed used in FreeSurfer removed
+these differences.
 
@@
 
-\begin{tabular}{cccccc}
+\begin{tabular}{cc|cc}
 \toprule
-       &                        &  \multicolumn{4}{c}{Image modalities in subject data}    \\ 
-\cmidrule(lr){3-6}       
-Type   &   \makecell{Number of \\ Subjects}   &  T1w\_1          & T1w\_2      & T2w\_1          & T2w\_2        \\ \midrule
-1      &               9                      &   \ding{51}      &   \ding{51} &   \ding{51}     &   \ding{51}   \\ 
-2      &               8                      &   \ding{51}      &             &   \ding{51}     &               \\ 
-3      &               1                      &   \ding{51}      &             &   \ding{51}     &   \ding{51}   \\ 
-4      &               2                      &   \ding{51}      &   \ding{51} &   \ding{51}     &               \\ 
+Type   &   \makecell{Number of \\ Subjects}   &  \makecell{Number of \\ T1w images}          & \makecell{Number of \\ T2w images}   \\ \midrule
+1      &               9                      &  2 & 2 \\
+2      &               8                      &  1 & 1 \\
+3      &               1                      &  1 & 2 \\
+4      &               2                      &  2 & 1\\
 \bottomrule
@@
 Differences were observed in linear registration 
-with FSL FLIRT (in ACPC-Alignment, Brain Extraction, Distortion Correction, and
-Atlas Registration), in non-linear registration with FSL FNIRT (in Brain Extraction 
+with FSL \flirt (in ACPC-Alignment, Brain Extraction, Distortion Correction, and
+Atlas Registration), in non-linear registration with FSL \fnirt (in Brain Extraction 
 and Atlas Registration), and in image warping with FSL \texttt{new\_invwarp} (in Brain Extraction 
@@
   number of subjects where the corresponding process appears. For example,
-  the FLIRT tool was invoked 6 times in step DC for each of the 20
+  the \flirt tool was invoked 6 times in step DC for each of the 20
   subjects: 2 instances weren't reproducible in 19 subjects, 3
@@
 To visually illustrate the observed differences,
-Figure~\ref{fig:fnirt_result} compares FNIRT results in Brain Extraction
+Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
 for a particular subject. Differences appear to be very substantial, in
@@
     \includegraphics[width=\columnwidth]{images/t2w_alignment.png} 
-    \caption{Differences between T2 FNIRT results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
+    \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
     CentOS7). An animated version of the comparison is available 
@@
 non-reproducible across operating systems starting from the motion
-correction step, a step that uses FSL's \texttt{flirt} tool internally. It
+correction step, a step that uses FSL's \flirt tool internally. It
 would be relevant to investigate if the observed instability of
@@
 a given tool may vary across subjects and across processing parameters. For
-instance, linear registration with FLIRT seems to be fully reproducible in
+instance, linear registration with \flirt seems to be fully reproducible in
 the Anatomical Average sub-pipeline, while it is highly non-reproducible in

### 2020-04-06 bin/code/heatmap_plot.py
Fixed heatmap script

@@ -338,7 +338,7 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
                             # str(float(a)/float(b))+' ['+str(b)+']')
                 if ch is True:
                     freq_list.append(0)
-                    freq_label.append('0') # ['+str(b)+']')
+                    freq_label.append('') # ['+str(b)+']')
             frequency_matrix.append(freq_list)
             heatmap_label.append(freq_label)
             xlabel_ = [str(x).split(' ')[0] for x in xlabel_]
@@ -383,7 +383,7 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
                 axes[j, i].set_ylabel(short_name[element], fontsize=40)
             if j == len(ylabels)-1:
                 axes[j, i].set_xlabel(emp, fontsize=40)
-    plt.savefig(output_folder + 'heatmapn.png', bbox_inches='tight')
+    plt.savefig(os.path.join(output_folder, 'heatmap.pdf'), bbox_inches='tight')
     # plt.show()
 
 
@@ -401,22 +401,13 @@ def plot_totall_items(total_freq, output_folder):
     plt.yticks(fontsize=18, rotation=0)
     # plt.title('Number of occurrences of process in PreFreeSurfer '
     #           'on 100 subjects')
-    plt.savefig(output_folder+'frequency_plot.png')
+    plt.savefig(os.path.join(output_folder, 'frequency_plot.png'))
     # plt.show()
 
 
 def main(args=None):
-    parser = argparse.ArgumentParser(description='Classification of the nodes'
-                                                 ' in the pipeline graph.')
-    parser.add_argument("input_folder",
-                        help='input folder of sqlite databases ')
-    parser.add_argument("output_folder",
-                        help='output folder to save clusters, plots, '
-                             'and figures.')
-
-    args = parser.parse_args(args)
-    input_folder = args.input_folder
-    output_folder = args.output_folder
+    input_folder = 'data/pfs-20sbj-input'
+    output_folder = 'images'
 
     # get frequency of the processes with differences
     total_freq, sep_item_freq = get_item_frequency(input_folder)

### 2020-04-07 paper.tex
Fixed heatmap script
update pfs-graph

@@
   \centering
-    \includegraphics[width=\columnwidth]{images/pfs-labeled.png} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline.
-     Full-resolution image is available 
-     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
+    \includegraphics[width=\columnwidth]{./images/pfs-labeled.pdf} 
+    \caption{A complete provenance graph from the PreFreesurfer pipeline.}
     \label{fig:complete_pfs}
@@
 \centering
-  \includegraphics[width=\textwidth]{./data/pfs-20sbj-output/heatmap.pdf}
+  \includegraphics[width=\textwidth]{./images/heatmap.pdf}
   \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 

### 2020-04-07 paper.tex
Added conclusion, fixed intro and abstract

--- 2020-04-06 paper.tex
@@
 
-\title{File-based localization of numerical perturbations in data analysis pipelines}
+\title{\toolname: file-based localization of numerical perturbations in data analysis pipelines}
 \begin{document}
@@
 \begin{abstract} 
-
-% Data analysis pipelines are known to be impacted by operating system
-% updates and other infrastructural factors, presumably due to the creation,
-% propagation and amplification of numerical errors. However, the precise
-% causes of such instabilities and the path along which they propagate in the
-% pipelines are unclear, while this process could play a major role in the
-% current reproducibility crisis. We present a technique to identify the
-% computational processes in a pipeline that create numerical differences
-% when executed in different computational conditions. Our technique
-% leverages system-call interception to reconstruct processing and file
-% graphs without pipeline instrumentation. By applying our tool to the
-% structural pre-processing pipelines released by the Human Connectome
-% Project, we find that linear and non-linear registration are the cause of
-% most numerical instabilities in such pipelines, which confirms previous
-% findings.
-
+Data analysis pipelines are known to be impacted by computational conditions, presumably due to the creation,
+propagation and amplification of numerical errors. However, the precise
+causes of such instabilities and the path along which they propagate in the
+pipelines are unclear, while this process could play a major role in the
+current reproducibility crisis. We present \toolname, a tool to identify the
+processes in a pipeline that create numerical differences
+when executed in different computational conditions. \toolname
+leverages system-call interception to reconstruct and compare provenance
+graphs without pipeline instrumentation. By applying \toolname to the
+structural pre-processing pipelines of the Human Connectome
+Project, we found that linear and non-linear registration are the cause of
+most numerical instabilities in such pipelines, which confirms previous
+findings. 
 \end{abstract}
 
-\begin{keywords}
-Computational reproducibility; Numerical instability; Neuroimaging.
-\end{keywords}
+% \begin{keywords}
+% Computational reproducibility; Numerical instability; Neuroimaging.
+% \end{keywords}
 
@@
 instance, the fMRIprep pipeline~\cite{esteban2019fmriprep} assembles software blocks
-from FSL~\cite{jenkinson2012fsl}, AFNI~\cite{cox2012afni}, Freesurfer~\cite{fischl2012freesurfer} and ANTs~\cite{avants2009advanced} to provide a state-of-the art
+from FSL~\cite{jenkinson2012fsl}, AFNI~\cite{cox2012afni}, FreeSurfer~\cite{fischl2012freesurfer} and ANTs~\cite{avants2009advanced} to provide a state-of-the art
 functional MRI processing tool with minimal user input. Another example are the pipelines of the Human
-Connectome Project~\cite{glasser2013} that combine tools from FSL and Freesurfer to pre-process
+Connectome Project~\cite{glasser2013} that combine tools from FSL and FreeSurfer to pre-process
 structural, functional and diffusion data from their
@@
 
-Numerical evaluations are traditionnally performed using techniques such as
+Numerical evaluations are traditionally performed using techniques such as
 interval arithmetics~\cite{hickey2001interval} that require complete code re-writes and are
@@
 complex pipelines without instrumentation. Using system-call interception
-through the Reprozip tool~\cite{chirigati2016reprozip}, Spot traverses graphs of processes and
+through the \reprozip tool~\cite{rampin2016reprozip}, Spot traverses graphs of processes and
 intermediary files to pinpoint the pipeline components that are unstable
@@
 disambiguate files that have been written by multiple processes. The remainder of this paper 
-presents the design of our tool, and its 
+presents the design of \toolname, and its 
 application to pre-processing pipelines of the HCP project.
-
-\todo{explain the issue with propagating differences and with transient files}
 
@@
 % Furthermore, we can stabilize the identified processes in the next steps.
-\todo{explain creation vs proSpagation}
-\todo{define pipeline}
-\todo{explain the need for reduced instrumentation}
 
@@
 % //If threshold is 0, why do we even need the edit distance?//
+\begin{table}[b]
+  \centering
+  \begin{threeparttable}
+  \caption{Types of provenance graphs in PreFreeSurfer.}
+  \label{table:data-clusters}
+  \begin{tabular}{cc|cc}
+  \toprule
+  Type   &   \makecell{Number of \\ Subjects}   &  \makecell{Number of \\ T1w images}          & \makecell{Number of \\ T2w images}   \\ \midrule
+  1      &               9                      &  2 & 2 \\
+  2      &               8                      &  1 & 1 \\
+  3      &               1                      &  1 & 2 \\
+  4      &               2                      &  2 & 1\\
+  \bottomrule
+  \end{tabular}
+  \end{threeparttable}
+  \end{table}
 
@@
 subject types are coming from different  numbers of T1 and T2 images in the
-raw data: subjects of type 1 have 2 T1 images and 2 T2 images, while
-subjects of type 2 only have 1 T1 image and 1 T2 image. We verified that
+raw data. We verified that
 the provenance graphs were identical for all subjects of the same type, for
@@
 
-\begin{table}
-\centering
-\begin{threeparttable}
-\caption{Types of provenance graphs in PreFreeSurfer.}
-\label{table:data-clusters}
-
-\begin{tabular}{cc|cc}
-\toprule
-Type   &   \makecell{Number of \\ Subjects}   &  \makecell{Number of \\ T1w images}          & \makecell{Number of \\ T2w images}   \\ \midrule
-1      &               9                      &  2 & 2 \\
-2      &               8                      &  1 & 1 \\
-3      &               1                      &  1 & 2 \\
-4      &               2                      &  2 & 1\\
-\bottomrule
-\end{tabular}
-\end{threeparttable}
-\end{table}
-
-Figure~\ref{fig:pfs_freq} shows the frequency of non-reproducible pipeline processes
+Figure~\ref{fig:pfs_heatmap} shows the frequency of non-reproducible pipeline processes
 in PreFreeSurfer. 
@@
 and Atlas Registration). Differences were also observed in image mean 
-computations with FSL maths  (in Anatomical Average). 
-A complete provenance graph of the PreFreeSurfer is illustrated in Figure~\ref{fig:complete_pfs}
-\begin{figure}
+computations with FSL \texttt{maths}  (in Anatomical Average). 
+Figure~\ref{fig:complete_pfs} shows a complete PreFreeSurfer labeled DAG, localizing 
+the observed differences in the entire pipeline, for a given subject.
+\begin{figure*}
+\begin{minipage}{\linewidth}
 \centering
-    \includegraphics[width=\columnwidth]{images/pfs-labeled.png} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline.
-     Full-resolution image is available 
-     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
-    \label{fig:complete_pfs}
-\end{figure}
-% Between subject differences can be a result of different subject types. 
-% For instance, the subjects contain two T1 images need more calculations 
-% to produce an average image between them.
-
-% make figure in two columns use {figure*}
-\begin{figure*}
-\centering
-  \includegraphics[width=\textwidth]{./data/pfs-20sbj-output/heatmap.pdf}
+  \includegraphics[width=\textwidth]{./images/heatmap.pdf}
   \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 
@@
         }
-  \label{fig:pfs_freq}
+  \label{fig:pfs_heatmap}
+\end{minipage}
+
+\vspace*{1cm}
+
+\begin{minipage}{\linewidth}
+  \centering
+    \includegraphics[width=\linewidth]{images/pfs-labeled.png} 
+    \caption{A complete provenance graph from the PreFreesurfer pipeline.
+     Full-resolution image is available 
+     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
+    \label{fig:complete_pfs}
+\end{minipage}
 \end{figure*}
 
-To visually illustrate the observed differences,
+To illustrate the magnitude of the differences,
 Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
@@
     \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
-    CentOS7). An animated version of the comparison is available 
-    \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}
-    {here}.
-} 
+    CentOS7). An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}{here}.}
     \label{fig:fnirt_result}
@@
 \texttt{mris\_make\_surfaces} (cortical and white matter surfaces
-generation), which is also the only dynamically-linked executable in this pipeline
-\tristan{Ali, could you check that this is indeed the only dynamic
-executable in the pipeline?
-Ali: There is another dynamically-linked executable named mris\_curvature 
-that doesn't create difference when set a seed for that.
-Also, another executables are in directory freesurfer/mni/bin/ such as volume\_hist, 
-mincinfo, mincresample, \dots
-that are dynamically-linked but do not create differences.}. 
-\texttt{mris\_make\_surfaces} produced different results for
+generation), a dynamically-linked executable 
+that produced different results for
 10 out of 20 subjects.
 
-While by design \toolname only identifies processes that create differences
-between conditions, pipeline results may still differ due to the
-propagation and amplification of differences created at previous processing
-steps, PreFreeSurfer in our case. We observed the effect of this
-propagation in FreeSurfer results, as shown in
-Figure~\ref{fig:tissue_class} for whole-brain segmentations. Furthermore,
-the Dice coefficients associated with the 44 regions segmented by
-FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing, for some subjects, Dice values
-below 0.9 in structures such as the corpus callosum, pallidum, amygdala,
-accumbens area, and ventricles.
-
+However, FreeSurfer results still differ between conditions, due to the
+propagation and amplification of differences created in PreFreeSurfer. We
+observed the effect of this propagation in FreeSurfer results, as shown in
+Figure~\ref{fig:tissue_class} for whole-brain segmentations. The
+Dice coefficients associated with the 44 regions segmented by
+FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing that many 
+structures have Dice coefficients below 0.9 for some subjects.
 \begin{figure}
@@
 associated with these processes. It also corroborates our previous findings
-in~\cite{Glatard2015} where fMRI pre-processing using FSL was found to be
-non-reproducible across operating systems starting from the motion
+in~\cite{Glatard2015}, where fMRI pre-processing with FSL was found to vary across operating systems 
+starting from the motion
 correction step, a step that uses FSL's \flirt tool internally. It
 would be relevant to investigate if the observed instability of
-registration processes generalizes to other toolkits, or if it is specific
+registration processes generalizes to other toolkits, or if it remains specific
 to FSL. In view of the effect of small data perturbations in a variety of
@@
 FreeSurfer and CIVET~\cite{Lewis2017-ll} or connectome estimation using
-Dipy~\cite{kiar2019comparing}, it is possible that this observation
+Dipy~\cite{kiar2019comparing}, it is probable that this observation
 generalizes widely across toolboxes and requires a deeper investigation of
@@
 While only a handful or processes were found non-reproducible across the
-tested operating systems, including most notably linear and non-linear
-registration, the effect of such instabilities were found to propagate
-widely in the pipelines, resulting in Dice coefficients frequently below
-0.9 for the amygdala or ventricles. This illustrates the need to conduct such 
-reproducibility studies for entire pipelines rather than isolated processes.
+tested operating systems, the effect of such instabilities were found to
+propagate widely in the pipelines, and to substantially impact FreeSurfer's
+segmentations. This illustrates the need to conduct reproducibility studies
+on entire pipelines rather than isolated processes. It also highlights the need 
+for a deeper stability analysis of pipeline processes.
 
-Furthermore, as shown in Figure~\ref{fig:pfs_freq}, the reproducibility of
+Furthermore, as shown in Figure~\ref{fig:pfs_heatmap}, the reproducibility of
 a given tool may vary across subjects and across processing parameters. For
@@
 for some subjects only. Therefore, reproducibility studies need to be
-performed on several representative subjects rather than only a couple of
-them. While this is common practice in neurosciences, it is common that
-software tests be executed only on a single dataset, to reduce the
-associated computational load. It is likely, however, that the observed
-instabilities stem from numerical edge cases that may not manifest in all
-datasets.
+performed on several subjects. While this is common practice in neurosciences, 
+software tests are often executed only on a single dataset, to reduce the
+associated computational load. Our results show that pipeline tests should 
+encompass enough subjects to cover execution paths adequately.
 
-While our findings give insight on the reasons why image analysis processes
-are unstable across operating systems, they do not provide a complete
-uncertainty quantification for neuroimaging pipelines across operating
-systems. Although CentOS is a popular distribution in
-neurosciences~\cite{hanke2011neuroscience}, different observations could be
-made on other operating systems. The numerical noise introduced by
-operating system updates is a realistic one, as such updates are likely to
-occur throughout the time span of a neuroscience study, but it is also
-uncontrolled, as it originates in updates of low-level libraries by
-third-party developers. A possible method to study this problem more
-comprehensively would be to introduce controlled numerical perturbations in
-pipelines, which could be done either using data noise, or through
-computational noise introduced by Monte-Carlo
-Arithmetic~\cite{Parker1997-qq}. The work in~\cite{kiar2019comparing}
-discusses and compares these two techniques.
+Our results illustrate the type of variability that can be introduced in
+ neuroimaging results due to operating system updates. The
+ numerical noise introduced by operating system updates is realistic, as
+ such updates are likely to occur throughout the time span of a
+ neuroscience study, but it is also uncontrolled, as it originates in
+ updates of low-level libraries by third-party developers. A possible
+ method to study this problem more comprehensively would be to introduce
+ controlled numerical perturbations in pipelines, which could be done by
+ introducing noise either in the data, or in floating-point computations
+ through Monte-Carlo Arithmetic~\cite{Parker1997-qq}. The work
+ in~\cite{kiar2019comparing} discusses and compares these two techniques.
 
@@
 associated with dynamically-linked executables. 
-This makes complete sense as statically-linked executables wouldn't
-be impacted by library updates. Moreover, the hypothetical effects of
+This makes complete sense as statically-linked executables are not
+impacted by library updates. Moreover, the hypothetical effects of
 hardware or Linux kernel updates were not measured, as the different
-operating sytems were deployed in Docker containers on the same host, that
+operating systems were deployed in Docker containers on the same host, that
 is, using the same kernel and hardware.
@@
 
-Summary of key findings.
+We presented \toolname, a tool to detect the source of numerical
+differences in complex pipelines executed in different computational
+conditions. \toolname leverages system-call interception through the
+\reprozip tool, and therefore can be applied to the most complex pipelines
+without requiring their instrumentation.
 
-Mention non-instrumentation / reprozip
+By applying \toolname to the pre-processing pipelines of the Human
+Connectome Project, compared in different operating systems, we showed that
+between-OS differences are mostly originating in linear and non-linear
+image registration tools. Moreover, differences introduced during image
+registration propagate widely in the pipelines, leading to important
+variability in whole-brain segmentations.
 
-Spot works well but is a first step toward deeper investigation of 
-particular processes.
+Future work will investigate in more details the numerical stability of
+registration algorithms. Besides, we plan on using Monte-Carlo arithmetic
+to inject controlled amounts of noise in the pipelines, and monitory
+uncertainty propagation and amplification in the pipelines.
 
@@
 
-This research was enabled in part by support provided by 
-Calcul Quebec (http://www.calculquebec.ca) and 
-Compute Canada (http://www.computecanada.ca).
-Also, data were provided by the Human Connectome Project, WU-Minn 
+We warmly thank Compute Canada (\url{http://www.computecanada.ca}) and Calcul
+Qu\'ebec (\url{http://www.calculquebec.ca}) for providing the infrastructure used in our experiments.
+Data and pipelines were provided by the Human Connectome Project, WU-Minn 
 Consortium (Principal Investigators: David Van Essen and Kamil Ugurbil; 
@@
 Center for Systems Neuroscience at Washington University.
-
-% \begin{figure*}
-% \centering
-%   \includegraphics[width=.9\textwidth]{images/graph}
-%   \caption{A complete process graph from the PreFreeSurfer pipeline.
-% Full-resolution image available at \url{https://drive.google.com/open?id=174yyn8SuVOUcK5aRVw0bagjDanLD0FLt}.}
-%   \label{fig:complete-graph}
-% \end{figure*}
-
-% \begin{figure*}
-% \centering
-%   \includegraphics[width=.8\textwidth]{images/hclusters}
-%   \caption{Different data types clustered among 100 subjects.}
-%   \label{fig:subj-clusters}
-% \end{figure*}
+\tristan{check ethics}
 

### 2020-04-07 paper.tex
Fixed missing file in Fig 1

@@
     \centering
-    \includegraphics[width=.5\linewidth]{images/p-graph.pdf}
+    \includegraphics[width=.5\linewidth]{figures/p-graph.pdf}
     \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
@@
     \centering
-    \includegraphics[width=.4\linewidth]{images/p-graph-dag.pdf}
+    \includegraphics[width=.4\linewidth]{figures/p-graph-dag.pdf}
     \caption{Provenance DAG, with disambiguated transient files.}
@@
       \centering
-      \includegraphics[width=.4\linewidth]{images/p-graph-dag-labelled.pdf}
+      \includegraphics[width=.4\linewidth]{figures/p-graph-dag-labelled.pdf}
      \caption{Labeled DAG comparing 2 execution conditions, showing 1 non-reproducible process.}
@@
   \end{subfigure}
-    % \includegraphics[width=.8\columnwidth]{images/provenance-graphs}
+    % \includegraphics[width=.8\columnwidth]{figures/provenance-graphs}
     \caption{Provenance graphs
@@
 \begin{listing}
-  \inputminted{bash}{"bin/example.sh"}
+  \inputminted{bash}{"data/example/example.sh"}
   \caption{Example pipeline}
@@
 \centering
-  \includegraphics[width=\textwidth]{./images/heatmap.pdf}
+  \includegraphics[width=\textwidth]{./figures/heatmap.pdf}
   \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 
@@
   \centering
-    \includegraphics[width=\linewidth]{images/pfs-labeled.png} 
+    \includegraphics[width=\linewidth]{figures/pfs-labeled.png} 
     \caption{A complete provenance graph from the PreFreesurfer pipeline.
@@
   \centering
-    \includegraphics[width=\columnwidth]{images/t2w_alignment.png} 
+    \includegraphics[width=\columnwidth]{figures/t2w_alignment.png} 
     \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
-    CentOS7). An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/images/pfs_t2w_alignment.gif}{here}.}
+    CentOS7). An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs_t2w_alignment.gif}{here}.}
     \label{fig:fnirt_result}
@@
 \centering
-  \includegraphics[width=\columnwidth]{images/brain_segmentation_mni.png} 
+  \includegraphics[width=\columnwidth]{figures/brain_segmentation_mni.png} 
   \caption{Sum of binarized differences between whole-brain FreeSurfer
@@
   template. An animated comparison of segmentations obtained for a particular subject is available
-\href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/images/fs_brain_segmentation.gif}
+\href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/fs_brain_segmentation.gif}
 {here}.
@@
   \hspace*{-1cm}
-  \includegraphics[width=1.2\linewidth]{images/dice_regions.pdf}
+  \includegraphics[width=1.2\linewidth]{figures/dice_regions.pdf}
     \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).}

### 2020-04-07 paper.tex
Added comment

@@
 \item Brain Extraction (BExt), 
-\item \tristan{Ali, why isn't BBR cross-modal registration here and in the heatmap?}
+\item \tristan{Ali, why isn't BBR cross-modal registration here and in the heatmap?
+Ali: I found this main sub-pipelines as the separate scripts in PreFreeSurfer github repo, 
+btw, we can add it here. 
+I only reprsent sub-pipelines in the heatmap that at leats include one red process.}
 \item Bias Field Correction (BFC), 
@@
   \centering
-    \includegraphics[width=\linewidth]{figures/pfs-labeled.png} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline.
-     Full-resolution image is available 
-     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
+    \includegraphics[width=\linewidth]{figures/pfs-labeled.pdf} 
+    \caption{A complete provenance graph from the PreFreesurfer pipeline.}
     \label{fig:complete_pfs}

### 2020-04-08 paper.tex
basic language edits pt1

--- 2020-04-07 paper.tex
@@
 Data analysis pipelines are known to be impacted by computational conditions, presumably due to the creation,
-propagation and amplification of numerical errors. However, the precise
-causes of such instabilities and the path along which they propagate in the
-pipelines are unclear, while this process could play a major role in the
-current reproducibility crisis. We present \toolname, a tool to identify the
-processes in a pipeline that create numerical differences
+propagation, and amplification of numerical errors. While this process could
+play a major role in the current reproducibility crisis, the precise
+causes of such instabilities and the path along which they propagate in
+pipelines are unclear. We present \toolname, a tool to identify which
+processes in a pipeline create numerical differences
 when executed in different computational conditions. \toolname
@@
 Project, we found that linear and non-linear registration are the cause of
-most numerical instabilities in such pipelines, which confirms previous
+most numerical instabilities in these pipelines, which confirms previous
 findings.
@@
 Numerical stability is believed to play a central role in the differences
-observed in data analysis pipelines across computational infrastructures.
+observed in data analysis pipelines across computational infrastructures. \note{Greg}{cite or soften claim}
 Variability resulting from operating system upgrades, parallelization, or
 even minor data perturbations may originate in errors that accumulate in
-numerical computations, with important consequences on the reproducibility
-of data analysis results. However, it remains challenging to identify the
+numerical computations with important consequences on the reproducibility
+of analytical results. However, it remains challenging to identify the
 exact elements in a complex pipeline that are prone to numerical instability,
 leaving pipeline developers and users with the daunting task to manually
-compare results obtained in different conditions.
+compare results obtained across different conditions.
 
 In neuroimaging, our primary application field, data analyses often consist
-of hundreds of computational processes, sometimes coming from multiple
-toolboxes, that are aggregated to perform a specific function. For
+of hundreds of computational processes -- often coming from multiple
+toolboxes -- that are aggregated to perform a specific function. For
 instance, the fMRIprep pipeline~\cite{esteban2019fmriprep} assembles software blocks
@@
 structural, functional and diffusion data from their
-open dataset. In both cases, pipelines leverage toolboxes that are
-widely trusted in the community. At the same time, substantial results
-variations have been observed in these toolboxes resulting from minor data
-or infrastructure perturbations~\cite{Glatard2015, Gronenschild2012, Lewis2017-ll}, which suggests that further investigation of their
+uniquely high-fidelity open dataset. In both cases, pipelines leverage toolboxes that are
+widely trusted in the community, yet, at the same time substantial variations in results
+have been observed in these toolboxes resulting from minor data
+or infrastructure perturbations~\cite{Glatard2015, Gronenschild2012, Lewis2017-ll}, suggesting that further investigation of their
 numerical conditioning is required. For such complex pipelines, a
@@
 
-We present Spot, a tool to identify the source of numerical differences in
+We present \toolname, a tool to identify the source of numerical differences in
 complex pipelines without instrumentation. Using system-call interception
-through the \reprozip tool~\cite{rampin2016reprozip}, Spot traverses graphs of processes and
+through the \reprozip tool~\cite{rampin2016reprozip}, \toolname traverses graphs of processes and
 intermediary files to pinpoint the pipeline components that are unstable
@@
 
-\toolname identifies the processes in a pipeline that produce different
+\toolname identifies the components in a pipeline, as the resolution level of a system process, that produce different
 results in different execution conditions. First, a directed
-bipartite provenance graph is recorded for each subject, where nodes
+bipartite provenance graph is recorded for each pipeline execution, where nodes
 represent application processes and files, and edges represent read and
@@
  The resulting labeled graph identifies the
-non-reproducible processes in the pipeline.
+non-reproducible processes in the pipeline. \note{Greg}{clarify why many processes don't appear in the graph as shown}
 
@@
 
-We use the \reprozip tool~\cite{rampin2016reprozip}
+We use \reprozip~\cite{rampin2016reprozip}
 to capture: (1) the set of processes created by the
@@
 
-Provenance graphs are often subject-dependent, due to variations in input data among subjects.
+Provenance graphs are often data-dependent, due to variations in input data that
+may trigger differing branching or looping patterns across executions, for example.
 %, which justifies the processing
@@
 % purposes. 
-Some of these differences can be neglected, for instance when a data
+Some of these differences can be neglected: for instance, when a data
 decompression step is present at the beginning of the execution for some
-subjects only, but other differences cannot, for instance when complete
-different processing paths are used in different subjects. \toolname
-includes helpers to identify different instances of provenance graphs. In
-particular, it can cluster process trees, where nodes are processes and
+subjects only. Other differences cannot: for instance, when entirely
+different processing paths are used for different datasets. \toolname
+includes helpers to identify different instances of provenance graphs, 
+such as supporting the clustering of process trees, where nodes are processes and
 edges are \texttt{fork()} or \texttt{clone()} system calls, using the tree
@@
      Full-resolution image is available 
-     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}.}
+     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}. \tristan{Ali, link points to the wrong figure.}}
     \label{fig:complete_pfs}

### 2020-04-08 paper.tex
finished methods edits/suggestions

@@
 We capture temporary files by replacing every
-process P by a wrapper that first calls P and then saves the produced
+process $P$ by a wrapper that first calls $P$ and then saves the produced
 temporary files to a read-only directory. This process replacement is done by pre-pending
@@
 variable a directory that contains a wrapper script named after the executable
-called by P.
+called by $P$.
 
 Files written by multiple processes are disambiguated using a similar technique. For a
- file F written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
+ file $F$ written by the processes in \textbf{P} = \{$P_{1}$, \ldots,
  $P_{n}$\}, we first check that processes in \textbf{P} do not
- write concurrently to F, which would violate our assumptions. Then, we
+ write concurrently to $F$, which would violate our assumptions. Then, we
  replace every process $P_{i}$ by a \texttt{PATH}-based wrapper that first
- calls $P_{i}$ and then saves F to a read-only directory. In this way,
- successive versions of F are preserved for comparison. We finally
+ calls $P_{i}$ and then saves $F$ to a read-only directory. In this way,
+ successive versions of $F$ are preserved for comparison. We finally
  update the provenance graph accordingly, so that all files in the graph
@@
 
-After capturing transient files in the first condition, we re-run the pipeline
+After capturing transient files in the first condition (i.e. operating system, library version, etc.),
+we re-run the pipeline
 step by step in the second one to label processes. The output files
-produced by a process in both conditions are compared: if no differences are found, the process is marked as
+created by a process in both conditions are compared: if no differences are found, the process is marked as
 reproducible; otherwise, the process is marked as non-reproducible, and the
@@
 that differences do not propagate further in the pipeline. Processes are
-instrumented transparently, through a modification of the \texttt{PATH}
+instrumented transparently through a modification of the \texttt{PATH}
 variable similar to the one described previously. By default, differences
@@
 (\href{https://www.humanconnectome.org}{HCP}), a leading initiative in
-neurosciences.
+neuroimaging.
 
@@
 pipelines is available in~\cite{glasser2013}. 
-In summary, PreFreeSurfer consists of the following steps: 
+In summary, the PreFreeSurfer pipeline consists of the following steps: 
 \begin{itemize}
@@
 \end{itemize}
-And FreeSurfer consists of the following ones:
+And the FreeSurfer pipeline consists of the following:
 \begin{itemize}
@@
 83K files and created 7.7K processes, and FreeSurfer opened 62K files and 
-created 4K processes in total for a single subject.
+created 4K processes in total for a single execution, in this case corresponding to a single subject.
 
@@
 available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
-For each subject, available data consisted of 1 or 2 T1 images, and 1 or 2 T2 images, with 
-256 x 320 x 320 voxels of size 0.7 × 0.7 × 0.7 mm.
+For each subject, available data consisted of 1 or 2 T1w images and 1 or 2 T2w images, with 
+$256 \times 320 \times 320$ voxels of size $0.7 \times 0.7 \times 0.7$ mm.
 Acquisition protocols and parameters are detailed in~\cite{van2013wu}.

### 2020-04-08 paper.tex
finished first pass

@@
 \texttt{mris\_curvature}. Fixing the random seed used in FreeSurfer removed
-these differences.
+these differences. \note{Greg}{can you make some statement on the magnitude of these differences?}
 
@@
 provenance graphs (Table~\ref{table:data-clusters}). Differences between
-subject types are coming from different  numbers of T1 and T2 images in the
+subject types came from different numbers of T1 and T2 images in the
 raw data. We verified that
@@
 in PreFreeSurfer. 
-Differences were observed in linear registration 
+The processes identified as non-reproducible were observed in linear registration 
 with FSL \flirt (in ACPC-Alignment, Brain Extraction, Distortion Correction, and
@@
   instances were always reproducible, and 1 instance wasn't reproducible in
-  17 subjects.
+  17 subjects. \note{Greg}{colorbar doesn't match the numbers: it's from 0-1 but the numbers are 
+  subject counts. I don't really have a preference on which, but harmonizing them to be the same unit should happen.}
         }
@@
 for a particular subject. Differences appear to be very substantial, in
-particular in the frontal and parietal lobes. 
+particular in the frontal and parietal lobes.\note{Greg}{This is very hard to see from the figure because the
+luminosity of the two colormaps are so different - the contrast moving from orange to grayscale is difficult.
+Perhaps use two colors, green and purple for instance, with the same intensity to see if this is more easily seen?}
 \begin{figure}
@@
 FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing that many 
-structures have Dice coefficients below 0.9 for some subjects.
+structures have Dice coefficients below 0.9 for some subjects. \note{Greg}{awkward "many" followed by "some". reword?}
 \begin{figure}
@@
   \includegraphics[width=1.2\linewidth]{figures/dice_regions.pdf}
-    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).}
+    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).
+\note{Greg}{I don't think the legend is important here. Focus on having values legible and removing those which
+don't really illustrate the "point".}}
     \label{fig:scatter_plot}
@@
 \subsection{Key findings}
-
+\note{Greg}{feel free to ignore, but a peeve of mine is one heading directly following another. Is there some
+kind of high-level summary you could put above this, or just omit the key-findings label?}
 Linear and non-linear registration with FSL were found to
@@
 tested operating systems, the effect of such instabilities were found to
-propagate widely in the pipelines, and to substantially impact FreeSurfer's
-segmentations. This illustrates the need to conduct reproducibility studies
+propagate widely in the pipelines, and to substantially impact the segmentations
+created by FreeSurfer. This illustrates the need to conduct reproducibility studies
 on entire pipelines rather than isolated processes. It also highlights the need 
@@
 
-Furthermore, as shown in Figure~\ref{fig:pfs_heatmap}, the reproducibility of
+As is shown in Figure~\ref{fig:pfs_heatmap}, the reproducibility of
 a given tool may vary across subjects and across processing parameters. For
@@
 for some subjects only. Therefore, reproducibility studies need to be
-performed on several subjects. While this is common practice in neurosciences, 
-software tests are often executed only on a single dataset, to reduce the
+performed on several subjects. While this is common practice to some extent in neuroimaging, 
+software tests are often executed only on a single dataset to reduce the
 associated computational load. Our results show that pipeline tests should 
@@
 one more execution than the theoretical minimum of 2. It might be possible
-to further reduce this overhead, by executing at step (2) only the
+to further reduce this overhead by executing at step (2) only the
 processes depending on transient files.
@@
 Future work will investigate in more details the numerical stability of
-registration algorithms. Besides, we plan on using Monte-Carlo arithmetic
-to inject controlled amounts of noise in the pipelines, and monitory
-uncertainty propagation and amplification in the pipelines.
+registration algorithms. Additionally, we plan on using Monte-Carlo arithmetic
+to inject controlled amounts of noise in pipelines and monitor
+uncertainty propagation and amplification in their results.
 

### 2020-04-08 paper.tex
Typo

@@
 
-\toolname identifies the components in a pipeline, as the resolution level of a system process, that produce different
+\toolname identifies the components in a pipeline, at the resolution level of a system process, that produce different
 results in different execution conditions. First, a directed

### 2020-04-10 paper.tex
Fixed notes format
Commented on @gkiar's comments.

@@
 \newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
-\newcommand{\note}[2]{\color{blue}Note: #1\color{black}}
+\newcommand{\note}[2]{\color{blue}Note from \textsc{#1}: #2\color{black}}
 \newcommand{\reprozip}[0]{ReproZip\xspace}
-\newcommand{\tristan}[1]{\color{blue}\textbf{From Tristan:}#1\color{black}}
+\newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
 \newcommand{\english}[1]{\uwave{#1}}
@@
 
-Numerical stability is believed to play a central role in the differences
+Computational infrastructures 
+are known to impact data analysis pipelines, but it remains difficult to 
+\tristan{I will finish this.}
+Numerical stability may play a central role in the differences
 observed in data analysis pipelines across computational infrastructures. \note{Greg}{cite or soften claim}
@@
 
-\newpage
-
 \section{Tool description}
@@
  The resulting labeled graph identifies the
-non-reproducible processes in the pipeline. \note{Greg}{clarify why many processes don't appear in the graph as shown}
+non-reproducible processes in the pipeline. \note{Greg}{clarify why many processes don't appear in the graph as shown} \tristan{Ali, 
+do you pre-process the reprozip trace to remove some processes, or does the figure show all the processes in this trace?}
 
@@
 \texttt{mris\_curvature}. Fixing the random seed used in FreeSurfer removed
-these differences. \note{Greg}{can you make some statement on the magnitude of these differences?}
+these differences. \note{Greg}{can you make some statement on the magnitude of these differences?} \tristan{Ali, that's for you.}
 
@@
   17 subjects. \note{Greg}{colorbar doesn't match the numbers: it's from 0-1 but the numbers are 
-  subject counts. I don't really have a preference on which, but harmonizing them to be the same unit should happen.}
+  subject counts. I don't really have a preference on which, but harmonizing them to be the same unit should happen.} \tristan{
+    Ali, it looks like this figure is still unclear, and that other people may still misunderstand it. To better 
+    show that the number is a number of subjects, I'd suggest to replace number "x" with "(n=x)",
+    write it in smaller font size, and at the bottom-right of the cell instead of in the middle. 
+  }
         }
@@
 Perhaps use two colors, green and purple for instance, with the same intensity to see if this is more easily seen?}
+\tristan{Ali, I can't find the script to generate this figure in the repo. Could you fix the colors and contrast as per Greg's suggestion?}
 \begin{figure}
@@
 structures have Dice coefficients below 0.9 for some subjects. \note{Greg}{awkward "many" followed by "some". reword?}
+\tristan{Will do.}
 \begin{figure}
@@
 
-\begin{figure}
+\begin{figure*}
   \hspace*{-1cm}
@@
 \note{Greg}{feel free to ignore, but a peeve of mine is one heading directly following another. Is there some
-kind of high-level summary you could put above this, or just omit the key-findings label?}
+kind of high-level summary you could put above this, or just omit the key-findings label?} \tristan{will do}
 Linear and non-linear registration with FSL were found to

### 2020-04-10 paper.tex

@@
   \centering
-    \includegraphics[width=\linewidth]{figures/pfs-labeled.png} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline.
-     Full-resolution image is available 
-     \href{https://drive.google.com/file/d/1BKq5ZXIVr3bkij88SmeW7ulWOExSCm7n/view?usp=sharing}{here}. \tristan{Ali, link points to the wrong figure.}}
+    \includegraphics[width=\linewidth]{figures/pfs-labeled.pdf} 
+    \caption{A complete provenance graph from the PreFreesurfer pipeline.}
     \label{fig:complete_pfs}

### 2020-04-10 paper.tex
Fixed intro, Fig 6 and corresponding text

--- 2020-04-10 paper.tex
@@
 when executed in different computational conditions. \toolname
-leverages system-call interception to reconstruct and compare provenance
+leverages system-call interception through \reprozip to reconstruct and compare provenance
 graphs without pipeline instrumentation. By applying \toolname to the
@@
 
-Computational infrastructures 
-are known to impact data analysis pipelines, but it remains difficult to 
-\tristan{I will finish this.}
-Numerical stability may play a central role in the differences
-observed in data analysis pipelines across computational infrastructures. \note{Greg}{cite or soften claim}
-Variability resulting from operating system upgrades, parallelization, or
-even minor data perturbations may originate in errors that accumulate in
-numerical computations with important consequences on the reproducibility
-of analytical results. However, it remains challenging to identify the
-exact elements in a complex pipeline that are prone to numerical instability,
-leaving pipeline developers and users with the daunting task to manually
-compare results obtained across different conditions.
+Vibrations in computational infrastructures impact data
+analyses in various fields, but identifying the origin of these effects in complex pipelines
+remains challenging.  In some cases, most likely due to numerical
+instabilities, small perturbations resulting from changes in operating
+system versions~\cite{Glatard2015}, hardware~\cite{jezequel2015estimation}, or parallelization parameters~\cite{diethelm2011limits},
+may result in substantially different analysis outcomes. To better understand and 
+correct these effects, efficient tools are 
+needed to assist pipeline developers
+in the comparison of results obtained across different conditions.
 
@@
 have been observed in these toolboxes resulting from minor data
-or infrastructure perturbations~\cite{Glatard2015, Gronenschild2012, Lewis2017-ll}, suggesting that further investigation of their
+or infrastructure perturbations~\cite{Gronenschild2012, Glatard2015, Lewis2017-ll}, suggesting that further investigation of their
 numerical conditioning is required. For such complex pipelines, a
@@
 Dice coefficients associated with the 44 regions segmented by
-FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing that many 
-structures have Dice coefficients below 0.9 for some subjects. \note{Greg}{awkward "many" followed by "some". reword?}
-\tristan{Will do.}
+FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing that Dice coefficents below 0.9 are 
+observed in most regions, and particularly in the smallest ones.
 \begin{figure}
@@
   \hspace*{-1cm}
-  \includegraphics[width=1.2\linewidth]{figures/dice_regions.pdf}
-    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20).
-\note{Greg}{I don't think the legend is important here. Focus on having values legible and removing those which
-don't really illustrate the "point".}}
+  \includegraphics[width=1.1\linewidth]{figures/dice_regions.pdf}
+    \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20), ordered by increasing 
+    median values. Boxes brightness is proportional to the logarithm of the corresponding brain region size.}
     \label{fig:scatter_plot}

### 2020-04-10 bin/code/regions.py

@@ -2,7 +2,7 @@ from matplotlib import pyplot as plt
 import matplotlib._color_data as mcd
 import random
 
-
+plt.rcParams.update({'font.size': 16})
 plt.figure(figsize=(15,15))
 
 with open('./data/fs_seg_dice_accumulated_20sbj.csv') as f:
@@ -35,14 +35,22 @@ data = [ dices[region] for region in sorted_regions ]
 mean_sizes = [ [ mean(sizes[region]) for x in sizes[region]] for region in sorted_regions ]
 
 from math import log
-plt.boxplot(data, showfliers=False,
+bplot = plt.boxplot(data, showfliers=False, patch_artist=True,
             labels = range(len(sorted_regions)))
-plt.gca().xaxis.grid(True)
+
+#print(bplot)
+# fill with colors
+for i, patch in enumerate(bplot['boxes']):
+    region = sorted_regions[i]
+    rs = mean(sizes[region])
+    bs = mean(sizes['Background'])
+    a = log(rs)/log(bs)
+    patch.set_facecolor((0, 0, 0, a))
 
 for i, region in enumerate(sorted_regions):
     plt.scatter( [i+1 for x in dices[region]], dices[region],
-                 s=30,
-                 #c=colors[region], 
+                 s=35,
+                 color='black',
                  label=f'{i} - {region}',
                  marker='.')
 
@@ -50,5 +58,27 @@ plt.yticks([i/10 for i in range(11)])
 # plt.ylim(0,1.1)
 plt.ylabel('Dice coefficient')
 plt.xlabel('Region')
-plt.legend(fontsize=10)
-plt.savefig('images/dice_regions.pdf')
+plt.legend(handlelength=0, handletextpad=0, 
+           fontsize=14,
+           bbox_to_anchor=(0.98, 0),
+           loc="lower right",
+           bbox_transform=plt.gcf().transFigure,
+           ncol=4)
+plt.subplots_adjust(bottom=0.25)
+leg = plt.gca().get_legend()
+for i in range(len(sorted_regions)):
+    leg.legendHandles[i].set_visible(False)
+plt.gca().xaxis.tick_top()
+plt.gca().xaxis.set_ticks_position('top')
+plt.gca().xaxis.set_label_position('top')
+plt.gca().xaxis.grid(True, alpha=0.3)
+
+import matplotlib.ticker as plticker
[…]

### 2020-04-11 paper.tex
A few fixes

@@
 from the HCP data release S500 
-available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. 
+available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. \tristan{Ali, I can't find any S500 
+data release there, please confirm}
 For each subject, available data consisted of 1 or 2 T1w images and 1 or 2 T2w images, with 
@@
 
+Results from our experiments provide insights on the reproducibility of
+neuroimaging pipelines. They also allow us to evaluate the approach adopted
+in \toolname.
+
 \subsection{Key findings}
-\note{Greg}{feel free to ignore, but a peeve of mine is one heading directly following another. Is there some
-kind of high-level summary you could put above this, or just omit the key-findings label?} \tristan{will do}
 Linear and non-linear registration with FSL were found to
@@
 \reprozip tool, and therefore can be applied to the most complex pipelines
-without requiring their instrumentation.
+without requiring their instrumentation. \toolname is available at \url{https://github.com/big-data-lab-team/spot} under MIT license.
 

### 2020-04-11 bin/code/regions.py

@@ -77,8 +77,9 @@ import matplotlib.ticker as plticker
 #locs, labels = plt.xticks()
 import numpy as np
 plt.xticks(np.arange(1, 45, step=4), labels=np.arange(1, 45, step=4)-1)
-yticks = [0.1, 0.3, 0.5, 0.7, 0.9]
-plt.yticks(yticks, labels=yticks)
+yticks = [0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1]
+ylabels = [0, "", 0.2, "", 0.4, "", 0.6, "", 0.8, "", 1]
+plt.yticks(yticks, labels=ylabels)
 
 
 plt.savefig('figures/dice_regions.pdf')

### 2020-04-11 bin/code/heatmap_plot.py
update fig.2 fig.4 and comments

@@ -334,7 +334,10 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
                         ch = False
                         a, b = subsets_dic[x][y][lab].split(' outof ')
                         freq_list.append(float(a)/float(b))
-                        freq_label.append(str(a))
+                        if a == '0':
+                            freq_label.append("")
+                        else:
+                            freq_label.append("(n={})".format(a))
                             # str(float(a)/float(b))+' ['+str(b)+']')
                 if ch is True:
                     freq_list.append(0)
@@ -343,7 +346,7 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
             heatmap_label.append(freq_label)
             xlabel_ = [str(x).split(' ')[0] for x in xlabel_]
             sns.set(font_scale=1.1)
-            my_annot_kws = {"size": 30}  # , "weight":"bold"
+            my_annot_kws = {"size": 14}  # , "weight":"bold"
             # counted_ = list(Counter(frequency_matrix[0]).values())
             # frequency_matrix = [list(Counter(frequency_matrix[0]).keys())]
             # heatmap_label = [Counter(heatmap_label[0]).keys()]
@@ -359,12 +362,19 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
             #                 heatmap_label2[0].append(val_c)
             #             break
 
-            sns.heatmap(frequency_matrix, cbar=first == 0,
+            ax = sns.heatmap(frequency_matrix, cbar=first == 0,
                         cbar_ax=None if first else cbar_ax, vmin=0, vmax=1,
                         annot=np.array(heatmap_label), fmt='',
                         annot_kws=my_annot_kws, linewidths=2,
                         xticklabels='', yticklabels='', cmap=own_cmap1,
                         ax=axes[j, i])
+
+            for t in ax.texts:
+                trans = t.get_transform()
+                offs = matplotlib.transforms.ScaledTranslation(0, 0.43,
+                                matplotlib.transforms.IdentityTransform())
+                t.set_transform( offs + trans )
+
             # linewidths=.1
             first = first + 1
             if i == 0:
@@ -407,7 +417,7 @@ def plot_totall_items(total_freq, output_folder):
 
 def main(args=None):
     input_folder = 'data/pfs-20sbj-input'
-    output_folder = 'images'
+    output_folder = 'figures'
 
     # get frequency of the processes with differences
     total_freq, sep_item_freq = get_item_frequency(input_folder)

### 2020-04-12 bin/code/heatmap_plot.py
fix heatmap plot annotation position

@@ -268,6 +268,23 @@ def get_item_frequency(input_folder):
     return total_freq, process_freq
 
 
+def set_annot_position(heatmap_label):
+    my_annot_kws = {"size": 14}  # , "weight":"bold"
+    ver_ = 0.43
+    hor_ = 0.45
+    if len(heatmap_label[0]) == 2:
+        hor_ = 0.40
+    if len(heatmap_label[0]) == 6:
+        hor_ = 0.20
+    if len(heatmap_label[0]) == 10:
+        hor_ = 0
+    if len(heatmap_label[0]) > 10:
+        ver_ = 0.30
+        hor_ = 0.2
+        my_annot_kws = {"size": 14, "rotation":90}  # , "weight":"bold"
+    return my_annot_kws, hor_, ver_
+
+
 def plot_heatmap(subsets_dic, interesting_p, output_folder):
     ylabels = list(subsets_dic.keys())
     for ind, item in enumerate(ylabels):
@@ -334,11 +351,8 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
                         ch = False
                         a, b = subsets_dic[x][y][lab].split(' outof ')
                         freq_list.append(float(a)/float(b))
-                        if a == '0':
-                            freq_label.append("")
-                        else:
-                            freq_label.append("(n={})".format(a))
-                            # str(float(a)/float(b))+' ['+str(b)+']')
+                        freq_label.append("(n={})".format(b))
+                        # str(float(a)/float(b))+' ['+str(b)+']')
                 if ch is True:
                     freq_list.append(0)
                     freq_label.append('') # ['+str(b)+']')
@@ -346,32 +360,18 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
             heatmap_label.append(freq_label)
             xlabel_ = [str(x).split(' ')[0] for x in xlabel_]
             sns.set(font_scale=1.1)
-            my_annot_kws = {"size": 14}  # , "weight":"bold"
-            # counted_ = list(Counter(frequency_matrix[0]).values())
-            # frequency_matrix = [list(Counter(frequency_matrix[0]).keys())]
-            # heatmap_label = [Counter(heatmap_label[0]).keys()]
-            # heatmap_label2 = [[]]
-            # for ind, val in enumerate(frequency_matrix[0]):
-            #     for val_c in heatmap_label[0]:
-            #         hh = val_c.split(' ')[0]
-            #         if val_c.split(' ')[0] == str(val):
-            #             if counted_[ind] > 1:
[…]

### 2020-04-13 paper.tex

@@
 do you pre-process the reprozip trace to remove some processes, or does the figure show all the processes in this trace?}
+\note{Ali}{It only shows processes that belong to the pipeline (e.g. /usr/local/src/freesurfer/) 
+by making a query on the traced DB file. 
+We exclude processes like [cp, awk, sed, tee, date\dots].}
 
@@
 available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. \tristan{Ali, I can't find any S500 
-data release there, please confirm}
+data release there, please confirm.}
+\note{Ali}{The Dataset is: \emph{WU-Minn HCP Data - 1200 Subjects} in the project of 
+\emph{HCP 1200 Subject Release}.
+There is a group name \emph{S500 Release Subjects} that my data is downloaded from.}
 For each subject, available data consisted of 1 or 2 T1w images and 1 or 2 T2w images, with 
@@
 these differences. \note{Greg}{can you make some statement on the magnitude of these differences?} \tristan{Ali, that's for you.}
+\note{Ali}{
+\texttt{mris\_curvature} produces surfaces like lh.inflated.K and lh.inflated.H, 
+I only can see differences using \texttt{mri\_diff} command that it shows:
+
+  Volumes differ in pixel data
+
+  maxdiff   0.18604827 at 130715 0 0 0
+
+for lh.inflated.H and
+
+  Volumes differ in pixel data
+
+  maxdiff   0.85469627 at 130682 0 0 0
+
+for lh.inflated.K
+
+If you know other ways to check magnitude of differences please let me know.
+\texttt{mri\_segreg} produces different sum files (ASCII file in which summary statistics are saved), 
+but differences belong to meta-data like hostname, date, and SynthSeed. We should remove \texttt{mri\_segreg}
+because it doesn't create data differences.
+The outputs of both executables are in Consider:/data/asalari/ali-tests/paper\_images/
+}
 
@@
 \tristan{Ali, I can't find the script to generate this figure in the repo. Could you fix the colors and contrast as per Greg's suggestion?}
+\note{Ali}{Image is updated.
+I made this figure manually using ImageMagick tool and other scripts that you can find them in 
+Consider:/data/asalari/ali-tests/paper\_images/pfs\_fnirt\_imgs.}
+
 \begin{figure}

### 2020-04-13 paper.tex
Edits in figures

--- 2020-04-11 paper.tex
@@
 do you pre-process the reprozip trace to remove some processes, or does the figure show all the processes in this trace?}
+\note{Ali}{It only shows processes that belong to the pipeline (e.g. /usr/local/src/freesurfer/) 
+by making a query on the traced DB file. 
+We exclude processes like [cp, awk, sed, tee, date\dots].
+\tristan{Several points here: (1) freesurfer is obviously not used in this example, so what filtering was actually done?
+(2) where is this implemented in the code?
+(3)
+how do you know that you're not breaking the graph by removing these processes?}}
 
@@
 from the HCP data release S500 
-available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}. \tristan{Ali, I can't find any S500 
-data release there, please confirm}
+available in \href{https://db.humanconnectome.org}{the ConnectomDB repository} as a subset of the 1200 Subject Release.
 For each subject, available data consisted of 1 or 2 T1w images and 1 or 2 T2w images, with 
@@
 \texttt{mris\_curvature}. Fixing the random seed used in FreeSurfer removed
-these differences. \note{Greg}{can you make some statement on the magnitude of these differences?} \tristan{Ali, that's for you.}
+these differences. 
+% \note{Greg}{can you make some statement on the magnitude of these differences?} \tristan{Ali, that's for you.}
+% \note{Ali}{
+% \texttt{mris\_curvature} produces surfaces like lh.inflated.K and lh.inflated.H, 
+% I only can see differences using \texttt{mri\_diff} command that it shows:
+
+%   Volumes differ in pixel data
+
+%   maxdiff   0.18604827 at 130715 0 0 0
+
+% for lh.inflated.H and
+
+%   Volumes differ in pixel data
+
+%   maxdiff   0.85469627 at 130682 0 0 0
+
+% for lh.inflated.K
+
+% If you know other ways to check magnitude of differences please let me know.
+% \texttt{mri\_segreg} produces different sum files (ASCII file in which summary statistics are saved), 
+% but differences belong to meta-data like hostname, date, and SynthSeed. We should remove \texttt{mri\_segreg}
+% because it doesn't create data differences.
+% The outputs of both executables are in Consider:/data/asalari/ali-tests/paper\_images/
+% }
 
@@
   \includegraphics[width=\textwidth]{./figures/heatmap.pdf}
-  \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps (N=20). 
+  \caption{Heatmap of non-reproducible processes across PreFreeSurfer pipeline steps.
   Each cell represents the occurrence of a particular command line in a
@@
   Correction (BFC), or Atlas-Registration (AR). Cell labels indicate the
-  number of subjects where the corresponding process appears. For example,
+  fraction of subjects for which the corresponding process wasn't reproducible. For example,
   the \flirt tool was invoked 6 times in step DC for each of the 20
@@
   instances were always reproducible, and 1 instance wasn't reproducible in
-  17 subjects. \note{Greg}{colorbar doesn't match the numbers: it's from 0-1 but the numbers are 
-  subject counts. I don't really have a preference on which, but harmonizing them to be the same unit should happen.} \tristan{
-    Ali, it looks like this figure is still unclear, and that other people may still misunderstand it. To better 
-    show that the number is a number of subjects, I'd suggest to replace number "x" with "(n=x)",
-    write it in smaller font size, and at the bottom-right of the cell instead of in the middle. 
-  }
-        }
+  17 subjects. }
   \label{fig:pfs_heatmap}
@@
     \includegraphics[width=\linewidth]{figures/pfs-labeled.pdf} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline.}
+    \caption{A complete provenance graph from the PreFreesurfer pipeline. \tristan{Ali, the process tree needs to be in dashed and the file
+    dependencies in plain.}}
     \label{fig:complete_pfs}
@@
 for a particular subject. Differences appear to be very substantial, in
-particular in the frontal and parietal lobes.\note{Greg}{This is very hard to see from the figure because the
-luminosity of the two colormaps are so different - the contrast moving from orange to grayscale is difficult.
-Perhaps use two colors, green and purple for instance, with the same intensity to see if this is more easily seen?}
-\tristan{Ali, I can't find the script to generate this figure in the repo. Could you fix the colors and contrast as per Greg's suggestion?}
+particular in areas framed in red.
 \begin{figure}

### 2020-04-13 bin/code/heatmap_plot.py

@@ -269,19 +269,19 @@ def get_item_frequency(input_folder):
 
 
 def set_annot_position(heatmap_label):
-    my_annot_kws = {"size": 14}  # , "weight":"bold"
-    ver_ = 0.43
-    hor_ = 0.45
-    if len(heatmap_label[0]) == 2:
-        hor_ = 0.40
-    if len(heatmap_label[0]) == 6:
-        hor_ = 0.20
-    if len(heatmap_label[0]) == 10:
-        hor_ = 0
-    if len(heatmap_label[0]) > 10:
-        ver_ = 0.30
-        hor_ = 0.2
-        my_annot_kws = {"size": 14, "rotation":90}  # , "weight":"bold"
+    my_annot_kws = {"size": 24}  # , "weight":"bold"
+    ver_ = 0.4
+    hor_ = 0
+    # if len(heatmap_label[0]) == 2:
+    #     hor_ = 0.20
+    # if len(heatmap_label[0]) == 6:
+    #     hor_ = 0.10
+    # if len(heatmap_label[0]) == 10:
+    #     hor_ = 0
+    if len(heatmap_label[0]) >= 10:
+        ver_ = 0.25
+        # hor_ = 0.2
+        my_annot_kws = {"size": 24, "rotation":90}  # , "weight":"bold"
     return my_annot_kws, hor_, ver_
 
 
@@ -351,7 +351,7 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
                         ch = False
                         a, b = subsets_dic[x][y][lab].split(' outof ')
                         freq_list.append(float(a)/float(b))
-                        freq_label.append("(n={})".format(b))
+                        freq_label.append(f"({a}/{b})")
                         # str(float(a)/float(b))+' ['+str(b)+']')
                 if ch is True:
                     freq_list.append(0)
@@ -367,7 +367,8 @@ def plot_heatmap(subsets_dic, interesting_p, output_folder):
                         annot=np.array(heatmap_label), fmt='',
                         annot_kws=my_annot_kws, linewidths=2, 
                         xticklabels='', yticklabels='', cmap=own_cmap1,
-                        ax=axes[j, i])
+                        ax=axes[j, i], cbar_kws={'label': 'Fraction of non-reproducible processes'})
+            ax.figure.axes[-1].yaxis.label.set_size(36)
 
             for t in ax.texts:
                 trans = t.get_transform()

### 2020-04-14 paper.tex
reply comments

@@
 how do you know that you're not breaking the graph by removing these processes?}}
+\note{Ali}{In fact, I excluded processes in /usr/bin/ like [cp, sed, awk, tee, ...] in \emph{get\_processes\_list} function in \emph{auto\_spot.py} script.
+These processes will not affect the outputs as the results are shown.
+Spot will stop the procedure of labeling if any excluded processes create differences as it cannot labels included processed that read and write files with differences.
+It didn't happen in my experiments.
+}
 
@@
     Processes are represented with circles, files with rectangles, and read/write accesses 
-    with plain edges. 
+    with dashed edges. 
     For visualization purposes, the process tree is represented too, with
-    dashed edges. Every node in the graph is labeled using (1) a process id
+    plain edges. Every node in the graph is labeled using (1) a process id
     created by our reconstruction, (2) the name of the executable run by the
@@
     \caption{A complete provenance graph from the PreFreesurfer pipeline. \tristan{Ali, the process tree needs to be in dashed and the file
-    dependencies in plain.}}
+    dependencies in plain.} \note{Ali}{This figure is consistent with Fig.1. plain edges show process tree and dashed edges are read/write accesses. I just changed the caption in Fig.1.}}
     \label{fig:complete_pfs}

### 2020-04-14 paper.tex
fix fig.3

--- 2020-04-13 paper.tex
@@
     \includegraphics[width=\linewidth]{figures/pfs-labeled.pdf} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline. \tristan{Ali, the process tree needs to be in dashed and the file
-    dependencies in plain.}}
+    \caption{A complete provenance graph from the PreFreesurfer pipeline.}
     \label{fig:complete_pfs}

### 2020-04-21 paper.tex
update fig.1 and fig.3

--- 2020-04-13 paper.tex
@@
  The resulting labeled graph identifies the
-non-reproducible processes in the pipeline. \note{Greg}{clarify why many processes don't appear in the graph as shown} \tristan{Ali, 
-do you pre-process the reprozip trace to remove some processes, or does the figure show all the processes in this trace?}
-\note{Ali}{It only shows processes that belong to the pipeline (e.g. /usr/local/src/freesurfer/) 
-by making a query on the traced DB file. 
-We exclude processes like [cp, awk, sed, tee, date\dots].
-\tristan{Several points here: (1) freesurfer is obviously not used in this example, so what filtering was actually done?
-(2) where is this implemented in the code?
-(3)
-how do you know that you're not breaking the graph by removing these processes?}}
+non-reproducible processes in the pipeline. 
+\note{The graphs show all executed processes captured by the \reprozip tool
+except those that their binaries are placed in directory \texttt{/usr/bin/}. 
+We excluded these processes for a better representation of the pipeline execution.}
 
@@
     \centering
-    \includegraphics[width=.5\linewidth]{figures/p-graph.pdf}
+    \includegraphics[width=.9\linewidth]{figures/p-graph.pdf}
     \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
@@
     \centering
-    \includegraphics[width=.4\linewidth]{figures/p-graph-dag.pdf}
+    \includegraphics[width=.7\linewidth]{figures/p-graph-dag.pdf}
     \caption{Provenance DAG, with disambiguated transient files.}
@@
       \centering
-      \includegraphics[width=.4\linewidth]{figures/p-graph-dag-labelled.pdf}
+      \includegraphics[width=.7\linewidth]{figures/p-graph-dag-labelled.pdf}
      \caption{Labeled DAG comparing 2 execution conditions, showing 1 non-reproducible process.}

### 2020-04-21 bin/code/provenance_graph.py

@@ -33,7 +33,7 @@ def get_the_processes_name(executed_cursor, pid):
     executed_cursor.execute(process_name_query % pid)
     process_name = executed_cursor.fetchall()
     if process_name != []:
-        process_name = str(process_name[0][0]).split('/')[-1:][0]
+        process_name = str(process_name[0][0]) #.split('/')[-1:][0]
     else:
         process_name = ""
     return process_name
@@ -63,65 +63,67 @@ def create_provenance_graph(db_path, pid, graph_, pipe_list=[], multi_list={},
     child_list = get_the_child_processes(process_cursor, pid)
     # Get the process name
     process_name = get_the_processes_name(executed_cursor, pid)
-    # Get the list of opened files (w/r) from pid
-    process_ofiles = get_opened_files_list(openfile_cursor, pid)
 
-    # graph_.attr('node', shape='circle', style='filled', fillcolor='white')
-    if pid in total_proc_diff:
-        graph_.node(str(pid), ''.join([str(pid), '#', process_name]),
-                    shape='circle', style='filled', fillcolor='red')
-    elif total_proc_diff != []:
-        graph_.node(str(pid), ''.join([str(pid), '#', process_name]),
-                    shape='circle', style='filled', fillcolor='green')
-    else:
-        graph_.node(str(pid), ''.join([str(pid), '#', process_name]),
-                    shape='circle', style='filled', fillcolor='white')
+    if "/usr/bin" not in process_name and process_name != "":
+        # Get the list of opened files (w/r) from pid
+        process_ofiles = get_opened_files_list(openfile_cursor, pid)
 
-    for file in process_ofiles:
-        fname_ = str(os.path.basename(file[1]))
-        file_code = str(file[1])
-        if dag and fname_ in multi_list.keys():
-            if file[2] == 2:
-                file_code = str(file[1]) + str(pid)
-            else:
-                sorted_ = sorted(multi_list[fname_])
-                set_pid = pid
-                for pid_lst in sorted_:
-                    if pid > pid_lst:
-                        set_pid = pid_lst
-                        continue
-                file_code = str(file[1]) + str(set_pid)
+        p_name = process_name.split('/')[-1:][0]
+        # graph_.attr('node', shape='circle', style='filled', fillcolor='white')
+        if pid in total_proc_diff:
+            graph_.node(str(pid), ''.join([str(pid), '#', p_name]),
+                        shape='circle', style='filled', fillcolor='red')
+        elif total_proc_diff != []:
[…]

### 2020-04-22 paper.tex

@@
 non-reproducible processes in the pipeline. 
-\note{The graphs show all executed processes captured by the \reprozip tool
+\note{Ali}{The graphs show all executed processes captured by the \reprozip tool
 except those that their binaries are placed in directory \texttt{/usr/bin/}. 
@@
     \includegraphics[width=\linewidth]{figures/pfs-labeled.pdf} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline.}
+    \caption{A complete provenance graph from the PreFreesurfer pipeline.
+    \note{Ali}Full-resolution image is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs-labeled-orig.pdf}{here}.}
     \label{fig:complete_pfs}

### 2020-04-25 paper.tex
update Fig.3 and Fig.1

--- 2020-04-21 paper.tex
@@
 non-reproducible processes in the pipeline. 
-\note{The graphs show all executed processes captured by the \reprozip tool
-except those that their binaries are placed in directory \texttt{/usr/bin/}. 
+\note{Ali}{The graphs show all executed processes captured by the \reprozip tool
+except those that their binaries are placed in directory \texttt{/usr/bin/} or \texttt{/bin/}.
 We excluded these processes for a better representation of the pipeline execution.}
@@
     For visualization purposes, the process tree is represented too, with
-    dashed edges. Every node in the graph is labeled using (1) a process id
-    created by our reconstruction, (2) the name of the executable run by the
-    process or the file name. Process 1 is the initial call to the
-    \texttt{sh} interpreter; the other processes are the calls to FSL bet,
-    stats and maths made in Listing~\ref{listing:sample-script}. Process 73
-    is forked by process 2: it was captured by \reprozip while it did not
+    dashed edges. Every node in the graph is labeled using the name of the executable run by the
+    process or the file name. All the processes are the calls to FSL bet,
+    stats and maths made in Listing~\ref{listing:sample-script}. Process \texttt{bet2}
+    is forked by process \texttt{bet}: it was captured by \reprozip while it did not
     appear in Listing~\ref{listing:sample-script}.
@@
   \centering
-    \includegraphics[width=\linewidth]{figures/pfs-labeled.pdf} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline.}
+    \includegraphics[width=.6\linewidth]{figures/pfs-labeled.pdf} 
+    \caption{A complete provenance graph from the PreFreesurfer pipeline.
+    We excluded six more processes in the pipeline for better visualization, including:
+\texttt{imtest}, \texttt{imcp}, \texttt{remove\_ext}, \texttt{fslval}, \texttt{avscale}, and \texttt{fslhd}.}
     \label{fig:complete_pfs}

### 2020-04-25 bin/code/provenance_graph.py

@@ -50,6 +50,45 @@ def get_opened_files_list(openfile_cursor, pid):
     return openfile_cursor.fetchall()
 
 
+def add_edge(graph_, from_, to_, mode):
+    if mode == 'file':
+        graph_.attr('edge', style='solid',color='black', penwidth='1')
+    elif mode == 'ptree':
+        graph_.attr('edge', style='dashed', color='grey', penwidth='1')
+    graph_.edge(str(from_), str(to_))
+    return graph_
+
+
+def add_node(graph_, id, name, mode, color):
+    # with graph_.subgraph(name='cluster_{}'.format(cluster)) as ngraph_:
+    if mode == 'process':
+        if color == 'red':
+            if name == 'fslmaths':
+                name = 'fsl\nmaths'
+            if name == 'new_invwarp':
+                name = 'new\ninvwarp'
+            graph_.node(str(id), name.upper(), width='1.5', fontsize='16',fontname="Helvetica bold",
+                        fontcolor='black', shape='circle',
+                        style='filled', fillcolor='#EDB9B8', penwidth='1')
+        elif color == 'green':
+            graph_.node(str(id), name, width='1.5', penwidth='1', fontsize='16', shape='circle',
+                        style='filled', fillcolor='#7FCDB1', fontname="Helvetica bold")
+        elif color == 'subgreen':
+            graph_.node(str(id), name, width='1.5', fontsize='16',fontname="Helvetica bold",
+                        shape='circle', style='filled', fillcolor='#7FCDB1')
+        elif color == 'white':
+            graph_.node(str(id), name, width='1.5', shape='circle', style='filled', fontsize='16',
+                        fontname="Helvetica bold", fillcolor='white')
+    elif mode == 'file':
+        if name != '':
+            graph_.node(str(id), name, shape='box', style='filled', fillcolor=color, penwidth='1',
+                        fontsize='16', fontname="Helvetica bold")
+        else:
+            graph_.node(str(id), name, shape='box', style='filled', fillcolor=color, penwidth='1')
+    # graph_.node(str(id), name) 
+    return graph_
+
+
 def create_provenance_graph(db_path, pid, graph_, pipe_list=[], multi_list={},
                             tmp_list=[], total_proc_diff=[], dag=False):
     try:
@@ -64,21 +103,32 @@ def create_provenance_graph(db_path, pid, graph_, pipe_list=[], multi_list={},
     # Get the process name
[…]

### 2020-04-28 paper.tex

@@
     For visualization purposes, the process tree is represented too, with
-    dashed edges. Every node in the graph is labeled using the name of the executable run by the
+    dashed edges. \note{Ali}{Every node in the graph is labeled using the name of the executable run by the
     process or the file name. All the processes are the calls to FSL bet,
@@
     is forked by process \texttt{bet}: it was captured by \reprozip while it did not
-    appear in Listing~\ref{listing:sample-script}.
+    appear in Listing~\ref{listing:sample-script}}.
   }
@@
     \caption{A complete provenance graph from the PreFreesurfer pipeline.
-    We excluded six more processes in the pipeline for better visualization, including:
-\texttt{imtest}, \texttt{imcp}, \texttt{remove\_ext}, \texttt{fslval}, \texttt{avscale}, and \texttt{fslhd}.}
+   \note{Ali}{ We excluded six more processes in the pipeline for better visualization, including:
+\texttt{imtest}, \texttt{imcp}, \texttt{remove\_ext}, \texttt{fslval}, \texttt{avscale}, and \texttt{fslhd}.}}
     \label{fig:complete_pfs}

### 2020-04-28 bin/code/provenance_graph.py
Fixed figure paths

@@ -200,7 +200,7 @@ def main(args=None):
             lst_files.append(elem)
     # Create provenance graphs
     graph = Di('Graph',
-               filename='images/p-graph',
+               filename='figures/p-graph',
                format='pdf',
                strict=True)
     # graph.attr(compound=True)
@@ -209,7 +209,7 @@ def main(args=None):
     graph.render()
 
     graph = Di('Graph',
-               filename='images/p-graph-dag',
+               filename='figures/p-graph-dag',
                format='pdf',
                strict=True)
     create_provenance_graph(db_file, 1, graph, lst_files, multi_list,
@@ -217,7 +217,7 @@ def main(args=None):
     graph.render()
 
     graph = Di('Graph',
-               filename='images/p-graph-dag-labelled',
+               filename='figures/p-graph-dag-labelled',
                format='pdf',
                strict=True)
     create_provenance_graph(db_file, 1, graph, lst_files, multi_list,

### 2020-04-28 paper.tex
Put figures back on 2 pages
Small fixes

--- 2020-04-25 paper.tex
@@
 %%%% Commands %%%%
-\newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
-\newcommand{\note}[2]{\color{blue}Note from \textsc{#1}: #2\color{black}}
+% \newcommand{\todo}[1]{\color{red}\textbf{TODO:}#1\color{black}}
+% \newcommand{\note}[2]{\color{blue}Note from \textsc{#1}: #2\color{black}}
 \newcommand{\reprozip}[0]{ReproZip\xspace}
 \newcommand{\tristan}[1]{\color{orange}\textbf{From Tristan:}#1\color{black}}
-\newcommand{\english}[1]{\uwave{#1}}
 \newcommand{\toolname}[0]{Spot\xspace}
@@
 
-\title{\toolname: file-based localization of numerical perturbations in data analysis pipelines}
+\title{File-based localization of numerical perturbations in data analysis pipelines}
 \begin{document}
@@
 non-reproducible processes in the pipeline.
-\note{Ali}{The graphs show all executed processes captured by the \reprozip tool
-except those that their binaries are placed in directory \texttt{/usr/bin/} or \texttt{/bin/}.
-We excluded these processes for a better representation of the pipeline execution.}
 
@@
     \includegraphics[width=.9\linewidth]{figures/p-graph.pdf}
-    \caption{Raw provenance graph (\reprozip output), with transient files shown in gray.}
+    \caption{Raw provenance graph (\reprozip output), with transient files shown in gray boxes.}
     \label{fig:provenance-graph}
@@
     % \includegraphics[width=.8\columnwidth]{figures/provenance-graphs}
-    \caption{Provenance graphs
-    created from the example pipeline in
-    Listing~\ref{listing:sample-script}.
-    Processes are represented with circles, files with rectangles, and read/write accesses 
-    with plain edges. 
-    For visualization purposes, the process tree is represented too, with
-    dashed edges. Every node in the graph is labeled using the name of the executable run by the
-    process or the file name. All the processes are the calls to FSL bet,
-    stats and maths made in Listing~\ref{listing:sample-script}. Process \texttt{bet2}
-    is forked by process \texttt{bet}: it was captured by \reprozip while it did not
-    appear in Listing~\ref{listing:sample-script}.
+    \caption{Provenance graphs created from the example pipeline in
+    Listing~\ref{listing:sample-script}. Processes are represented with
+    circles, files with rectangles, and read/write accesses with plain edges. For convenience, the process tree 
+    is also shown, with gray dashed edges. Processes forked by
+    \texttt{bet} were captured by \reprozip while they did not appear in
+    Listing~\ref{listing:sample-script}. Processed associated with
+    executables located in \texttt{/usr/bin/} or \texttt{/bin/} are not shown.
   }
@@
 labeling of the example in Listing~\ref{listing:sample-script}. Process
-\texttt{73\#bet2} is labeled as non-reproducible (red) as it produces files
+\texttt{bet2} is labeled as non-reproducible (red) as it produces files
 with differences. To prevent the propagation of these differences, the
-files produced by \texttt{73\#bet2} in Condition 2 are replaced with the
-files produced by \texttt{73\#bet2} in Condition 1. Processes
-\texttt{75\#fslmaths} and \texttt{76\#fslstats} are then executed and
+files produced by \texttt{bet2} in Condition 2 are replaced with the
+files produced by \texttt{bet2} in Condition 1. Processes
+\texttt{fslmaths} and \texttt{fslstats} are then executed and
 labeled as reproducible (green) as they produce files without differences.
@@
 \item Brain Extraction (BExt), 
-\item \tristan{Ali, why isn't BBR cross-modal registration here and in the heatmap?
-Ali: I found this main sub-pipelines as the separate scripts in PreFreeSurfer github repo, 
-btw, we can add it here. 
-I only reprsent sub-pipelines in the heatmap that at leats include one red process.}
 \item Bias Field Correction (BFC), 
@@
 \end{itemize}
-The average processing time per subject is approximately 2~hours for
-PreFreeSurfer and 8~hours for FreeSurfer. The average output file size is
-approximately 2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
-%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
-The \reprozip traced files show that approximately PreFreeSurfer opened  
-83K files and created 7.7K processes, and FreeSurfer opened 62K files and 
-created 4K processes in total for a single execution, in this case corresponding to a single subject.
 
-We randomly selected 20 unprocessed subjects 
-from the HCP data release S500 
-available in \href{https://db.humanconnectome.org}{the ConnectomDB repository} as a subset of the 1200 Subject Release.
-For each subject, available data consisted of 1 or 2 T1w images and 1 or 2 T2w images, with 
-$256 \times 320 \times 320$ voxels of size $0.7 \times 0.7 \times 0.7$ mm.
-Acquisition protocols and parameters are detailed in~\cite{van2013wu}.
+We randomly selected 20 unprocessed subjects from the HCP data release S500
+available in \href{https://db.humanconnectome.org}{the ConnectomDB
+repository} as a subset of the 1200 Subject Release. For each subject,
+available data consisted of 1 or 2 T1-weighted images and 1 or 2
+T2-weighted images, with $256 \times 320 \times 320$ voxels of size $0.7
+\times 0.7 \times 0.7$ mm. Acquisition protocols and parameters are
+detailed in~\cite{van2013wu}.
 
@@
 system to detect within-OS variability coming from pseudo-random
-operations. We ignored execution-specific information such as file path or
-timestamps by using FreeSurfer tools \texttt{mri\_diff},
-\texttt{mris\_diff}, and \texttt{lta\_diff} that determine if MRI volumes
-or transformations differ. To compare segmentations $X$ and $Y$, we used the Dice coefficient defined as follows:
+operations. We compared pipeline results using  FreeSurfer tools \texttt{mri\_diff},
+\texttt{mris\_diff}, and \texttt{lta\_diff}, to ignore execution-specific information such as file path or
+timestamps. To compare segmentations $X$ and $Y$, we used the Dice coefficient defined as follows:
 \[DICE=\frac{2|X \cap Y|}{|X| + |Y|}\]
@@
 \section{Results}
+
+The average processing time per subject was approximately 2~hours for
+PreFreeSurfer and 8~hours for FreeSurfer. The average output file size was
+2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
+%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
+For each subject, PreFreeSurfer accessed
+83K files and created 7.7K processes, and FreeSurfer accessed 62K files and 
+created 4K processes.
 
@@
 \begin{figure*}
-\begin{minipage}{\linewidth}
 \centering
@@
   \label{fig:pfs_heatmap}
-\end{minipage}
+\end{figure*}
 
-\vspace*{1cm}
-
-\begin{minipage}{\linewidth}
+\begin{figure*} 
    \centering
-    \includegraphics[width=.6\linewidth]{figures/pfs-labeled.pdf} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline.
-    We excluded six more processes in the pipeline for better visualization, including:
+    \includegraphics[width=\linewidth]{figures/pfs-labeled.pdf} 
+    \caption{A complete provenance graph from the PreFreesurfer pipeline. For better visualization, processes associated with commands 
+    in \texttt{/bin} or \texttt{/usr/bin} were ommitted, as well as 
   \texttt{imtest}, \texttt{imcp}, \texttt{remove\_ext}, \texttt{fslval}, \texttt{avscale}, and \texttt{fslhd}.}
     \label{fig:complete_pfs}
-\end{minipage}
 \end{figure*}
@@
 Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
-for a particular subject. Differences appear to be very substantial, in
-particular in areas framed in red.
+for a particular subject. Differences appear to be important, in
+particular in the areas framed in red.
 \begin{figure}
@@
 
-Results from our experiments provide insights on the reproducibility of
-neuroimaging pipelines. They also allow us to evaluate the approach adopted
-in \toolname.
+Our results provide insights on the reproducibility of
+neuroimaging pipelines, and on the relevance of the approach implemented
+in \toolname for reproducibility studies.
 
@@
 
-However, file-based analyses also have limitations related to the
+File-based analyses also have limitations related to the
 granularity at which they operate. Indeed, differences can only be
@@
 another approach. Similarly, \toolname would not be able to detect
-differences in data not saved in files, and passed to subsequent processes
+differences in data not saved in files but instead passed to subsequent processes
 in memory. A common scenario in neuroimaging pipelines is that tools return
 results in their standard output, which is parsed by the calling process
-and passed to subsequent ones through variables. Differences that occur in data 
-that is processed in this way would not be detected by \toolname.
+and passed to subsequent ones through variables.
 
@@
 \reprozip tool, and therefore can be applied to the most complex pipelines
-without requiring their instrumentation. \toolname is available at \url{https://github.com/big-data-lab-team/spot} under MIT license.
+without requiring their instrumentation. It is available at
+\url{https://github.com/big-data-lab-team/spot} under MIT license.
 

### 2020-05-02 paper.tex
Minor fixes

@@
 have been observed in these toolboxes resulting from minor data
-or infrastructure perturbations~\cite{Gronenschild2012, Glatard2015, Lewis2017-ll}, suggesting that further investigation of their
+or infrastructure perturbations~\cite{Gronenschild2012, Glatard2015, Lewis2017-ll, kennedy2019everything}, suggesting that further investigation of their
 numerical conditioning is required. For such complex pipelines, a
@@
 Center for Systems Neuroscience at Washington University.
-\tristan{check ethics}
+%\tristan{check ethics}
 

### 2020-06-30 supplementary-files/file1.tex
add supplementary files

\documentclass[11pt,a4paper]{osa-supplemental}
\setboolean{shortarticle}{false}
\usepackage{threeparttable}

\title{File-based localization of numerical perturbations in data analysis pipelines}

\author{Ali Salari*, Gregory Kiar, Lindsay Lewis, Alan C. Evans, Tristan Glatard}

\begin{document}

\maketitle

\section{supplementary material}

We used unprocessed data from 20 subjects from the HCP data releases of the 1200 Subject project available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}.
Table \ref{tab:dataset} shows a summary of the subjects.

\begin{table}[htbp]
\centering
\begin{threeparttable}
\caption{\bf Summary of the subjects used in the experiments.}
\begin{tabular}{ccccc}
\hline
Subject & Release\tnote{1} & Acquisition\tnote{2} & Gender & Age \\
\hline
101309 & S500 & Q06 & M & 26-30 \\
102008 & S500 & Q06 & M & 22-25 \\ 
102311 & S500 & Q06 & F & 26-30 \\
103414 & Q2 & Q02 & F & 22-25 \\
103515 & Q1 & Q02 & F & 26-30 \\
103818 & Q1 & Q01 & F & 31-35 \\
105014 & S500 & Q05 & F & 26-30 \\
105115 & Q2 & Q02 & M & 31-35 \\
106319 & Q3 & Q03 & M & 26-30 \\
106521 & S500 & Q06 & F & 26-30 \\
107321 & S500 & Q04 & F & 22-25 \\
107422 & S500 & Q07 & M & 22-25 \\
108121 & S500 & Q04 & F & 26-30 \\
108323 & S500 & Q04 & F & 26-30 \\
113922 & S500 & Q04 & M & 31-35 \\
140420 & Q2 & Q02 & F & 26-30 \\
140824 & Q3 & Q04 & M & 31-35 \\
140925 & S500 & Q04 & F & 22-25 \\
142424 & Q3 & Q03 & M & 26-30 \\
142828 & Q1 & Q01 & M & 31-35 \\
\hline
\end{tabular}
  \label{tab:dataset}
  \begin{tablenotes}
  \footnotesize
    \item[1] Release refers to the HCP data release in which this subject's data was initially published to ConnectomeDB such as the initial Quarter releases (Q1, Q2, Q3)
    \item[2] Acquisition indicates the Quarter in which this subject's data was initially acquired
  \end{tablenotes}
\end{threeparttable}
\end{table}

% Quarter Released(Release)
% Acquisition Quarter(Acquisition)

\end{document}

### 2020-07-22 paper.tex

@@
 % \newcommand{\note}[2]{\color{blue}Note from \textsc{#1}: #2\color{black}}
+\newcommand{\revision}[1]{\color{purple}\textbf{REVISION:}#1\color{black}}
+\newcommand{\revised}[1]{\color{blue}\textbf{REVISED:}#1\color{black}}
 \newcommand{\reprozip}[0]{ReproZip\xspace}
@@
 \begin{abstract} 
-Data analysis pipelines are known to be impacted by computational conditions, presumably due to the creation,
-propagation, and amplification of numerical errors. While this process could
+Data analysis pipelines are known to be impacted by computational conditions, presumably due to the creation 
+and propagation of numerical errors. While this process could
 play a major role in the current reproducibility crisis, the precise
@@
 
+\revised{
+The labeled graph can be different depending on the condition that we obtained \reprozip trace as it can produce 
+different DAGs. Even using the same DAGs, it depends on the order of executions in which condition 
+we execute pipeline to pinpoint the propagation of differences. 
+Therefore, any permutation of these two factors can produce different results.
+% Pinpointing differences in one condition can be different than other condition.
+
+% add a paragraph to the methods, to explain that the order matters, due to:
+% 1. possible differences in the reprozip trace
+% 2. the way the comparison is done
+}
+
 \begin{figure*}
@@
     executables located in \texttt{/usr/bin/} or \texttt{/bin/} are not shown.
-  }
+    \revised{Node labels are enlarged.}}
     \label{fig:spot-example}
@@
   \inputminted{bash}{"data/example/example.sh"}
-  \caption{Example pipeline}
+  \caption{\revised{Example pipeline: computes the volume and number 
+  of voxels of the brain from a T1 image.}}
   \label{listing:sample-script}
@@
 labeled as reproducible (green) as they produce files without differences.
+
+\subsection{Implementation}
+\revised{ 
+The tool is implemented in Python programming language and supports the Python version 3.5 and higher.
+It utilizes the Python modules including NumPy~\cite{oliphant2006guide} and Pandas~\cite{mckinney2011pandas}, 
+which allow data processing, 
+SciPy~\cite{2020SciPy-NMeth} and Scikit-learn~\cite{pedregosa2011scikit} packages 
+for clustering the subjects, and Zss~\cite{zhang1989simple} package to calculate the tree distances.
+Also, the Python APIs of Docker and Boutiques~\cite{glatard2018boutiques} tools were used to create the container 
+images and simplify the pipeline executions.
+
+We used the latest version of Boutiques tool v0.5.25 for creating descriptors 
+to execute pipelines across the operating systems automatically.
+In a JSON schema, the descriptors define inputs, outputs, error codes, containers, and the command line for running pipelines.
+
+Besides, we used \reprozip tool version 1.0.16 to collect the pipeline system call processes and 
+files in an SQLite file format.
+}
 
@@
 We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2
-CentOS versions. Each subject was processed twice on the same operating
+CentOS versions. 
+\revised{Using the \reprozip trace obtained from CentOS7, 
+we captured transient files in CentOS6 and labeled pipeline processes in CentOS7.}
+Each subject was processed twice on the same operating
 system to detect within-OS variability coming from pseudo-random
@@
 
+\revised{
+The Dice coefficient~\cite{dice1945measures} is a commonly used metric in validating medical image segmentation.
+The Dice value ranges from 0 to 1, with 1 indicating a complete overlap between two segmentation results.
+Besides, the Jaccard coefficient~\cite{jaccard1912distribution} can be considered an equivalent metric 
+to measure the overlap between images.
+
+% Also, we used the Pearson correlation coefficient to establish the relationship between Dice value and segmented region size.
+% Like other correlation coefficients, this one varies between -1 and +1 with 0 implying no correlation. 
+% Correlations of -1 or +1 imply an exact linear relationship. 
+}
+
 % Finally, to cluster the subjects, the threshold value is set to zero in the clustering method. 
@@
 \section{Results}
-
+\revised{
+All experiments were run on the machine with a 3.4GHz Intel Core i7 processor 
+and 32GB memory under the CentOS7 operating system.
 The average processing time per subject was approximately 2~hours for
@@
 created 4K processes.
-
+The scripts and analyses used to create the figures in this section are available at 
+\url{https://github.com/big-data-lab-team/HCP-reproducibility-paper}.
+}
 % \subsection{Subject clustering}
@@
   instances were always reproducible, and 1 instance wasn't reproducible in
-  17 subjects. }
+  17 subjects. 
+  \revised{Grey cells show that any command line of the process did not occur in the pipeline step.}}
   \label{fig:pfs_heatmap}
@@
     \includegraphics[width=\linewidth]{figures/pfs-labeled.pdf} 
-    \caption{A complete provenance graph from the PreFreesurfer pipeline. For better visualization, processes associated with commands 
-    in \texttt{/bin} or \texttt{/usr/bin} were ommitted, as well as 
+    \caption{A complete provenance graph from the PreFreesurfer pipeline. 
+    \revised{Node labels show the same abbreviations as in Figure~\ref{fig:pfs_heatmap}.}
+    For better visualization, processes associated with commands 
+    in \texttt{/bin} or \texttt{/usr/bin} were omitted, as well as 
   \texttt{imtest}, \texttt{imcp}, \texttt{remove\_ext}, \texttt{fslval}, \texttt{avscale}, and \texttt{fslhd}.}
@@
 
-To illustrate the magnitude of the differences,
+\revised{
 Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
-for a particular subject. Differences appear to be important, in
-particular in the areas framed in red.
+for a particular subject using the checkerboard pattern.
+Differences appear to be visually important, in particular in the areas framed in red.
+The checkerboard representation is a common method to illustrate the magnitude of the differences in registration results.
+Although differences are more clearly represented in the animations linked in this figure.}
+
 \begin{figure}
@@
     \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
-    CentOS7). An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs_t2w_alignment.gif}{here}.}
+    CentOS7). 
+    \revised{
+    The colored squares indicate result images obtained from CentOS6 (in purple) and CentOS7 (in green), 
+    and the red boxes highlight regions with significant differences.}
+    % Result images obtained from CentOS6 and CentOS7 are colored in purple and green squares.
+    % The red boxes highlight regions with significant differences.}
+    An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs_t2w_alignment.gif}{here}.}
     \label{fig:fnirt_result}
@@
 However, FreeSurfer results still differ between conditions, due to the
-propagation and amplification of differences created in PreFreeSurfer. We
+propagation of differences created in PreFreeSurfer. We
 observed the effect of this propagation in FreeSurfer results, as shown in
@@
 observed in most regions, and particularly in the smallest ones.
+\revised{
+The correlation between the Dice values and region sizes computed as the Pearson's correlation coefficient 
+was 0.12 (\emph{P}=.43). 
+The positive correlation implies that as region size increases, Dice value increases. 
+% This is an average Dice values over 20 subjects for each 44 regions.
+}
+
 \begin{figure}
@@
    (N=20). Segmentations were resampled and overlaid to the MNI152 volume
-  template. An animated comparison of segmentations obtained for a particular subject is available
+  template. 
+  \revised{Each voxel shows the number of subjects that created different results. 
+  The color bar ranges from 1 to 7 as the minimum and the maximum number of subjects in each voxel.
+  Voxels are not colored if the identical results were produced in all 20 subjects.}
+  An animated comparison of segmentations obtained for a particular subject is available
 \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/fs_brain_segmentation.gif}
-{here}.
-} 
+{here}.} 
   \label{fig:tissue_class}
@@
     \caption{Dice coefficients between regions segmented by FreeSurfer in CentOS6 vs CentOS7 (N=20), ordered by increasing 
-    median values. Boxes brightness is proportional to the logarithm of the corresponding brain region size.}
+    median values. 
+    \revised{Each point indicates the Dice value in a particular region for a single subject.}
+    Boxes brightness is proportional to the logarithm of the corresponding brain region size.}
     \label{fig:scatter_plot}
@@
 \section{Discussion}
+
+\revision{
+Discuss irreproducibility in 2 parts:
+1. code coverage 
+2. statistics, see eq 1 in https://ieeexplore.ieee.org/abstract/document/7152570
+}
 

### 2020-07-22 supplementary-files/file1.tex

@@
 
-We used unprocessed data from 20 subjects from the HCP data releases of the 1200 Subject project available in \href{https://db.humanconnectome.org}{the ConnectomDB repository}.
+We used unprocessed data from 20 subjects from the HCP data releases of the 1200 Subject project available in \href{https://db.humanconnectome.org}{the ConnectomeDB repository}.
 Table \ref{tab:dataset} shows a summary of the subjects.
@@
   \footnotesize
-    \item[1] Release refers to the HCP data release in which this subject's data was initially published to ConnectomeDB such as the initial Quarter releases (Q1, Q2, Q3)
+    \item[1] Release refers to the HCP data release in which this subject's data was initially published to ConnectomeDB
     \item[2] Acquisition indicates the Quarter in which this subject's data was initially acquired

### 2020-07-29 paper.tex
Answers to some review comments

@@
 \newcommand{\revision}[1]{\color{purple}\textbf{REVISION:}#1\color{black}}
-\newcommand{\revised}[1]{\color{blue}\textbf{REVISED:}#1\color{black}}
+\newcommand{\revised}[1]{\color{blue}#1\color{black}\xspace}
 \newcommand{\reprozip}[0]{ReproZip\xspace}
@@
 
-Vibrations in computational infrastructures impact data
-analyses in various fields, but identifying the origin of these effects in complex pipelines
-remains challenging.  In some cases, most likely due to numerical
-instabilities, small perturbations resulting from changes in operating
-system versions~\cite{Glatard2015}, hardware~\cite{jezequel2015estimation}, or parallelization parameters~\cite{diethelm2011limits},
-may result in substantially different analysis outcomes. To better understand and 
+\revised{Numerical perturbations resulting from variations in computational
+environments} impact data analyses in various fields, but identifying the
+origin of these perturbations in complex pipelines remains challenging.  In
+some cases, small perturbations resulting from changes in operating system
+versions~\cite{Glatard2015}, hardware~\cite{jezequel2015estimation}, or
+parallelization parameters~\cite{diethelm2011limits}, result in
+substantially different analysis outcomes, \revised{due to the propagation
+and amplification of floating-point errors. While the existence of such
+numerical errors is well known~\cite{stoer2013introduction}, their impact
+on scientific computations has multiplied with the rise of the Big Data
+era, due to the sustained growth of data sets, the increasing complexity of
+analysis pipelines, and the diversification of computing infrastructures.}
+To better understand and 
 correct these effects, efficient tools are 
@@
 
+\revised{
+Computational environments are only one of many factors contributing to the
+on-going reproducibility crisis. In fact, sample size selection,
+publication bias, or methodological flexibility in the analysis are likely
+to have a stronger effect than numerical perturbations, although
+to our knowledge no evidence of this is available. We refer to the
+studies in~\cite{botvinik2020variability,bowring2019exploring,bhagwat2020understanding,kennedy2019everything} for deeper
+analyses of the associated effects on neuroimaging analyses. It should also be
+noted that the effects of computational environments and these other
+factors manifest at different levels: referring to the terminology used
+in~\cite{peng2011reproducible}, computational environments are associated with
+\emph{reproducibility}, the minimal standard by which identical results
+should be obtainable from identical data and parameters, while the other
+aforementioned factors belong to \emph{replicability}, the ultimate
+standard by which independent experimenters should be able to draw similar
+conclusions from similar experiments. In practice, variability resulting
+from computational environments manifests during software testing (test
+results depend on execution platform), deployment on HPC systems (results
+obtained on local vs HPC systems differ), or software version updates
+(results obtained before vs after the update differ), while factors related
+to replicability impact the community more broadly. Ultimately, both
+reproducibility and replicability should be understood and improved.
+}
+
 \section{Conclusion}

### 2020-07-29 paper.tex
Update revisions

--- 2020-07-22 paper.tex
@@
 \newcommand{\revision}[1]{\color{purple}\textbf{REVISION:}#1\color{black}}
-\newcommand{\revised}[1]{\color{blue}\textbf{REVISED:}#1\color{black}}
+\newcommand{\revised}[1]{\color{blue}#1\color{black}}
 \newcommand{\reprozip}[0]{ReproZip\xspace}
@@
 
-\revised{
-The labeled graph can be different depending on the condition that we obtained \reprozip trace as it can produce 
-different DAGs. Even using the same DAGs, it depends on the order of executions in which condition 
-we execute pipeline to pinpoint the propagation of differences. 
-Therefore, any permutation of these two factors can produce different results.
-% Pinpointing differences in one condition can be different than other condition.
-
-% add a paragraph to the methods, to explain that the order matters, due to:
-% 1. possible differences in the reprozip trace
-% 2. the way the comparison is done
-}
-
 \begin{figure*}
@@
     Listing~\ref{listing:sample-script}. Processed associated with
-    executables located in \texttt{/usr/bin/} or \texttt{/bin/} are not shown.
-    \revised{Node labels are enlarged.}}
+    executables located in \texttt{/usr/bin/} or \texttt{/bin/} are not shown.}
     \label{fig:spot-example}
@@
 to file reads and writes (Figure~\ref{fig:provenance-graph}).
+\revised{We assume that provenance graphs are identical for the \reprozip traces obtained from 
+the same subjects in different operating systems.}
 
@@
 labeled as reproducible (green) as they produce files without differences.
+
+\revised{
+The labeled graph can be different depending on the order of executions in which condition 
+we execute pipeline to pinpoint the propagation of differences. 
+Therefore, we label processes in two order of executions, and then aggregate findings so that label processes 
+as non-reproducible (red) if it creates different results in both order of executions.
+}
 
@@
 
+% \revision{Using the \reprozip trace obtained from CentOS7, 
+% we captured transient files in CentOS6 and labeled pipeline processes in CentOS7.}
 We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2
 CentOS versions. 
-\revised{Using the \reprozip trace obtained from CentOS7, 
-we captured transient files in CentOS6 and labeled pipeline processes in CentOS7.}
 Each subject was processed twice on the same operating
@@
 to measure the overlap between images.
-
-% Also, we used the Pearson correlation coefficient to establish the relationship between Dice value and segmented region size.
-% Like other correlation coefficients, this one varies between -1 and +1 with 0 implying no correlation. 
-% Correlations of -1 or +1 imply an exact linear relationship. 
 }
@@
 \revised{
-All experiments were run on the machine with a 3.4GHz Intel Core i7 processor 
-and 32GB memory under the CentOS7 operating system.
-The average processing time per subject was approximately 2~hours for
-PreFreeSurfer and 8~hours for FreeSurfer. The average output file size was
-2.7~GB for PreFreeSurfer and 4.1~GB for FreeSurfer.
-%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
-For each subject, PreFreeSurfer accessed
-83K files and created 7.7K processes, and FreeSurfer accessed 62K files and 
-created 4K processes.
+All experiments were run on a machine with a 3.4GHz Intel Core i7 processor and 32GB memory 
+under the CentOS7 operating system. 
+The details of the computation of the PreFreeSurfer and FreeSurfer pipelines are illustrated 
+in~Table~\ref{table:pipeline-stats}.
 The scripts and analyses used to create the figures in this section are available at 
 \url{https://github.com/big-data-lab-team/HCP-reproducibility-paper}.
+
+\begin{table*}[ht]
+  \centering
+  \begin{threeparttable}
+  \caption{The computation details of the pipelines per subject.}
+  \label{table:pipeline-stats}
+  \begin{tabular}{ccc|cc}
+  \toprule
+  \multicolumn{1}{c}{} & \multicolumn{2}{c}{PreFreeSurfer } &   \multicolumn{2}{c}{FreeSurfer } \\
+
+   &        \makecell{Mean}   &  \makecell{Standard error}  & \makecell{Mean}   &  \makecell{Standard error}   \\ \midrule
+  Processing time (mins)      &     106.67     &     2.68     &     nd     &     nd     \\
+  Output file size (GB)       &     2.8        &     0.10     &     nd     &     nd     \\
+  Number of file accesses     &     94089      &     2644.99  &     nd     &     nd     \\
+  Number of processes         &     8731       &     198.18   &     nd     &     nd     \\
+  \bottomrule
+  \end{tabular}
+  \end{threeparttable}
+  \end{table*}
 }
+% All experiments were run on a machine with a 3.4GHz Intel Core i7 processor 
+% and 32GB memory under the CentOS7 operating system.
+% The average processing time per subject was approximately 2~hours (with standard error X) 
+% for PreFreeSurfer and 8~hours (with standard error X) for FreeSurfer. 
+% The average output file size was 2.7~GB (with standard error Y) for PreFreeSurfer 
+% and 4.1~GB (with standard error Y) for FreeSurfer.
+%\tristan{Ali, are these numbers per subject or for 20 subjects? it's an average of 20 subjects}
+% For each subject, PreFreeSurfer accessed 83K files and created 7.7K processes, and FreeSurfer accessed 62K files and 
+% created 4K processes.
+
 % \subsection{Subject clustering}
@@
   17 subjects. 
-  \revised{Grey cells show that any command line of the process did not occur in the pipeline step.}}
+  \revised{Grey cells show that the process did not occur in the pipeline step.}}
   \label{fig:pfs_heatmap}
@@
 Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
-for a particular subject using the checkerboard pattern.
-Differences appear to be visually important, in particular in the areas framed in red.
-The checkerboard representation is a common method to illustrate the magnitude of the differences in registration results.
-Although differences are more clearly represented in the animations linked in this figure.}
+for a particular subject using the checkerboard pattern, 
+a common method to illustrate the magnitude of the differences in registration results.
+Differences appear to be visually important, in particular in the areas framed in red.}
 
@@
 observed in most regions, and particularly in the smallest ones.
-\revised{
-The correlation between the Dice values and region sizes computed as the Pearson's correlation coefficient 
-was 0.12 (\emph{P}=.43). 
-The positive correlation implies that as region size increases, Dice value increases. 
-% This is an average Dice values over 20 subjects for each 44 regions.
-}
+\revised{However, there is no significant correlation between Dice value and region size.}
 
@@
   template. 
-  \revised{Each voxel shows the number of subjects that created different results. 
-  The color bar ranges from 1 to 7 as the minimum and the maximum number of subjects in each voxel.
+  \revised{
+  The binary images are created based on the intensity differences between result images in CentOS6 and CentOS7. 
+  The voxel is 1 if the intensity value of the corresponding voxel in the images is different, otherwise it is 0.
+  Each voxel shows the number of subjects that created different results. 
+  % The color bar ranges from 1 to 7 as the minimum and the maximum number of subjects in each voxel.
   Voxels are not colored if the identical results were produced in all 20 subjects.}
@@
 
-To evaluate the reproducibility of a pipeline, \toolname needs to execute
+\revision{To evaluate the reproducibility of a pipeline, \toolname needs to execute
 it 3 times in order to (1) record a first \reprozip trace, (2) save
@@
 to further reduce this overhead by executing at step (2) only the
-processes depending on transient files.
+processes depending on transient files.}
 

### 2020-07-29 paper.tex
solve the conflict

@@
 \newcommand{\revision}[1]{\color{purple}\textbf{REVISION:}#1\color{black}}
-\newcommand{\revised}[1]{\color{blue}#1\color{black}}
+\newcommand{\revised}[1]{\color{blue}#1\color{black}\xspace}
 \newcommand{\reprozip}[0]{ReproZip\xspace}

### 2020-07-29 paper.tex
minor edit

@@
   \inputminted{bash}{"data/example/example.sh"}
-  \caption{\revised{Example pipeline: computes the volume and number 
-  of voxels of the brain from a T1 image.}}
+  \caption{\revised{Example pipeline that computes the volume of the brain from a T1 image.}}
   \label{listing:sample-script}
@@
   17 subjects. 
-  \revised{Grey cells show that the process did not occur in the pipeline step.}}
+  \revised{Grey cells indicate that the process did not occur in the corresponding pipeline step.}}
   \label{fig:pfs_heatmap}
@@
     \caption{A complete provenance graph from the PreFreesurfer pipeline. 
-    \revised{Node labels show the same abbreviations as in Figure~\ref{fig:pfs_heatmap}.}
+    \revised{Node labels use the same abbreviations as in Figure~\ref{fig:pfs_heatmap}.}
     For better visualization, processes associated with commands 
@@
     median values. 
-    \revised{Each point indicates the Dice value in a particular region for a single subject.}
+    \revised{Each point represents the Dice coefficient between segmentations of a particular region obtained in CentOS 6 vs CentOS 7 for a given subject.}
     Boxes brightness is proportional to the logarithm of the corresponding brain region size.}

### 2020-08-12 paper.tex
Revise comments

@@
 \revised{
-The labeled graph can be different depending on the order of executions in which condition 
-we execute pipeline to pinpoint the propagation of differences. 
-Therefore, we label processes in two order of executions, and then aggregate findings so that label processes 
-as non-reproducible (red) if it creates different results in both order of executions.
-}
+The labeled graph can differ depending on the order of executions in which condition we capture transient files 
+or execute the pipeline to pinpoint the propagation of differences. Therefore, we label processes in two orders 
+of execution, and then aggregate findings so that label processes as non-reproducible (red) if it creates 
+different results in either order of executions.}
 
@@
 \revised{
-The tool is implemented in Python programming language and supports the Python version 3.5 and higher.
-It utilizes the Python modules including NumPy~\cite{oliphant2006guide} and Pandas~\cite{mckinney2011pandas}, 
-which allow data processing, 
-SciPy~\cite{2020SciPy-NMeth} and Scikit-learn~\cite{pedregosa2011scikit} packages 
-for clustering the subjects, and Zss~\cite{zhang1989simple} package to calculate the tree distances.
-Also, the Python APIs of Docker and Boutiques~\cite{glatard2018boutiques} tools were used to create the container 
-images and simplify the pipeline executions.
+\toolname is implemented in Python (>=3.5). In this work we used \toolname version 0.2 and the following version of
+the Python package dependencies: NumPy v1.19.0~\cite{oliphant2006guide} and Pandas v1.0.5~\cite{mckinney2011pandas},
+for data manipulations, SciPy v1.5.1~\cite{2020SciPy-NMeth} and Scikit-learn v.0.23.1~\cite{pedregosa2011scikit}
+for the clustering of provenance graphs, Zss v1.2.0~\cite{zhang1989simple} for tree distances,
+\reprozip v1.0.11 for the capture of provenance traces, Docker v17.05~\cite{merkel2014docker} for
+the edition of container images, and Boutiques v0.5.25~\cite{glatard2018boutiques} for uniform pipeline executions.
 
-We used the latest version of Boutiques tool v0.5.25 for creating descriptors 
-to execute pipelines across the operating systems automatically.
-In a JSON schema, the descriptors define inputs, outputs, error codes, containers, and the command line for running pipelines.
-
-Besides, we used \reprozip tool version 1.0.16 to collect the pipeline system call processes and 
-files in an SQLite file format.
-}
+Software users will mostly have to interact with the Boutiques and \reprozip packages.
+Boutiques is a flexible description framework for containerized pipelines, required by
+the pipelines analyzed in \toolname. It provides a JSON schema to describe inputs, outputs and their dependencies.
+Examples, tutorials and usage documentation are available at \url{http://boutiques.github.io}. \reprozip intercepts
+system calls to identify the files and processes involved in a pipeline execution.
+Before using \toolname, users have to collect \reprozip traces of their pipeline executions.
+Examples in the \toolname documentation include \reprozip provenance capture.
+More documentation on \reprozip is available at \url{https://www.reprozip.org}.}
 
@@
 available in \href{https://db.humanconnectome.org}{the ConnectomDB
-repository} as a subset of the 1200 Subject Release. For each subject,
-available data consisted of 1 or 2 T1-weighted images and 1 or 2
+repository} as a subset of the 1200 Subject Release \revised{(see Supplementary Table S1)}.
+For each subject, available data consisted of 1 or 2 T1-weighted images and 1 or 2
 T2-weighted images, with $256 \times 320 \times 320$ voxels of size $0.7
@@
 % we captured transient files in CentOS6 and labeled pipeline processes in CentOS7.}
-We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2
-CentOS versions. 
+We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2 CentOS versions.
+\revised{First, PreFreeSurfer was utilized to analyze the unprocessed data, and then its results 
+from CentOS6 were used as the input of FreeSurfer in both conditions. 
+We also used the ReproZip trace file captured in CentOS6 for labeling the processes in both pipelines.}
 Each subject was processed twice on the same operating
@@
 \revised{
-The Dice coefficient~\cite{dice1945measures} is a commonly used metric in validating medical image segmentation.
-The Dice value ranges from 0 to 1, with 1 indicating a complete overlap between two segmentation results.
-Besides, the Jaccard coefficient~\cite{jaccard1912distribution} can be considered an equivalent metric 
-to measure the overlap between images.
-}
+The Dice coefficient~\cite{dice1945measures} is a commonly used metric to validate medical image segmentation.
+Dice values range from 0 to 1, with 1 indicating a perfect overlap between two segmentation results and 0 indicates no overlap.
+Alternatively, the Jaccard coefficient~\cite{jaccard1912distribution} could be used;
+there is a direct correspondence between both metrics.}
 
@@
 \revised{
-All experiments were run on a machine with a 3.4GHz Intel Core i7 processor and 32GB memory 
-under the CentOS7 operating system. 
-The details of the computation of the PreFreeSurfer and FreeSurfer pipelines are illustrated 
-in~Table~\ref{table:pipeline-stats}.
+All experiments were run on a machine with a 3.4GHz Intel Core i7 processor using 8 cores and 32GB of RAM running
+the CentOS7 operating system and Linux kernel version 3.10.
+The processing time, output file size, number of file accesses and number of processes observed in PreFreeSurfer
+and FreeSurfer are shown in~Table~\ref{table:pipeline-stats}.
 The scripts and analyses used to create the figures in this section are available at
@@
   \begin{threeparttable}
-  \caption{The computation details of the pipelines per subject.}
+  \caption{Execution statistics of the pipelines per subject.}
   \label{table:pipeline-stats}
@@
   Output file size (GB)       &     2.8        &     0.10     &     nd     &     nd     \\
-  Number of file accesses     &     94089      &     2644.99  &     nd     &     nd     \\
-  Number of processes         &     8731       &     198.18   &     nd     &     nd     \\
+  Number of file accesses     &     94,089      &     2,645  &     nd     &     nd     \\
+  Number of processes         &     8731       &     198   &     nd     &     nd     \\
   \bottomrule
@@
     \includegraphics[width=\columnwidth]{figures/t2w_alignment.png}
-    \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs 
-    CentOS7). 
+    \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs CentOS7).
     \revised{
-    The colored squares indicate result images obtained from CentOS6 (in purple) and CentOS7 (in green), 
-    and the red boxes highlight regions with significant differences.}
-    % Result images obtained from CentOS6 and CentOS7 are colored in purple and green squares.
-    % The red boxes highlight regions with significant differences.}
-    An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs_t2w_alignment.gif}{here}.}
+    The colored squares indicate results obtained with CentOS6 (in purple) and CentOS7 (in green).
+    The red boxes highlight regions with significant differences between the two OSes. }
+    An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs_t2w_alignment.gif}{here} for better visualization.}
     \label{fig:fnirt_result}
@@
 observed in most regions, and particularly in the smallest ones.
-\revised{However, there is no significant correlation between Dice value and region size.}
+\revised{However, no significant correlation between the Dice values and the region sizes was found
+(Pearson's coefficient = 0.12, p-value = 0.43).}
 
@@
   \revised{
-  The binary images are created based on the intensity differences between result images in CentOS6 and CentOS7. 
-  The voxel is 1 if the intensity value of the corresponding voxel in the images is different, otherwise it is 0.
-  Each voxel shows the number of subjects that created different results. 
-  % The color bar ranges from 1 to 7 as the minimum and the maximum number of subjects in each voxel.
-  Voxels are not colored if the identical results were produced in all 20 subjects.}
+  Each voxel shows the number of subjects for which different results were observed between CentOS 6 and CentOS 7.}
   An animated comparison of segmentations obtained for a particular subject is available
 \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/fs_brain_segmentation.gif}
-{here}.} 
+{here} for better visualization.}
   \label{fig:tissue_class}
@@
 \section{Discussion}
-
-\revision{
-Discuss irreproducibility in 2 parts:
-1. code coverage 
-2. statistics, see eq 1 in https://ieeexplore.ieee.org/abstract/document/7152570
-}
 
@@
 
-\revision{To evaluate the reproducibility of a pipeline, \toolname needs to execute
-it 3 times in order to (1) record a first \reprozip trace, (2) save
-transient files, and (3) compare results in the second condition. This is
-one more execution than the theoretical minimum of 2. It might be possible
-to further reduce this overhead by executing at step (2) only the
-processes depending on transient files.}
+\revised{
+To evaluate the reproducibility of a pipeline, \toolname needs to execute it 5 times 
+in order to (1) record a first \reprozip trace, (2) save transient files in the first condition, 
+(3) compare results in the second condition, and repeat steps (2) and (3) for the other order of execution. 
+It might be possible to further reduce this overhead by executing at step (2) only the processes 
+depending on transient files, and capturing the transient files for the second condition simultaneously at step (3).}
 
@@
 neuroimaging pipelines.
+
+\revised{
+% Furthermore, we can use methods to ensure that enough tests have been performed for labeling a tool as reproducible.
+% We approved the irreproducibility of the tools, but we cannot say a tool is reproducible since
+\toolname labels pipeline processes based on the tested dataset, 
+while the reproducibility of the tools can be changed over the new datasets. 
+In this regard, we need to confirm the reproducibility of the tools over all the possible 
+subject diversity and processing parameters. 
+We can rely on the code coverage to measure the proportion of the code covered by the tests. 
+Ideally, the code coverage as close to 100\% as possible ensures that most of the code is fully tested. 
+This is a key principle of reproducibility to guarantee the appropriate testing of the tools. 
+Having sufficient subject tests so that covers most of the code could increase confidence to confirm 
+the reproducibility of the tools.
+As another solution, we can estimate the probability of the tool reproducibility based on 
+the number of subjects that the tool was tested as reproducible. To this order, 
+we refer to the probability measurement used in~\cite{glatard2015classifications}, which can state that the probability 
+to observe all reproducible results among $N’$ future tests given that $n$ non-reproducible results happened among $N$ 
+previous tests is:
+
+\[p(N',n,N)=\frac{N+1}{N+N'+1}\frac{\binom Nn}{\binom {N+N'}n}\]
+
+This formula gives a high probability to see reproducible results from the tools have been tested on a high series of 
+subjects as reproducible.
+}
 
@@
 \url{https://github.com/big-data-lab-team/spot} under MIT license.
+% Moreover, we registered the spot tool in the bio.tools with the identifier biotools:spottool,
+% and in the SciCrunch.org database with RRID (Research Resource Identifier) SCR\_018915.
 
@@
 uncertainty propagation and amplification in their results.
+
+\revised{
+\section{Availability of Source Code and Requirements}
+\begin{itemize}
+  \item Project name: \toolname
+  \item Project home page:\url{https://github.com/big-data-lab-team/spot}
+  \item Operating system: Linux
+  \item Programming language: Python
+  \item Other requirements: Python 3.6 or higher, \reprozip, Docker, and Boutiques
+  \item License: MIT License
+  \item Biotools identifier: \href{https://bio.tools/spottool}{spottool}
+  \item SciCrunch ID: RRID:SCR\_018915
+
+\end{itemize}
+
+\section{Supplementary Data}
+Supplementary materials include Supplementary Table S1. Summary of the subjects used in the experiments.
+}
 

### 2020-08-20 paper.tex
Update region.py script

@@
                           & \makecell{Mean}  &  \makecell{Standard error}  & \makecell{Mean}   &  \makecell{Standard error}   \\ \midrule
-  Processing time (mins)      &     106.67     &     2.68     &     nd     &     nd     \\
-  Output file size (GB)       &     2.8        &     0.10     &     nd     &     nd     \\
-  Number of file accesses     &     94,089      &     2,645  &     nd     &     nd     \\
-  Number of processes         &     8731       &     198   &     nd     &     nd     \\
+  Processing time (mins)  &     106.67       &     2.68                    &     650.25        &     8.88     \\
+  Output file size (GB)   &     2.8          &     0.10                    &     4.15          &     0.15     \\
+  Number of file accesses &     94,089       &     2,645                   &     62,729        &     984     \\
+  Number of processes     &     8731         &     198                     &     4031          &     47     \\
   \bottomrule

### 2020-08-21 paper.tex
Update revisions

--- 2020-08-12 paper.tex
@@
 \usepackage{pifont}
+\usepackage{pdfpages}
 
@@
 The labeled graph can differ depending on the order of executions in which condition we capture transient files 
-or execute the pipeline to pinpoint the propagation of differences. Therefore, we label processes in two orders 
-of execution, and then aggregate findings so that label processes as non-reproducible (red) if it creates 
-different results in either order of executions.}
+or execute the pipeline to pinpoint the propagation of differences. Therefore, we run the comparison in both condition orders, 
+and we label a process as non-reproducible (red) if it creates different results in at least one condition order.}
 
@@
 \revised{
-\toolname is implemented in Python (>=3.5). In this work we used \toolname version 0.2 and the following version of
+\toolname is implemented in Python (>=3.6). In this work we used \toolname version 0.2 and the following version of
 the Python package dependencies: NumPy v1.19.0~\cite{oliphant2006guide} and Pandas v1.0.5~\cite{mckinney2011pandas},
@@
 We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2 CentOS versions.
-\revised{First, PreFreeSurfer was utilized to analyze the unprocessed data, and then its results 
-from CentOS6 were used as the input of FreeSurfer in both conditions. 
-We also used the ReproZip trace file captured in CentOS6 for labeling the processes in both pipelines.}
+\revised{The PreFreesurfer results obtained in CentOS6 were used as the input of FreeSurfer in both conditions. 
+We also used the \reprozip trace file captured in CentOS6 for labeling the processes in both pipelines.}
 Each subject was processed twice on the same operating
@@
 The Dice coefficient~\cite{dice1945measures} is a commonly used metric to validate medical image segmentation.
-Dice values range from 0 to 1, with 1 indicating a perfect overlap between two segmentation results and 0 indicates no overlap.
+Dice values range from 0 to 1, with 1 indicating a perfect overlap between two segmentation results and 0 indicating no overlap.
 Alternatively, the Jaccard coefficient~\cite{jaccard1912distribution} could be used;
@@
 \revised{
-All experiments were run on a machine with a 3.4GHz Intel Core i7 processor using 8 cores and 32GB of RAM running
-the CentOS7 operating system and Linux kernel version 3.10.
+All experiments were run on a machine with a 3.4GHz, 8-core  Intel Core i7 processor, 32GB of RAM, 
+CentOS 7.3.1611, and Linux kernel version 3.10.
 The processing time, output file size, number of file accesses and number of processes observed in PreFreeSurfer
@@
                           & \makecell{Mean}  &  \makecell{Standard error}  & \makecell{Mean}   &  \makecell{Standard error}   \\ \midrule
-  Processing time (mins)      &     106.67     &     2.68     &     nd     &     nd     \\
-  Output file size (GB)       &     2.8        &     0.10     &     nd     &     nd     \\
-  Number of file accesses     &     94,089      &     2,645  &     nd     &     nd     \\
-  Number of processes         &     8731       &     198   &     nd     &     nd     \\
+  Processing time (mins)  &     106.67       &     2.68                    &     650.25        &     8.88     \\
+  Output file size (GB)   &     2.8          &     0.10                    &     4.15          &     0.15     \\
+  Number of file accesses &     94,089       &     2,645                   &     62,729        &     984     \\
+  Number of processes     &     8,731         &     198                     &     4,031          &     47     \\
   \bottomrule
@@
 a common method to illustrate the magnitude of the differences in registration results.
-Differences appear to be visually important, in particular in the areas framed in red.}
+Differences appear to be visually important, in particular in the areas framed in red, 
+to the point that most experimenters would likely reject such a registration following visual quality control.}
 
@@
 Dice coefficients associated with the 44 regions segmented by
-FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing that Dice coefficents below 0.9 are
+FreeSurfer are shown in Figure~\ref{fig:scatter_plot}, showing that Dice coefficients below 0.9 are
 observed in most regions, and particularly in the smallest ones.
@@
 \revised{
-% Furthermore, we can use methods to ensure that enough tests have been performed for labeling a tool as reproducible.
-% We approved the irreproducibility of the tools, but we cannot say a tool is reproducible since
-\toolname labels pipeline processes based on the tested dataset, 
-while the reproducibility of the tools can be changed over the new datasets. 
-In this regard, we need to confirm the reproducibility of the tools over all the possible 
-subject diversity and processing parameters. 
-We can rely on the code coverage to measure the proportion of the code covered by the tests. 
-Ideally, the code coverage as close to 100\% as possible ensures that most of the code is fully tested. 
-This is a key principle of reproducibility to guarantee the appropriate testing of the tools. 
-Having sufficient subject tests so that covers most of the code could increase confidence to confirm 
-the reproducibility of the tools.
-As another solution, we can estimate the probability of the tool reproducibility based on 
-the number of subjects that the tool was tested as reproducible. To this order, 
-we refer to the probability measurement used in~\cite{glatard2015classifications}, which can state that the probability 
-to observe all reproducible results among $N’$ future tests given that $n$ non-reproducible results happened among $N$ 
-previous tests is:
-
-\[p(N',n,N)=\frac{N+1}{N+N'+1}\frac{\binom Nn}{\binom {N+N'}n}\]
-
-This formula gives a high probability to see reproducible results from the tools have been tested on a high series of 
-subjects as reproducible.
-}
+\toolname only tests pipeline reproducibility in the scope of a particular dataset. However, it is very plausible 
+for pipeline processes to exhibit different reproducibility behaviors when executed on different datasets. 
+Therefore, only the lack of reproducibility of a pipeline process could be guaranteed from an analysis with \toolname, 
+since proving reproducibility would require testing the pipeline on all possible datasets, in all possible environments, 
+which is not feasible. Two elements could be considered in future work to address this issue. 
+First, similar to conventional software testing, a code coverage metric could be developed to assess 
+the fraction of the pipeline code involved in the tested dataset and parameters. This would quantify 
+the representativity of the dataset and pipeline parameters used in the evaluation. Second, statistical 
+risk models could be used to estimate the probability for a process to be reproducible, given a set of 
+observations with no numerical differences. For instance, models similar to the work in~\cite{beauzamy2004methodes} 
+could be leveraged for this purpose.}
 
@@
   \item Biotools identifier: \href{https://bio.tools/spottool}{spottool}
-  \item SciCrunch ID: RRID:SCR\_018915
+  \item SciCrunch ID: \href{https://scicrunch.org/scicrunch/Resources/record/nlx_144509-1/SCR_018915/resolver}{RRID:SCR\_018915}
+  \item DOI: \href{https://doi.org/10.5281/zenodo.3873219}{10.5281/zenodo.3873219}
 
@@
 
+\includepdf[pages=-]{supplementary-files/subject-details.pdf}
+
 \end{document}

### 2020-08-27 paper.tex
minor fixes

@@
 % \newcommand{\note}[2]{\color{blue}Note from \textsc{#1}: #2\color{black}}
-\newcommand{\revision}[1]{\color{purple}\textbf{REVISION:}#1\color{black}}
 \newcommand{\revised}[1]{\color{blue}#1\color{black}\xspace}
@@
 
-% \revision{Using the \reprozip trace obtained from CentOS7,
-% we captured transient files in CentOS6 and labeled pipeline processes in CentOS7.}
 We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2 CentOS versions.
@@
   \item Project name: \toolname
-  \item Project home page:\url{https://github.com/big-data-lab-team/spot}
+  \item Project home page: \url{https://github.com/big-data-lab-team/spot}
   \item Operating system: Linux
-  \item Programming language: Python
-  \item Other requirements: Python 3.6 or higher, \reprozip, Docker, and Boutiques
+  \item Programming language: Python (3.6 or higher)
+  \item Main dependencies: \reprozip, Docker, and Boutiques
+  \item Other dependencies: see \href{https://github.com/big-data-lab-team/spot}{\texttt{setup.py}}
   \item License: MIT License
@@
 \section{Supplementary Data}
-Supplementary materials include Supplementary Table S1. Summary of the subjects used in the experiments.
+Supplementary materials include Supplementary Table S1 that contains a summary of the subjects used in the experiments.
 }
@@
 \section{Acknowledgments}
-
+\revised{We thank the reviewers for their thorough and insightul review which greatly contributed to the improvement 
+of the paper.}
 We warmly thank Compute Canada (\url{http://www.computecanada.ca}) and Calcul
@@
 
-\includepdf[pages=-]{supplementary-files/subject-details.pdf}
+\includepdf[pages=-]{supplementary-files/file1.pdf}
 

### 2020-08-28 paper.tex
update readme and bib files

@@
 therefore barely applicable to complex pipelines. Recently, Monte-Carlo
-Arithmetic~\cite{Parker1997-qq, Denis2016-wo} provided a practical way to
+Arithmetic~\cite{Parker1997-qq, denis2015verificarlo} provided a practical way to
 evaluate the uncertainty of numerical results without the need to rewrite
@@
 FreeSurfer and CIVET~\cite{Lewis2017-ll} or connectome estimation using
-Dipy~\cite{kiar2019comparing}, it is probable that this observation
+Dipy~\cite{kiar2020comparing}, it is probable that this observation
 generalizes widely across toolboxes and requires a deeper investigation of
@@
  through Monte-Carlo Arithmetic~\cite{Parker1997-qq}. The work
- in~\cite{kiar2019comparing} discusses and compares these two techniques.
+ in~\cite{kiar2020comparing} discusses and compares these two techniques.
 
@@
 risk models could be used to estimate the probability for a process to be reproducible, given a set of 
-observations with no numerical differences. For instance, models similar to the work in~\cite{beauzamy2004methodes} 
+observations with no numerical differences. For instance, models described in~\cite{beauzamy2004methodes} 
 could be leveraged for this purpose.}

### 2020-09-10 paper.tex

@@
 
+\revised{The target users of the \toolname tool are primarily pipeline developers and users who have 
+technical skills for creating Docker containers and Boutiques JSON files.}
 We demonstrated the applicability of our approach by evaluating two of the

### 2020-09-11 bin/code/binarized_heatmap_img.py
Update README for generating figures

@@ -18,27 +18,21 @@ def log_error(message):
 
 
 def main():
-    parser = argparse.ArgumentParser(description="Computes a \
-                                     binarized heatmap of the images.")
-    parser.add_argument("input_imgs",
-                        help='Input folder of binarized difference images')
-    parser.add_argument("output_image_file",
-                        help='output folder')
-    parser.add_argument("ref_image_file",
-                        help='reference image like MNI-152')
+    input_imgs = 'data/fs-20sbj-output/in_bin_img/'
+    output_image_file = 'figures/brain_segmentation_mni.png'
+    ref_image_file = 'data/fs-20sbj-output/mni_reference.nii'
 
-    args = parser.parse_args()
     # Load images using nibabel
     img_sum = ''
-    for img in os.listdir(args.input_imgs):
+    for img in os.listdir(input_imgs):
         if os.path.splitext(os.path.basename(img))[1] in ['.nii', '.gz']:
-            img1 = nibabel.load(os.path.join(args.input_imgs, img))
+            img1 = nibabel.load(os.path.join(input_imgs, img))
             data1 = img1.get_data()
             if img_sum == '':
                 img_sum = data1
                 continue
             img_sum = img_sum + data1
-    im_ref = nibabel.load(args.ref_image_file)
+    im_ref = nibabel.load(ref_image_file)
     im_data_ref = im_ref.get_data()
 
     # Check that both images have the same dimensions
@@ -78,7 +72,7 @@ def main():
     hmax3 = sns.heatmap(np.rot90(axi_view), cmap=own_cmap1, xticklabels='', yticklabels='', cbar=False, ax=ax1, vmin=1, vmax=8)
     hmax3.imshow(np.rot90(axi_view_ref), cmap='gray', aspect='equal', extent=hmax3.get_xlim() + hmax3.get_ylim())
     plt.rcParams['axes.facecolor'] = 'black'
-    plt.savefig(args.output_image_file, facecolor=fig.get_facecolor(), bbox_inches='tight')
+    plt.savefig(output_image_file, facecolor=fig.get_facecolor(), bbox_inches='tight')
     # plt.show()
 
 if __name__ == '__main__':

### 2020-09-23 paper.tex
Add the section of Availability of Supporting Data
Remove highlights of the revised sections

--- 2020-08-28 paper.tex
@@
 % \newcommand{\note}[2]{\color{blue}Note from \textsc{#1}: #2\color{black}}
-\newcommand{\revised}[1]{\color{blue}#1\color{black}\xspace}
+% \newcommand{\revised}[1]{\color{blue}#1\color{black}\xspace}
 \newcommand{\reprozip}[0]{ReproZip\xspace}
@@
 
-\revised{Numerical perturbations resulting from variations in computational
-environments} impact data analyses in various fields, but identifying the
+Numerical perturbations resulting from variations in computational
+environments impact data analyses in various fields, but identifying the
 origin of these perturbations in complex pipelines remains challenging.  In
@@
 parallelization parameters~\cite{diethelm2011limits}, result in
-substantially different analysis outcomes, \revised{due to the propagation
+substantially different analysis outcomes, due to the propagation
 and amplification of floating-point errors. While the existence of such
@@
 era, due to the sustained growth of data sets, the increasing complexity of
-analysis pipelines, and the diversification of computing infrastructures.}
+analysis pipelines, and the diversification of computing infrastructures.
 To better understand and
@@
   \inputminted{bash}{"data/example/example.sh"}
-  \caption{\revised{Example pipeline that computes the volume of the brain from a T1 image.}}
+  \caption{Example pipeline that computes the volume of the brain from a T1 image.}
   \label{listing:sample-script}
@@
 to file reads and writes (Figure~\ref{fig:provenance-graph}).
-\revised{We assume that provenance graphs are identical for the \reprozip traces obtained from
-the same subjects in different operating systems.}
+We assume that provenance graphs are identical for the \reprozip traces obtained from
+the same subjects in different operating systems.
 
@@
 
-\revised{
 The labeled graph can differ depending on the order of executions in which condition we capture transient files 
 or execute the pipeline to pinpoint the propagation of differences. Therefore, we run the comparison in both condition orders, 
-and we label a process as non-reproducible (red) if it creates different results in at least one condition order.}
+and we label a process as non-reproducible (red) if it creates different results in at least one condition order.
 
 \subsection{Implementation}
-\revised{
 \toolname is implemented in Python (>=3.6). In this work we used \toolname version 0.2 and the following version of
@@
 Examples in the \toolname documentation include \reprozip provenance capture.
-More documentation on \reprozip is available at \url{https://www.reprozip.org}.}
+More documentation on \reprozip is available at \url{https://www.reprozip.org}.
 
@@
 available in \href{https://db.humanconnectome.org}{the ConnectomDB
-repository} as a subset of the 1200 Subject Release \revised{(see Supplementary Table S1)}.
+repository} as a subset of the 1200 Subject Release (see Supplementary Table S1).
 For each subject, available data consisted of 1 or 2 T1-weighted images and 1 or 2
@@
 We processed the 20 subjects with PreFreeSurfer and FreeSurfer, using the 2 CentOS versions.
-\revised{The PreFreesurfer results obtained in CentOS6 were used as the input of FreeSurfer in both conditions. 
-We also used the \reprozip trace file captured in CentOS6 for labeling the processes in both pipelines.}
+The PreFreesurfer results obtained in CentOS6 were used as the input of FreeSurfer in both conditions. 
+We also used the \reprozip trace file captured in CentOS6 for labeling the processes in both pipelines.
 Each subject was processed twice on the same operating
@@
 
-\revised{
 The Dice coefficient~\cite{dice1945measures} is a commonly used metric to validate medical image segmentation.
@@
 Alternatively, the Jaccard coefficient~\cite{jaccard1912distribution} could be used;
-there is a direct correspondence between both metrics.}
+there is a direct correspondence between both metrics.
 
@@
 \section{Results}
-\revised{
 All experiments were run on a machine with a 3.4GHz, 8-core  Intel Core i7 processor, 32GB of RAM, 
@@
   \end{table*}
-}
+
 % All experiments were run on a machine with a 3.4GHz Intel Core i7 processor
@@
   17 subjects.
-  \revised{Grey cells indicate that the process did not occur in the corresponding pipeline step.}}
+  Grey cells indicate that the process did not occur in the corresponding pipeline step.}
   \label{fig:pfs_heatmap}
@@
     \caption{A complete provenance graph from the PreFreesurfer pipeline.
-    \revised{Node labels use the same abbreviations as in Figure~\ref{fig:pfs_heatmap}.}
+    Node labels use the same abbreviations as in Figure~\ref{fig:pfs_heatmap}.
     For better visualization, processes associated with commands
@@
 
-\revised{
 Figure~\ref{fig:fnirt_result} compares \fnirt results in Brain Extraction
@@
 Differences appear to be visually important, in particular in the areas framed in red, 
-to the point that most experimenters would likely reject such a registration following visual quality control.}
+to the point that most experimenters would likely reject such a registration following visual quality control.
 
@@
     \caption{Differences between T2 \fnirt results in PreFreeSurfer's Brain Extraction (CentOS6 vs CentOS7).
-    \revised{
     The colored squares indicate results obtained with CentOS6 (in purple) and CentOS7 (in green).
-    The red boxes highlight regions with significant differences between the two OSes. }
+    The red boxes highlight regions with significant differences between the two OSes. 
     An animated version of the comparison is available \href{https://github.com/big-data-lab-team/HCP-reproducibility-paper/blob/master/figures/pfs_t2w_alignment.gif}{here} for better visualization.}
@@
 observed in most regions, and particularly in the smallest ones.
-\revised{However, no significant correlation between the Dice values and the region sizes was found
-(Pearson's coefficient = 0.12, p-value = 0.43).}
+However, no significant correlation between the Dice values and the region sizes was found
+(Pearson's coefficient = 0.12, p-value = 0.43).
 
@@
   template.
-  \revised{
-  Each voxel shows the number of subjects for which different results were observed between CentOS 6 and CentOS 7.}
+  Each voxel shows the number of subjects for which different results were observed between CentOS 6 and CentOS 7.
   An animated comparison of segmentations obtained for a particular subject is available
@@
     median values.
-    \revised{Each point represents the Dice coefficient between segmentations of a particular region obtained in CentOS 6 vs CentOS 7 for a given subject.}
+    Each point represents the Dice coefficient between segmentations of a particular region obtained in CentOS 6 vs CentOS 7 for a given subject.
     Boxes brightness is proportional to the logarithm of the corresponding brain region size.}
@@
 
-\revised{
 To evaluate the reproducibility of a pipeline, \toolname needs to execute it 5 times 
@@
 It might be possible to further reduce this overhead by executing at step (2) only the processes 
-depending on transient files, and capturing the transient files for the second condition simultaneously at step (3).}
+depending on transient files, and capturing the transient files for the second condition simultaneously at step (3).
 
+The target users of the \toolname tool are primarily pipeline developers and users who have 
+technical skills for creating Docker containers and Boutiques JSON files.
 We demonstrated the applicability of our approach by evaluating two of the
@@
 
-\revised{
 \toolname only tests pipeline reproducibility in the scope of a particular dataset. However, it is very plausible 
@@
 observations with no numerical differences. For instance, models described in~\cite{beauzamy2004methodes} 
-could be leveraged for this purpose.}
+could be leveraged for this purpose.
 
@@
 
-\revised{
 Computational environments are only one of many factors contributing to the
@@
 reproducibility and replicability should be understood and improved.
-}
 
@@
 
-\revised{
 \section{Availability of Source Code and Requirements}
@@
 
+\section{Availability of Supporting Data}
+Snapshots of our code and other supporting data are openly available in the GigaScience repository, GigaDB~\cite{salari2020global}.
+
 \section{Supplementary Data}
 Supplementary materials include Supplementary Table S1 that contains a summary of the subjects used in the experiments.
-}
 
 \section{Acknowledgments}
-\revised{We thank the reviewers for their thorough and insightul review which greatly contributed to the improvement 
-of the paper.}
+We thank the reviewers for their thorough and insightul review which greatly contributed to the improvement 
+of the paper.
 We warmly thank Compute Canada (\url{http://www.computecanada.ca}) and Calcul
